### 2021-10-19 paper/markdown/figures_and_tables.Rmd
Add files via upload

---
title: "Hyposmia as a standalone persistent symptom of long COVID"
subtitle: "Figures and Tables"
author: "Health after COVID-19 in Tyrol study team"
date: "`r format(Sys.time(), '%Y-%m-%d')`"

output: 
  pdf_document:
    latex_engine: xelatex
bibliography: cov_biblio.bib
csl: american-medical-association.csl
header-includes:
  \usepackage{longtable}
  \usepackage{tabu}
  \usepackage{caption}
  \usepackage{makecell}
  \usepackage{pdflscape}
  \usepackage{array}
  \usepackage{booktabs}
  \usepackage{threeparttable}
  \usepackage{threeparttablex}
  \usepackage{wrapfig}
  \usepackage{multirow}
  \usepackage[normalem]{ulem}
  \usepackage{colortbl}
  \usepackage{xcolor}
  \usepackage{float} \floatplacement{figure}{H} \floatplacement{table}{H}
  \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}}       \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
---

```{r, setup, include = FALSE}

library(bookdown)

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, dev = "cairo_pdf")

```

\newpage

# Tables {#sec:tbl}

```{r tab-covild}

paper_tables$cohort_covild %>% 
  format_line %>%  
  kable(caption = '\\textbf{Characteristic of the CovILD study cohort.}\\\\Ambulatory, Moderate, Severe: severity of acute SARS-CoV-2 infection, Test: statistical test used for the comparison between the severity strata (KW: Kruskal-Wallis), Significance: test p value corrected for multiple comparisons with Benjamini-Hochberg method.', 
        format = 'latex', 
        escape = F, 
        booktabs = T, 
        longtable = T) %>% 
  column_spec(1, width = '3cm') %>% 
  column_spec(2:4, width = '3cm') %>% 
  column_spec(5, width = '0.6cm') %>% 
  column_spec(6, width = '2cm') %>% 
  collapse_rows(columns = 1) %>% 
  kable_styling(full_width = F, 
                latex_options = c('striped', 
                                  'hold_position', 
                                  'repeat_header'))

```

\newpage

```{r tab-survey}

paper_tables$cohort_survey %>% 
  format_line %>%  
  kable(caption = '\\textbf{Characteristic of the Health after COVID-19 in Tyrol study cohorts.}\\\\Test: statistical test used for the comparison between the severity strata (U: Mann-Whitney U test), Significance: test p value corrected for multiple comparisons with Benjamini-Hochberg method.', 
        format = 'latex', 
        escape = F, 
        booktabs = T, 
        longtable = T) %>% 
  column_spec(1, width = '3.5cm') %>% 
  column_spec(2:3, width = '4cm') %>% 
  column_spec(4, width = '0.6cm') %>% 
  column_spec(5, width = '2cm') %>% 
  collapse_rows(columns = 1) %>% 
  kable_styling(full_width = F, 
                latex_options = c('striped', 
                                  'hold_position', 
                                  'repeat_header'))

```

\newpage

# Figures {#sec:figures}

```{r fig-consort, fig.width = mm_inch(paper_figures$consort$w), fig.height = mm_inch(paper_figures$consort$h), dpi = 600, fig.cap = 'CONSORT flow diagrams for the CoVILD and HACT study cohorts.'}

paper_figures$consort$plot

```

__Figure \@ref(fig:fig-consort). CONSORT flow diagrams for the CoVILD and HACT study cohorts.__

\newpage

```{r fig-kinetic, fig.width = mm_inch(paper_figures$kinetic_modeling$w), fig.height = mm_inch(paper_figures$kinetic_modeling$h), dpi = 600, fig.cap = 'Second order kinetic modeling of symptom recovery course in the HACT study.'}

paper_figures$kinetic_modeling$plot

```

__Figure \@ref(fig:fig-kinetic). Second order kinetic modeling of symptom recovery course in the HACT study.__

Symptom frequency during acute COVID-19 and recovery in the Austria and Italy HACT cohorts (0, 7, 14, 28 and 90 days post symptom onset) were modeled with mixed-effect logistic regression (__Supplementary Table S2__). Significance of the first and second order model terms was determined by step-wise likelihood ratio test (LRT). P values were corrected for multiple comparisons with Benjamini-Hochberg method. N number of complete observations is provided under the plots in (__B__). Dim.: diminished, imp.: impaired.

__(A)__ Significance of the first and second order model term in LRT test. Points represent single symptoms, color codes for significance (gray: non-significant). Top 10 symptoms with the most significant second order terms as candidate long-term persistent features are labeled with their names.

__(B)__ Median recovery times for top 10 symptoms with the most significant second order terms as candidate long-term persistent features. Whiskers represent interquartile ranges.

\newpage

```{r fig-hyposmia-kinetic, fig.width = mm_inch(paper_figures$hyposmia_kinetic$w), fig.height = mm_inch(paper_figures$hyposmia_kinetic$h), dpi = 600, fig.cap = 'Actual and predicted frequency of self-reported hyposmia in course of COVID-19 recovery.'}

paper_figures$hyposmia_kinetic$plot

```

__Figure \@ref(fig:fig-hyposmia-kinetic). Actual and predicted frequency of self-reported hyposmia in course of COVID-19 recovery.__

Frequency of self-reported hyposmia during acute COVID-19 and recovery in the HACT study (__A__, 0, 7, 14, 28 and 90 days post symptom onset) and severity subsets of the CovILD cohorts (__B__, acute COVID: 0, 60-, 100 and 180-day follow-up visits) was modeled with mixed-effect logistic regression (__Supplementary Table S2__). Points and solid lines represent the actual frequencies. The model predictions are displayed as dashed lines. The ranges of complete observations per time point are shown under the plots.

\newpage

```{r fig-hyposmia-overlap, fig.width = mm_inch(paper_figures$hyposmia_overlap$w), fig.height = mm_inch(paper_figures$hyposmia_overlap$h), dpi = 600, fig.cap = 'Co-occurrence of self-reported hyposmia and other symptoms in course of COVID-19 recovery.'}

paper_figures$hyposmia_overlap$plot

```

__Figure \@ref(fig:fig-hyposmia-overlap). Co-occurrence of self-reported hyposmia and other symptoms in course of COVID-19 recovery.__

Cosine similarity coefficients between hyposmia and other self-reported symptoms of COVID-19 in the HACT 
(__A__, `r length(hypo_overlap$variables$survey) + 1` symptoms in total, Austria: n = `r hypo_overlap$analysis_tbl$survey$survey_at$ID %>% unique %>% length`, Italy: `r hypo_overlap$analysis_tbl$survey$survey_it$ID %>% unique %>% length`) 
and COVILD study 
(__B__, `r length(hypo_overlap$variables$covild) + 1` symptoms, ambulatory: n = `r hypo_overlap$analysis_tbl$covild$A$ID %>% unique %>% length`, moderate: n = `r hypo_overlap$analysis_tbl$covild$HM$ID %>% unique %>% length`, severe COVID-19: `r hypo_overlap$analysis_tbl$covild$HS$ID %>% unique %>% length`) were calculated for the indicated time points (__Supplementary Table S3__) and presented in radial plots. Each bar represents a single symptom. Bar length and color code for the value of Cosine similarity coefficient. Dashed lines represent cosine similarity coefficient of 0.5.

\newpage

```{r fig-clustering, fig.width = mm_inch(paper_figures$clustering$w), fig.height = mm_inch(paper_figures$clustering$h), dpi = 600, fig.cap = 'Long-term COVID-19 recovery phenotypes in the HACT study.'}

paper_figures$clustering$plot

```

__Figure \@ref(fig:fig-clustering). Long-term COVID-19 recovery phenotypes in the HACT study.__

HACT participants with COVID-19 symptoms present for at least three months post clinical onset (Austria: n = `r nrow(pheno$analysis_tbl$survey_at)`, Italy: n = `r nrow(pheno$analysis_tbl$survey_it)`) were clustered in respect to manifestations present in at least 25% of the cohort for 28 days or longer with a combined algorithm employing self-organizing map (`r pheno$som_grid$xdim` $\times$ `r pheno$som_grid$ydim` hexagonal units, Jaccard distance) and hierarchical clustering (Ward D2 algorithm, Euclidean distance). Numbers of individuals assigned to particular clusters are presented under the plot in (__A__). Imp.: impaired.

(__A__) Presence of the features used for cluster definition in the participants assigned to the 'non-specific', 'hyposmia' and 'neuro/fatigue' subset.

(__B__) Frequency of self-reported hyposmia, taste disorders and tiredness at day in the participant subsets. Statistical significance of the frequency differences was assessed with $\chi^2$ test corrected for multiple comparisons with Benjamini-Hochberg method. P values are presented in the plot captions.

\newpage

```{r fig-risk, fig.width = mm_inch(paper_figures$hyposmia_risk$w), fig.height = mm_inch(paper_figures$hyposmia_risk$h), dpi = 600, fig.cap = 'Risk factors for long-term persistent self-reported post-COVID-19 hyposmia.'}

paper_figures$hyposmia_risk$plot

```

__Figure \@ref(fig:fig-risk). Risk factors for long-term persistent self-reported post-COVID-19 hyposmia.__

Association of candidate risk factors (__Supplementary Table S1__) with long-term persistent self-reported hyposmia in the CovILD (__A__, hyposmia declared at the 100-day follow-up visit) and HACT cohorts (__B__, self-reported hyposmia present for at least three months post clinical onset) was investigated with univariable logistic regression (__Suppmenetary Table S4__). For the HACT cohorts, the models were weighted for age and sex distribution in the general COVID-19 convalescent populations and included stratified the observation time variable as a confounder. Significance of model estimates (odds ratio, OR) was determined by Wald Z test corrected for multiple comparisons with Benjamini-Hochberg method. OR estimate values with 95% confidence intervals for significant and nearly significant (adjusted p < 0.1) co-variates are presented as points and whiskers in the Forest plots. The ranges of complete observations per time point are shown under the plots.

### 2021-10-19 paper/markdown/supplementary_material.Rmd
Add files via upload

---
title: "Hyposmia as a standalone persistent symptom of long COVID"
subtitle: "Supplementary Material"
author: "Health after COVID-19 in Tyrol study team"
date: "`r format(Sys.time(), '%Y-%m-%d')`"

output: 
  pdf_document:
    latex_engine: xelatex
bibliography: cov_biblio.bib
csl: american-medical-association.csl
header-includes:
  \usepackage{longtable}
  \usepackage{tabu}
  \usepackage{caption}
  \usepackage{makecell}
  \usepackage{pdflscape}
  \usepackage{array}
  \usepackage{booktabs}
  \usepackage{threeparttable}
  \usepackage{threeparttablex}
  \usepackage{wrapfig}
  \usepackage{multirow}
  \usepackage[normalem]{ulem}
  \usepackage{colortbl}
  \usepackage{xcolor}
  \usepackage{float} \floatplacement{figure}{H} \floatplacement{table}{H}
  \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}}       \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
---

```{r, setup, include = FALSE}

library(bookdown)

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, dev = "cairo_pdf")

```

\newpage
\beginsupplement

# Supplementary Methods

## Study design and participants

In the current project, the data obtained in the 'Health After COVID-19 in Tyrol' (HACT) online survey [@Sahanic2021; @Hufner2021] and during acute COVID-19, 60-, 100- and 180-day post-COVID-19 follow-up of the CovILD study [@Sonnweber2021; @Sonnweber2020; @Rass2021] were analyzed.

The HACT study collective encompassed two independently recruited cohorts of ambulatory COVID-19 convalescents in the Tyrol state in Austria and the South Tyrol province in Italy. The study inclusion criteria were laboratory-confirmed SARS-CoV-2 infection (PCR or seropositivity), adult age (Austria: $\geq$ 16 years, Italy: $\geq$ 18 years) and residence in the study regions [@Sahanic2021; @Hufner2021]. The analysis inclusion criterion was a minimal observation time of 90 days between the positive SARS-CoV-2 test and the survey completion (__Figure 1__). A total of `r nrow(hyposmia$survey_at) + nrow(hyposmia$survey_it)` questionnaires was eligible for the current analysis (Austria: n = `r nrow(hyposmia$survey_at)`, Italy: `r nrow(hyposmia$survey_it)`). All participants gave digitally signed informed consent before the survey start. The study was approved by the institutional review boards of the Medical University of Innsbruck (Austria, approval
number: 1257/2020) and of the Autonomous Province of Bolzano - South Tyrol (Italy, 0150701). The HACT study was registered at ClinicalTrials.gov (NCT: NCT04661462)

The CovILD cohort included a longitudinally investigated multi-center collective of convalescents of ambulatory (WHO categories 1 - 3, n = `r unique(kinet$analysis_tbl$covild$A$ID) %>% length`), moderate (WHO categories 4 - 5, n = `r unique(kinet$analysis_tbl$covild$HM$ID) %>% length`) and severe COVID-19 (WHO categories 6 - 9, n = `r unique(kinet$analysis_tbl$covild$HS$ID) %>% length`) recruited at the Department of Internal Medicine II at the Medical University of Innsbruck (Austria), St. Vinzenz Hospital in Zams (Austria) and the acute rehabilitation facility in Muenster (Austria) (__Figure 1__) [@Sonnweber2021; @Sonnweber2020; @Rass2021]. The study was approved by the institutional review board at the Medical University of Innsbruck (approval number: 1103/2020), and registered at ClinicalTrials.gov (NCT04416100).

## Measures, definitions and data transformation

In the HACT study, `r length(globals$symptoms)` symptoms of COVID-19 were surveyed. The participants were asked to classify them as 'absent', 'present for 1 - 3 days', 'present for up to 1 week', 'present for up to 2 weeks', 'present for up to 3 months', 'present for up to 6 months' or 'present for more than 6 months'. Based on that responses, individual symptom duration was calculated [@Sahanic2021]. In the CovILD study, `r length(kinet$variables$covild[kinet$variables$covild != 'sympt_present'])` self-reported symptoms were surveyed at each follow-up visit [@Sonnweber2020]. For both studies, detailed characteristic of demography, socioeconomic status, medical history before COVID-19 as well as the disease course and recovery was recorded as described [@Sonnweber2021; @Sonnweber2020; @Rass2021; @Sahanic2021; @Hufner2021]. The full list of the variables available for the current analysis is provided in __Supplementary Table \@ref(tab:tab-study-vars)__.

Self-reported symptoms were classified as acute when present in the fist weeks after clinical manifestation of COVID-19, persistent when lasting for at least 28 days and long-term persistent complaints when present for longer than 3 months (HACT) or present at the 100-day follow-up visit. For analysis, symptom counts and anti-SARS-CoV-2 S1/S2 antibody levels were stratified by quartiles [@Sahanic2021; @Sonnweber2021]. Laboratory readouts of inflammation, iron metabolism and hematopoiesis in the CovILD cohort were longitudinally determined by the local hospital laboratory and binarized with standard cutoffs [@Sonnweber2021; @Sonnweber2020]. Lung function and lung CT abnormality assessment in the CovILD cohort were performed as described [@Sonnweber2020; @Sonnweber2021]. Assessment of hyposmia by the 16 point sniffing stick test at the 100-day visit in the CovILD collective was done essentially as described before [@Rass2021]. For variable stratification scheme, see __Supplementary Table \@ref(tab:tab-study-vars)__.

## Statistical analysis
### Data transformation, visualization, descriptive statistic, hypothesis testing and symptom co-occurrence

Data transformation, analysis and visualization was accomplished with R 4.0.5, _tidyverse_ environment and _cowplot_ package [@Wickham2019; @Wickham2016; @Wilke2019]. As most of the investigated numeric variables were non-normally distributed as assessed by Shapiro-Wilk test, non-parametric tests: Mann-Whitney U and Kruskal-Wallis tests were used to determine statistical significance between analysis groups. Statistical significance for differences in frequency of categorical variables between analysis groups, $\chi^2$ test was applied.

The concordance of sniffing stick test-measured and self-reported hyposmia, was investigated with the Cohen's $\kappa$ statistic (package _vcd_, function _Kappa()_) [@Fleiss1969]. Co-occurrence of self-reported hyposmia and other symptoms (re-coded as 0: absent, 1: present) was measured by cosine similarity coefficient and cosine distance (package _philentropy_, function _distance_) [@Drost2018].

Test and modeling p values for each analysis task and cohort were corrected for multiple testing with Benjamini-Hochberg method [@Benjamini1995].

### Kinetic modeling

Symptom recovery kinetics was modeled by mixed-effect logistic regression (fixed terms: first and second time term, random term: individual, packages _lme4_ and _lmerTest_) [@Bates2015; @Kuznetsova2017] with the following general model formula:

$$Symptom \sim \beta_0 + \beta_1 \times t + \beta_2 \times t^2$$

where $t$ stands for time since symptom onset, $\beta_1$ was interpreted as the estimate of linear recovery and $\beta_2$ was interpreted as a parabolic 'plateau' term [@Box1978]. Significance of the 2^nd^ and 1^st^ order model terms was assessed by step-wise likelihood ratio test versus the lower order model (function _anova()_). Significance of the _beta_ estimates was determined by Wald Z test.

### Definition of the long-term COVID-19 recovery phenotypes

Clusters of HACT study participants with COVID-19 complaints present for at least 3 months (Austria: n = `r nrow(pheno$analysis_tbl$survey_at)`, Italy: n = `r nrow(pheno$analysis_tbl$survey_it)`) were defined in respect to the most frequent persistent features (present in at least 25% of participants for at least 28 days[@Sudre2021; @Sahanic2021]: impaired concentration, tachypnea, hyposmia, hypogeusia, fatigue, tiredness and forgetfulness; re-coded as 0: absent, 1: present). The clustering procedure employed a combined self-organizing map (SOM, `r pheno$som_grid$xdim` $\times$ `r pheno$som_grid$ydim` hexagonal grid, gaussian neighborhood function, Jaccard distance between the symptoms) and hierarchical clustering (HCL, Ward D2 algorithm, manhattan distance between the SOM nodes) [@Kohonen1995; @Vesanto2000; @Vesanto1999; @Hufner2021]. The clustering was accomplished with tools provided by _kohonen_ package [@Wehrens2018] and home-developed functions (https://github.com/PiotrTymoszuk/SOM_tools). The optimal number of final clusters was determined by the bend of the within-sum-of-square curve. Graphical representation of the clustering procedure training and cluster dentition in presented in __Supplementary Figure \@ref(fig:fig-clustering)__,

### Univariable modeling

To identify factors correlating with the risk of persistent long-term self-reported hyposmia (3 months or longer in the HACT study or present at the 100-day follow-up visit in the CovILD cohort) within the set of candidate parameters recorded during acute COVID-19 or, for the CovILD study, at the 60-day follow-up visit as well, a series of univariable logistic models was used. for the HACT study, the models were weighted for age and sex weights based on the age and sex distribution of the general COVID-19 convalescent population in Tyrol/Austria and Italy [@Sahanic2021; @IstitutoSuperiorediSanitaISS; @AustrianAgencyforHealthandFoodSafetyAGES]. To account for the effects of varying diagnosis - survey time in the HACT study, the stratified observation time variable was included as a confounder in each model [@Sahanic]. Odds ratio significance was determined by Wald Z test. Logistic modeling and visual model quality control was accomplished with home-developed wrappers (https://github.com/PiotrTymoszuk/lm_qc_tools) around base R modeling tools and diagnostic functions provided by _broom_ package [@Robinson2014].

# Data availability

The study data is available at reasonable request to the corresponding authors. The source code of the R analysis pipeline is available at GitHub (https://github.com/PiotrTymoszuk/hyposmia_analsis_pipeline).

\newpage

# Supplementary Tables {#sec:tables}

```{r tab-study-vars}

suppl_tables$study_vars[1, 1] %>% 
  mutate(`Variable in R code` = '') %>% 
  kable(caption = '\\textbf{Study variables.}\\\\Variable in R code: variable name in the analyis pipeline, Label: variable name used in plots and tables, Independent modeling variable: indicator if the variable was used for univariable hyposmia risk modeling.\\\\The table is available online.', 
        format = 'latex', 
        escape = T, 
        booktabs = T, 
        longtable = F, 
        col.names = '')

```

```{r tab-kinetic}

suppl_tables$kinetic_modeling[1, 1] %>% 
  mutate(Study = '') %>% 
  kable(caption = '\\textbf{Results of symptom kinetic modeling.}\\\\OR: odds ratio with 95$\\%$ confidence interval, Significance OR: significance of the OR estimate obtained by Benjamini-Hochberg-corrected Wald Z test, Significance LRT: model term significance obtained by Benjamini-Hochberg-corrected likelihood ratio test.\\\\The table is available online.', 
        format = 'latex', 
        escape = T, 
        booktabs = T, 
        longtable = F, 
        col.names = '')

```

```{r tab-similarity}

suppl_tables$hyposmia_similarity[1, 1] %>% 
  mutate(Symptom = '') %>% 
  kable(caption = '\\textbf{Values of cosine similarity coefficients for self-reported hyposmia and other COVID-19 symptoms.}\\\\The table is available online.', 
        format = 'latex', 
        escape = T, 
        booktabs = T, 
        longtable = F, 
        col.names = '')

```

```{r tab-uni-modeling}

suppl_tables$uni_modeling[1, 1] %>% 
  mutate(Study = '') %>% 
  kable(caption = '\\textbf{Univariable modeling of long-term persistent hyposmia risk.}\\\\OR: odds ratio with 95$\\%$ confidence interval, Significance OR: significance of the OR estimate obtained by Benjamini-Hochberg-corrected Wald Z test.\\\\The table is available online.', 
        format = 'latex', 
        escape = T, 
        booktabs = T, 
        longtable = F, 
        col.names = '')

```

\newpage

# Supplementary Figures {#sec:figures}

```{r fig-beta, fig.width = mm_inch(suppl_figures$kinetic_beta$w), fig.height = mm_inch(suppl_figures$kinetic_beta$h), dpi = 600, fig.cap = 'Symptom recovery and chronicity term estimates obtained by kinetic modeling in the HACT study.'}

suppl_figures$kinetic_beta$plot

```

__Supplementary Figure \@ref(fig:fig-beta). Symptom recovery and chronicity term estimates obtained by kinetic modeling in the HACT study.__

Symptom frequency during acute COVID-19 and recovery (0, 7, 14, 28 and 90 days post symptom onset) in the Austria (`r kinet$models$survey$survey_at %>% map_dbl(length) %>% paste_range`) and Italy (`r kinet$models$survey$survey_at %>% map_dbl(length) %>% paste_range`) HACT cohorts were modeled with mixed-effect logistic regression (__Supplementary Table \@ref(tab:tab-kinetic)__). Significance of the OR estimates was determined by Wald Z test. P values were corrected for multiple comparisons with Benjamini-Hochberg method. OR estimate values with 95% confidence intervals for the top 10 symptoms with the most significant second order terms as candidate long-term persistent features are presented as points and whiskers in the Forest plots. Dim.: diminished, imp.: impaired.

\newpage

```{r fig-low-chronicity, fig.width = mm_inch(suppl_figures$low_chronicity$w), fig.height = mm_inch(suppl_figures$low_chronicity$h), dpi = 600, fig.cap = 'Actual and predicted frequency of self-reported shivering and fever as symptoms characterized by an instant recovery in the HACT study.'}

suppl_figures$low_chronicity$plot

```

__Supplementary Figure \@ref(fig:fig-low-chronicity). Actual and predicted frequency of self-reported shivering and fever as symptoms characterized by an instant recovery in the HACT study.__

Frequencies of self-reported shivering (__A__) and fever (__B__) during acute COVID-19 and recovery in the HACT study (0, 7, 14, 28 and 90 days post symptom onset) were modeled with mixed-effect logistic regression (__Supplementary Table \@ref(tab:tab-kinetic)__). Points and solid lines represent the actual frequencies. The model predictions are displayed as dashed lines. The numbers of complete observations are shown under the plots.

\newpage

```{r fig-high-chronicity, fig.width = mm_inch(suppl_figures$high_chronicity$w), fig.height = mm_inch(suppl_figures$high_chronicity$h), dpi = 600, fig.cap = 'Actual and predicted frequency of self-reported fatigue, tiredness and taste disorders as symptoms characterized by a protracted recovery in the HACT study.'}

suppl_figures$high_chronicity$plot

```

__Supplementary Figure \@ref(fig:fig-high-chronicity). Actual and predicted frequency of self-reported fatigue, tiredness and taste disorders as symptoms characterized by a protracted recovery in the HACT study.__

Frequencies of self-reported fatigue (__A__), tiredness (__B__) and hypo- or ageusia (__C__) during acute COVID-19 and recovery in the HACT study (0, 7, 14, 28 and 90 days post symptom onset) were modeled with mixed-effect logistic regression (__Supplementary Table \@ref(tab:tab-kinetic)__). Points and solid lines represent the actual frequencies. The model predictions are displayed as dashed lines. The numbers of complete observations are shown under the plots.

\newpage

```{r fig-hypo-sniffing, fig.width = mm_inch(suppl_figures$hypo_sniffing$w), fig.height = mm_inch(suppl_figures$hypo_sniffing$h), dpi = 600, fig.cap = 'Frequencies and co-occurrence of self-reported and sniffing stick test hyposmia in the CovILD study.'}

suppl_figures$hypo_sniffing$plot

```

__Supplementary Figure \@ref(fig:fig-hypo-sniffing). Frequencies and co-occurrence of self-reported and sniffing stick test hyposmia in the CovILD study.__

Frequencies of self-reported hyposmia, moderate-to-severe (< 13 points) and severe hyposmia (< 8 points) detected by the sniffing stick test in the CovILD study at the 100-day follow-up visit were compared for the convalescents of ambulatory, moderate and severe COVID-19.

__(A, B)__ Frequencies of moderate-to-severe and severe sniffing test hyposmia in the participants declaring self-reported hyposmia or claiming hyposmia-free. Bars are labeled with the frequencies within the investigated subset.

__(C)__ Values of Cohen's $\kappa$ statistic for the concordance of self-reported and sniffing-test hyposmia in the COVID-19 severity strata and the entire CovILD cohort. Whiskers represent 95$\%$ confidence intervals.

\newpage

```{r fig-hypo-knn, fig.width = mm_inch(suppl_figures$hyposmia_knn$w), fig.height = mm_inch(suppl_figures$hyposmia_knn$h), dpi = 600, fig.cap = 'Five closest co-occurring symptoms of the self-reported hyposmia during COVID-19 recovery.'}

suppl_figures$hyposmia_knn$plot

```

__Supplementary Figure \@ref(fig:fig-hypo-knn). Five closest co-occurring symptoms of the self-reported hyposmia during COVID-19 recovery.__

Cosine similarity distances (1 - Cosine similarity coefficient) between hyposmia and other self-reported symptoms of COVID-19 in the HACT 
(__A__, `r length(hypo_overlap$variables$survey) + 1` symptoms in total, Austria: n = `r hypo_overlap$analysis_tbl$survey$survey_at$ID %>% unique %>% length`, Italy: `r hypo_overlap$analysis_tbl$survey$survey_it$ID %>% unique %>% length`) 
and COVILD study 
(__B__, `r length(hypo_overlap$variables$covild) + 1` symptoms, ambulatory: n = `r hypo_overlap$analysis_tbl$covild$A$ID %>% unique %>% length`, moderate: n = `r hypo_overlap$analysis_tbl$covild$HM$ID %>% unique %>% length`, severe COVID-19: `r hypo_overlap$analysis_tbl$covild$HS$ID %>% unique %>% length`) were calculated for the indicated time points (__Supplementary Table \@ref(tab:tab-similarity)__). Cosine distances for five closest co-occurring symptoms of the self-reported hyposmia at the particular time points are presented in radial plots. Line length, point and label color code for cosine distance. Dashed lines represent cosine similarity coefficient of 0.5.

\newpage

```{r fig-clustering, fig.width = mm_inch(suppl_figures$clustering$w), fig.height = mm_inch(suppl_figures$clustering$h), dpi = 600, fig.cap = 'Development of long-term recovery clusters.'}

suppl_figures$clustering$plot

```

__Supplementary Figure \@ref(fig:fig-hypo-knn). Development of long-term recovery clusters.__

HACT participants with COVID-19 symptoms present for at least three months post clinical onset (Austria: n = `r nrow(pheno$analysis_tbl$survey_at)`, Italy: n = `r nrow(pheno$analysis_tbl$survey_it)`) were clustered in respect to manifestations present in at least 25% of the cohort for 28 days or longer with a combined algorithm employing self-organizing map (SOM, `r pheno$som_grid$xdim` $\times$ `r pheno$som_grid$ydim` hexagonal units, Jaccard distance) and hierarchical clustering (HCL, Ward D2 algorithm, Euclidean distance).

__(A)__ Progress of the SOM training procedure visualized as the drop of the mean distance to the winning unit
with the algorithm iterations.

__(B)__ Determination of the optimal cluster number in hierarchical clustering of the SOM nodes by finding the
bend of the total within sum of square curve.

__(C)__ Assignment of the SOM nodes to the clusters defined by hierarchical clustering presented in dendrograms.

\newpage

# References

### 2021-10-19 analysis scripts/chronicity_phenotypes.R

@@ -1,0 +1,178 @@
+# This script investigates the phenotypes of chronic COVID-19 (day 90 on) in the survey study
+# clustering in respect to the symptoms present post day 28 in at least 25% participants
+# participants: only those with symptoms
+
+  insert_head()
+  
+# data container -----
+  
+  pheno <- list()
+  
+# globals: clustering variables, analysis tables -----
+  
+  ## clustering variables
+  
+  pheno$variables <- rec_time$median_duration$survey %>% 
+    map(filter, perc75 >= 28) %>% 
+    map(~.x$symptom) %>% 
+    reduce(union)
+  
+  ## analysis tables
+  
+  pheno$analysis_tbl <- kinet$analysis_tbl$survey %>% 
+    map(select, 
+        ID, 
+        time, 
+        all_of(pheno$variables)) %>% 
+    map(filter, 
+        time == 90) %>% 
+    map(function(x) filter(x, complete.cases(x)))
+  
+  pheno$id_vec <- pheno$analysis_tbl %>% 
+    map(~.x$ID)
+  
+  pheno$analysis_tbl <- pheno$analysis_tbl %>% 
+    map(select, 
+        - time, 
+        - ID) %>% 
+    map(function(x) x %>% 
+          map_dfc(function(y) as.numeric(y) - 1)) %>% 
+    map(data.frame) %>% 
+    map2(., pheno$id_vec, set_rownames)
+  
+  ## filtering the individuals with chronic covid
+  
+  pheno$analysis_tbl <- c('survey_at', 'survey_it') %>% 
+    map(function(x) pheno$id_vec[[x]] %>% 
+          map_dfr(function(y) if(all(pheno$analysis_tbl[[x]][y, ] == 0)) NULL else pheno$analysis_tbl[[x]][y, ])) %>% 
+    set_names(c('survey_at', 'survey_it'))
+    
+  
+# Clustering by the SOM/HCL combi method ------
+  
+  insert_msg('Clustering')
+  
+  ## grid definition, see Vasento et al.
+  
+  pheno$grid_size <- pheno$analysis_tbl %>% 
+    map(ncol) %>% 
+    map_dbl(function(x) 5*sqrt(x)) %>% 
+    min %>% 
+    sqrt %>% 
+    ceiling
+
+  source('./tools/kohonen_tools.R')
+  
+  pheno$som_grid <- somgrid(xdim = pheno$grid_size, 
+                            ydim = pheno$grid_size, 
+                            topo = 'hexagonal', 
[…]

### 2021-10-19 analysis scripts/hyposmia_kappa.R

@@ -1,0 +1,137 @@
+# This script checks for overlap between the sniffing-test hyposmia and the self-reported one in the COVILD cohort
+# Timepoint: day 100
+
+  insert_head()
+  
+# data container ----
+  
+  hypo_kappa <- list()
+  
+# globals -----
+  
+  insert_msg('Globals setup')
+
+  ## analysis tables
+  
+  hypo_kappa$analysis_tbl$cohort <-  hyposmia$covild
+  
+  hypo_kappa$analysis_tbl <- hyposmia$covild %>% 
+    select(ID, 
+           cat_WHO, 
+           anosmia_sympt_V2, ## self-declared hyposmia
+           hyposmia_mod_severe_V2, ## moderate-severe test hyposmia
+           hyposmia_severe) %>%  ## severe test hyposmia
+    dlply(.(cat_WHO)) %>%
+    c(hypo_kappa$analysis_tbl)
+  
+  ## variable labels
+  
+  hypo_kappa$var_labs <- c('hyposmia_mod_severe_V2' = 'Sniffing test < 13 pt', 
+                           'hyposmia_severe' = 'Sniffing test < 8 pt', 
+                           'anosmia_sympt_V2' = 'Self-reported hyposmia')
+
+# calculating the kappas -----
+  
+  insert_msg('Kappa calculation')
+
+  hypo_kappa$kappa <- c('hyposmia_mod_severe_V2', 
+                        'hyposmia_severe') %>% 
+    map(function(x) hypo_kappa$analysis_tbl %>% 
+          map(get_kappa, 
+              variable1 = 'anosmia_sympt_V2', 
+              variable2 = x, 
+              kappa_only = F) %>% 
+          transpose) %>% 
+    set_names(c('hyposmia_mod_severe_V2', 
+                'hyposmia_severe'))
+  
+  ## summary tables
+  
+  hypo_kappa$summary <- hypo_kappa$kappa %>% 
+    map(~.x$kappa) %>% 
+    transpose %>% 
+    map(reduce, rbind) %>% 
+    map2_dfr(., names(.), ~mutate(.x, subset = .y)) %>% 
+    mutate(p_adj = p.adjust(p_value, 'BH'), 
+           kappa = ifelse(kappa < 0, 0, kappa), 
+           plot_lab = paste0(signif(kappa, 2), 
+                             ' [', 
+                             signif(lower_ci, 2), 
+                             ' - ', 
+                             signif(upper_ci, 2), 
[…]

### 2021-10-19 analysis scripts/hyposmia_overlap.R

@@ -1,0 +1,100 @@
+# This script calculates overlap between the self-reported hyposmia and other symptoms
+# as a function of time
+
+  insert_head()
+  
+  library(philentropy)
+  
+# data container -----
+  
+  hypo_overlap <- list()
+  
+# globals: analysis tables an variables ----
+  
+  ## reference variables
+  
+  hypo_overlap$ref_vars <- c('covild' = 'anosmia_sympt', 
+                             'survey' = 'anosmia')
+  
+  ## variables 
+  
+  hypo_overlap$variables <- kinet$variables %>% 
+    map(function(x) x[!x %in% c('sympt_present', 'anosmia_sympt', 'anosmia')])
+  
+  ## analysis tables, recoding the symptoms to numerics (0: absent, 1: present)
+
+  hypo_overlap$analysis_tbl$covild <- kinet$analysis_tbl$covild %>% 
+    map(function(x) x %>% 
+          map_dfc(function(y) if(is.factor(y)) as.numeric(y) - 1 else y))
+  
+  hypo_overlap$analysis_tbl$survey <- kinet$analysis_tbl$survey %>% 
+    map(function(x) x %>% 
+          map_dfc(function(y) if(is.factor(y)) as.numeric(y) - 1 else y))
+  
+# calculation of cosine similarity coefficients, variable translation ------
+  
+  insert_msg('Calculation of cosine distances')
+
+  hypo_overlap$distances <- c('covild', 'survey') %>% 
+    map(function(x) hypo_overlap$analysis_tbl[[x]] %>% 
+          map(get_ref_dist, 
+              ref_var = hypo_overlap$ref_vars[[x]], 
+              overlap_vars = hypo_overlap$variables[[x]], 
+              method = 'cosine')) %>% 
+    set_names(c('covild', 'survey'))
+  
+  ## COVILD
+  
+  hypo_overlap$distances <- map2(hypo_overlap$distances, 
+                                 globals[c('var_lexicon', 'survey_var_lexicon')], 
+                                 function(x, y) map(x, 
+                                                    mutate, 
+                                                    var_label = translate_var(overlap_var, 
+                                                                              time_lab = F, 
+                                                                              dict = y)))
+  
+# visualization of the distances in a time lapse ----
+  
+  insert_msg('Plotting the cosine similarities')
+
+  hypo_overlap$bar_plots <- map2(c('covild', 'survey'), 
+                                 globals[c('sev_labels', 'cohort_labels')], 
[…]

### 2021-10-19 analysis scripts/hyposmia_uni_modeling.R

@@ -1,0 +1,128 @@
+# This script performs univariate logistic modeling of the self-reported and test hyposmia
+
+  insert_head()
+  
+# data container -----
+  
+  uni_model <- list()
+  
+# globals: independent variables and model responses -----
+  
+  insert_msg('Globals setup')
+  
+  ## responses 
+  
+  uni_model$responses$covild <- c('anosmia_sympt_V2', 
+                                  'hyposmia_mod_severe_V2') %>% 
+    set_names(c('self', 'test'))
+    
+  
+  uni_model$responses$survey <- 'anosmia_chronic'
+  
+  ## independent variables
+  
+  uni_model$variables$covild <- globals$variables$new_name
+  
+  uni_model$variables$survey <- globals$survey_var_lexicon %>% 
+    filter(modeling_variable == 'yes', 
+           variable != 'obs_time_strata') %>% 
+    .$variable
+  
+# serial modeling -----
+  
+  insert_msg('Serial modeling')
+  
+  ## covild: seld-reported and test hyposmia
+  
+  uni_model$models$covild <- uni_model$responses$covild %>% 
+    map(make_lm_model, 
+        data = hyposmia$covild, 
+        indep_variable = uni_model$variables$covild, 
+        mod_fun = glm, 
+        family = 'binomial', 
+        est_transf = exp, 
+        error_resistant = T) 
+
+  ## survey
+  
+  uni_model$models$survey <- hyposmia[c('survey_at', 'survey_it')] %>% 
+    map(make_lm_model, 
+        response = uni_model$responses$survey, 
+        indep_variable = uni_model$variables$survey, 
+        weight_variable = 'freq_weight', 
+        confounder = 'obs_time_strata', 
+        mod_fun = glm, 
+        family = 'quasibinomial', 
+        est_transf = exp, 
+        error_resistant = T)
+  
+# model summaries -----
+  
+  insert_msg('Modeling summaries')
+
+  ## summary, all models
+  
+  uni_model$summary <- uni_model$models %>% 
+    map(function(x) x %>% 
+          map(get_model_summary) %>% 
+          map(filter, 
+              variable != 'obs_time_strata') %>% 
+          map(mutate, 
+              significant = ifelse(p_adj < 0.05, 'yes', 'no'), 
+              correlation = ifelse(significant == 'no', 
+                                   'ns', 
+                                   ifelse(estimate > 1, 'positive', 'negative'))))
+  
+  ## variable translation
+  
[…]

### 2021-10-19 analysis scripts/symptom_kinetic.R

@@ -1,0 +1,286 @@
+# This script analyses symptom kinetic in the severity groups in the COVILD study and the Health after COVID cohorts
+
+  insert_head()
+  
+# data container ----
+  
+  kinet <- list()
+  
+# COVILD analysis table and variables ----
+  
+  insert_msg('Setting up the COVILD globals')
+  
+  ## variables 
+  
+  kinet$variables$covild <- globals$var_lexicon %>% 
+    filter(stri_detect(new_name, regex = 'sympt_.*'), 
+           !stri_detect(new_name, fixed = 'number')) %>% 
+    .$new_name
+  
+  ## analysis table
+  
+  kinet$analysis_tbl$covild <- hyposmia$covild %>% 
+    select(ID, 
+           all_of(kinet$variables$covild)) %>% 
+    gather(key = symptom, 
+           value = present,
+           all_of(kinet$variables$covild)) %>% 
+    mutate(time = stri_extract(symptom, regex = 'V.*'), 
+           time = car::recode(time, "'V0' = 0; 'V1' = 60; 'V2' = 100; 'V3' = 180") %>% 
+             as.numeric, 
+           symptom = stri_replace(symptom, regex = '_V.*', replacement = ''), 
+           present = factor(present)) %>% 
+    spread(key = 'symptom',
+           value = 'present')
+  
+  kinet$variables$covild <- stri_replace(kinet$variables$covild, 
+                                         regex = '_V.*', 
+                                         replacement = '') %>% 
+    unique
+
+  ## splitting the analysis table according to the WHO severity
+  
+  kinet$analysis_tbl$covild <- levels(hyposmia$covild$cat_WHO) %>% 
+    map(function(x) left_join(kinet$analysis_tbl$covild, 
+                              hyposmia$covild[c('ID', 'cat_WHO')], 
+                              by = 'ID') %>% 
+          filter(cat_WHO == x)) %>% 
+    set_names(levels(hyposmia$covild$cat_WHO))
+  
+# HACT analysis table and variables -----
+  
+  insert_msg('Setting up the survey globals')
+  
+  ## variables
+  
+  kinet$variables$survey <- globals$symptoms %>% 
+    map(paste, 
+        c('acute', 'immediate', 'subacute', 'long', 'chronic'), 
+        sep = '_') %>% 
+    reduce(c)
+  
+  ## analysis table
+  
[…]

### 2021-10-19 analysis scripts/time_to_recovery.R

@@ -1,0 +1,83 @@
+# This script calculates and compares times to recovery for particular symptoms -----
+
+  insert_head()
+
+# data container ----
+
+  rec_time <- list()
+  
+# globals: symptom variables, analysis tables -----
+  
+  insert_msg('Globals setup')
+  
+  ## symptom variables
+  
+  rec_time$variables <- kinet$variables %>% 
+    map(function(x) x[x != 'sympt_present'])
+  
+  ## analysis tables
+  ## re-coding the time variable to account for the maximal symptom duration according to the questionnaire
+
+  rec_time$analysis_tbl <- kinet$analysis_tbl
+  
+  rec_time$analysis_tbl$survey <- rec_time$analysis_tbl$survey %>% 
+    map(mutate, 
+        time = car::recode(time, "0 = 7; 7 = 14; 14 = 28; 28 = 90; 90 = 180"))
+  
+# serial analysis -----
+  
+  insert_msg('Calculating the times to recovery for particular individuals and symptoms')
+
+  rec_time$sympt_duration <- c('covild', 'survey') %>% 
+    map(function(x) rec_time$analysis_tbl[[x]] %>% 
+          map(function(y) rec_time$variables[[x]] %>% 
+                map(calculate_duration, 
+                    data = y) %>% 
+                reduce(left_join, by = 'ID'))) %>% 
+    set_names(c('covild', 'survey'))
+
+  
+# Calculating median time to recovery -----
+  
+  insert_msg('Median time to recovery for particular symptoms')
+  
+  rec_time$median_duration <- rec_time$sympt_duration %>% 
+    map(function(x) map(x, get_duration_stats, plot = F))
+  
+# visualization as a Forest plots, top factors with the most significant second order terms in kinetic modeling -----
+  
+  insert_msg('Plotting the duration distribution')
+  
+  ## CovILD
+  
+  rec_time$duration_plots$covild <- list(duration_tbl = rec_time$sympt_duration$covild %>% 
+                                           map2(., kinet$top_second_order$covild, 
+                                                ~select(.x, ID, any_of(.y))), 
+                                         plot_title = paste0('CovILD: ', 
+                                                             globals$sev_labels[names(rec_time$sympt_duration$covild)]), 
+                                         plot_tag = paste('\nn =', map_chr(rec_time$sympt_duration$covild, nrow)), 
+                                         color = globals$sev_colors[names(rec_time$sympt_duration$covild)]) %>% 
+    pmap(get_duration_stats, 
+         cust_geom = geom_blank, 
[…]

### 2021-10-19 data_analysis.R

@@ -1,0 +1,25 @@
+# A master script executing the project analyses
+
+  c('./tools/sys_tools.R', 
+    './tools/kinetic_tools.R', 
+    './tools/lm_qc_tools.R', 
+    './tools/project_tools.R',
+    './tools/counting_tools.R', 
+    './tools/clust_tools.R') %>% 
+    walk(source)
+  
+# Executing the analysis scripts -----
+  
+  insert_msg('Executing the analysis scripts')
+  
+  c('./analysis scripts/symptom_kinetic.R',
+    './analysis scripts/hyposmia_kappa.R', 
+    './analysis scripts/hyposmia_overlap.R', 
+    './analysis scripts/hyposmia_uni_modeling.R', 
+    './analysis scripts/time_to_recovery.R', 
+    './analysis scripts/chronicity_phenotypes.R') %>% 
+    walk(source)
+  
+# END ----
+  
+  insert_tail()

### 2021-10-19 data_import.R

@@ -1,0 +1,49 @@
+# This script reads and clears the data from the Health after COVID-19 survey
+# ant the COVILD-19 studies which concern hyposmia and COVID-19 course
+
+# data containers -----
+
+  globals <- list()
+  covild_data <- list()
+  survey_data <- list()
+  hyposmia <- list()
+
+# tools ------
+
+  library(plyr)
+  library(tidyverse)
+  library(foreign)
+  library(readxl)
+
+  c('./tools/sys_tools.R', 
+    './tools/survey_globals.R', 
+    './tools/survey_import_tools.R', 
+    './tools/project_tools.R') %>% 
+    walk(source, 
+         encoding = 'UTF-8')
+  
+# reading the survey and covild data -----
+  
+  c('./import scripts/survey.R', 
+    './import scripts/covild.R') %>% 
+    walk(source, 
+         encoding = 'UTF-8')
+  
+# merging the data sets into a common list ------
+  
+  insert_msg('A common list with the data sets')
+
+  hyposmia[c('covild', 
+             'survey_at', 
+             'survey_it')] <- c(covild_data['mod_tbl'], 
+                                survey_data[c('north', 'south')])
+  
+  
+# END -----
+  
+  rm(covild_data, 
+     survey_data, 
+     model_wrangling, 
+     raw_data)
+  
+  insert_tail()

### 2021-10-19 exec_all.R

@@ -1,0 +1,58 @@
+# This program executes the scripts of the COVID hyposmia project
+
+# libraries -----
+
+  require(plyr)
+  require(tidyverse)
+
+# listing figure scripts ----
+
+  executable_scripts <- c('data_import.R', 
+                          'data_analysis.R', 
+                          'render_paper.R')
+  
+# error container ----
+  
+  exec_errors <- list()
+  
+# executing the script list ----
+  
+  for(file in executable_scripts) {
+    
+    diagn_output <- try(source(file, encoding = 'UTF-8'), silent = T) ## UTF-8 encoding cause of greek letters 
+                                                                      ##in plots and tables
+    
+    if(class(diagn_output) == 'try-error') {
+      
+      exec_errors <- c(exec_errors, 'execution failed')
+      
+    } else {
+      
+      exec_errors <- c(exec_errors, 'execution successful')
+      
+    }
+
+  }
+  
+  exec_errors <- exec_errors %>% 
+    reduce(c) %>% 
+    tibble(execution_result = .)
+  
+  exec_errors <- exec_errors %>% 
+    mutate(file = executable_scripts) %>% 
+    select(file, execution_result)
+  
+  rm(diagn_output, executable_scripts)
+  
+  message(rep('-', 60))
+  print(exec_errors)
+  message(rep('-', 60))
+  
+# saving execution results on the disc ----
+
+  exec_errors %>% 
+    write_tsv('exec_log.log')
+  
+  save.image()
+  
+# END ----

### 2021-10-19 import scripts/covild.R

@@ -1,0 +1,245 @@
+# This script imports the data of the CovILD study
+
+
+  insert_head()
+  
+# Data containers ----
+  
+  prev_visits <- list()
+  neuro_data <- list()
+  qol_data <- list()
+  
+  covild_data <- list()
+
+# Globals ----
+  
+  insert_msg('Globals setup')
+
+  ## graphics
+  
+  globals$corr_colors <- c('negative' = 'steelblue4', 
+                           'positive' = 'firebrick4', 
+                           'ns' = 'gray60')
+  
+  globals$common_text <- element_text(size = 8, face = 'plain', color = 'black')
+  
+  globals$common_margin <- ggplot2::margin(t = 4, l = 3, r = 2, unit = 'mm')
+  
+  globals$common_theme <- theme_classic() + theme(axis.text = globals$common_text, 
+                                                  axis.title = globals$common_text, 
+                                                  plot.title = element_text(size = 8, 
+                                                                            face = 'bold', 
+                                                                            color = 'black', 
+                                                                            hjust = 0), 
+                                                  plot.subtitle = globals$common_text, 
+                                                  plot.tag = element_text(size = 8, 
+                                                                          face = 'plain', 
+                                                                          color = 'black', 
+                                                                          hjust = 0), 
+                                                  plot.tag.position = 'bottom', 
+                                                  legend.text = globals$common_text, 
+                                                  legend.title = globals$common_text, 
+                                                  strip.text = globals$common_text,
+                                                  strip.background = element_rect(fill = 'gray95', color = 'gray80'), 
+                                                  plot.margin = globals$common_margin)
+  
+  ## patient severity group colors and labels
+  
+  globals$sev_colors <- c(A = 'steelblue', 
+                          HM = 'indianred2', 
+                          HS = 'firebrick4', 
+                          cohort = 'coral3')
[…]

### 2021-10-19 import scripts/data_clearing_survey.R

@@ -1,0 +1,255 @@
+# this script performs additional data wrangling tasks: stratification of the symptom sums
+# for the risk modeling (acute and persistent)
+# source of weights: AGES (https://covid19-dashboard.ages.at/dashboard.html), Ministerio di Saluto
+# (https://www.epicentro.iss.it/coronavirus/bollettino/Bollettino-sorveglianza-integrata-COVID-19_7-luglio-2021.pdf)
+
+  insert_head()
+  
+# data container -----
+  
+  model_wrangling <- list()
+
+# acute symptom sum stratification by quartiles, merging with the main study tables ----
+  
+  insert_msg('Stratification of the acute symptom sum')
+  
+  model_wrangling$sympt_sum_tbls$acute <- survey_data %>% 
+    map(strat_quartile, 
+        numeric_variable = 'sum_symptoms_acute', 
+        new_var_name = 'sum_symptoms_acute_class',
+        labels = c('1Q', '2Q', '3Q', '4Q'))
+  
+# long symptom sum stratification by median and 75th percentile -----
+  
+  insert_msg('Stratification of the persistent symptom sum')
+  
+  model_wrangling$sympt_sum_tbls$long <- survey_data %>% 
+    map(strat_quartile, 
+        numeric_variable = 'sum_symptoms_long', 
+        quant_vec = c(0.5, 0.75), 
+        new_var_name = 'sum_symptoms_long_class',
+        labels = c('0', '1 - 3', '>3'))
+  
+# merging with the main table ------
+  
+  insert_msg('Merging with the modeling tables')
+  
+  survey_data <- map2(survey_data, 
+                   model_wrangling$sympt_sum_tbls$acute,
+                   left_join, 
+                   by = 'ID') %>% 
+    map2(., 
+         model_wrangling$sympt_sum_tbls$long, 
+         left_join, 
+         by = 'ID')
+  
+# reading the data and defining the weights based on age and sex of all Tyrolean and Italian convalescents ------
+  
+  insert_msg('Defining the modeling weights')
+
+  ## north
+  
+  model_wrangling$weights$north <- read_delim('./input data/CovidFaelle_Altersgruppe.csv', 
+                                              delim = ';')  %>% 
+    filter(Bundesland == 'Tirol') %>% 
+    mutate(age_strata = Altersgruppe , 
+           age_weight_class = paste('age', AltersgruppeID , sep = '_'), 
[…]

### 2021-10-19 import scripts/survey.R

@@ -1,0 +1,501 @@
+# This script reads and clears the complete survey data
+# For the North and South Tyrol cohorts
+
+  insert_head()
+
+# data containers -----
+  
+  raw_data <- list()
+  survey_data <- list()
+  
+  consort_data <- list()
+
+# reading the spss data -----
+  
+  insert_msg('Reading the SPSS data tables')
+  
+  raw_data[c('north', 
+             'south')] <- c('./input data/nordtirol_05072021.sav', 
+                            './input data/suedtirol_05072021.sav') %>% 
+    map(read.spss) %>% 
+    map(as_tibble)
+  
+# Clearing the demographics data -----
+  
+  insert_msg('Clearing the demographic data')
+  
+  survey_data$demo_tbl <- list(raw_survey = raw_data, 
+                            form_fill_var = c('creation_timestamp_gesundheit_nach_covid_19_1_0', 
+                                              'creation_timestamp_gesundheit_nach_covid_19__suedtirol__@'), 
+                            responder_var = c('ausf_llende_person', 
+                                              'gesundheit_nach_covid_19__suedtirol__1_0_item_001'), 
+                            region_var = c('bezirk', 
+                                           'gesundheit_nach_covid_19__suedtirol__1_0_item_005'), 
+                            lang_var = c('German', 
+                                         'gesundheit_nach_covid_19__suedtirol__1_0_item_004'), 
+                            age_var = c('gesundheit_nach_covid_19_1_0_item_002', 
+                                        'gesundheit_nach_covid_19__suedtirol__1_0_item_002'), 
+                            cohabitant_var = c('gesundheit_nach_covid_19_1_0_item_005', 
+                                               'gesundheit_nach_covid_19__suedtirol__1_0_item_006'), 
+                            pos_cohabitant_var = c('gesundheit_nach_covid_19_1_0_item_006', 
+                                                   'gesundheit_nach_covid_19__suedtirol__1_0_item_007'), 
+                            height_var = c('gesundheit_nach_covid_19_1_0_item_007', 
+                                           'gesundheit_nach_covid_19__suedtirol__1_0_item_008'), 
+                            w_before_var = c('gesundheit_nach_covid_19_1_0_item_008', 
+                                             'gesundheit_nach_covid_19__suedtirol__1_0_item_009'), 
+                            w_recent_var = c('gesundheit_nach_covid_19_1_0_item_009', 
+                                             'gesundheit_nach_covid_19__suedtirol__1_0_item_010'),
[…]

### 2021-10-19 paper scripts/cohort_characteristic.R

@@ -1,0 +1,138 @@
+# This script stitches together the cohort characteristic 
+
+
+  insert_head()
+  
+# data container ----
+  
+  cohort <- list()
+  
+# globals: variables of interest -----
+  
+  insert_msg('Globals setup')
+  
+  ## CovILD
+  
+  cohort$variables$covild <- c('sex', 
+                               'age', 
+                               'age_class',
+                               'smoking', 
+                               'comorb_present', 
+                               'no_comorb', 
+                               'weight_class',
+                               'cardiovascular_comorb', 
+                               'hypertension_comorb', 
+                               'pulmonary_comorb', 
+                               'endometabolic_comorb', 
+                               'hyperchol_comorb', 
+                               'diabetes_comorb', 
+                               'malingancy_comorb', 
+                               'sympt_present_V0', 
+                               'sympt_number_V0')
+  
+  ## Survey
+  
+  cohort$variables$survey <- c('obs_time', 
+                               'obs_time_strata', 
+                               'sex', 
+                               'age', 
+                               'age_class', 
+                               'smoking', 
+                               'comorb_absent', 
+                               'comorb_sum', 
+                               'bmi_class_before', 
+                               'heart_circulation', 
+                               'hypertension', 
+                               'lung', 
+                               'hay_fever', 
+                               'diabetes', 
+                               'malignancy', 
+                               'acute_covid', 
+                               'sum_symptoms_acute', 
+                               'sum_symptoms_acute_class')
+  
+# serial analysis -----
+  
+  insert_msg('Serial analysis')
+  
+  ## covild
+  
+  cohort$analysis$covild <- cohort$variables$covild %>% 
+    map(analyze_feature, 
+        inp_tbl = hyposmia$covild, 
+        split_var = 'cat_WHO') %>% 
+    set_names(cohort$variables$covild)
+  
+  ## survey
+  
+  cohort$analysis$survey <- cohort$variables$survey %>% 
+    map(analyze_feature, 
+        inp_tbl = map2_dfr(hyposmia[c('survey_at', 'survey_it')],
+                           c('Austria', 'Italy'), 
+                           ~mutate(.x, cohort = .y)) %>% 
+          mutate(cohort = factor(cohort)),  
+        split_var = 'cohort') %>% 
+    set_names(cohort$variables$survey)
[…]

### 2021-10-19 paper scripts/deploy_paper.R

@@ -1,0 +1,30 @@
+# this script renders a word file with the figures and supplementary material
+
+  insert_head()
+
+# tools -----
+
+  library(bookdown)
+  library(knitr)
+  library(rmarkdown)
+  library(kableExtra)
+  
+# rendering the figures and tables ------
+
+  insert_msg('Rendering the figures and tables')
+  
+  render('./paper/markdown/figures_and_tables.Rmd', 
+         output_format = pdf_document2(number_sections = F), 
+         output_dir = './paper/') 
+
+# rendering the supplementary material -----
+  
+  insert_msg('Rendering the supplementary material')
+  
+  render('./paper/markdown/supplementary_material.Rmd', 
+         output_format = pdf_document2(number_sections = F), 
+         output_dir = './paper/') 
+  
+# END ------
+  
+  insert_tail()

### 2021-10-19 paper scripts/paper_figures.R

@@ -1,0 +1,388 @@
+# This script generates paper figures 
+
+  insert_head()
+  
+# data containers -----
+  
+  paper_figures <- list()
+  suppl_figures <- list()
+  
+# Figure 1: CONSORT ----
+  
+  insert_msg('Figure 1: CONSORT diagrams')
+  
+  paper_figures$consort <- plot_grid(ggdraw() + 
+                                       draw_image('./study consort/consort_diagram.png')) %>% 
+    as_figure_object(figure_label = 'figure_1_consort', 
+                     w = 180, 
+                     h = 100)
+  
+# Figure 2: COVID-19 symptom presence and resolution time -----
+  
+  insert_msg('Figure 2: symptom presence and resolution time')
+  
+  paper_figures$kinetic_modeling <- plot_grid(plot_grid(kinet$lrt_summ_plots$survey$survey_at + 
+                                                          theme(plot.subtitle = element_blank()) + 
+                                                          labs(title = 'HACT: Austria') + 
+                                                          expand_limits(x = 320), 
+                                                        kinet$lrt_summ_plots$survey$survey_it + 
+                                                          theme(plot.subtitle = element_blank()) + 
+                                                          labs(title = 'HACT: Italy') + 
+                                                          expand_limits(x = 320), 
+                                                        ncol = 2, 
+                                                        align = 'hv'), 
+                                              plot_grid(rec_time$duration_plots$survey$survey_at + 
+                                                          scale_x_continuous(limits = c(0, 100)),
+                                                        rec_time$duration_plots$survey$survey_it + 
+                                                          scale_x_continuous(limits = c(0, 100)), 
+                                                        ncol = 2, 
+                                                        align = 'hv'), 
+                                              nrow = 2, 
+                                              rel_heights = c(0.45, 0.55), 
+                                              labels = LETTERS, 
+                                              label_size = 10) %>% 
+    as_figure_object(figure_label = 'figure_2_kinetic_modeling', 
+                     w = 180, 
+                     h = 165)
+  
+# Figure 3: Kinetic plots for self-reported hyposmia -----
+  
+  insert_msg('Figure 3: knetics of self-reported hyposmia')
+  
+  paper_figures$hyposmia_kinetic <- plot_grid(kinet$kinet_plots$survey$survey_at$anosmia + 
+                                                labs(title = 'HACT: Austria, hypo/anosmia') + 
+                                                theme(legend.position = 'none') + 
[…]

### 2021-10-19 paper scripts/paper_tables.R

@@ -1,0 +1,190 @@
+# This script generates supplementary tables 
+
+  insert_head()
+  
+# data container -----
+  
+  paper_tables <- list()
+  suppl_tables <- list()
+  
+# Tables 1 and 2: characteristic of the CovILD and HACT cohorts ------
+  
+  insert_msg('Tables 1 and 2: study cohort characteristic')
+  
+  paper_tables[c('cohort_covild', 
+                 'cohort_survey')] <- cohort$raw_tbl %>% 
+    map2(., list(c('Parameter', 'Ambulatory', 'Moderate', 'Severe', 'Test', 'Significance'), 
+                 c('Parameter', 'Austria', 'Italy', 'Test', 'Significance')), 
+         set_names)
+  
+  paper_tables$cohort_covild <- paper_tables$cohort_covild %>% 
+    mutate(Test = ifelse(Test == 'U', 'KW', Test))
+  
+# Supplementary Table S1: study variables -----
+  
+  insert_msg('Table S1: study variables')
+  
+  suppl_tables$study_vars$survey <- globals$survey_var_lexicon %>% 
+    mutate(study = 'HACT', 
+           timepoint = NA)
+  
+  suppl_tables$study_vars$covild <- globals$var_lexicon %>% 
+    mutate(variable = new_name, 
+           timepoint = car::recode(reference, 
+                                   "0 = 'acute COVID-19'; 
+                                   1 = '60 day FUP'; 
+                                   2 = '100 day FUP'; 
+                                   3 = '180 day FUP'; 
+                                   4 = '100 day FUP'; 
+                                   5 = '100 day FUP'"), 
+           study = 'CovILD', 
+           description = label, 
+           cutpoints = NA, 
+           levels = NA, 
+           modeling_variable = ifelse(var_type != 'independent' | is.na(var_type), 'no', 'yes'))
+  
+  suppl_tables$study_vars <- suppl_tables$study_vars %>% 
+    map(select, 
+        variable, 
+        label, 
+        unit, 
+        description, 
+        cutpoints, 
+        levels, 
+        modeling_variable, 
+        study, 
+        timepoint) %>% 
+    reduce(rbind) %>% 
+    set_names('Variable in R code', 
+              'Label', 
+              'Unit', 
+              'Description', 
+              'Cutoffs', 
+              'Levels', 
+              'Independent modeling variable', 
+              'Study', 
+              'Time point')
+
+# Supplementary Table S2: results of kinetic modeling -----
+  
[…]

### 2021-10-19 render_paper.R

@@ -1,0 +1,25 @@
+# This script generates figures, tables and a report file
+
+  c('./tools/sys_tools.R',
+    './tools/plotting_tools.R', 
+    './tools/counting_tools.R', 
+    './tools/project_tools.R') %>% 
+    walk(source)
+  
+  library(writexl)
+  
+  insert_head()
+  
+# Rendering scripts -----
+  
+  insert_msg('Executing the paper scripts')
+  
+  c('./paper scripts/paper_figures.R', 
+    './paper scripts/cohort_characteristic.R', 
+    './paper scripts/paper_tables.R', 
+    './paper scripts/deploy_paper.R') %>%  
+    walk(source)
+  
+# END ----
+  
+  insert_tail()

### 2021-10-19 tools/clust_tools.R

@@ -1,0 +1,494 @@
+## a script providing tools for hierarchical and kmeans or PAM clustering
+
+# tools -----
+
+  require(tidyverse)
+  require(philentropy)
+  require(factoextra)
+  require(dendextend)
+  require(scrime)
+  require(cluster)
+  require(pcaPP)
+  require(dbscan)
+
+# helper functions -----
+
+  calculate_dist <- function(inp_tbl, method) {
+    
+    ## calculates a dissimilarity matrix
+    
+    if(method == 'smc') {
+      
+      dist_mtx <- inp_tbl %>% 
+        as.matrix %>% 
+        smc(dist = T)
+      
+      
+    } else {
+      
+      dist_mtx <- inp_tbl %>% 
+        as.matrix %>% 
+        distance(method = method, 
+                 use.row.names = T)
+      
+      if(method %in% c('cosine', 'ruzicka')) {
+        
+        ### handiling the 0 - 1 similarity coefficients
+        
+        dist_mtx <- 1 - dist_mtx
+        
+      }
+      
+    }
+    
+    return(dist_mtx)
+    
+  }
+  
+  plot_knn_distance <- function(diss_obj, k, eps = NULL) {
+    
+    ## makes k-NN distance plot
+    
+    sort_distances <- kNNdist(x = diss_obj, 
+                              k = k) %>% 
+      sort %>% 
+      tibble(sample = names(.), 
+             knn_dist = unname(.)) %>% 
+      select(sample, 
+             knn_dist) %>% 
+      mutate(point = 1:nrow(.))
+    
+    dist_plot <- sort_distances %>% 
+      ggplot(aes(x = point, 
+                 y = knn_dist)) + 
+      geom_area(color = 'cornsilk4', 
+                fill = 'cornsilk2') + 
+      theme_classic() + 
+      labs(x = 'Sample', 
+           y = paste(k, 'NN distance', sep = '-'))
+    
+    if(!is.null(eps)) {
+      
+      dist_plot <- dist_plot + 
+        geom_hline(yintercept = eps, 
+                   linetype = 'dashed', 
+                   color = 'black')
+      
+    }
+    
+    return(dist_plot)
+    
+  }
+
+# dendrogram plotting -----
+
+  plot_dendro <- function(clust_str, k, labels = T, cluster_colors = NULL, cluster_labels = NULL, 
[…]

### 2021-10-19 tools/counting_tools.R

@@ -1,0 +1,1078 @@
+# This simple script contains some tools for counting variable levels
+# And calculating the characteristic of numeric variables
+
+
+# libraries -----
+
+  require(plyr)
+  require(tidyverse)
+  require(stringi)
+  require(ggsignif)
+  require(ggpubr)
+  
+# functions for factors -------
+
+  complete_cases <- function(inp_tbl, index_var = 'ID') {
+    
+    ## excludes individuals with the incomplete longitudinal record
+    
+    out_tbl <- inp_tbl %>% 
+      dlply(index_var, function(x) if(all(complete.cases(x))) x else NULL) %>% 
+      compact %>% 
+      reduce(rbind) %>% 
+      as_tibble
+    
+    return(out_tbl)
+    
+  }
+
+  count_feature <- function(inp_tbl, var_to_count, remove_na = T, .drop = T) {
+    
+    ## calculates the percentage and number of participants with/without the given feature
+    ## the .drop argument specifies if the empty factor levels will be included in the output
+    
+    if(remove_na) {
+      
+      count_tbl <- inp_tbl %>% 
+        filter(!is.na(.data[[var_to_count]]), 
+               .data[[var_to_count]] != 'no_answer')
+      
+    } else {
+      
+      count_tbl <- inp_tbl
+      
+    }
+    
+    feature_counts <- count_tbl %>% 
+      count(.data[[var_to_count]], 
+            .drop = .drop) %>% 
+      mutate(percent = n/sum(n) * 100, 
+             total_n = sum(n))
+    
+    return(feature_counts)
+    
+  }
+
+  count_feature_lst <- function(inp_tbl, var_to_count_vec, remove_na = T, positive_only = T) {
+    
+    ## wrapper for variable vectors
+    
+    count_tbl <- var_to_count_vec %>% 
+      map(count_feature, 
+          inp_tbl = inp_tbl, 
+          remove_na = remove_na) %>% 
+      map(set_names, 
+          c('feature_strata', 
+            'n', 
+            'percent', 
+            'total_n')) %>% 
+      set_names(var_to_count_vec) %>% 
+      map2_dfr(., names(.), 
+               function(x, y) mutate(x, feature = y))
+    
+    if(positive_only) {
+      
+      count_tbl <- count_tbl %>% 
+        filter(feature_strata == 'yes')
+      
+    }
+    
+    return(count_tbl)
+    
+  }
+  
[…]

### 2021-10-19 tools/kinetic_tools.R

@@ -1,0 +1,1016 @@
+# This scrip provided tool for first and second term linear modling of the variable's kinetics
+
+# toolbox -----
+
+  require(tidyverse)
+  require(rlang)
+  require(lme4)
+  require(lmerTest)
+  require(stringi)
+  require(ggrepel)
+  require(furrr)
+
+# helper functions -----
+
+  identity_matrix_ <- function(dimension) {
+    
+    ## creates an identity matrix with a given dimension
+    
+    out_matrix <- matrix(0, dimension, dimension)
+    
+    for (i in 1:dimension) {
+      
+      out_matrix[i, i] <- 1
+      
+    }
+    
+    return(out_matrix)
+    
+  }
+
+  get_lme_results_ <- function(model_mix, L_matrix = NULL, anova = F, ...){
+    
+    require(lmerTest)
+    
+    ## a function returning confidence intervals and significances for particular coefficents of the mixed model with one fixed
+    ## effect. The coefficients to display are defined by the L_matrix. Uses the contest function from the lmerTest package
+    ## If the L_matrix is not given, stats for all coefficients are calculated
+
+    if(is.null(L_matrix)) {
+      
+      L_matrix <- identity_matrix_(length(model_mix@beta))
+      
+    }
+    
+    if(anova) {
+      
+      output <- contest(model_mix, L_matrix, ...)
+      
+    } else {
+      
+      output <- contest(model_mix, L_matrix, joint = F, ...) %>% 
+        as_tibble
+
+      coef_names <- rownames(summary(model_mix)$coefficients) %>% 
+        as_tibble
+      
+      output <- cbind(coef_names, output) %>%
+        set_names(c('parameter', 
+                    'estimate', 
+                    'se', 
+                    'df', 
+                    't', 
+                    'lower_ci', 
+                    'upper_ci', 
+                    'p_value'))
+      
+    }
+    
+    return(output)
+  }
+  
+  get_glm_results_ <- function(glm_model, exponentiate = F, default_signif = 2) {
+    
+    # retrieves regression estimate stats from a glm
+    
+    ## model coefficients
+    
+    coefs <- glm_model %>% 
+      summary %>% 
+      coefficients %>% 
+      data.frame()
+    
+    estimates <- coefs[, c('Estimate', 'Std..Error')] %>% 
+      as_tibble %>% 
+      set_names(c('estimate', 'se'))
+    
[…]

### 2021-10-19 tools/project_tools.R

@@ -1,0 +1,1401 @@
+# A medley of functional project tools
+
+  require(plyr)
+  require(tidyverse)
+  require(furrr)
+  require(caret)
+  require(doParallel)
+  require(randomForest)
+  require(grDevices)
+  require(pcaPP)
+  require(rlang)
+  require(cowplot)
+  require(gmodels)
+  require(vcd)
+  
+# data import and transformation -----
+  
+  recode_var <- function(data, old_var, new_var, ID_var = 'ID', time_var = NULL, trans_fun = 'my_identity', args = NULL) {
+    
+    ## renames a variable, if time and id variable names provided, they're included
+    ## in  the output as well
+    ## the transformation function enables
+    ## args specifies the arguments to the transforming function
+    
+    trans_call <- call2(trans_fun, 
+                        data[[old_var]], 
+                        !!!args)
+
+    new_data <- data %>% 
+      mutate(!!ensym(new_var) := eval(trans_call))
+
+    if(!is.null(time_var)) {
+      
+      return(new_data[c(enexpr(new_var), ID_var, time_var)])
+      
+    } else {
+      
+      return(new_data[c(enexpr(new_var), ID_var)])
+      
+    }
+    
+  }
+  
+  recode_vec <- function(vector, recodes, ...) {
+    
+    return(car::recode(vector, 
+                       recodes = recodes, 
+                       ...))
+    
+  }
+  
+  recode_yn <- function(vector, reverse = F, as.factor = T, ...) {
+    
+    ## recodes a 0/1 vector no/yes or the other way round
+    
+    if(reverse) {
+      
+      new_vec <- recode_vec(vector, 
+                            "'yes' = 1;
+                             'no' = 0",
+                            as.factor = as.factor)
+      
+      #if(!as_factor) {
+        
+       # new_vec <- as.numeric(new_vec)
+        
+      #}
+      
+    } else {
+      
+      new_vec <- recode_vec(vector, 
+                            "1 = 'yes';
+                             0 = 'no'", 
+                            as.factor = as.factor)
+      
+    }
+    
+    #if(as_factor) {
+      
+     # new_vec <- factor(new_vec)
+      
+    #}
+    
+    return(new_vec)
+    
+  }
+  
+  binarize <- function(vector, cutoff, labels = "'yes';'no'") {
+    
+    return(cut(vector, 
+               c(-Inf, cutoff, Inf), 
+               unlist(stri_split_fixed(labels, ';'))))
+    
+    
[…]

### 2021-10-19 tools/survey_globals.R

@@ -1,0 +1,112 @@
+# This script provides globals and import functions 
+
+# toolbox ----
+
+  source('./tools/sys_tools.R')
+
+  library(readr)
+  library(readxl)
+  library(tibble)
+  library(ggplot2)
+  library(stringi)
+
+# Co-morbidities -----
+  
+  globals$comorb <- c('autoimmunity', 
+                      'anemia', 
+                      'hypertension', 
+                      'depression_burnout',
+                      'diabetes', 
+                      'pre_cov_epilepsy',
+                      'dementia', 
+                      'surgery_6m_before', 
+                      'frequent_flu_like', 
+                      'heart_circulation', 
+                      'hay_fever', 
+                      'malignancy', 
+                      'pins_needles_feet', 
+                      'gastrointenstinal', 
+                      'lung', 
+                      'two_plus_infections_antibiotics', 
+                      'multiple_sclerosis', 
+                      'night_dyspnoe', 
+                      'bruxism', 
+                      'kidney', 
+                      'parkinson', 
+                      'insomnia', 
+                      'stroke', 
+                      'embolism', 
+                      'transplantation')
+
+# Daily medication -----
+  
+  globals$daily_medication <- c('daily_immunosuppression', 
+                                'daily_cortison', 
+                                'daily_ace_drugs', 
+                                'daily_pain_killers', 
+                                'daily_anti_coagulation')
+  
+# self-reported symptoms -------
+  
+  globals$symptoms <- c('fever', 
+                        'ague', 
+                        'sore_throat', 
+                        'running_nose', 
+                        'fatigue', 
+                        'dry_cough', 
+                        'wet_cough', 
+                        'breath_short', 
+                        'dyspnoe', 
+                        'chest_pain', 
+                        'tachycardia', 
+                        'extrasystole', 
+                        'joint_pain', 
+                        'bone_pain', 
+                        'muscle_pain', 
+                        'abdominal_pain', 
+                        'nausea', 
+                        'vomiting', 
+                        'dim_appetite', 
+                        'diarrhea', 
+                        'dizziness', 
+                        'headache', 
+                        'anosmia', 
+                        'taste_loss', 
+                        'confusion', 
+                        'tingle_feet', 
+                        'tingle_hands', 
+                        'ache_feet', 
+                        'ache_hands', 
+                        'numb_feet', 
[…]

### 2021-10-19 tools/survey_import_tools.R

@@ -1,0 +1,1518 @@
+# this script provides tools for wrangling the survey data
+
+# tools ------
+
+  require(plyr)
+  require(tidyverse)
+
+# basic formatting functions -----
+
+  pss2date <- function(x) {
+    
+    ## converts the SPSS date to R
+    
+    return(as.Date(x/86400, origin = '1582-10-14') )
+    
+  }
+  
+  rm_space <- function(x) {
+    
+    ## removes extra trailing spaces
+    
+    return(stri_replace(x, 
+                        regex = '\\s+$', 
+                        replacement = ''))
+    
+  }
+  
+  format_na_no <- function(x) {
+    
+    ## formats the comorbidity (and other data) data
+    ## where NA is assumed 'no'
+    
+    return(ifelse(is.na(x), 'no', 'yes') %>% 
+             factor(c('no', 'yes')))
+    
+    
+  }
+  
+  yn_variable <- function(inp_vec) {
+    
+    ## recodes 0 to no, 1 to yes
+    
+    out_vec <- car::recode(inp_vec, 
+                           "'0' = 'no'; 
+                              '1' = 'yes'") %>% 
+      factor(c('no', 'yes'))
+    
+    return(out_vec)
+    
+  }
+  
+  sum_variables <- function(inp_tbl, vars_to_sum) {
+    
+    ## calculates a sum of the given variables
+    ## coded as yes/no (changed to 1 and 0, respectively)
+    
+    var_sum_vec <- vars_to_sum %>% 
+      map(function(x) car::recode(inp_tbl[[x]], "'no' = 0; 'yes' = 1") %>% 
+            as.character %>% 
+            as.numeric) %>% 
+      reduce(function(x, y) x + y)
+    
+    return(var_sum_vec)
+    
+  }
+  
+  binarize_variables <- function(inp_tbl, vars_to_recode) {
+    
+    ## recodes yes/no to 1/0
+    
+    tbl_list <- vars_to_recode %>% 
+      map(function(x) tibble(ID = inp_tbl[['ID']], 
+                             sympt_recoded = car::recode(inp_tbl[[x]], 
+                                                                    "'no' = 0; 
+                                                                        'yes' = 1") %>% 
+                               as.character %>% 
+                               as.numeric)) %>% 
+      set_names(vars_to_recode)
+    
+    rec_tbl <- tbl_list %>% 
+      map2(., names(.), 
[…]

### 2021-10-19 paper/markdown/figures_and_tables.Rmd
Add files via upload

@@
 title: "Hyposmia as a standalone persistent symptom of long COVID"
-subtitle: "Figures and Tables"
+subtitle: "Methods, Results, Figures and Tables"
 author: "Health after COVID-19 in Tyrol study team"
@@
 \newpage
+
+# Methods
+
+## Study design and approval
+
+Data generated in the 'Health After COVID-19 in Tyrol' (HACT, ClinicalTrials.gov: NCT04661462) online survey [@Sahanic2021; @Hufner2021] and in the longitudinal, multi-center cohort observation CovILD study
+CovILD study (ClnicalTrials.gov: NCT04416100, acute COVID-19 and 60-, 100- and 180-day post-COVID-19 follow-up data) [@Sonnweber2021; @Sonnweber2020; @Rass2021] were re-analyzed in the current report.
+The studies were conducted in accordance with the Declaration of Helsinki and European Data Policy.
+The participant's data were analyzed in anonymized form.
+All participants gave digitally signed or written informed consent prior to enrollment.
+The study protocols were reviewed and approved by the institutional review boards of the Medical University of Innsbruck (HACT, Austria cohort, approval number: 1257/2020, CovILD: 1103/2020) and of the Autonomous Province of Bolzano - South Tyrol (HACT Italy cohort, approval number: 0150701).
+
+The HACT study encompassed two independently recruited cohorts of ambulatory COVID-19 convalescents in the Tyrol state in Austria and the South Tyrol province in Italy.
+The study exclusion criterion was hospitalization because of COVID-19.
+The study inclusion criteria were laboratory-confirmed SARS-CoV-2 infection (PCR or seropositivity), adult age (Austria: $\geq$ 16 years, Italy: $\geq$ 18 years) and residence in the study regions [@Sahanic2021; @Hufner2021].
+The analysis inclusion criterion was a minimal observation time of 90 days between the positive SARS-CoV-2 test and the survey completion (__Figure \@ref(fig:fig-consort)__).
+A total of `r nrow(hyposmia$survey_at) + nrow(hyposmia$survey_it)` questionnaires was eligible for the current analysis (Austria: n = `r nrow(hyposmia$survey_at)`, Italy: `r nrow(hyposmia$survey_it)`).
+
+The CovILD cohort consists of longitudinally investigated multi-center collective of convalescents of ambulatory (WHO categories 1 - 3, n = `r unique(kinet$analysis_tbl$covild$A$ID) %>% length`), moderate (WHO categories 4 - 5, n = `r unique(kinet$analysis_tbl$covild$HM$ID) %>% length`) and severe COVID-19 (WHO categories 6 - 9, n = `r unique(kinet$analysis_tbl$covild$HS$ID) %>% length`) recruited at the Department of Internal Medicine II at the Medical University of Innsbruck (Austria), St.
+Vinzenz Hospital in Zams (Austria) and the acute rehabilitation facility in Muenster (Austria) (__Figure 1__) [@Sonnweber2021; @Sonnweber2020; @Rass2021].
+
+## Procedures and definitions
+
+In the HACT study, the duration classes of `r length(globals$symptoms)` symptoms of COVID-19 were surveyed ('absent', 'present for 1 - 3 days', 'present for up to 1 week', 'present for up to 2 weeks', 'present for up to 3 months', 'present for up to 6 months' or 'present for more than 6 months') and duration of individual symptoms was calculated [@Sahanic2021].
+In the CovILD study, `r length(kinet$variables$covild[kinet$variables$covild != 'sympt_present'])` self-reported symptoms were surveyed at each follow-up visit [@Sonnweber2020].
+For both studies, detailed characteristic of demography, socioeconomic status, medical history before COVID-19 as well as the disease course and recovery was recorded as described [@Sonnweber2021; @Sonnweber2020; @Rass2021; @Sahanic2021; @Hufner2021].
+The full list of the variables available for the current analysis is provided in __Supplementary Table S1__.
+
+Self-reported symptoms were classified as acute when present for $\geq$ 14 days, persistent when present $\geq$ 28 days and long-term persistent complaints when lasting for $\geq$ 3 months post clinical onset (HACT) or present at the 100-day follow-up visit (CovILD).
+Numeric variables (symptoms, anti-SARS-CoV-2 S1/S2 antibody titers) were stratified by quartiles, laboratory parameters recorded during the CovILD visits were binarized with standard cutoffs as described [@Sonnweber2020; @Sonnweber2021; @Sahanic2021; @Hufner2021; @Rass2021].
+Assessment of hyposmia by the sniffing stick test, lung CT imaging and lung function measurements were conducted as reported before [@Sonnweber2020; @Sonnweber2021; @Rass2021].
+Details on study variables and their stratification are provided in __Supplementary Methods__ and __Supplementary Table S1__.
+
+## Statistical analysis
+
+Statistical analysis was performed with R 4.0.5 [@Wickham2019; @Wickham2016; @Wilke2019].
+Statistical significance for differences in numeric variables between analysis groups as determined by Mann-Whitney U and Kruskal-Wallis test, as appropriate.
+Statistical significance for differences in frequency of categorical variables was assessed by $\chi^2$ test.
+Concordance between self-reported hyposmia and the results of sniffing stick test was measured with Cohen's $\kappa$ statistic [@Fleiss1969].
+Overlap between symptoms was expressed as cosine similarity coefficient [@Drost2018].
+Modeling of symptom resolution was accomplished by mixed-effect second-order logistic regression (random effect of the individual); model term and odds ratio estimate significances were assessed by likelihood ratio and Wald Z test, respectively [@Bates2015; @Kuznetsova2017; @Box1978].
+Clustering of the HACT study participants in respect to long-term persistent symptoms was done with a combined self-organizing map and hierarchical clustering approach [@Kohonen1995; @Vesanto2000; @Vesanto1999; @Hufner2021].
+Univariable modeling of long-term persistent hyposmia risk was done with a series of logistic regression models.
+In the HACT study, the risk models included age and sex weights and the stratified diagnosis - survey time as a confounder [@Sahanic2021; @Hufner2021].
+Statistical testing results were corrected for multiple testing with Benjamini-Hochberg method [@Benjamini1995].
+Details of the statistical analysis are provided in __Supplementary Methods__.
+
+# Data availability
+
+The study data is available at reasonable request to the corresponding authors.
+The source code of the R analysis pipeline is available at GitHub (https://github.com/PiotrTymoszuk/hyposmia_analsis_pipeline).
+
+\newpage
+
+# Results
+
+## Characteristic of the study collective
+
+## Post-COVID-19 hyposmia is a long-term persistent symptom
+
+In order to delineate the recovery kinetics, overlap of hyposmia and other COVID-19 symptoms and risk factors for delayed long-term post-COVID-19 olfactory disorders, we first re-analyzed data gathered in the population of ambulatory COVID-19 convalescents enrolled in the bi-national 'Health After COVID-19 in Tyrol' study (HACT) [@Sahanic2021; @Hufner2021].
+By means of mixed-effect logistic modeling of the recovery curves [@Box1978] for `r length(globals$symptoms)` COVID-19-related complaints (__Supplementary Table S2__) gauged by the online HACT survey [@Sahanic2021], we identified a set of symptoms characterized by a highly significant second-order time term suggestive of strong 'plateauing' effect and hence a substantial chronicity potential.
+This set of candidate chronic manifestations was highly consistent in both Austria and Italy HACT cohorts and included among others headache, muscle and joint pain, tiredness and fatigue as well as olfactory and taste disorders (__Figure \@ref(fig:fig-kinetic)A__, __Supplementary Figure S1__, __Supplementary Table S2__).
+Among them, tiredness, fatigue and self-reported hyposmia were the symptoms with the longest median recovery times (14 days) still present in at least 25% of the participants beyond 28 days post clinical onset (__Figure \@ref(fig:fig-kinetic)B__).
+Of note, whereas the duration of non-specific infection symptoms like fever or shivering in course of COVID-19 was virtually limited to the first two weeks of the disease (__Supplementary Figure S2__), fatigue, tiredness and taste disorders were still present in more than 10% the Austria or Italy HACT cohort 3 months post clinical onset (__Supplementary Figure S3__).
+Importantly, an analogically retarded recovery was observed for self-declared hyposmia in these collectives (__Figure \@ref(fig:fig-hyposmia-kinetic)A__).
+To explore, if a similar hyposmia kinetic may be observed in individuals hospitalized during acute COVID-19, we resorted to the longitudinally investigated CovILD cohort [@Sonnweber2020; @Sonnweber2021; @Rass2021].
+Surprisingly, self-reported smell disorders displayed nearly complete resolution in the subset of severe COVID-19 survivors (WHO class 6 - 9) within 180 days post COVID-19 diagnosis.
+However, this disease symptom was still reported by 13 to 15% of ambulatory (WHO class 1 - 3) and moderate (WHO class 4 - 6) COVID-19 patients at the 180-day follow-up (__Figure \@ref(fig:fig-hyposmia-kinetic)B__).
+Hence, hyposmia persisting beyond the contraction of the non-specific acute infection and acute upper airway symptoms belongs to the complaints most frequently declared by ambulatory and moderate COVID-19 convalescents.
+
+## Sniffing stick test detected hyposmia is far more frequent than self-reported olfactory disorders
+
+Next, we sought to investigate the strike differences in self-perceived olfactory deficits between the survivors of severe, moderate and ambulatory COVID-19 in the CovILD cohort at the 100-day post-COVID-19 follow-up.
+As shown in __Supplementary Figure S4A__, the frequency of moderate-to-severe hyposmia defined by < 13 points in the 16-point sniffing stick test was way higher that the self-reported smell disorders.
+In particular, up to `r signif(hypo_kappa$proportions$hyposmia_mod_severe_V2$HS[1, 2], 3)*100`% of the severe COVID-19 survivors declaring no hyposmia, were affected by at least moderate olfactory impairment suggestive of a distorted symptom perception (__Supplementary Figure S4A__).
+Notably, the prevalence of severe hyposmia characterized by < 8 test points was detected only in a minute fraction of the participants (up to `r signif(sum(hypo_kappa$proportions$hyposmia_severe$HM[, 2]), 3)*100`% in the moderate COVID-19 subset, __Supplementary Figure S4B__).
+In general, the concordance between the test-detected moderate-to-severe long-term hyposmia and the self-perceived symptom was fair for the moderate ($\kappa$ = `r signif(hypo_kappa$kappa$hyposmia_mod_severe_V2$kappa$A$kappa[1], 3)`) and weak ($\kappa$ = `r signif(hypo_kappa$kappa$hyposmia_mod_severe_V2$kappa$HM$kappa[1], 3)`) for the severe COVID-19 convalescent subset [@Fleiss1969].
+No statistically valid overlap was observed for the severe COVID-19 survivors ($\kappa$ = `r signif(hypo_kappa$kappa$hyposmia_mod_severe_V2$kappa$HS$kappa[1], 3)`) (__Supplementary Figure S4C__).
+Collectively, the true prevalence of hyposmia may be greatly underestimated by the patient's report, especially in severe COVID-19
+
+## Limited co-occurrence of hyposmia and other persistent symptoms in long-term COVID-19 recovery
+
+As suggested by our previous report, isolated smell and/or taste disorders were sole long COVID symptoms in a substantial fraction of ambulatory COVID-19 convalescents [@Sahanic2021].
+In acute COVID-19, self-reported hyposmia co-occurred with numerous other manifestations in the HACT study cohorts (20 of `r length(globals$symptoms)` symptoms with cosine similarity coefficient > 0.5) and in the CovILD study (ambulatory: 8, moderate: 6, severe: 5 symptoms of `r length(kinet$variables$covild[kinet$variables$covild != 'sympt_present'])` features surveyed with cosine similarity coefficient > 0.5) including non-specific infection symptoms such as fever, muscle and joint pain and fatigue or respiratory manifestations like tachy- and dyspnea (__Figure \@ref(fig:fig-hyposmia-overlap)__, __Supplementary Table S3__).
+However, in course of recovery, the association with the majority of symtoms with faster resolution kinetic (__Figure \@ref(fig:fig-kinetic)__, __Supplementary Figure S1__) was lost and solely self-reported taste disorders showed moderate co-occurrence (cosine similarity coefficient > 0.5) with self-reported hyposmia in the HACT cohort (__Figure \@ref(fig:fig-hyposmia-overlap)A__, Supplementary Figure S4A).
+In the ambulatory and moderate COVID-19 subsets of the CovILD study, such moderate association could by discerned only for dyspnea (ambulatory) and pain (moderate) at the 180-day follow-up (__Figure \@ref(fig:fig-hyposmia-overlap)B__, Supplementary Figure S4B).
+
+Of interest, two other highly persistent symptoms frequently reported during the long-term COVID-19 recovery: fatigue and tiredness (__Figure \@ref(fig:fig-kinetic)__, __Supplementary Figure S1__ and __S3__ ) demonstrated only a weak-to-fair co-existence with self-reported hyposmia in the HACT (90-day Austria, fatigue: cosine similarity = `r hypo_overlap$distances$survey$survey_at %>% filter(time == 90, overlap_var == 'fatigue') %>% .$stat %>% signif(2)`, tiredness: `r hypo_overlap$distances$survey$survey_at %>% filter(time == 90, overlap_var == 'fatigue_day') %>% .$stat %>% signif(2)`, Italy, fatigue: cosine similarity = `r hypo_overlap$distances$survey$survey_it %>% filter(time == 90, overlap_var == 'fatigue') %>% .$stat %>% signif(2)`, tiredness: `r hypo_overlap$distances$survey$survey_it %>% filter(time == 90, overlap_var == 'fatigue_day') %>% .$stat %>% signif(2)`) and CovILD study (100-day follow-up, fatigue, ambulatory: cosine similarity = `r hypo_overlap$distances$covild$A %>% filter(time == 100, overlap_var == 'fatigue_sympt') %>% .$stat %>% signif(2)`, moderate: `r hypo_overlap$distances$covild$HM %>% filter(time == 100, overlap_var == 'fatigue_sympt') %>% .$stat %>% signif(2)`) (Supplementary Figure S4).
+
+To corroborate this hypothesis we resorted to an association analysis of the HACT participants with long-term persisting COVID-19 symptoms beyond 3 months post clinical disease onset in respect to the most frequent long COVID manifestations (present in at least 25% of the participants at $\geq$ 28 days post symptom onset: impaired concentration, tachypnea, hypogeusia, hyposmia, fatigue, tiredness and forgetfulness) (__Figure \@ref(fig:fig-kinetic)B__).
+By applying a combined self-organizing map and hierarchical clustering algorithm [@Kohonen1995; @Vesanto2000; @Hufner2021], three subsets of the participants could be discerned, further referred to as 'non-specific', 'hyposmia' and 'neuro/fatigue' clusters (__Figure \@ref(fig:fig-clustering)A__, __Supplementary Figure S5__).
+The hyposmia cluster comprising approximately one-third of the individuals affected by the highly protracted COVID-19 recovery (Austria: `r signif(pheno$n_numbers$survey_at$percent[2], 3)`%, Italy: `r signif(pheno$n_numbers$survey_it$percent[2], 3)`%) was characterized by a 100% prevalence and of self-reported hyposmia, approximately 60% frequency od taste disorders and, remarkably, by minimal occurrence of other symptoms (up to `r extract_counts(pheno$clust_ft$analyses$survey_it$fatigue_day)[4, 'percent'] %>% signif(3)`% for tiredness) (__Figure \@ref(fig:fig-clustering)B__).
+In summary, the results of the overlap and association analyses strongly suggest that in the late course of COVID-19 recovery, self-reported hyposmia may pose the sole protracted disease symptom occurring independently of post-COVID-19 fatigue, tiredness, memory and concentration deficits.
+
+## Male sex is associated with lower risk of persistent long-term hyposmia
+
+Finally, we intended to find demographic and clinical features identifiable during acute disease and early recovery which correlate with the risk of development of long-term persistent hyposmia present for 3 months or longer post clinical COVID-19 onset.
+By means of logistic risk modeling, hyposmia and taste disorders in the course of acute COVID-19 were identified as the strongest correlates of the long-term hyposmia risk (__Figure \@ref(fig:fig-risk)__).
+Furthermore, in the HACT study, highly poly-symptomatic acute disease (50 - 100^th^ symptom count percentile) was linked to a substantially higher protracted hyposmia risk.
+As demonstrated above (__Figure \@ref(fig:fig-hyposmia-kinetic)B__), severe acute COVID-19 was associated with an 80% lower risk of hyposmia present at the 100-day follow-up in the CovILD cohort, however, this effect was not significant when adjusted for multiple testing (__Figure \@ref(fig:fig-risk)A__).
+Of note, independently of the investigated study, male sex was a significant favorable co-variate associated with over 50% lower risk of long-term hyposmia (__Figure \@ref(fig:fig-risk)__).
+
+A more detailed investigation and modeling of recovery from post-COVID-19 hyposmia in female and male participants of the HACT study revealed no significant differences in pace of the resolution and comparable 'plateauing' effect between the sex strata in both the Austria and Italy cohort (__Supplementary Figure S7AB__).
+Furthermore, no differences in the median recovery time in females and males suffering from olfactory deficits during acute COVID-19 could be observed (__Supplementary Figure S7C__).
+Instead, the prevalence of hyposmia during acute COVID-19 was significantly higher in females than in males in both HACT cohorts ($\chi^2$ test; Austria, females: `r signif(kinet_sex$acute_covid$counts$survey_at$percent[1], 3)`%, males: `r signif(kinet_sex$acute_covid$counts$survey_at$percent[2], 3)`%, p = `r signif(kinet_sex$acute_covid$test$survey_at$p_value, 2)`; Italy, females: `r signif(kinet_sex$acute_covid$counts$survey_it$percent[1], 3)`%, males: `r signif(kinet_sex$acute_covid$counts$survey_it$percent[2], 3)`%, p = `r signif(kinet_sex$acute_covid$test$survey_it$p_value, 2)`) (__Supplementary Figure S7A__).
 

### 2021-10-19 paper/markdown/supplementary_material.Rmd
Add files via upload

@@
 The HACT study collective encompassed two independently recruited cohorts of ambulatory COVID-19 convalescents in the Tyrol state in Austria and the South Tyrol province in Italy.
+The study exclusion criterion was hospitalization because of COVID-19.
 The study inclusion criteria were laboratory-confirmed SARS-CoV-2 infection (PCR or seropositivity), adult age (Austria: $\geq$ 16 years, Italy: $\geq$ 18 years) and residence in the study regions [@Sahanic2021; @Hufner2021].
@@
 The study was approved by the institutional review boards of the Medical University of Innsbruck (Austria, approval
-number: 1257/2020) and of the Autonomous Province of Bolzano - South Tyrol (Italy, 0150701). The HACT study was registered at ClinicalTrials.gov (NCT: NCT04661462)
+number: 1257/2020) and of the Autonomous Province of Bolzano - South Tyrol (Italy, 0150701). The HACT study was registered at ClinicalTrials.gov (NCT04661462)
 
@@
 
-## Measures, definitions and data transformation
+## Procedures and definitions
 
@@
 
-Self-reported symptoms were classified as acute when present in the fist weeks after clinical manifestation of COVID-19, persistent when lasting for at least 28 days and long-term persistent complaints when present for longer than 3 months (HACT) or present at the 100-day follow-up visit.
-For analysis, symptom counts and anti-SARS-CoV-2 S1/S2 antibody levels were stratified by quartiles [@Sahanic2021; @Sonnweber2021].
+Self-reported symptoms were classified as acute when present in the fist weeks after clinical manifestation of COVID-19, persistent when lasting for at least 28 days and long-term persistent complaints when present for longer than 3 months (HACT) or present at the 100-day follow-up visit (CovILD).
+For analysis, symptom counts and anti-SARS-CoV-2 S1/S2 antibody levels were stratified by quartiles [@Sahanic2021; @Sonnweber2021; @Hufner2021].
 Laboratory readouts of inflammation, iron metabolism and hematopoiesis in the CovILD cohort were longitudinally determined by the local hospital laboratory and binarized with standard cutoffs [@Sonnweber2021; @Sonnweber2020].
@@
 The concordance of sniffing stick test-measured and self-reported hyposmia, was investigated with the Cohen's $\kappa$ statistic (package _vcd_, function _Kappa()_) [@Fleiss1969].
-Co-occurrence of self-reported hyposmia and other symptoms (re-coded as 0: absent, 1: present) was measured by cosine similarity coefficient and cosine distance (package _philentropy_, function _distance_) [@Drost2018].
+Co-occurrence of self-reported hyposmia and other symptoms (re-coded as 0: absent, 1: present) was measured by cosine similarity coefficient and cosine distance (package _philentropy_, function _distance_, for the complete calculation result see __Supplementary Table \@ref(tab:tab-similarity)__) [@Drost2018].
 
@@
 Significance of the 2^nd^ and 1^st^ order model terms was assessed by step-wise likelihood ratio test versus the lower order model (function _anova()_).
-Significance of the _beta_ estimates was determined by Wald Z test.
+Significance of the $\beta$ estimates was determined by Wald Z test.
+For the summary of the kinetic modeling results, see: __Supplementary Table \@ref(tab:tab-kinetic)__.
 
@@
 for the HACT study, the models were weighted for age and sex weights based on the age and sex distribution of the general COVID-19 convalescent population in Tyrol/Austria and Italy [@Sahanic2021; @IstitutoSuperiorediSanitaISS; @AustrianAgencyforHealthandFoodSafetyAGES].
-To account for the effects of varying diagnosis - survey time in the HACT study, the stratified observation time variable was included as a confounder in each model [@Sahanic].
+To account for the effects of varying diagnosis - survey time in the HACT study, the stratified observation time variable was included as a confounder in each model [@Sahanic2021].
 Odds ratio significance was determined by Wald Z test.
 Logistic modeling and visual model quality control was accomplished with home-developed wrappers (https://github.com/PiotrTymoszuk/lm_qc_tools) around base R modeling tools and diagnostic functions provided by _broom_ package [@Robinson2014].
+For the list of independent variables modeling variables, see __Supplementary Table \@ref(tab:tab-study-vars)__.
+For the summary of modeling results, refer to __Supplementary Table \@ref(tab:tab-uni-modeling)__.
 
@@
 
-Symptom frequency during acute COVID-19 and recovery (0, 7, 14, 28 and 90 days post symptom onset) in the Austria (`r kinet$models$survey$survey_at %>% map_dbl(length) %>% paste_range`) and Italy (`r kinet$models$survey$survey_at %>% map_dbl(length) %>% paste_range`) HACT cohorts were modeled with mixed-effect logistic regression (__Supplementary Table \@ref(tab:tab-kinetic)__).
+Symptom frequency during acute COVID-19 and recovery (0, 7, 14, 28 and 90 days post symptom onset) in the Austria (`r kinet$models$survey$survey_at %>% map_dbl(length) %>% paste_range`) and Italy (`r kinet$models$survey$survey_it %>% map_dbl(length) %>% paste_range`) HACT cohorts were modeled with mixed-effect logistic regression (__Supplementary Table \@ref(tab:tab-kinetic)__).
 Significance of the OR estimates was determined by Wald Z test.
@@
 
+```{r fig-gender, fig.width = mm_inch(suppl_figures$gender$w), fig.height = mm_inch(suppl_figures$gender$h), dpi = 600, fig.cap = 'Sex-dependent kinetics of self-reported hyposmia resolution in the HACT study.'}
+
+suppl_figures$gender$plot
+
+```
+
+__Supplementary Figure \@ref(fig:fig-gender). Sex-dependent kinetics of self-reported hyposmia resolution in the HACT study.__
+
+Frequency of self-reported hyposmia during acute COVID-19 and recovery (0, 7, 14, 28 and 90 days post symptom onset) in females and males from the Austria and Italy HACT cohorts were modeled with mixed-effect logistic regression.
+Significance of the OR estimates was determined by Wald Z test.
+P values were corrected for multiple comparisons with Benjamini-Hochberg method.
+The numbers of complete observations are shown under the plots in (__A__).
+
+__(A)__ Frequency of self-reported hyposmia in time post clinical onset. Points and solid lines represent the actual frequencies. The model predictions are displayed as dashed lines.
+
+__(B)__ OR estimate values with 95% confidence intervals for the first and second order terms in females and males presented as points and whiskers in the Forest plots.
+
+__(C)__ Time to recovery from hyposmia in females and males reporting hyposmia during acute COVID-19 presented as violin plots.
+Diamonds and whiskers represent medians with înterquartile ranges.
+Statistical significance was assessed with Mann-Whitney U test corrected for multiple comparisons with Benjamini-Hochberg method.
+
+\newpage
+
 # References

### 2021-10-19 analysis scripts/chronicity_phenotypes.R

@@ -103,7 +103,9 @@
   ## n numbers and plot tags
   
   pheno$n_numbers <- pheno$analysis_tbl %>% 
-    map(count, clust_name)
+    map(count, clust_name) %>% 
+    map(mutate, 
+        percent = n/sum(n) * 100)
   
   pheno$plot_tags <- pheno$n_numbers %>% 
     map(function(x) map2_chr(x$clust_name, 

### 2021-10-19 analysis scripts/symptom_kinetic_gender.R

@@ -1,0 +1,217 @@
+# This script checks for differences in hyposmia resolution kinetic between the sexes in the HACT study
+
+  insert_head()
+  
+# data container -----
+  
+  kinet_sex <- list()
+  
+# globals: analysis table ------
+  
+  insert_msg('Analysis table')
+  
+  kinet_sex$analysis_tbl <- kinet$analysis_tbl$survey %>% 
+    map(select, 
+        ID, time, anosmia) %>% 
+    map2(., 
+         map(hyposmia[c('survey_at', 'survey_it')], 
+             ~.x[c('ID', 'sex')]), 
+         left_join, 
+         by = 'ID') %>% 
+    map(dlply, 'sex', as_tibble) %>% 
+    unlist(recursive = F)
+  
+# kinetic models -----
+  
+  insert_msg('Kinetic models')
+  
+  kinet_sex$models <- kinet_sex$analysis_tbl %>% 
+    map(model_kinetic, 
+        response = 'anosmia', 
+        time_var = 'time', 
+        family = 'binomial')
+  
+# LRT -----
+  
+  insert_msg('LRT')
+  
+  kinet_sex$lrt <- kinet_sex$models %>% 
+    lrt_list(.parallel = T) %>% 
+    mutate(cohort = c(rep('survey_at', 6), 
+                      rep('survey_it', 6)), 
+           sex = c(rep('female', 3), 
+                   rep('male', 3), 
+                   rep('female', 3), 
+                   rep('male', 3)))
+  
+# inference -----
+  
+  insert_msg('Inference')
+
+  kinet_sex$beta <- kinet_sex$models %>% 
+    inference_kinetic_list(.parallel = T) %>% 
+    mutate(cohort = c(rep('survey_at', 6), 
+                      rep('survey_it', 6)), 
+           sex = c(rep('female', 3), 
+                   rep('male', 3), 
+                   rep('female', 3), 
+                   rep('male', 3)))
+  
+# single kinetic plots ------
+  
+  insert_msg('Kinetic plots')
+  
+  kinet_sex$plots <- list(kinetic_object = kinet_sex$models, 
+                          plot_title = c('HACT Austria, female', 
+                                         'HACT Austria, male', 
+                                         'HACT Italy, female', 
+                                         'HACT Italy, male'), 
+                          point_color = c('indianred', 
+                                          'cornflowerblue', 
+                                          'indianred', 
+                                          'cornflowerblue'), 
+                          resp_color = c('indianred', 
[…]

### 2021-10-19 data_analysis.R

@@ -17,7 +17,8 @@
     './analysis scripts/hyposmia_overlap.R', 
     './analysis scripts/hyposmia_uni_modeling.R', 
     './analysis scripts/time_to_recovery.R', 
-    './analysis scripts/chronicity_phenotypes.R') %>% 
+    './analysis scripts/chronicity_phenotypes.R', 
+    './analysis scripts/symptom_kinetic_gender.R') %>% 
     walk(source)
   
 # END ----

### 2021-10-19 paper scripts/paper_figures.R

@@ -369,6 +369,37 @@
                      h = 210)
   
   
+# Supplementary Figure S7: differences in hyposmia resolution between males and females -----
+  
+  insert_msg('Figure S7: gender and hyposmia resolution in the HACT cohort')
+  
+  suppl_figures$gender$top_panel <- kinet_sex$plots %>% 
+    map(function(x) x + theme(legend.position = 'none')) %>% 
+    plot_grid(plotlist = ., 
+              ncol = 2, 
+              align = 'hv', 
+              labels = 'A', 
+              label_size = 10)
+  
+  suppl_figures$gender$bottom_panel <- plot_grid(kinet_sex$beta_plot + 
+                                                   theme(plot.subtitle = element_blank()), 
+                                                 kinet_sex$duration$plots$survey_at + 
+                                                   theme(plot.tag = element_blank()), 
+                                                 kinet_sex$duration$plots$survey_it + 
+                                                   theme(plot.tag = element_blank()), 
+                                                 ncol = 3, 
+                                                 rel_widths = c(1.8, 0.6, 0.6), 
+                                                 labels = c('B', 'C', ''), 
+                                                 label_size = 10)
+  
+  suppl_figures$gender <- plot_grid(suppl_figures$gender$top_panel, 
+                                    suppl_figures$gender$bottom_panel, 
+                                    nrow = 2, 
+                                    rel_heights = c(1.8, 1)) %>% 
+    as_figure_object(figure_label = 'figure_s7_hyposmia_gender', 
+                     w = 180, 
+                     h = 200)
+  
 # Saving the figures on the disc -----
   
   insert_msg('Saving the figures on the disc')

### 2022-06-02 tools/survey_globals.R
General analysis update
Replaced by a development package

@@ -1,112 +1,0 @@
-# This script provides globals and import functions 
-
-# toolbox ----
-
-  source('./tools/sys_tools.R')
-
-  library(readr)
-  library(readxl)
-  library(tibble)
-  library(ggplot2)
-  library(stringi)
-
-# Co-morbidities -----
-  
-  globals$comorb <- c('autoimmunity', 
-                      'anemia', 
-                      'hypertension', 
-                      'depression_burnout',
-                      'diabetes', 
-                      'pre_cov_epilepsy',
-                      'dementia', 
-                      'surgery_6m_before', 
-                      'frequent_flu_like', 
-                      'heart_circulation', 
-                      'hay_fever', 
-                      'malignancy', 
-                      'pins_needles_feet', 
-                      'gastrointenstinal', 
-                      'lung', 
-                      'two_plus_infections_antibiotics', 
-                      'multiple_sclerosis', 
-                      'night_dyspnoe', 
-                      'bruxism', 
-                      'kidney', 
-                      'parkinson', 
-                      'insomnia', 
-                      'stroke', 
-                      'embolism', 
-                      'transplantation')
-
-# Daily medication -----
-  
-  globals$daily_medication <- c('daily_immunosuppression', 
-                                'daily_cortison', 
-                                'daily_ace_drugs', 
-                                'daily_pain_killers', 
-                                'daily_anti_coagulation')
-  
-# self-reported symptoms -------
-  
-  globals$symptoms <- c('fever', 
-                        'ague', 
-                        'sore_throat', 
-                        'running_nose', 
-                        'fatigue', 
-                        'dry_cough', 
-                        'wet_cough', 
-                        'breath_short', 
-                        'dyspnoe', 
-                        'chest_pain', 
-                        'tachycardia', 
-                        'extrasystole', 
-                        'joint_pain', 
-                        'bone_pain', 
-                        'muscle_pain', 
-                        'abdominal_pain', 
-                        'nausea', 
-                        'vomiting', 
-                        'dim_appetite', 
-                        'diarrhea', 
-                        'dizziness', 
-                        'headache', 
-                        'anosmia', 
-                        'taste_loss', 
-                        'confusion', 
-                        'tingle_feet', 
-                        'tingle_hands', 
-                        'ache_feet', 
-                        'ache_hands', 
-                        'numb_feet', 
[…]

### 2022-06-02 tools/survey_import_tools.R
General analysis update
Redundant file

@@ -1,1518 +1,0 @@
-# this script provides tools for wrangling the survey data
-
-# tools ------
-
-  require(plyr)
-  require(tidyverse)
-
-# basic formatting functions -----
-
-  pss2date <- function(x) {
-    
-    ## converts the SPSS date to R
-    
-    return(as.Date(x/86400, origin = '1582-10-14') )
-    
-  }
-  
-  rm_space <- function(x) {
-    
-    ## removes extra trailing spaces
-    
-    return(stri_replace(x, 
-                        regex = '\\s+$', 
-                        replacement = ''))
-    
-  }
-  
-  format_na_no <- function(x) {
-    
-    ## formats the comorbidity (and other data) data
-    ## where NA is assumed 'no'
-    
-    return(ifelse(is.na(x), 'no', 'yes') %>% 
-             factor(c('no', 'yes')))
-    
-    
-  }
-  
-  yn_variable <- function(inp_vec) {
-    
-    ## recodes 0 to no, 1 to yes
-    
-    out_vec <- car::recode(inp_vec, 
-                           "'0' = 'no'; 
-                              '1' = 'yes'") %>% 
-      factor(c('no', 'yes'))
-    
-    return(out_vec)
-    
-  }
-  
-  sum_variables <- function(inp_tbl, vars_to_sum) {
-    
-    ## calculates a sum of the given variables
-    ## coded as yes/no (changed to 1 and 0, respectively)
-    
-    var_sum_vec <- vars_to_sum %>% 
-      map(function(x) car::recode(inp_tbl[[x]], "'no' = 0; 'yes' = 1") %>% 
-            as.character %>% 
-            as.numeric) %>% 
-      reduce(function(x, y) x + y)
-    
-    return(var_sum_vec)
-    
-  }
-  
-  binarize_variables <- function(inp_tbl, vars_to_recode) {
-    
-    ## recodes yes/no to 1/0
-    
-    tbl_list <- vars_to_recode %>% 
-      map(function(x) tibble(ID = inp_tbl[['ID']], 
-                             sympt_recoded = car::recode(inp_tbl[[x]], 
-                                                                    "'no' = 0; 
-                                                                        'yes' = 1") %>% 
-                               as.character %>% 
-                               as.numeric)) %>% 
-      set_names(vars_to_recode)
-    
-    rec_tbl <- tbl_list %>% 
-      map2(., names(.), 
[…]

### 2022-06-02 data_analysis.R
General analysis update
Redundant file

@@ -1,26 +1,0 @@
-# A master script executing the project analyses
-
-  c('./tools/sys_tools.R', 
-    './tools/kinetic_tools.R', 
-    './tools/lm_qc_tools.R', 
-    './tools/project_tools.R',
-    './tools/counting_tools.R', 
-    './tools/clust_tools.R') %>% 
-    walk(source)
-  
-# Executing the analysis scripts -----
-  
-  insert_msg('Executing the analysis scripts')
-  
-  c('./analysis scripts/symptom_kinetic.R',
-    './analysis scripts/hyposmia_kappa.R', 
-    './analysis scripts/hyposmia_overlap.R', 
-    './analysis scripts/hyposmia_uni_modeling.R', 
-    './analysis scripts/time_to_recovery.R', 
-    './analysis scripts/chronicity_phenotypes.R', 
-    './analysis scripts/symptom_kinetic_gender.R') %>% 
-    walk(source)
-  
-# END ----
-  
-  insert_tail()

### 2022-06-02 data_import.R
Redundant file

@@ -1,49 +1,0 @@
-# This script reads and clears the data from the Health after COVID-19 survey
-# ant the COVILD-19 studies which concern hyposmia and COVID-19 course
-
-# data containers -----
-
-  globals <- list()
-  covild_data <- list()
-  survey_data <- list()
-  hyposmia <- list()
-
-# tools ------
-
-  library(plyr)
-  library(tidyverse)
-  library(foreign)
-  library(readxl)
-
-  c('./tools/sys_tools.R', 
-    './tools/survey_globals.R', 
-    './tools/survey_import_tools.R', 
-    './tools/project_tools.R') %>% 
-    walk(source, 
-         encoding = 'UTF-8')
-  
-# reading the survey and covild data -----
-  
-  c('./import scripts/survey.R', 
-    './import scripts/covild.R') %>% 
-    walk(source, 
-         encoding = 'UTF-8')
-  
-# merging the data sets into a common list ------
-  
-  insert_msg('A common list with the data sets')
-
-  hyposmia[c('covild', 
-             'survey_at', 
-             'survey_it')] <- c(covild_data['mod_tbl'], 
-                                survey_data[c('north', 'south')])
-  
-  
-# END -----
-  
-  rm(covild_data, 
-     survey_data, 
-     model_wrangling, 
-     raw_data)
-  
-  insert_tail()

### 2022-06-02 render_paper.R
General analysis update
Redundant file

@@ -1,25 +1,0 @@
-# This script generates figures, tables and a report file
-
-  c('./tools/sys_tools.R',
-    './tools/plotting_tools.R', 
-    './tools/counting_tools.R', 
-    './tools/project_tools.R') %>% 
-    walk(source)
-  
-  library(writexl)
-  
-  insert_head()
-  
-# Rendering scripts -----
-  
-  insert_msg('Executing the paper scripts')
-  
-  c('./paper scripts/paper_figures.R', 
-    './paper scripts/cohort_characteristic.R', 
-    './paper scripts/paper_tables.R', 
-    './paper scripts/deploy_paper.R') %>%  
-    walk(source)
-  
-# END ----
-  
-  insert_tail()

### 2022-06-02 paper/markdown/figure_chunks.Rmd
Major analysis update
Corresponds now to the analyses presented in the pre-print

```{r fig-consort, fig.width = 7.086614166, fig.height = 4.330708657, fig.cap = 'Flow diagram of the analysis inclusion process for the CovILD and HACT studies.'}

figures$consort$plot

```

__Figure \@ref(fig:fig-consort). Flow diagram of the analysis inclusion process for the CovILD and HACT studies.__

```{r fig-sympt-recovery, fig.width = 7.086614166, fig.height = 8.661417314, fig.cap = 'Symptom-specific recovery times in ambulatory COVID-19.'}

figures$sympt_recovery$plot

```

__Figure \@ref(fig:fig-sympt-recovery). Symptom-specific recovery times in ambulatory COVID-19.__

```{r fig-sympt-mds, fig.width = 7.086614166, fig.height = 7.086614166, fig.cap = 'Self-reported smell and taste disorders are isolated symptoms of long COVID and PASC.'}

figures$sympt_mds$plot

```

__Figure \@ref(fig:fig-sympt-mds). Self-reported smell and taste disorders are isolated symptoms of long COVID and PASC.__

```{r fig-apriori, fig.width = 7.086614166, fig.height = 7.480314953, fig.cap = 'High extent of co-occurrence of smell and taste disorders in long COVID and PASC.'}

figures$apriori$plot

```

__Figure \@ref(fig:fig-apriori). High extent of co-occurrence of smell and taste disorders in long COVID and PASC.__

```{r fig-clustering, fig.width = 7.086614166, fig.height = 7.086614166, fig.cap = 'Clustering of ambulatory COVID-19 individuals by symptom-specific recovery times.'}

figures$clustering$plot

```

__Figure \@ref(fig:fig-clustering). Clustering of ambulatory COVID-19 individuals by symptom-specific recovery times.__

```{r fig-sympt-clusters, fig.width = 7.086614166, fig.height = 7.086614166, fig.cap = 'Duration of neurocognitive and respiratory symptoms, fatigue, smell and taste disorders differs between the COVID-19 recovery clusters.'}

figures$sympt_clusters$plot

```

__Figure \@ref(fig:fig-sympt-clusters). Duration of neurocognitive and respiratory symptoms, fatigue, smell and taste disorders differs between the COVID-19 recovery clusters.__

```{r fig-base-clusters, fig.width = 7.086614166, fig.height = 8.267716527, fig.cap = 'COVID-19 recovery clusters differ in sex distribution, comorbidity and daily medication rates.'}

figures$base_clusters$plot

```

__Figure \@ref(fig:fig-base-clusters). COVID-19 recovery clusters differ in sex distribution, comorbidity and daily medication rates.__

```{r fig-recovery-clusters, fig.width = 7.086614166, fig.height = 4.724409444, fig.cap = 'Good clinical, mental and psychosocial recovery in individuals with persistent smell and taste disorders'}

figures$recovery_clusters$plot

```

__Figure \@ref(fig:fig-recovery-clusters). Good clinical, mental and psychosocial recovery in individuals with persistent smell and taste disorders__

### 2022-06-02 paper/markdown/figures_and_tables.Rmd
Major analysis update
Corresponds now to the analyses presented in the pre-print

--- 2021-10-19 paper/markdown/figures_and_tables.Rmd
@@
 ---
-title: "Hyposmia as a standalone persistent symptom of long COVID"
-subtitle: "Figures and Tables"
-author: "Health after COVID-19 in Tyrol study team"
+title: "Distinct smell and taste disorder phenotype of post-acute COVID-19 sequelae"
+subtitle: "Tables and figures"
+author: "HACT and CovILD study teams"
 date: "`r format(Sys.time(), '%Y-%m-%d')`"
@@
 output: 
-  pdf_document:
-    latex_engine: xelatex
-bibliography: cov_biblio.bib
-csl: american-medical-association.csl
+  bookdown::word_document2:   
+    reference_docx: ms_template.docx
+
+bibliography: hyposmia_biblio.bib
+csl: frontiers_medical.csl
 header-includes:
@@
 
-knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, dev = "cairo_pdf")
+knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, dev = "png", dpi = 600)
 
@@
 
-# Tables {#sec:tbl}
+# Tables
 
-```{r tab-covild}
+```{r tab-hact, tab.cap = 'Baseline characteristic of the HACT study Austria (AT) and Italy (IT) cohorts.'}
 
-paper_tables$cohort_covild %>% 
-  format_line %>%  
-  kable(caption = '\\textbf{Characteristic of the CovILD study cohort.}\\\\Ambulatory, Moderate, Severe: severity of acute SARS-CoV-2 infection, Test: statistical test used for the comparison between the severity strata (KW: Kruskal-Wallis), Significance: test p value corrected for multiple comparisons with Benjamini-Hochberg method.',
-        format = 'latex', 
-        escape = F, 
-        booktabs = T, 
-        longtable = T) %>% 
-  column_spec(1, width = '3cm') %>% 
-  column_spec(2:4, width = '3cm') %>% 
-  column_spec(5, width = '0.6cm') %>% 
-  column_spec(6, width = '2cm') %>% 
-  collapse_rows(columns = 1) %>% 
-  kable_styling(full_width = F, 
-                latex_options = c('striped', 
-                                  'hold_position', 
-                                  'repeat_header'))
+paper_tbl$hact_baseline %>% 
+  flextable %>% 
+  width(1, width = 3.7, unit = 'cm') %>% 
+  width(2:3, width = 4, unit = 'cm') %>% 
+  width(4, width = 2.6, unit = 'cm') %>% 
+  width(5, width = 2.2, unit = 'cm') %>% 
+  footnote(1, 4:5, value = as_paragraph('Categorical variables: \u03C7\u00B2 test with Cramer V effect size statistic.
+Numeric variables: Mann-Whitney U test with wilcoxon r effect size statistic.
+P values corrected form multiple testing with Benjamini-Hochberg method.'), ref_symbols = 'a', part = 'header') %>%
+  footnote(3, 1, value = as_paragraph('BMI: body mass index, overweight > 25 kg/m\u00B2, obesity > 30 kg/m\u00B2,'), ref_symbols = 'b', part = 'body') %>% 
+  footnote(15, 1, value = as_paragraph('Frequent respiratory infections, > 2 per year.'), ref_symbols = 'c', part = 'body') %>% 
+  footnote(16, 1, value = as_paragraph('Frequent bacterial infections with antibiotic therapy, > 2 per year.'), ref_symbols = 'd', part = 'body') %>% 
+  theme_vanilla %>%
+  font(part = 'all', fontname = 'Cambria') %>% 
+  fontsize(size = 10, part = 'all')
 
@@
 
-```{r tab-survey}
+```{r tab-covild, tab.cap = 'Baseline characteristic of the CovILD study cohort and the study participants stratified by COVID-19 severity.'}
 
-paper_tables$cohort_survey %>% 
-  format_line %>%  
-  kable(caption = '\\textbf{Characteristic of the Health after COVID-19 in Tyrol study cohorts.}\\\\Test: statistical test used for the comparison between the severity strata (U: Mann-Whitney U test), Significance: test p value corrected for multiple comparisons with Benjamini-Hochberg method.',
-        format = 'latex', 
-        escape = F, 
-        booktabs = T, 
-        longtable = T) %>% 
-  column_spec(1, width = '3.5cm') %>% 
-  column_spec(2:3, width = '4cm') %>% 
-  column_spec(4, width = '0.6cm') %>% 
-  column_spec(5, width = '2cm') %>% 
-  collapse_rows(columns = 1) %>% 
-  kable_styling(full_width = F, 
-                latex_options = c('striped', 
-                                  'hold_position', 
-                                  'repeat_header'))
+paper_tbl$covild_baseline %>% 
+  flextable %>% 
+  width(1, width = 3, unit = 'cm') %>% 
+  width(2:5, width = 2.7, unit = 'cm') %>% 
+  width(6, width = 2.6, unit = 'cm') %>% 
+  width(7, width = 2.2, unit = 'cm') %>% 
+  footnote(1, 6:7, value = as_paragraph('Comparison of ambulatory, moderate and severe COVID-19 individuals.
+Categorical variables: \u03C7\u00B2 test with Cramer V effect size statistic.
+Numeric variables: Kruskal-Wallis test with \u03B7\u00B2 effect size statistic.
+P values corrected form multiple testing with Benjamini-Hochberg method.'), ref_symbols = 'a', part = 'header') %>%
+  footnote(3, 1, value = as_paragraph('BMI: body mass index, overweight > 25 kg/m\u00B2, obesity > 30 kg/m\u00B2,'), ref_symbols = 'b', part = 'body') %>% 
+  theme_vanilla %>%
+  font(part = 'all', fontname = 'Cambria') %>% 
+  fontsize(size = 10, part = 'all')
 
@@
 
-```{r fig-consort, fig.width = mm_inch(paper_figures$consort$w), fig.height = mm_inch(paper_figures$consort$h), dpi = 600, fig.cap = 'CONSORT flow diagrams for the CoVILD and HACT study cohorts.'}
+```{r fig-consort, fig.width = mm_inch(figures$consort$w), fig.height = mm_inch(figures$consort$h), fig.cap = 'Flow diagram of the analysis inclusion process for the CovILD and HACT studies.'}
 
-paper_figures$consort$plot
+figures$consort$plot
 
@@
 
-__Figure \@ref(fig:fig-consort). CONSORT flow diagrams for the CoVILD and HACT study cohorts.__
+__Figure \@ref(fig:fig-consort). Flow diagram of the analysis inclusion process for the CovILD and HACT studies..__
 
@@
 
-```{r fig-kinetic, fig.width = mm_inch(paper_figures$kinetic_modeling$w), fig.height = mm_inch(paper_figures$kinetic_modeling$h), dpi = 600, fig.cap = 'Second order kinetic modeling of symptom recovery course in the HACT study.'}
+```{r fig-sympt-recovery, fig.width = mm_inch(figures$sympt_recovery$w), fig.height = mm_inch(figures$sympt_recovery$h), fig.cap = 'Symptom-specific recovery times in ambulatory COVID-19.'}
 
-paper_figures$kinetic_modeling$plot
+figures$sympt_recovery$plot
 
@@
 
-__Figure \@ref(fig:fig-kinetic). Second order kinetic modeling of symptom recovery course in the HACT study.__
+__Figure \@ref(fig:fig-sympt-recovery). Symptom-specific recovery times in ambulatory COVID-19.__
 
-Symptom frequency during acute COVID-19 and recovery in the Austria and Italy HACT cohorts (0, 7, 14, 28 and 90 days post symptom onset) were modeled with mixed-effect logistic regression (__Supplementary Table S2__).
-Significance of the first and second order model terms was determined by step-wise likelihood ratio test (LRT).
-P values were corrected for multiple comparisons with Benjamini-Hochberg method.
-N number of complete observations is provided under the plots in (__B__).
-Dim.: diminished, imp.: impaired.
+Symptom-specific recovery times were calculated for each participants of the HACT study cohorts (Austria: AT, Italy: IT).
 
-__(A)__ Significance of the first and second order model term in LRT test.
-Points represent single symptoms, color codes for significance (gray: non-significant).
-Top 10 symptoms with the most significant second order terms as candidate long-term persistent features are labeled with their names.
+__(A)__ Distribution of the recovery times in the individuals with the indicated symptoms present during acute COVID-19.
+Diamonds represent median recovery times, tinted ellipses code for interquartile ranges.
+Numbers of complete observations are indicated in the plot captions.
 
-__(B)__ Median recovery times for top 10 symptoms with the most significant second order terms as candidate long-term persistent features. Whiskers represent interquartile ranges.
+__(B)__ Percentages of individuals with smell and taste disorders in the AT (Austria) and IT (Italy) HACT study cohorts at particular time points after clinical onset.
+Numbers of complete observations are indicated under the plots.
 
@@
 
-```{r fig-hyposmia-kinetic, fig.width = mm_inch(paper_figures$hyposmia_kinetic$w), fig.height = mm_inch(paper_figures$hyposmia_kinetic$h), dpi = 600, fig.cap = 'Actual and predicted frequency of self-reported hyposmia in course of COVID-19 recovery.'}
+```{r fig-sympt-mds, fig.width = mm_inch(figures$sympt_mds$w), fig.height = mm_inch(figures$sympt_mds$h), fig.cap = 'Self-reported smell and taste disorders are isolated symptoms of long COVID and PASC.'}
 
-paper_figures$hyposmia_kinetic$plot
+figures$sympt_mds$plot
 
@@
 
-__Figure \@ref(fig:fig-hyposmia-kinetic). Actual and predicted frequency of self-reported hyposmia in course of COVID-19 recovery.__
+__Figure \@ref(fig:fig-sympt-mds). Self-reported smell and taste disorders are isolated symptoms of long COVID and PASC.__
 
-Frequency of self-reported hyposmia during acute COVID-19 and recovery in the HACT study (__A__, 0, 7, 14, 28 and 90 days post symptom onset) and severity subsets of the CovILD cohorts (__B__, acute COVID: 0, 60-, 100 and 180-day follow-up visits) was modeled with mixed-effect logistic regression (__Supplementary Table S2__).
-Points and solid lines represent the actual frequencies.
-The model predictions are displayed as dashed lines.
-The ranges of complete observations per time point are shown under the plots.
+Binary symptom occurrence data during for long COVID ($\geq$ 28 days after clinical onset, __A__) and PASC ($\geq$ 90 days, __B__) in the HACT Austria (AT) and Italy (IT) cohorts were subjected to two-dimensional multi-dimensional scaling (MDS) with simple matching distance (SMD) between the symptoms.
+MDS coordinates are presented in point plots.
+Selected data points are labeled with the symptom names.
+Percentages of the data set variance associated with the MDS dimensions are indicated in the plot axes.
+Numbers of complete observations are indicated under the plots.
 
@@
 
-```{r fig-hyposmia-overlap, fig.width = mm_inch(paper_figures$hyposmia_overlap$w), fig.height = mm_inch(paper_figures$hyposmia_overlap$h), dpi = 600, fig.cap = 'Co-occurrence of self-reported hyposmia and other symptoms in course of COVID-19 recovery.'}
+```{r fig-apriori, fig.width = mm_inch(figures$apriori$w), fig.height = mm_inch(figures$apriori$h), fig.cap = 'Frequent co-occurrence of smell and taste disorders in long COVID and PASC.'}
 
-paper_figures$hyposmia_overlap$plot
+figures$apriori$plot
 
@@
 
-__Figure \@ref(fig:fig-hyposmia-overlap). Co-occurrence of self-reported hyposmia and other symptoms in course of COVID-19 recovery.__
+__Figure \@ref(fig:fig-apriori). Frequent co-occurrence of smell and taste disorders in long COVID and PASC.__
 
-Cosine similarity coefficients between hyposmia and other self-reported symptoms of COVID-19 in the HACT 
-(__A__, `r length(hypo_overlap$variables$survey) + 1` symptoms in total, Austria: n = `r hypo_overlap$analysis_tbl$survey$survey_at$ID %>% unique %>% length`, Italy: `r hypo_overlap$analysis_tbl$survey$survey_it$ID %>% unique %>% length`)
-and COVILD study 
-(__B__, `r length(hypo_overlap$variables$covild) + 1` symptoms, ambulatory: n = `r hypo_overlap$analysis_tbl$covild$A$ID %>% unique %>% length`, moderate: n = `r hypo_overlap$analysis_tbl$covild$HM$ID %>% unique %>% length`, severe COVID-19: `r hypo_overlap$analysis_tbl$covild$HS$ID %>% unique %>% length`) were calculated for the indicated time points (__Supplementary Table S3__) and presented in radial plots.
-Each bar represents a single symptom.
-Bar length and color code for the value of Cosine similarity coefficient.
-Dashed lines represent cosine similarity coefficient of 0.5.
+Frequent combinations of symptoms of long COVID (__A__) and PASC (__B__) in the HACT study Austria (AT) and Italy (IT) cohorts were identified with the apriori algorithm.
+Symptom combination support and confidence are presented in the plots.
+Point color codes for lift values.
+The hypo/anosmia and hypo/ageusia combinations are labeled.
 
@@
 
-```{r fig-clustering, fig.width = mm_inch(paper_figures$clustering$w), fig.height = mm_inch(paper_figures$clustering$h), dpi = 600, fig.cap = 'Long-term COVID-19 recovery phenotypes in the HACT study.'}
+```{r fig-clustering, fig.width = mm_inch(figures$clustering$w), fig.height = mm_inch(figures$clustering$h), fig.cap = 'Clustering of ambulatory COVID-19 individuals by symptom-specific recovery times.'}
 
-paper_figures$clustering$plot
+figures$clustering$plot
 
@@
 
-__Figure \@ref(fig:fig-clustering). Long-term COVID-19 recovery phenotypes in the HACT study.__
+__Figure \@ref(fig:fig-clustering). Clustering of ambulatory COVID-19 individuals by symptom-specific recovery times.__
 
-HACT participants with COVID-19 symptoms present for at least three months post clinical onset (Austria: n = `r nrow(pheno$analysis_tbl$survey_at)`, Italy: n = `r nrow(pheno$analysis_tbl$survey_it)`) were clustered in respect to manifestations present in at least 25% of the cohort for 28 days or longer with a combined algorithm employing self-organizing map (`r pheno$som_grid$xdim` $\times$ `r pheno$som_grid$ydim` hexagonal units, Jaccard distance) and hierarchical clustering (Ward D2 algorithm, Euclidean distance).
-Numbers of individuals assigned to particular clusters are presented under the plot in (__A__).
-Imp.: impaired.
+Individuals of the training Austria (AT) cohort of the HACT study were subjected to clustering in respect to symptom-specific recovery times with the PAM (partitioning around medoids) algorithm and Euclidean distance measure (__Supplementary Figure S8__).
+Cluster assignment in the test Italy (IT) HACT cohort was done with k-NN label propagation algorithm.
+Recovery times for particular COVID-19 symptoms in the COVID-19 recovery clusters are presented as heat maps.
+Numbers of individuals assigned to the recovery clusters are indicated under the plots.
 
-(__A__) Presence of the features used for cluster definition in the participants assigned to the 'non-specific', 'hyposmia' and 'neuro/fatigue' subset.
-
-(__B__) Frequency of self-reported hyposmia, taste disorders and tiredness at day in the participant subsets.
-Statistical significance of the frequency differences was assessed with $\chi^2$ test corrected for multiple comparisons with Benjamini-Hochberg method.
-P values are presented in the plot captions.
+Dim. appetite: diminished appetite, Imp. concentration: impaired concentration, Imp. walk: impaired walk, Imp. FMS: impaired fine motor skills.
 
@@
 
-```{r fig-risk, fig.width = mm_inch(paper_figures$hyposmia_risk$w), fig.height = mm_inch(paper_figures$hyposmia_risk$h), dpi = 600, fig.cap = 'Risk factors for long-term persistent self-reported post-COVID-19 hyposmia.'}
+```{r fig-sympt-clusters, fig.width = mm_inch(figures$sympt_clusters$w), fig.height = mm_inch(figures$sympt_clusters$h), fig.cap = 'Duration of neurocognitive and respiratory symptoms, fatigue, smell and taste disorders differs between the COVID-19 recovery clusters.'}
 
-paper_figures$hyposmia_risk$plot
+figures$sympt_clusters$plot
 
@@
 
-__Figure \@ref(fig:fig-risk). Risk factors for long-term persistent self-reported post-COVID-19 hyposmia.__
+__Figure \@ref(fig:fig-sympt-clusters). Duration of neurocognitive and respiratory symptoms, fatigue, smell and taste disorders differs between the COVID-19 recovery clusters.__
 
-Association of candidate risk factors (__Supplementary Table S1__) with long-term persistent self-reported hyposmia in the CovILD (__A__, hyposmia declared at the 100-day follow-up visit) and HACT cohorts (__B__, self-reported hyposmia present for at least three months post clinical onset) was investigated with univariable logistic regression (__Suppmenetary Table S4__).
-For the HACT cohorts, the models were weighted for age and sex distribution in the general COVID-19 convalescent populations and included stratified the observation time variable as a confounder.
-Significance of model estimates (odds ratio, OR) was determined by Wald Z test corrected for multiple comparisons with Benjamini-Hochberg method.
-OR estimate values with 95% confidence intervals for significant and nearly significant (adjusted p < 0.1) co-variates are presented as points and whiskers in the Forest plots.
-The ranges of complete observations per time point are shown under the plots.
+Semi-supervised clustering was performed as presented in __Figure \@ref(fig:fig-clustering)__ and __Supplementary Figure S8__.
+Mean recovery times in the recovery clusters are presented as lines, 2 $\times$ SEM intervals are displayed as tinted regions.
+Numbers of individuals assigned to the recovery clusters are indicated under the plots.
+
+Dim. appetite: diminished appetite, Imp. concentration: impaired concentration, Imp. walk: impaired walk, Imp. FMS: impaired fine motor skills.
+
+\newpage
+
+```{r fig-base-clusters, fig.width = mm_inch(figures$base_clusters$w), fig.height = mm_inch(figures$base_clusters$h), fig.cap = 'COVID-19 recovery clusters differ in sex distribution, comorbidity and daily medication rates.'}
+
+figures$base_clusters$plot
+
+```
+
+__Figure \@ref(fig:fig-base-clusters). COVID-19 recovery clusters differ in sex distribution, comorbidity and daily medication rates.__
+
+Differences in sex distribution (__A__), frequency of comorbidity (__B__) and daily medication (__C__) between the recovery clusters in the training Austria (AT) and the test Italy (IT) HACT study cohorts (__Figure \@ref(fig:fig-clustering)__, __Supplementary Figure S8__) were assessed by $\chi^2$ test with Cramer V effect size statistic.
+P values were corrected for multiple testing with Benjamini-Hochberg method.
+The frequencies are presented as bar plots.
+Effect size statistics and p values are indicated in the plot caption.
+Numbers of complete observations are displayed under the plots.
+
+\newpage
+
+```{r fig-recovery-clusters, fig.width = mm_inch(figures$recovery_clusters$w), fig.height = mm_inch(figures$recovery_clusters$h), fig.cap = 'Good clinical and mental health status and quality of life at the survey completion in individuals with persistent smell and taste disorders.'}
+
+figures$recovery_clusters$plot
+
+```
+
+__Figure \@ref(fig:fig-recovery-clusters). Good clinical and mental health status and quality of life at the survey completion in individuals with persistent smell and taste disorders.__
+
+Differences in minimum/maximum scaled readouts of clinical and physical recovery, mental health and quality of life scoring at the time of survey completion in the recovery clusters (__Figure \@ref(fig:fig-clustering)__, __Supplementary Figure S8__) are presented in radial plots.
+Yes/no items (incomplete convalescence, weight loss, new medication and need for rehabilitation) were binarized (yes: 1, no: 0) prior to visualization.
+Lines represent mean values, 2 $\times$ SEM intervals are displayed as tinted regions.
+Numbers of individuals assigned to the recovery clusters are indicated under the plots.
+
+incomplete recovery: self-reported incomplete recovery, # persist.
+symptoms: number of long COVID symptoms ($\geq$ 28 days after clinical onset), phys.
+performance loss: physical performance loss as compared with the time before COVID-19, QoL score: quality of life impairment score, OMH score: overall mental health impairment score, ANX score: anxiety score, DPR: depression score.

### 2022-06-02 paper/markdown/supplement_chunks.Rmd
Major analysis update
Corresponds now to the analyses presented in the pre-print

```{r fig-hact-freq, fig.width = 7.086614166, fig.height = 7.480314953, fig.cap = 'Frequency of COVID-19 symptoms in acute COVID-19, long COVID and PASC in the HACT study.'}

suppl$hact_freq$plot

```

__Figure \@ref(fig:fig-hact-freq). Frequency of COVID-19 symptoms in acute COVID-19, long COVID and PASC in the HACT study.__

```{r fig-hact-recovery, fig.width = 7.086614166, fig.height = 7.086614166, fig.cap = 'Kinetic of recovery from leading acute COVID-19 symptoms in the HACT study.'}

suppl$hact_recovery$plot

```

__Figure \@ref(fig:fig-hact-recovery). Kinetic of recovery from leading acute COVID-19 symptoms in the HACT study.__

```{r fig-covild-freq, fig.width = 7.086614166, fig.height = 8.661417314, fig.cap = 'Symptom frequency in ambulatory, moderate and severe COVID-19 subsets of the CovILD study.'}

suppl$covild_freq$plot

```

__Figure \@ref(fig:fig-covild-freq). Symptom frequency in ambulatory, moderate and severe COVID-19 subsets of the CovILD study.__

```{r fig-covild-recovery, fig.width = 7.086614166, fig.height = 7.086614166, fig.cap = 'Kinetic of recovery from smell disorders, reduced performance and dyspnea in ambulatory, moderate and severe COVID-19 subsets of the CovILD study.'}

suppl$covild_recovery$plot

```

__Figure \@ref(fig:fig-covild-recovery). Kinetic of recovery from smell disorders, reduced performance and dyspnea in ambulatory, moderate and severe COVID-19 subsets of the CovILD study.__

```{r fig-hyposmia-rater100, fig.width = 7.086614166, fig.height = 6.299212592, fig.cap = 'Rates of self-reported hyposmia and hyposmia in sniffing stick test at 3-month post COVID-19 follow-up in the ambulatory, moderate and severe COVID-19 subsets of the CovILD study.'}

suppl$hyposmia_rater100$plot

```

__Figure \@ref(fig:fig-hyposmia-rater100). Rates of self-reported hyposmia and hyposmia in sniffing stick test at 3-month post COVID-19 follow-up in the ambulatory, moderate and severe COVID-19 subsets of the CovILD study.__

```{r fig-hyposmia-rater360, fig.width = 7.086614166, fig.height = 6.299212592, fig.cap = 'Rates of self-reported hyposmia and hyposmia in sniffing stick test at 1-year post COVID-19 follow-up in the ambulatory, moderate and severe COVID-19 subsets of the CovILD study.'}

suppl$hyposmia_rater360$plot

```

__Figure \@ref(fig:fig-hyposmia-rater360). Rates of self-reported hyposmia and hyposmia in sniffing stick test at 1-year post COVID-19 follow-up in the ambulatory, moderate and severe COVID-19 subsets of the CovILD study.__

```{r fig-mds-acute, fig.width = 7.086614166, fig.height = 3.93700787, fig.cap = 'Multi-dimensional scaling analysis of acute COVID-19 symptoms in the HACT study.'}

suppl$mds_acute$plot

```

__Figure \@ref(fig:fig-mds-acute). Multi-dimensional scaling analysis of acute COVID-19 symptoms in the HACT study.__

```{r fig-clust-devel, fig.width = 7.086614166, fig.height = 7.086614166, fig.cap = 'Definition of the COVID-19 recovery clusters and clustering feature importance in the HACT study.'}

suppl$clust_devel$plot

```

__Figure \@ref(fig:fig-clust-devel). Definition of the COVID-19 recovery clusters and clustering feature importance in the HACT study.__

```{r fig-cov-clusters, fig.width = 7.086614166, fig.height = 5.511811018, fig.cap = 'Numbers of acute COVID-19 and long COVID symptoms in the HACT study recovery clusters.'}

suppl$cov_clusters$plot

```

__Figure \@ref(fig:fig-cov-clusters). Numbers of acute COVID-19 and long COVID symptoms in the HACT study recovery clusters.__

```{r fig-anosmia-clusters, fig.width = 7.086614166, fig.height = 7.086614166, fig.cap = 'Recovery from fatigue, tiredness and tachypnea in the HACT study recovery clusters.'}

suppl$fatigue_clsters$plot

```

__Figure \@ref(fig:fig-anosmia-clusters). Recovery from fatigue, tiredness and tachypnea in the HACT study recovery clusters.__

```{r fig-fatigue-clsters, fig.width = 7.086614166, fig.height = 7.086614166, fig.cap = 'Recovery from neurocognitive symptoms of COVID-19 in the HACT study recovery clusters.'}

suppl$neurocog_clusters$plot

```

__Figure \@ref(fig:fig-fatigue-clsters). Recovery from neurocognitive symptoms of COVID-19 in the HACT study recovery clusters.__

```{r fig-neurocog-clusters, fig.width = 7.086614166, fig.height = 4.724409444, fig.cap = 'Recovery from smell and taste disorders in the HACT study recovery clusters'}

suppl$anosmia_clusters$plot

```

__Figure \@ref(fig:fig-neurocog-clusters). Recovery from smell and taste disorders in the HACT study recovery clusters__

```{r fig-rec-clusters, fig.width = 7.086614166, fig.height = 8.267716527, fig.cap = 'Physical performance loss, impairment of quality of life and self-perceived complete convalescence in the HACT study recovery clusters.'}

suppl$rec_clusters$plot

```

__Figure \@ref(fig:fig-rec-clusters). Physical performance loss, impairment of quality of life and self-perceived complete convalescence in the HACT study recovery clusters.__

```{r fig-psych-clusters, fig.width = 7.086614166, fig.height = 9.055118101, fig.cap = 'Rating of anxiety, depression, overall mental health impairment and stress in the HACT study recovery clusters.'}

suppl$psych_clusters$plot

```

__Figure \@ref(fig:fig-psych-clusters). Rating of anxiety, depression, overall mental health impairment and stress in the HACT study recovery clusters.__

### 2022-06-02 paper/markdown/supplementary_material.Rmd
Major analysis update
Corresponds now to the analyses presented in the pre-print

--- 2021-10-19 paper/markdown/supplementary_material.Rmd
@@
 ---
-title: "Hyposmia as a standalone persistent symptom of long COVID"
+title: "Distinct smell and taste disorder phenotype of post-acute COVID-19 sequelae"
 subtitle: "Supplementary Material"
-author: "Health after COVID-19 in Tyrol study team"
+author: "HACT and CovILD study teams"
 date: "`r format(Sys.time(), '%Y-%m-%d')`"
@@
 \newpage
+
 \beginsupplement
@@
 # Supplementary Methods
+## Study procedures and variables
 
-## Study design and participants
+The complete list of study variables and stratification scheme is provided in __Supplementary Table S\@ref(tab:tab-hact-vars)__ for the HACT study and in __Supplementary Table S\@ref(tab:tab-covild-vars)__ for the CovILD study.
 
-In the current project, the data obtained in the 'Health After COVID-19 in Tyrol' (HACT) online survey [@Sahanic2021; @Hufner2021] and during acute COVID-19, 60-, 100- and 180-day post-COVID-19 follow-up of the CovILD study [@Sonnweber2021; @Sonnweber2020; @Rass2021] were analyzed.
+### COVID-19 symptoms
 
-The HACT study collective encompassed two independently recruited cohorts of ambulatory COVID-19 convalescents in the Tyrol state in Austria and the South Tyrol province in Italy.
-The study inclusion criteria were laboratory-confirmed SARS-CoV-2 infection (PCR or seropositivity), adult age (Austria: $\geq$ 16 years, Italy: $\geq$ 18 years) and residence in the study regions [@Sahanic2021; @Hufner2021].
-The analysis inclusion criterion was a minimal observation time of 90 days between the positive SARS-CoV-2 test and the survey completion (__Figure 1__).
-A total of `r nrow(hyposmia$survey_at) + nrow(hyposmia$survey_it)` questionnaires was eligible for the current analysis (Austria: n = `r nrow(hyposmia$survey_at)`, Italy: `r nrow(hyposmia$survey_it)`).
-All participants gave digitally signed informed consent before the survey start.
-The study was approved by the institutional review boards of the Medical University of Innsbruck (Austria, approval
-number: 1257/2020) and of the Autonomous Province of Bolzano - South Tyrol (Italy, 0150701). The HACT study was registered at ClinicalTrials.gov (NCT: NCT04661462)
+A total of `r length(globals$hact_symptoms)` was surveyed in the HACT study cohorts (__Supplementary Figure S\@ref(fig:fig-hact-freq)__ and __Supplementary Table S\@ref(tab:tab-hact-vars)__).
+The symptom duration was coded as follows: absent: 0 days, 1 - 3 days: 3 days, up to 1 week: 7 days, up to 2 weeks: 14 days, up to 4 weeks: 28 days, up to 3 months: 90 days, up to 6 months: 90 days, over 6 months: 90 days.
+Acute symptoms were defined as complaints present during the first 14 days after clinical onset of COVID-19.
+Long COVID symptoms were defined as complaints present for $\geq$ 28 days.
+PASC symptoms were defined as manifestations present for $\geq$ 90 days [@Venkatesan2021; @Shah2021].
 
-The CovILD cohort included a longitudinally investigated multi-center collective of convalescents of ambulatory (WHO categories 1 - 3, n = `r unique(kinet$analysis_tbl$covild$A$ID) %>% length`), moderate (WHO categories 4 - 5, n = `r unique(kinet$analysis_tbl$covild$HM$ID) %>% length`) and severe COVID-19 (WHO categories 6 - 9, n = `r unique(kinet$analysis_tbl$covild$HS$ID) %>% length`) recruited at the Department of Internal Medicine II at the Medical University of Innsbruck (Austria), St.
-Vinzenz Hospital in Zams (Austria) and the acute rehabilitation facility in Muenster (Austria) (__Figure 1__) [@Sonnweber2021; @Sonnweber2020; @Rass2021].
-The study was approved by the institutional review board at the Medical University of Innsbruck (approval number: 1103/2020), and registered at ClinicalTrials.gov (NCT04416100).
+In the CovILD study, a total of `r length(globals$covild_symptoms)` symptoms (reduced physical performance, hyposmia/anosmia, dyspnea, sleep problems, cough, fever, night sweating, gastrointestinal symptoms) were prospectively recorded with a standardized questionnaire at each of 60-, 100-, 180- and 360-day post COVID-19 follow-up (__Supplementary Figure S\@ref(fig:fig-covild-freq)__ and __Supplementary Table S\@ref(tab:tab-covild-vars)__).
+Acute COVID-19 symptoms were assessed retrospectively [@Sonnweber2020; @Sonnweber2022].
 
-## Measures, definitions and data transformation
+### Rating of physical recovery, mental health and quality of life in the HACT study
 
-In the HACT study, `r length(globals$symptoms)` symptoms of COVID-19 were surveyed.
-The participants were asked to classify them as 'absent', 'present for 1 - 3 days', 'present for up to 1 week', 'present for up to 2 weeks', 'present for up to 3 months', 'present for up to 6 months' or 'present for more than 6 months'.
-Based on that responses, individual symptom duration was calculated [@Sahanic2021].
-In the CovILD study, `r length(kinet$variables$covild[kinet$variables$covild != 'sympt_present'])` self-reported symptoms were surveyed at each follow-up visit [@Sonnweber2020].
-For both studies, detailed characteristic of demography, socioeconomic status, medical history before COVID-19 as well as the disease course and recovery was recorded as described [@Sonnweber2021; @Sonnweber2020; @Rass2021; @Sahanic2021; @Hufner2021].
-The full list of the variables available for the current analysis is provided in __Supplementary Table \@ref(tab:tab-study-vars)__.
+Self-perceived complete recovery, rehabilitation need and new medication since COVID-19 at the time of study participation were surveyed as single yes/no items.
+Percentage of physical performance loss as compared with the time before COVID-19 was rated with a 0 - 100% scale [@Hufner2022; @Sahanic2021].
+Quality of life impairment (QoL) and overall mental health impairment (OMH) were rated with a 4 item Likert scale each (possible answers: “excellent,” “good,” “fair,” “poor,” scored: 0, 1, 2, and 3) [@Hufner2022; @Sahanic2021].
+Anxiety (ANX) and depression (DPR) were assessed with the Patient Health Questionnaire (PHQ-4) [@Hufner2022; @Sahanic2021; @Lowe2010].
+Psychosocial stress was scored with a modified 7 item PHQ stress module as described before [@Hufner2022; @Sahanic2021; @Grafe2004].
 
-Self-reported symptoms were classified as acute when present in the fist weeks after clinical manifestation of COVID-19, persistent when lasting for at least 28 days and long-term persistent complaints when present for longer than 3 months (HACT) or present at the 100-day follow-up visit.
-For analysis, symptom counts and anti-SARS-CoV-2 S1/S2 antibody levels were stratified by quartiles [@Sahanic2021; @Sonnweber2021].
-Laboratory readouts of inflammation, iron metabolism and hematopoiesis in the CovILD cohort were longitudinally determined by the local hospital laboratory and binarized with standard cutoffs [@Sonnweber2021; @Sonnweber2020].
-Lung function and lung CT abnormality assessment in the CovILD cohort were performed as described [@Sonnweber2020; @Sonnweber2021].
-Assessment of hyposmia by the 16 point sniffing stick test at the 100-day visit in the CovILD collective was done essentially as described before [@Rass2021].
-For variable stratification scheme, see __Supplementary Table \@ref(tab:tab-study-vars)__.
+### Rating of hyposmia with sniffing stick test
+
+Objective hyposmia at the 100-day and 360-day follow-up in the CovILD study participants was investigated with the 16-item sniffing stick test as described [@Rass2021].
+Clinically relevant hyposmia was defined as $\leq$ 12 correct answers [@Rass2021; @Hummel2007].
+In the analysis, participants with the complete answers concerning self-reported hyposmia and complete test results were included.
 
 ## Statistical analysis
-### Data transformation, visualization, descriptive statistic, hypothesis testing and symptom co-occurrence
+### Data transformation, descriptive statistic
 
-Data transformation, analysis and visualization was accomplished with R 4.0.5, _tidyverse_ environment and _cowplot_ package [@Wickham2019; @Wickham2016; @Wilke2019].
-As most of the investigated numeric variables were non-normally distributed as assessed by Shapiro-Wilk test, non-parametric tests: Mann-Whitney U and Kruskal-Wallis tests were used to determine statistical significance between analysis groups.
-Statistical significance for differences in frequency of categorical variables between analysis groups, $\chi^2$ test was applied.
+Data transformation and statistical analysis was done with R version 4.0.5 with _tidyverse_ data science environment [@Wickham2019].
+Analysis results were visualized with _ggplot2_ [@Wickham2016], _cowplot_ [@Wilke2019] as well as in-house developed _ExDA_ (https://github.com/PiotrTymoszuk/ExDA) and _figur_ (https://github.com/PiotrTymoszuk/figur) packages.
 
-The concordance of sniffing stick test-measured and self-reported hyposmia, was investigated with the Cohen's $\kappa$ statistic (package _vcd_, function _Kappa()_) [@Fleiss1969].
-Co-occurrence of self-reported hyposmia and other symptoms (re-coded as 0: absent, 1: present) was measured by cosine similarity coefficient and cosine distance (package _philentropy_, function _distance_) [@Drost2018].
+Descriptive statistics including median with interquartile ranges and frequency of complete answers for numeric and categorical variables were calculated with base R functions and _ExDA_ package.
 
-Test and modeling p values for each analysis task and cohort were corrected for multiple testing with Benjamini-Hochberg method [@Benjamini1995].
+### Statistical hypothesis testing
 
-### Kinetic modeling
+Since multiple study variables were non-normally distributed as assessed by Shapiro-Wilk test and visual assessment of their distribution (quantile-quantile plots), statistical significance for differences in outcome numeric variables were assessed with Mann-Whitney U test by Wilcoxon r effect size statistic (two groups) or Kruskal-Wallis test with $\eta^2$ effect size statistic.
+Differences in frequency distribution for categorical outcome variables were assessed by $\chi^2$ test with Cramer V effect size statistic.
+Interrater assessment of self-reported and sniffing test hyposmia was accomplished with Cohen's $\kappa$ statistic [@Fleiss1969].
+$\kappa$ significance ($\kappa\neq0$) was estimated with Wald Z test.
+P values were adjusted for multiple testing with Benjamini-Hochberg method [@Benjamini1995] separately for each analysis task and cohort.
+Packages _rstatix_ [@Kassambara2021], _vcd_ [@Meyer2021] and _Exda_ (https://github.com/PiotrTymoszuk/ExDA) were used for statistical hypothesis testing.
 
-Symptom recovery kinetics was modeled by mixed-effect logistic regression (fixed terms: first and second time term, random term: individual, packages _lme4_ and _lmerTest_) [@Bates2015; @Kuznetsova2017] with the following general model formula:
+### Modeling of symptom recovery kinetic
 
-$$Symptom \sim \beta_0 + \beta_1 \times t + \beta_2 \times t^2$$
+To model recovery kinetics for binary symptom variables (0: absent, 1: present), second-order mixed-effect logistic (categorical features) modeling was applied (packages: _lme4_, _lmerTest_ and development package _kinet_ [https://github.com/PiotrTymoszuk/kinet]) [@Bates2015; @Kuznetsova2017; @Box1978].
+Each model followed the general formula:
 
-where $t$ stands for time since symptom onset, $\beta_1$ was interpreted as the estimate of linear recovery and $\beta_2$ was interpreted as a parabolic 'plateau' term [@Box1978].
-Significance of the 2^nd^ and 1^st^ order model terms was assessed by step-wise likelihood ratio test versus the lower order model (function _anova()_).
-Significance of the _beta_ estimates was determined by Wald Z test.
+$$Response \sim time + time^2 + (1|individual)$$
 
-### Definition of the long-term COVID-19 recovery phenotypes
+where $(1|individual)$ indicates the random effect of the individual and $time$ and $time^2$ indicate the first- and second-order time effect terms.
+The first-order term estimate was interpreted as a measure of the recovery speed and the second-order term estimate was used to assess the plateau/rebound effect.
+Significance of the accuracy gain of the full second-order model compared with the nested null model was determined by likelihood ratio test (LRT) versus the nested first-order and null models, respectively.
+Likelihood ratio $\lambda$ statistic (full versus null model) was used as an effect size measure.
+Individuals from the HACT or CovILD studies with the complete longitudinal symptom record were included in the kinetic modeling tasks.
+Results of the kinetic modeling were adjusted for multiple comparisons with Benjamini-Hochberg method [@Benjamini1995].
 
-Clusters of HACT study participants with COVID-19 complaints present for at least 3 months (Austria: n = `r nrow(pheno$analysis_tbl$survey_at)`, Italy: n = `r nrow(pheno$analysis_tbl$survey_it)`) were defined in respect to the most frequent persistent features (present in at least 25% of participants for at least 28 days[@Sudre2021; @Sahanic2021]: impaired concentration, tachypnea, hyposmia, hypogeusia, fatigue, tiredness and forgetfulness; re-coded as 0: absent, 1: present).
-The clustering procedure employed a combined self-organizing map (SOM, `r pheno$som_grid$xdim` $\times$ `r pheno$som_grid$ydim` hexagonal grid, gaussian neighborhood function, Jaccard distance between the symptoms) and hierarchical clustering (HCL, Ward D2 algorithm, manhattan distance between the SOM nodes) [@Kohonen1995; @Vesanto2000; @Vesanto1999; @Hufner2021].
-The clustering was accomplished with tools provided by _kohonen_ package [@Wehrens2018] and home-developed functions (https://github.com/PiotrTymoszuk/SOM_tools).
-The optimal number of final clusters was determined by the bend of the within-sum-of-square curve.
-Graphical representation of the clustering procedure training and cluster dentition in presented in __Supplementary Figure \@ref(fig:fig-clustering)__,
+## Symptom-symptom distances and multi-dimensional scaling
 
-### Univariable modeling
+To assess co-occurrence or exclusivity of symptoms, simple matching distances between manifestations of each acute COVID-19, long COVID and PASC in the HACT study cohorts were calculated (package _scrime_ and development package _clustTools_ [https://github.com/PiotrTymoszuk/clustTools]) [@Boriah2008; @Schwender2018].
+Subsequently, the distance matrix was subjected to multi-dimensional scaling (MDS, k = 2 dimensions, package _stats_, function _cmdscale()_).
+Association of specific symptoms was assessed by visual analysis of MDS coordinate plots.
 
-To identify factors correlating with the risk of persistent long-term self-reported hyposmia (3 months or longer in the HACT study or present at the 100-day follow-up visit in the CovILD cohort) within the set of candidate parameters recorded during acute COVID-19 or, for the CovILD study, at the 60-day follow-up visit as well, a series of univariable logistic models was used.
-for the HACT study, the models were weighted for age and sex weights based on the age and sex distribution of the general COVID-19 convalescent population in Tyrol/Austria and Italy [@Sahanic2021; @IstitutoSuperiorediSanitaISS; @AustrianAgencyforHealthandFoodSafetyAGES].
-To account for the effects of varying diagnosis - survey time in the HACT study, the stratified observation time variable was included as a confounder in each model [@Sahanic].
-Odds ratio significance was determined by Wald Z test.
-Logistic modeling and visual model quality control was accomplished with home-developed wrappers (https://github.com/PiotrTymoszuk/lm_qc_tools) around base R modeling tools and diagnostic functions provided by _broom_ package [@Robinson2014].
+### Apriori analysis of acute COVID-19, long COVID and PASC symptoms in the HACT study
 
-# Data availability
+Frequent combinations of symptoms of long COVID and PASC in the HACT study cohorts were identified with the apriori algorithm (package _arules_) [@Agrawal1993; @Hahsler2005] with the minimal support cutoff of 0.1, 2 - 10 item transaction length, confidence > 0.8 and lift > 2.
+The support statistic were used to estimate the symptom combination frequency.
+The confidence value was treated as an estimate of conditional probability of the symptom co-occurrence.
+The lift statistic was interpreted as a measure of the symptom dependence (lift = 1, symptoms are independent).
+The complete analysis results are presented in __Supplementary Tables S\@ref(tab:tab-apriori-long)__ and __S\@ref(tab:tab-apriori-pasc)__.
 
-The study data is available at reasonable request to the corresponding authors.
-The source code of the R analysis pipeline is available at GitHub (https://github.com/PiotrTymoszuk/hyposmia_analsis_pipeline).
+### Clustering analysis
+
+COVID-19 recovery clusters of the training Autria (AT) HACT cohort participants in respect to symptom-specific recovery times (__Figure 1A__) were defined with the PAM (partitioning around medoids) algorithm and Euclidean distance statistic (packages _cluster_, _philentropy_ and development package _clustTools_ [https://github.com/PiotrTymoszuk/clustTools]) [@Drost2018; @Schubert2019].
+The set of participants with the complete clustering variable set was included in the analysis.
+The symptom recovery times were not subjected to any type of pre-processing.
+The choice of the clustering procedure was motivated by the analysis of the clustering variance (ratio of the total between-cluster to total sum of squares) and clustering structure stability in 10-fold cross-validation (metric: rate of correct cluster assignment, cluster assignment predicted by k = 5 nearest neighbors label propagation algorithm, package _clustTools_) [@Lange2004; @Leng2014] for several clustering algorithms as presented in __Supplementary Figure S\@ref(fig:fig-clust-devel)A__.
+The optimal number of clusters was determined by the bend of the total within-cluster sum of squares curve (__Supplementary Figure S\@ref(fig:fig-clust-devel)B__, package _factoextra_) [@Kassambara2020].
+Permutation importance of specific clustering variables was investigated by calculating difference in clustering variance (ratio of total between-cluster sum of squares to total sum of squares) between the initial clustering object and the clustering object with the given variable reshuffled at random (package _clustTools_).
+Assignment of the Italy HACT cohort participants to the recovery clusters was accomplished with k-nearest neighbors label propagation algorithm (k = 5) [@Leng2014].
+The clustering efficacy in the training AT cohort and the test IT cohort measured by clustering variance statistic defined above was similar (AT: `r clustTools::var(part_clust$clust_north)$frac_var %>% signif(2)`, IT: `r clustTools::var(part_clust$clust_south)$frac_var %>% signif(2)`).
+
+## Data and source code availability
+
+The raw data files will be made available upon request. The entire analysis pipeline was published at https://github.com/PiotrTymoszuk/hyposmia_analsis_pipeline.
 
@@
 
-```{r tab-study-vars}
+```{r tab-hact-vars, tab.cap = 'HACT study variables.'}
 
-suppl_tables$study_vars[1, 1] %>% 
-  mutate(`Variable in R code` = '') %>% 
-  kable(caption = '\\textbf{Study variables.}\\\\Variable in R code: variable name in the analyis pipeline, Label: variable name used in plots and tables, Independent modeling variable: indicator if the variable was used for univariable hyposmia risk modeling.\\\\The table is available online.',
-        format = 'latex', 
-        escape = T, 
-        booktabs = T, 
-        longtable = F, 
-        col.names = '')
+suppl_tbl$hact_vars %>% 
+  select(- Stratification) %>% 
+  flextable %>% 
+  width(1, width = 3.7, unit = 'cm') %>% 
+  width(2, width = 3.9, unit = 'cm') %>% 
+  width(3, width = 1.8, unit = 'cm')  %>% 
+  width(4, width = 8.7, unit = 'cm') %>% 
+  theme_vanilla %>% 
+  font(part = 'all', fontname = 'Cambria') %>% 
+  fontsize(size = 10, part = 'all')
 
@@
 
-```{r tab-kinetic}
+\newpage
 
-suppl_tables$kinetic_modeling[1, 1] %>% 
-  mutate(Study = '') %>% 
-  kable(caption = '\\textbf{Results of symptom kinetic modeling.}\\\\OR: odds ratio with 95$\\%$ confidence interval, Significance OR: significance of the OR estimate obtained by Benjamini-Hochberg-corrected Wald Z test, Significance LRT: model term significance obtained by Benjamini-Hochberg-corrected likelihood ratio test.\\\\The table is available online.',
-        format = 'latex', 
-        escape = T, 
-        booktabs = T, 
-        longtable = F, 
-        col.names = '')
+```{r tab-covild-vars, tab.cap = 'CovILD study variables.'}
+
+suppl_tbl$covild_vars %>% 
+  flextable %>% 
+  width(1, width = 4.5, unit = 'cm') %>% 
+  width(2, width = 4.5, unit = 'cm') %>% 
+  width(3, width = 2, unit = 'cm')  %>% 
+  theme_vanilla %>% 
+  font(part = 'all', fontname = 'Cambria') %>% 
+  fontsize(size = 10, part = 'all')
 
@@
 
-```{r tab-similarity}
+\newpage
 
-suppl_tables$hyposmia_similarity[1, 1] %>% 
-  mutate(Symptom = '') %>% 
-  kable(caption = '\\textbf{Values of cosine similarity coefficients for self-reported hyposmia and other COVID-19 symptoms.}\\\\The table is available online.', 
-        format = 'latex', 
-        escape = T, 
-        booktabs = T, 
-        longtable = F, 
-        col.names = '')
+```{r tab-apriori-long, tab.cap = 'Results of apriori analysis for long COVID-19 symptoms in the HACT study .
+Top 10 most frequent symptom combinations (highest support value) in the AT (Austria) and IT (Italy) cohort are presented.
+The full results are available as a supplementary Excel table.'}
+
+suppl_tbl$apriori_long %>% 
+  group_by(Cohort) %>% 
+  top_n(10, Support) %>% 
+  ungroup %>% 
+  arrange(Cohort, Support) %>% 
+  select(-Time) %>% 
+  flextable %>% 
+  width(1, width = 1.5, unit = 'cm') %>% 
+  width(2, width = 9.5, unit = 'cm') %>% 
+  width(3:5, width = 2.2, unit = 'cm')  %>% 
+  theme_vanilla %>% 
+  font(part = 'all', fontname = 'Cambria') %>% 
+  fontsize(size = 10, part = 'all')
 
@@
 
-```{r tab-uni-modeling}
+\newpage
 
-suppl_tables$uni_modeling[1, 1] %>% 
-  mutate(Study = '') %>% 
-  kable(caption = '\\textbf{Univariable modeling of long-term persistent hyposmia risk.}\\\\OR: odds ratio with 95$\\%$ confidence interval, Significance OR: significance of the OR estimate obtained by Benjamini-Hochberg-corrected Wald Z test.\\\\The table is available online.',
-        format = 'latex', 
-        escape = T, 
-        booktabs = T, 
-        longtable = F, 
-        col.names = '')
+```{r tab-apriori-pasc, tab.cap = 'Results of apriori analysis for PASC symptoms in the HACT study.
+Top 10 most frequent symptom combinations (highest support value) in the AT (Austria) and IT (Italy) cohort are presented.
+The full results are available as a supplementary Excel table.'}
+
+suppl_tbl$apriori_pasc %>% 
+  group_by(Cohort) %>% 
+  top_n(10, Support) %>% 
+  ungroup %>% 
+  arrange(Cohort, Support) %>% 
+  select(-Time) %>% 
+  flextable %>% 
+  width(1, width = 1.5, unit = 'cm') %>% 
+  width(2, width = 9.5, unit = 'cm') %>% 
+  width(3:5, width = 2.2, unit = 'cm')  %>% 
+  theme_vanilla %>% 
+  font(part = 'all', fontname = 'Cambria') %>% 
+  fontsize(size = 10, part = 'all')
+
+```
+
+\newpage
+
+```{r tab-clust-base-at, tab.cap = 'Demographic and baseline clinical characteristic at the COVID-19 onset of the HACT study participants assigned to the recovery clusters, Austria (AT) cohort.'}
+
+suppl_tbl$baseline_clusters_north %>% 
+  flextable %>% 
+  width(1, width = 3.5, unit = 'cm') %>% 
+  width(2:4, width = 3.7, unit = 'cm') %>% 
+  width(5, width = 2.6, unit = 'cm') %>% 
+  width(6, width = 2.2, unit = 'cm') %>% 
+  footnote(1, 5:6, value = as_paragraph('Categorical variables: \u03C7\u00B2 test with Cramer V effect size statistic.
+Numeric variables: Kruskal-Wallis test with \u03B7\u00B2 effect size statistic.
+P values corrected form multiple testing with Benjamini-Hochberg method.'), ref_symbols = 'a', part = 'header') %>%
+  footnote(3, 1, value = as_paragraph('BMI: body mass index, overweight > 25 kg/m\u00B2, obesity > 30 kg/m\u00B2,'), ref_symbols = 'b', part = 'body') %>% 
+  footnote(15, 1, value = as_paragraph('Frequent respiratory infections, > 2 per year.'), ref_symbols = 'c', part = 'body') %>% 
+  footnote(16, 1, value = as_paragraph('Frequent bacterial infections with antibiotic therapy, > 2 per year.'), ref_symbols = 'd', part = 'body') %>% 
+  theme_vanilla %>%
+  font(part = 'all', fontname = 'Cambria') %>% 
+  fontsize(size = 10, part = 'all')
+
+```
+
+\newpage
+
+```{r tab-clust-base-it, tab.cap = 'Demographic and baseline clinical characteristic at the COVID-19 onset of the HACT study participants assigned to the recovery clusters, Italy (IT) cohort.'}
+
+suppl_tbl$baseline_clusters_south %>% 
+  flextable %>% 
+  width(1, width = 3.5, unit = 'cm') %>% 
+  width(2:4, width = 3.7, unit = 'cm') %>% 
+  width(5, width = 2.6, unit = 'cm') %>% 
+  width(6, width = 2.2, unit = 'cm') %>% 
+  footnote(1, 5:6, value = as_paragraph('Categorical variables: \u03C7\u00B2 test with Cramer V effect size statistic.
+Numeric variables: Kruskal-Wallis test with \u03B7\u00B2 effect size statistic.
+P values corrected form multiple testing with Benjamini-Hochberg method.'), ref_symbols = 'a', part = 'header') %>%
+  footnote(3, 1, value = as_paragraph('BMI: body mass index, overweight > 25 kg/m\u00B2, obesity > 30 kg/m\u00B2,'), ref_symbols = 'b', part = 'body') %>% 
+  footnote(15, 1, value = as_paragraph('Frequent respiratory infections, > 2 per year.'), ref_symbols = 'c', part = 'body') %>% 
+  footnote(16, 1, value = as_paragraph('Frequent bacterial infections with antibiotic therapy, > 2 per year.'), ref_symbols = 'd', part = 'body') %>% 
+  theme_vanilla %>%
+  font(part = 'all', fontname = 'Cambria') %>% 
+  fontsize(size = 10, part = 'all')
+
+```
+
+\newpage
+
+```{r tab-clust-reco-at, tab.cap = 'COVID-19 course and recovery in the HACT study participants assigned to the recovery clusters, Austria (AT) cohort.'}
+
+suppl_tbl$cov_clusters_north %>% 
+  flextable %>% 
+  width(1, width = 3.5, unit = 'cm') %>% 
+  width(2:4, width = 3.7, unit = 'cm') %>% 
+  width(5, width = 2.6, unit = 'cm') %>% 
+  width(6, width = 2.2, unit = 'cm') %>% 
+  footnote(1, 5:6, value = as_paragraph('Categorical variables: \u03C7\u00B2 test with Cramer V effect size statistic.
+Numeric variables: Kruskal-Wallis test with \u03B7\u00B2 effect size statistic.
+P values corrected form multiple testing with Benjamini-Hochberg method.'), ref_symbols = 'a', part = 'header') %>%
+  footnote(3, 1, value = as_paragraph('Number of acute COVID-19 symptoms.'), ref_symbols = 'b', part = 'body') %>% 
+  footnote(4, 1, value = as_paragraph('Number of long COVID symptoms.'), ref_symbols = 'c', part = 'body') %>% 
+  footnote(7, 1, value = as_paragraph('Self-reported incomplete recovery.'), ref_symbols = 'd', part = 'body') %>% 
+  footnote(11, 1, value = as_paragraph('ANX: anxiety.'), ref_symbols = 'e', part = 'body') %>% 
+  footnote(12, 1, value = as_paragraph('DPR: depression.'), ref_symbols = 'f', part = 'body') %>% 
+  footnote(14, 1, value = as_paragraph('OMH score: overall mental health impairment score'), ref_symbols = 'g', part = 'body') %>% 
+  footnote(15, 1, value = as_paragraph('QoL score: quality of life impairment score'), ref_symbols = 'h', part = 'body') %>%
+  theme_vanilla %>%
+  font(part = 'all', fontname = 'Cambria') %>% 
+  fontsize(size = 10, part = 'all')
+
+```
+
+\newpage
+
+```{r tab-clust-reco-it, tab.cap = 'COVID-19 course and recovery in the HACT study participants assigned to the recovery clusters, Italy (AT) cohort.'}
+
+suppl_tbl$cov_clusters_south %>% 
+  flextable %>% 
+  width(1, width = 3.5, unit = 'cm') %>% 
+  width(2:4, width = 3.7, unit = 'cm') %>% 
+  width(5, width = 2.6, unit = 'cm') %>% 
+  width(6, width = 2.2, unit = 'cm') %>% 
+  footnote(1, 5:6, value = as_paragraph('Categorical variables: \u03C7\u00B2 test with Cramer V effect size statistic.
+Numeric variables: Kruskal-Wallis test with \u03B7\u00B2 effect size statistic.
+P values corrected form multiple testing with Benjamini-Hochberg method.'), ref_symbols = 'a', part = 'header') %>%
+  footnote(3, 1, value = as_paragraph('Number of acute COVID-19 symptoms.'), ref_symbols = 'b', part = 'body') %>% 
+  footnote(4, 1, value = as_paragraph('Number of long COVID symptoms.'), ref_symbols = 'c', part = 'body') %>% 
+  footnote(7, 1, value = as_paragraph('Self-reported incomplete recovery.'), ref_symbols = 'd', part = 'body') %>% 
+  footnote(11, 1, value = as_paragraph('ANX: anxiety.'), ref_symbols = 'e', part = 'body') %>% 
+  footnote(12, 1, value = as_paragraph('DPR: depression.'), ref_symbols = 'f', part = 'body') %>% 
+  footnote(14, 1, value = as_paragraph('OMH score: overall mental health impairment score'), ref_symbols = 'g', part = 'body') %>% 
+  footnote(15, 1, value = as_paragraph('QoL score: quality of life impairment score'), ref_symbols = 'h', part = 'body') %>%
+  theme_vanilla %>%
+  font(part = 'all', fontname = 'Cambria') %>% 
+  fontsize(size = 10, part = 'all')
 
@@
 
-```{r fig-beta, fig.width = mm_inch(suppl_figures$kinetic_beta$w), fig.height = mm_inch(suppl_figures$kinetic_beta$h), dpi = 600, fig.cap = 'Symptom recovery and chronicity term estimates obtained by kinetic modeling in the HACT study.'}
+```{r fig-hact-freq, fig.width = mm_inch(suppl$hact_freq$w), fig.height = mm_inch(suppl$hact_freq$h), fig.cap = 'Frequency of COVID-19 symptoms in acute COVID-19, long COVID and PASC in the HACT study.'}
 
-suppl_figures$kinetic_beta$plot
+suppl$hact_freq$plot
 
@@
 
-__Supplementary Figure \@ref(fig:fig-beta). Symptom recovery and chronicity term estimates obtained by kinetic modeling in the HACT study.__
+__Supplementary Figure S\@ref(fig:fig-hact-freq). Frequency of COVID-19 symptoms in acute COVID-19, long COVID and PASC in the HACT study.__
 
-Symptom frequency during acute COVID-19 and recovery (0, 7, 14, 28 and 90 days post symptom onset) in the Austria (`r kinet$models$survey$survey_at %>% map_dbl(length) %>% paste_range`) and Italy (`r kinet$models$survey$survey_at %>% map_dbl(length) %>% paste_range`) HACT cohorts were modeled with mixed-effect logistic regression (__Supplementary Table \@ref(tab:tab-kinetic)__).
-Significance of the OR estimates was determined by Wald Z test.
-P values were corrected for multiple comparisons with Benjamini-Hochberg method.
-OR estimate values with 95% confidence intervals for the top 10 symptoms with the most significant second order terms as candidate long-term persistent features are presented as points and whiskers in the Forest plots.
-Dim.: diminished, imp.: impaired.
+Frequency of symptoms of acute COVID-19 (first 14 days post clinical onset), long COVID ($\geq$ 28 days) and PASC (post-acute sequelae of COVID-19, $\geq$ 90 days) in the AT (Austria, __A__) and IT (Italy, __B__) HACT study cohorts expressed as percentages of the cohort.
+Point size and color represents the percentage.
+Numbers of complete observations are indicated under the plot.
+
+Dim. appetite: diminished appetite, Imp. concentration: impaired concentration, Imp. walk: impaired walk, Imp. FMS: impaired fine motor skills.
 
@@
 
-```{r fig-low-chronicity, fig.width = mm_inch(suppl_figures$low_chronicity$w), fig.height = mm_inch(suppl_figures$low_chronicity$h), dpi = 600, fig.cap = 'Actual and predicted frequency of self-reported shivering and fever as symptoms characterized by an instant recovery in the HACT study.'}
+```{r fig-hact-recovery, fig.width = mm_inch(suppl$hact_recovery$w), fig.height = mm_inch(suppl$hact_recovery$h), fig.cap = 'Kinetic of recovery from leading acute COVID-19 symptoms in the HACT study.'}
 
-suppl_figures$low_chronicity$plot
+suppl$hact_recovery$plot
 
@@
 
-__Supplementary Figure \@ref(fig:fig-low-chronicity). Actual and predicted frequency of self-reported shivering and fever as symptoms characterized by an instant recovery in the HACT study.__
+__Supplementary Figure S\@ref(fig:fig-hact-recovery). Kinetic of recovery from leading acute COVID-19 symptoms in the HACT study.__
 
-Frequencies of self-reported shivering (__A__) and fever (__B__) during acute COVID-19 and recovery in the HACT study (0, 7, 14, 28 and 90 days post symptom onset) were modeled with mixed-effect logistic regression (__Supplementary Table \@ref(tab:tab-kinetic)__).
-Points and solid lines represent the actual frequencies.
-The model predictions are displayed as dashed lines.
-The numbers of complete observations are shown under the plots.
+Percentages of individuals with fever (__A__), diminished appetite (__B__), joint pain (__C__), muscle pain (__D__), fatigue (__E__) and tachypnea (__F__) in the AT (Austria) and IT (Italy) HACT study cohorts at particular time points after clinical onset.
+Numbers of complete observations are indicated under the plots.
 
@@
 
-```{r fig-high-chronicity, fig.width = mm_inch(suppl_figures$high_chronicity$w), fig.height = mm_inch(suppl_figures$high_chronicity$h), dpi = 600, fig.cap = 'Actual and predicted frequency of self-reported fatigue, tiredness and taste disorders as symptoms characterized by a protracted recovery in the HACT study.'}
+```{r fig-covild-freq, fig.width = mm_inch(suppl$covild_freq$w), fig.height = mm_inch(suppl$covild_freq$h), fig.cap = 'Symptom frequency in ambulatory, moderate and severe COVID-19 subsets of the CovILD study.'}
 
-suppl_figures$high_chronicity$plot
+suppl$covild_freq$plot
 
@@
 
-__Supplementary Figure \@ref(fig:fig-high-chronicity).
-Actual and predicted frequency of self-reported fatigue, tiredness and taste disorders as symptoms characterized by a protracted recovery in the HACT study.__
+__Supplementary Figure S\@ref(fig:fig-covild-freq). Symptom frequency in ambulatory, moderate and severe COVID-19 subsets of the CovILD study.__
 
-Frequencies of self-reported fatigue (__A__), tiredness (__B__) and hypo- or ageusia (__C__) during acute COVID-19 and recovery in the HACT study (0, 7, 14, 28 and 90 days post symptom onset) were modeled with mixed-effect logistic regression (__Supplementary Table \@ref(tab:tab-kinetic)__).
-Points and solid lines represent the actual frequencies.
-The model predictions are displayed as dashed lines.
-The numbers of complete observations are shown under the plots.
+Frequency of symptoms during acute COVID-19 and at the 60-, 100-, 180- and 360-day follow-ups in ambulatory (__A__), moderate (__B__) and severe COVID-19 (__C__) participants of the CovILD study expressed as percentages of individuals with the complete longitudinal data set.
+Point size and color represents the percentage.
+Numbers of complete observations are indicated under the plots.
 
@@
 
-```{r fig-hypo-sniffing, fig.width = mm_inch(suppl_figures$hypo_sniffing$w), fig.height = mm_inch(suppl_figures$hypo_sniffing$h), dpi = 600, fig.cap = 'Frequencies and co-occurrence of self-reported and sniffing stick test hyposmia in the CovILD study.'}
+```{r fig-covild-recovery, fig.width = mm_inch(suppl$covild_recovery$w), fig.height = mm_inch(suppl$covild_recovery$h), fig.cap = 'Kinetic of recovery from smell disorders, reduced performance and dyspnea in ambulatory, moderate and severe COVID-19 subsets of the CovILD study.'}
 
-suppl_figures$hypo_sniffing$plot
+suppl$covild_recovery$plot
 
@@
 
-__Supplementary Figure \@ref(fig:fig-hypo-sniffing). Frequencies and co-occurrence of self-reported and sniffing stick test hyposmia in the CovILD study.__
+__Supplementary Figure S\@ref(fig:fig-covild-recovery).
+Kinetic of recovery from smell disorders, reduced performance and dyspnea in ambulatory, moderate and severe COVID-19 subsets of the CovILD study.__
 
-Frequencies of self-reported hyposmia, moderate-to-severe (< 13 points) and severe hyposmia (< 8 points) detected by the sniffing stick test in the CovILD study at the 100-day follow-up visit were compared for the convalescents of ambulatory, moderate and severe COVID-19.
-
-__(A, B)__ Frequencies of moderate-to-severe and severe sniffing test hyposmia in the participants declaring self-reported hyposmia or claiming hyposmia-free.
-Bars are labeled with the frequencies within the investigated subset.
-
-__(C)__ Values of Cohen's $\kappa$ statistic for the concordance of self-reported and sniffing-test hyposmia in the COVID-19 severity strata and the entire CovILD cohort.
-Whiskers represent 95$\%$ confidence intervals.
+Percentages of individuals with the complete longitudinal data set suffering from smell disorders (__A__), reduced physical performance (__B__) and dyspnea (__C__) in the ambulatory, moderate and severe COVID-19 subsets of the CovILD study during acute COVID-19 and at the 60-, 100-, 180- and 360-day follow-ups.
+Numbers of complete observations are indicated under the plots.
 
@@
 
-```{r fig-hypo-knn, fig.width = mm_inch(suppl_figures$hyposmia_knn$w), fig.height = mm_inch(suppl_figures$hyposmia_knn$h), dpi = 600, fig.cap = 'Five closest co-occurring symptoms of the self-reported hyposmia during COVID-19 recovery.'}
+```{r fig-hyposmia-rater100, fig.width = mm_inch(suppl$hyposmia_rater100$w), fig.height = mm_inch(suppl$hyposmia_rater100$h), fig.cap = 'Rates of self-reported hyposmia and hyposmia in sniffing stick test at 3-month post COVID-19 follow-up in the ambulatory, moderate and severe COVID-19 subsets of the CovILD study.'}
 
-suppl_figures$hyposmia_knn$plot
+suppl$hyposmia_rater100$plot
 
@@
 
-__Supplementary Figure \@ref(fig:fig-hypo-knn). Five closest co-occurring symptoms of the self-reported hyposmia during COVID-19 recovery.__
+__Supplementary Figure S\@ref(fig:fig-hyposmia-rater100).
+Rates of self-reported hyposmia and hyposmia in sniffing stick test at 3-month post COVID-19 follow-up in the ambulatory, moderate and severe COVID-19 subsets of the CovILD study.__
 
-Cosine similarity distances (1 - Cosine similarity coefficient) between hyposmia and other self-reported symptoms of COVID-19 in the HACT 
-(__A__, `r length(hypo_overlap$variables$survey) + 1` symptoms in total, Austria: n = `r hypo_overlap$analysis_tbl$survey$survey_at$ID %>% unique %>% length`, Italy: `r hypo_overlap$analysis_tbl$survey$survey_it$ID %>% unique %>% length`)
-and COVILD study 
-(__B__, `r length(hypo_overlap$variables$covild) + 1` symptoms, ambulatory: n = `r hypo_overlap$analysis_tbl$covild$A$ID %>% unique %>% length`, moderate: n = `r hypo_overlap$analysis_tbl$covild$HM$ID %>% unique %>% length`, severe COVID-19: `r hypo_overlap$analysis_tbl$covild$HS$ID %>% unique %>% length`) were calculated for the indicated time points (__Supplementary Table \@ref(tab:tab-similarity)__).
-Cosine distances for five closest co-occurring symptoms of the self-reported hyposmia at the particular time points are presented in radial plots.
-Line length, point and label color code for cosine distance.
-Dashed lines represent cosine similarity coefficient of 0.5.
+Association of self-reported and sniffing stick test hyposmia rates was investigated with Cohen's $\kappa$ statistic.
+Statistical significance was assessed with Wald's Z test corrected for multiple testing with Benjamini-Hochberg method.
+Percentages of individuals with self-reported and test hyposmia within the entire cohort (__A__), the ambulatory (__B__), moderate (__C__) and severe (__D__) COVID-19 subsets of the CovILD-19 subsets are presented in bubble plots.
+Point size and color represents the percentage.
+$\kappa$ values with 95$\%$ confidence intervals, p values and numbers of complete observations are indicated in the plot captions.
 
@@
 
-```{r fig-clustering, fig.width = mm_inch(suppl_figures$clustering$w), fig.height = mm_inch(suppl_figures$clustering$h), dpi = 600, fig.cap = 'Development of long-term recovery clusters.'}
+```{r fig-hyposmia-rater360, fig.width = mm_inch(suppl$hyposmia_rater360$w), fig.height = mm_inch(suppl$hyposmia_rater360$h), fig.cap = 'Rates of self-reported hyposmia and hyposmia in sniffing stick test at 1-year post COVID-19 follow-up in the ambulatory, moderate and severe COVID-19 subsets of the CovILD study.'}
 
-suppl_figures$clustering$plot
+suppl$hyposmia_rater360$plot
 
@@
 
-__Supplementary Figure \@ref(fig:fig-hypo-knn). Development of long-term recovery clusters.__
+__Supplementary Figure S\@ref(fig:fig-hyposmia-rater360).
+Rates of self-reported hyposmia and hyposmia in sniffing stick test at 1-year post COVID-19 follow-up in the ambulatory, moderate and severe COVID-19 subsets of the CovILD study.__
 
-HACT participants with COVID-19 symptoms present for at least three months post clinical onset (Austria: n = `r nrow(pheno$analysis_tbl$survey_at)`, Italy: n = `r nrow(pheno$analysis_tbl$survey_it)`) were clustered in respect to manifestations present in at least 25% of the cohort for 28 days or longer with a combined algorithm employing self-organizing map (SOM, `r pheno$som_grid$xdim` $\times$ `r pheno$som_grid$ydim` hexagonal units, Jaccard distance) and hierarchical clustering (HCL, Ward D2 algorithm, Euclidean distance).
+Association of self-reported and sniffing stick test hyposmia rates was investigated with Cohen's $\kappa$ statistic.
+Statistical significance was assessed with Wald's Z test corrected for multiple testing with Benjamini-Hochberg method.
+Percentages of individuals with self-reported and test hyposmia within the entire cohort (__A__), the ambulatory (__B__), moderate (__C__) and severe (__D__) COVID-19 subsets of the CovILD-19 subsets are presented in bubble plots.
+Point size and color represents the percentage.
+$\kappa$ values with 95$\%$ confidence intervals, p values and numbers of complete observations are indicated in the plot captions.
 
-__(A)__ Progress of the SOM training procedure visualized as the drop of the mean distance to the winning unit
-with the algorithm iterations.
+\newpage
 
-__(B)__ Determination of the optimal cluster number in hierarchical clustering of the SOM nodes by finding the
-bend of the total within sum of square curve.
+```{r fig-mds-acute, fig.width = mm_inch(suppl$mds_acute$w), fig.height = mm_inch(suppl$mds_acute$h), fig.cap = 'Multi-dimensional scaling analysis of acute COVID-19 symptoms in the HACT study.'}
 
-__(C)__ Assignment of the SOM nodes to the clusters defined by hierarchical clustering presented in dendrograms.
+suppl$mds_acute$plot
+
+```
+
+__Supplementary Figure S\@ref(fig:fig-mds-acute). Multi-dimensional scaling analysis of acute COVID-19 symptoms in the HACT study.__
+
+Binary symptom occurrence data for acute COVID-19 (first 14 days after clinical onset) in the HACT study Austria (AT) and Italy (IT) cohorts were subjected to two-dimensional multi-dimensional scaling (MDS) with simple matching distance (SMD) between the symptoms.
+MDS coordinates are presented in point plots.
+Selected data points are labeled with the symptom names.
+Percentages of the data set variance associated with the MDS dimensions are indicated in the plot axes.
+Numbers of complete observations are indicated under the plots.
+
+\newpage
+
+```{r fig-clust-devel, fig.width = mm_inch(suppl$clust_devel$w), fig.height = mm_inch(suppl$clust_devel$h), fig.cap = 'Definition of the COVID-19 recovery clusters and clustering feature importance in the HACT study.'}
+
+suppl$clust_devel$plot
+
+```
+
+__Supplementary Figure S\@ref(fig:fig-clust-devel). Definition of the COVID-19 recovery clusters and clustering feature importance in the HACT study.__
+
+Individuals of the training Austria (AT) cohort of the HACT study were subjected to clustering in respect to symptom-specific recovery times with the PAM (partitioning around medoids) algorithm and Euclidean distance measure.
+
+__(A)__ Comparison of performance of various algorithms (HCl: hierarchical clustering, SOM + HCl: combined self-organizing map and hierarchical clustering, k-means) and distance statistic in clustering of the training data set investigated by clustering variance (ratio of total between-cluster sum of squares to total sum of squares) and cluster assignment accuracy in 10-fold cross-validation (CV).
+
+__(B)__ Determination of the optimal cluster number in the PAM clustering of the training cohort by the bend of the total within-cluster sum of squares curve.
+
+__(C)__ Permutation importance of the clustering features (symptoms) for clustering of the training cohort expressed as the difference in clustering variance (ratio of total between-cluster sum of squares to total sum of squares) between the initial clustering object and the clustering object with the given variable reshuffled at random.
+
+\newpage
+
+```{r fig-cov-clusters, fig.width = mm_inch(suppl$cov_clusters$w), fig.height = mm_inch(suppl$cov_clusters$h), fig.cap = 'Numbers of acute COVID-19 and long COVID symptoms in the HACT study recovery clusters.'}
+
+suppl$cov_clusters$plot
+
+```
+
+__Supplementary Figure S\@ref(fig:fig-cov-clusters). Numbers of acute COVID-19 and long COVID symptoms in the HACT study recovery clusters.__
+
+The recovery clusters were defined in the training Austria cohort (AT, __Supplementary Figure S\@ref(fig:fig-clust-devel)__) and the cluster assignment predicted in the test Italy cohort (IT, __Figure 5__).
+Differences in numbers of acute COVID-19 (first 14 days after clinical onset, __A__) and long COVID symptoms ($\geq$ 28 days, __B__) between the clusters were assessed by Kruskal-Wallis test and $\eta^2$ effect size statistic.
+P values were corrected for multiple testing with Benjamini-Hochberg method.
+Symptom counts are presented in violin plots.
+Points represent single observations, orange diamonds with whiskers code for medians and interquartile ranges.
+Effect size statistics and p values are indicated in the plot caption.
+Numbers of complete observations are displayed under the plots.
+
+\newpage
+
+```{r fig-anosmia-clusters, fig.width = mm_inch(suppl$anosmia_clusters$w), fig.height = mm_inch(suppl$anosmia_clusters$h), fig.cap = 'Recovery from smell and taste disorders in the HACT study recovery clusters'}
+
+suppl$anosmia_clusters$plot
+
+```
+
+__Supplementary Figure S\@ref(fig:fig-anosmia-clusters). Recovery from smell and taste disorders in the HACT study recovery clusters__
+
+The recovery clusters were defined in the training Austria cohort (AT, __Supplementary Figure S\@ref(fig:fig-clust-devel)__) and the cluster assignment predicted in the test Italy cohort (IT, __Figure 5__).
+Differences in recovery times for smell (__A__) and taste disorders (__D__) between the clusters were assessed by Kruskal-Wallis test and $\eta^2$ effect size statistic.
+P values were corrected for multiple testing with Benjamini-Hochberg method.
+Symptom counts are presented in violin plots.
+Points represent single observations, orange diamonds with whiskers code for medians and interquartile ranges.
+Effect size statistics and p values are indicated in the plot caption.
+Numbers of complete observations are displayed under the plots.
+
+\newpage
+
+```{r fig-fatigue-clsters, fig.width = mm_inch(suppl$fatigue_clsters$w), fig.height = mm_inch(suppl$fatigue_clsters$h), fig.cap = 'Recovery from fatigue, tiredness and tachypnea in the HACT study recovery clusters.'}
+
+suppl$fatigue_clsters$plot
+
+```
+
+__Supplementary Figure S\@ref(fig:fig-fatigue-clsters). Recovery from fatigue, tiredness and tachypnea in the HACT study recovery clusters.__
+
+The recovery clusters were defined in the training Austria cohort (AT, __Supplementary Figure S\@ref(fig:fig-clust-devel)__) and the cluster assignment predicted in the test Italy cohort (IT, __Figure 5__).
+Differences in recovery times for fatigue (__A__), day tiredness (__B__) and tachypnea (__C__) between the clusters were assessed by Kruskal-Wallis test and $\eta^2$ effect size statistic.
+P values were corrected for multiple testing with Benjamini-Hochberg method.
+Symptom counts are presented in violin plots.
+Points represent single observations, orange diamonds with whiskers code for medians and interquartile ranges.
+Effect size statistics and p values are indicated in the plot caption.
+Numbers of complete observations are displayed under the plots.
+
+\newpage
+
+```{r fig-neurocog-clusters, fig.width = mm_inch(suppl$neurocog_clusters$w), fig.height = mm_inch(suppl$neurocog_clusters$h), fig.cap = 'Recovery from neurocognitive symptoms of COVID-19 in the HACT study recovery custers.'}
+
+suppl$neurocog_clusters$plot
+
+```
+
+__Supplementary Figure S\@ref(fig:fig-neurocog-clusters). Recovery from neurocognitive symptoms of COVID-19 in the HACT study recovery clusters.__
+
+The recovery clusters were defined in the training Austria cohort (AT, __Supplementary Figure S\@ref(fig:fig-clust-devel)__) and the cluster assignment predicted in the test Italy cohort (IT, __Figure 5__).
+Differences in recovery times for forgetfulness (__A__), concentration impairment (__B__) and confusion (__C__) between the clusters were assessed by Kruskal-Wallis test and $\eta^2$ effect size statistic.
+P values were corrected for multiple testing with Benjamini-Hochberg method.
+Symptom counts are presented in violin plots.
+Points represent single observations, orange diamonds with whiskers code for medians and interquartile ranges.
+Effect size statistics and p values are indicated in the plot caption.
+Numbers of complete observations are displayed under the plots.
+
+\newpage
+
+```{r fig-rec-clusters, fig.width = mm_inch(suppl$rec_clusters$w), fig.height = mm_inch(suppl$rec_clusters$h), fig.cap = 'Physical performance loss, impairment of quality of life and self-perceived complete convalescence in the HACT study recovery clusters.'}
+
+suppl$rec_clusters$plot
+
+```
+
+__Supplementary Figure S\@ref(fig:fig-rec-clusters).
+Physical performance loss, impairment of quality of life and self-perceived complete convalescence at the survey completion in the HACT study recovery clusters.__
+
+The recovery clusters were defined in the training Austria cohort (AT, __Supplementary Figure S\@ref(fig:fig-clust-devel)__) and the cluster assignment predicted in the test Italy cohort (IT, __Figure 5__).
+Differences in self-reported physical performance loss (__A__) and scoring of impairment of quality of life (QoL, __B__) at the survey completion between the clusters were assessed by Kruskal-Wallis test and $\eta^2$ effect size statistic.
+The scoring data were presented in violin plots.
+Points represent single observations, orange diamonds with whiskers code for medians and interquartile ranges.
+Differences in frequency of self-perceived complete convalescence (__C__) between the clusters were investigated by $\chi^2$ test and Cramer V effect size statistic.
+The frequencies are presented as bar plots.
+P values were corrected for multiple testing with Benjamini-Hochberg method.
+Effect size statistics and p values are indicated in the plot caption.
+Numbers of complete observations are displayed under the plots.
+
+\newpage
+
+```{r fig-psych-clusters, fig.width = mm_inch(suppl$psych_clusters$w), fig.height = mm_inch(suppl$psych_clusters$h), fig.cap = 'Rating of anxiety, depression, overall mental health impairment and stress in the HACT study recovery clusters.'}
+
+suppl$psych_clusters$plot
+
+```
+
+__Supplementary Figure S\@ref(fig:fig-psych-clusters). Rating of anxiety, depression, overall mental health impairment and stress at the survey completion in the HACT study recovery clusters.__
+
+The recovery clusters were defined in the training Austria cohort (AT, __Supplementary Figure S\@ref(fig:fig-clust-devel)__) and the cluster assignment predicted in the test Italy cohort (IT, __Figure 5__).
+Differences in rating of anxiety (ANX, __A__), depression (DPR, __B__), stress (__C__) and self-reported overall mental health impairment (OMH, __D__) at the survey completion between the clusters were assessed by Kruskal-Wallis test and $\eta^2$ effect size statistic.
+P values were corrected for multiple testing with Benjamini-Hochberg method.
+The scoring data were presented in violin plots.
+Points represent single observations, orange diamonds with whiskers code for medians and interquartile ranges.
+Effect size statistics and p values are indicated in the plot caption.
+Numbers of complete observations are displayed under the plots.
 

### 2022-06-02 analysis scripts/apriori.R

@@ -1,0 +1,141 @@
+# Apriori analysis of the symptoms of acute, sub-acute. long, pasc and chronic 
+# COVID-19 in the HACT cohorts. Limits for the rule calculation, see code.
+
+  insert_head()
+  
+# container list ----
+  
+  ap_sympt <- list()
+  
+# globals ------
+  
+  insert_msg('Globals: analysis tables')
+  
+  ## analysis tables, HACT
+  
+  ap_sympt$hact_tbl <- mod_tbl[c('north', 'south')] %>% 
+    map(filter, time %in% c(3, 14, 28, 90)) %>% 
+    map(dlply, 'time', select, -ID, -time) %>% 
+    unlist(recursive = FALSE) %>% 
+    map(~filter(.x, complete.cases(.x))) %>% 
+    map(as.matrix)
+  
+  ## transaction objects
+  
+  ap_sympt$hact_trasactions <- ap_sympt$hact_tbl %>% 
+    map(as, 'transactions')
+  
+# Calculating the rules -----
+  
+  insert_msg('Rule calculation')
+  
+  plan('multisession')
+  
+  ap_sympt$rules <- ap_sympt$hact_trasactions %>% 
+    future_map(apriori, 
+               parameter = list(support = 0.1, 
+                                maxlen = 10, 
+                                minlen = 2, 
+                                confidence = 0.8), 
+               .options = furrr_options(seed = TRUE)) %>% 
+    map(~.x[!is.redundant(.x)]) %>% 
+    map(~subset(.x, subset = lift > 2))
+
+# Exporting the results into a data frame, adding nicer symptom names -----
+  
+  insert_msg('Exporting the results to data frames')
+  
+  ap_sympt$rules_tbl <- ap_sympt$rules %>% 
+    future_map(DATAFRAME, 
+               .options = furrr_options(seed = TRUE)) %>% 
+    future_map(clear_rule_tbl)
+  
+  plan('sequential')
+  
+# Plotting the confidence versus support ------
+  
+  insert_msg('Confidence versus support plots')
+  
+  ap_sympt$conf_supp_plots <- list(data = ap_sympt$rules_tbl %>% 
+                                     map(mutate, 
+                                         trans_lab = stri_replace(trans_lab, 
+                                                                  fixed = ' \u2192 ', 
+                                                                  replacement = '\u2192\n')), 
+                                   plot_title = rep(c('Acute CoV', 
+                                                      'Sub-acute CoV', 
+                                                      'Long CoV', 
[…]

### 2022-06-02 analysis scripts/clust_characteristic.R

@@ -1,0 +1,261 @@
+# Demographic, cinical and recovery characteristic of the participant clusters
+
+  insert_head()
+  
+# container list ------
+  
+  clust_chara <- list()
+  
+# globals ------
+  
+  insert_msg('Globals setup')
+  
+  ## types of the tests and plots
+  
+  clust_chara$var_types <- globals[c('demo_vars', 
+                                     'clinic_vars', 
+                                     'cov_vars', 
+                                     'psych_var')] %>% 
+    reduce(c) %>% 
+    map_dfr(~tibble(variable = .x, 
+                    type_test = ifelse(is.numeric(hact$north[[.x]]), 
+                                       'kruskal_test', 'chisq_test'), 
+                    type_plot = ifelse(is.numeric(hact$north[[.x]]), 
+                                       'violin', 'stack'), 
+                    y_lab = ifelse(is.numeric(hact$north[[.x]]), 
+                                   translate_var(.x, 
+                                                 out_value = 'axis_lab', 
+                                                 dict = hact$dict), 
+                                   '% of cluster')))
+  
+  clust_chara$num_vars <- clust_chara$var_types %>% 
+    filter(type_test == 'kruskal_test') %>% 
+    .$variable
+  
+  ## analysis tables 
+  
+  clust_chara$analysis_tbl <- clust_ft$analysis_tbl
+  
+  ## recovery variables
+  
+  clust_chara$rec_vars <- c('sum_symptoms_long', 
+                            'weight_loss_kg', 
+                            'incomplete_covelescence', 
+                            'new_medication_fup', 
+                            'rehabilitation_fup_needed', 
+                            'perf_impairment', 
+                            'life_quality_score', 
+                            'mental_health_score', 
+                            'phq_anxiety_score', 
+                            'phq_depression_score', 
+                            'stress_score')
+  
+# descriptive stats -------
+  
+  insert_msg('Descriptive stats')
+  
+  clust_chara$desc_stats <- clust_chara$analysis_tbl %>% 
+    map(~explore(.x, 
+                 split_factor = 'clust_id', 
+                 variables = clust_chara$var_types$variable, 
+                 what = 'table', 
+                 pub_styled = TRUE)) %>% 
+    map(reduce, left_join, by = 'variable') %>% 
+    map(set_names, c('variable', 'clust_1', 'clust_2', 'clust_3'))
+  
+# testing ------
+  
[…]

### 2022-06-02 analysis scripts/clust_devel.R

@@ -1,0 +1,139 @@
+# Choice of the best clustering algorithm for the participant clustering 
+# by the symptom-specific recovery time. The clustering is done in the AT cohort
+# of the HACT study. Pre-processing: mean-centered normalization.
+
+  insert_head()
+  
+# container list -----
+  
+  cl_devel <- list()
+  
+# globals: analysis tables -----
+  
+  insert_msg('Analysis tables')
+  
+  cl_devel$analysis_tbl <- rec_time[c('north', 'south')] %>% 
+    map(~filter(.x, complete.cases(.x))) %>% 
+    map(column_to_rownames, 'ID')
+  
+  cl_devel$test_dist <- c('euclidean', 'manhattan', 'sumofsquares', 'cosine')
+  
+  cl_devel$som_x <- floor(sqrt(5 * sqrt(nrow(cl_devel$analysis_tbl$north))))
+  
+# generating clust_analysis objects ----
+  
+  insert_msg('Creating the clustering objects')
+  
+  ## hierarchical clustering
+  
+  cl_devel$algos[paste0('hcl_', cl_devel$test_dist)] <- 
+    cl_devel$test_dist %>% 
+    map(~hcluster(data = cl_devel$analysis_tbl$north, 
+                  distance_method = .x, 
+                  hc_method = 'ward.D2', 
+                  k = 3, 
+                  seed = 1234))
+  
+  ## K-Means
+  
+  cl_devel$algos[paste0('kmeans_', cl_devel$test_dist)] <- 
+    cl_devel$test_dist %>% 
+    map(~kcluster(data = cl_devel$analysis_tbl$north, 
+                  distance_method = .x, 
+                  clust_fun = 'kmeans', 
+                  k = 3, 
+                  seed = 1234))
+  
+  ## PAM
+  
+  cl_devel$algos[paste0('pam_', cl_devel$test_dist)] <- 
+    cl_devel$test_dist %>% 
+    map(~kcluster(data = cl_devel$analysis_tbl$north, 
+                  distance_method = .x, 
+                  clust_fun = 'pam', 
+                  k = 3, 
+                  seed = 1234))
+  
+  ## combi SOM + HCl clustering
+  
+  cl_devel$algos[paste0('combi_', cl_devel$test_dist)] <- 
+    cl_devel$test_dist %>% 
+    map(~combi_cluster(data = cl_devel$analysis_tbl$north, 
+                       distance_som = .x, 
+                       xdim = cl_devel$som_x, 
[…]

### 2022-06-02 analysis scripts/clust_features.R

@@ -1,0 +1,173 @@
+# Characterized the clustering feature levels in the participant clusters.
+
+  insert_head()
+  
+# container list ----
+  
+  clust_ft <- list()
+  
+# globals -----
+  
+  insert_msg('Globals setup')
+  
+  ## analysis tables
+  
+  clust_ft$analysis_tbl <- part_clust[c('clust_north', 
+                                        'clust_south')] %>% 
+    map(~left_join(model.frame(.x) %>% 
+                     rownames_to_column('ID'), 
+                   .x$clust_assignment %>% 
+                     mutate(ID = observation), 
+                   by = 'ID')) %>% 
+    map(as_tibble)
+  
+  ## appending them with the demographic, clinical and recovery features
+  
+  clust_ft$analysis_tbl <- map2(clust_ft$analysis_tbl, 
+                                map(hact[c('north', 'south')], 
+                                    select, 
+                                    ID, 
+                                    all_of(globals$demo_vars), 
+                                    all_of(globals$clinic_vars), 
+                                    all_of(globals$cov_vars), 
+                                    all_of(globals$psych_var)), 
+                                left_join, by = 'ID')
+  
+  ## n numbers
+  
+  clust_ft$n_numbers <- part_clust[c('clust_north', 
+                                     'clust_south')] %>% 
+    map(ngroups)
+  
+  clust_ft$n_tags <- clust_ft$n_numbers %>% 
+    map(~map2_chr(.x[[1]], .x[[2]], paste, sep = ': n = ')) %>% 
+    map(paste, collapse = ', ') %>% 
+    map(~paste0('\n', .x))
+  
+# Descriptive stats ----
+  
+  insert_msg('Descriptive stats')
+
+  clust_ft$clust_ft_stats <- clust_ft$analysis_tbl %>% 
+    map(~explore(.x, 
+                 split_factor = 'clust_id', 
+                 variables = globals$hact_symptoms, 
+                 what = 'table', 
+                 pub_styled = TRUE)) %>% 
+    map(reduce, left_join, by = 'variable') %>% 
+    map(set_names, c('variable', 'clust_1', 'clust_2', 'clust_3'))
+  
+# Kruskal-Wallis tests ----
+  
+  insert_msg('Serial testing')
+  
+  clust_ft$clust_ft_test <- clust_ft$analysis_tbl %>% 
+    map(~compare_variables(.x, 
+                           split_factor = 'clust_id', 
+                           variables = globals$hact_symptoms, 
[…]

### 2022-06-02 analysis scripts/clustering.R

@@ -1,0 +1,146 @@
+# semi-supervised clustering of the HACT data sets by the symptom recovery times
+# AT: training cohort (PAM + eucllidean as tested), IT: test cohort.
+# kNN label propagation algorithm.
+
+  insert_head()
+  
+# container list -----
+  
+  part_clust <- list()
+  
+# globals -----
+  
+  insert_msg('Globals setup')
+  
+  part_clust$analysis_tbl <- cl_devel$analysis_tbl
+
+  part_clust$clust_north <- cl_devel$algos$pam_euclidean
+  
+  part_clust$clust_north$clust_assignment <- 
+    part_clust$clust_north$clust_assignment %>% 
+    mutate(clust_id = car::recode(clust_id, 
+                                  "'1' = 'SR'; '2' = 'RR'; '3' = 'STDR'"), 
+           clust_id = factor(clust_id, c('STDR', 'RR', 'SR'))) 
+  
+# cluster projection -----
+  
+  insert_msg('Predicting the cluster assignment in the IT cohort')
+  
+  part_clust$clust_south <- predict(part_clust$clust_north, 
+                                    newdata = part_clust$analysis_tbl$south, 
+                                    type = 'propagation', 
+                                    kNN = 5)
+  
+  part_clust$clust_south$clust_assignment <- 
+    part_clust$clust_south$clust_assignment %>% 
+    mutate(clust_id = factor(clust_id, c('STDR', 'RR', 'SR'))) 
+  
+# Diagnostic of the clustering objects ------
+  
+  insert_msg('Characteristic of the clustering objects')
+  
+  ## diagnostic plots
+  
+  part_clust$diagnostic_plots <- part_clust$clust_north %>% 
+    plot(type = 'diagnostic', 
+         cust_theme = globals$common_theme)
+  
+  ## diastance heat maps
+  
+  part_clust$heat_maps <- part_clust[c('clust_north', 
+                                       'clust_south')] %>% 
+    map(plot, 
+        type = 'heat_map', 
+        cust_theme = globals$common_theme) %>% 
+    map2(., c('AT, HACT study', 'IT, HACT study'), 
+         ~.x + 
+           labs(subtitle = paste(.y, .x$labels$subtitle, sep = ', ')) + 
+           theme(axis.text.x = element_blank(), 
+                 axis.text.y = element_blank(), 
+                 axis.ticks = element_blank()))
+  
+  ## PCA score plots
+  
[…]

### 2022-06-02 analysis scripts/hyposmia_rater.R

@@ -1,0 +1,94 @@
+# Snbiffing stick versus self-reported hyposmia
+
+  insert_head()
+  
+# container list -----
+  
+  rater <- list()
+  
+# globals: analysis tables ----
+  
+  insert_msg('Globals setup')
+  
+  rater$analysis_tbl <- list(fup100 = sst$data_100_fup, 
+                             fup360 = sst$data_360_fup) %>% 
+    map(~dlply(.x, 'cat_WHO', as_tibble)) %>% 
+    map2(., sst[c('data_100_fup', 'data_360_fup')], 
+         ~c(.x, list(cohort = .y))) %>% 
+    unlist(recursive = FALSE)
+  
+# calculation of the kappas -----
+  
+  insert_msg('Calculation of Kohen kappas')
+  
+  rater$test_results <- rater$analysis_tbl %>% 
+    map(~correlate_variables(.x, 
+                             variables = c('anosmia_sympt', 'sniff_hyposmia'), 
+                             what = 'correlation', 
+                             type = 'kappa', 
+                             ci = TRUE, 
+                             pub_styled = TRUE)) %>% 
+    map2_dfr(., names(.), ~mutate(.x, 
+                                  time = stri_split_fixed(.y, pattern = '.', simplify = TRUE)[, 1], 
+                                  subset = stri_split_fixed(.y, pattern = '.', simplify = TRUE)[, 2])) %>% 
+    re_adjust %>% 
+    mutate(plot_cap = paste(eff_size, significance, sep = ', '))
+
+# Plotting the interrater results -------
+  
+  insert_msg('Plotting the interrater results, bar plots')
+  
+  rater$bar_plots <- list(data = rater$analysis_tbl, 
+                          plot_title = rep(c('Ambulatory, CovILD study', 
+                                             'Moderate, CovILD study', 
+                                             'Severe, CovILD study', 
+                                             'CovILD cohort'), 2), 
+                          plot_subtitle = rater$test_results$plot_cap) %>% 
+    pmap(plot_correlation, 
+         variables = c('anosmia_sympt', 'sniff_hyposmia'), 
+         scale = 'percent', 
+         type = 'bar', 
+         x_lab = 'Hypo/anosmia', 
+         cust_theme = globals$common_theme) %>% 
+    map2(.,map(rater$analysis_tbl, nrow), 
+         ~.x + 
+           labs(subtitle = paste0(.x$labels$subtitle, ', n = ', .y)) + 
+           scale_fill_manual(values = c('bisque3', 'steelblue'), 
[…]

### 2022-06-02 analysis scripts/kinetic_modeling.R

@@ -1,0 +1,178 @@
+# Modeling of the recovery kinetic for particular symptoms: top 10 most common
+# symptoms of acute COVID-19 in the  HACT study and for all symptoms in the 
+# CovILD study.
+# Second order model with the participant random effect.
+# Only participants with the complete answer sets are included in the analysis.
+
+  insert_head()
+  
+# container list ----
+  
+  kin_mod <- list()
+  
+# globals ------
+  
+  insert_msg('Globals')
+  
+  kin_mod$covild_tbl <- mod_tbl$covild %>% 
+    complete_cases %>% 
+    dlply('cat_WHO', select, -cat_WHO) %>% 
+    map(as_tibble)
+  
+  kin_mod$hact_top_symptoms <- eda_sympt$hact_top$acute
+  
+# model construction for the HACST study ------
+  
+  insert_msg('Building the models, HACT')
+  
+  plan('multisession')
+
+  kin_mod$models_hact$north <- kin_mod$hact_top_symptoms %>% 
+    future_map(~safely(model_kinetic)(data = mod_tbl$north[c('ID', 'time', .x)] %>% 
+                                        complete_cases, 
+                                      response = .x, 
+                                      time = 'time', 
+                                      ID = 'ID', 
+                                      family = 'binomial', 
+                                      order = 2), 
+               .options = furrr_options(seed = TRUE))
+  
+  kin_mod$models_hact$south <- kin_mod$hact_top_symptoms %>% 
+    future_map(~safely(model_kinetic)(data = mod_tbl$south[c('ID', 'time', .x)] %>% 
+                                        complete_cases, 
+                                      response = .x, 
+                                      time = 'time', 
+                                      ID = 'ID', 
+                                      family = 'binomial', 
+                                      order = 2), 
+               .options = furrr_options(seed = TRUE))
+  
+  kin_mod$models_hact <- kin_mod$models_hact %>% 
+    map(~map(.x, ~.x$result)) %>% 
+    map(set_names, kin_mod$hact_top_symptoms) %>% 
+    map(compact)
+  
+# Builing the models for the CovILD study ------
+  
+  insert_msg('Building the models, CovILD')
+  
+  kin_mod$models_covild$A <- globals$covild_symptoms %>% 
+    future_map(~safely(model_kinetic)(data = kin_mod$covild_tbl$A[c('ID', 'time_numeric', .x)] %>% 
+                                        complete_cases, 
+                                      response = .x, 
[…]

### 2022-06-02 analysis scripts/pca.R

@@ -1,0 +1,157 @@
+# PCA and investigation of the clustering tendency by the symptom recovery times
+# done in the AT cohort of the HACT study. Done separately for acute, sub-acute
+# long, PASC and chronic CoV.
+# no pre-processing applied since the symptom duration data are anyway on the 
+# same scale (days post CoV)
+
+  insert_head()
+  
+# container list ----
+  
+  pca <- list()
+  
+# globals: analysis tables ------
+  
+  insert_msg('Globals setup')
+  
+  pca$analysis_tbl <- rec_time[c('north', 'south')] %>% 
+    map(~filter(.x, complete.cases(.x))) %>% 
+    map(column_to_rownames, 'ID')
+  
+# PCA -----
+  
+  insert_msg('PCA and visualizations')
+  
+  pca$pca_obj <- pca$analysis_tbl %>% 
+    map(reduce_data, 
+        kdim = 8, 
+        red_fun = 'pca')
+  
+  pca$pca_obj$north$loadings <- pca$pca_obj$north$loadings %>% 
+    mutate(variable = translate_var(variable, dict = hact$dict))
+  
+  pca$pca_obj$south$loadings <- pca$pca_obj$south$loadings %>% 
+    mutate(variable = translate_var(variable, dict = hact$dict))
+  
+  ## Scree plot
+  
+  pca$pca_scree_plot <- list(x = pca$pca_obj, 
+                             segment_color = globals$hact_colors) %>% 
+    pmap(plot,
+         type = 'scree', 
+         cust_theme = globals$common_theme) %>% 
+    map2(c('AT, HACT study', 
+           'IT, HACT study'), 
+         ~.x + labs(subtitle = .y))
+  
+  ## loadings
+  
+  pca$pca_loadings_plot <- list(x = pca$pca_obj, 
+                                plot_subtitle = c('AT, HACT study', 
+                                                  'IT, HACT study'), 
+                                point_color = globals$hact_colors) %>% 
+    pmap(plot,
+         type = 'loadings', 
+         cust_theme = globals$common_theme, 
+         segment_color = 'gray60')
+  
+  ## score plots
+  
+  pca$pca_score_plot <- list(x = pca$pca_obj, 
+                             plot_subtitle = c('AT, HACT study', 
+                                               'IT, HACT study'), 
+                             point_color = globals$hact_colors) %>% 
+    pmap(plot,
+         type = 'scores', 
+         cust_theme = globals$common_theme)
+  
+  ## global eigenvector plots
+  
[…]

### 2022-06-02 analysis scripts/recovery_distribution.R

@@ -1,0 +1,179 @@
+# Analyzes and visualizes the time to recovery for particular symptoms
+
+  insert_head()
+
+# container list ------
+
+  rec_dist <- list()
+  
+# globals -------
+  
+  insert_msg('Globals setup')
+  
+  rec_dist$tbl_hact <- rec_time[c('north', 'south')] %>% 
+    map2_dfr(., c('AT', 'IT'), ~mutate(.x, cohort = .y))
+  
+# descriptive stats of the recovery: individuals with the symptom present -----  
+  
+  insert_msg('Distribution stats for the rec. times, symptomatic  participants')
+  
+  rec_dist$stats_hact <- globals$hact_symptoms %>% 
+    map_dfr(~explore(rec_dist$tbl_hact[c('cohort', .x)] %>% 
+                       filter(.data[[.x]] != 0), 
+                     split_factor = 'cohort', 
+                     variables = .x, 
+                     what = 'table', 
+                     pub_styled = TRUE)) %>% 
+    reduce(left_join, by = 'variable') %>% 
+    set_names(c('variable', 'north', 'south'))
+  
+  rec_dist$stats_covild <- globals$covild_symptoms %>% 
+    map_dfr(~explore(rec_time$covild[c('cat_WHO', .x)] %>% 
+                       filter(.data[[.x]] != 0), 
+                     split_factor = 'cat_WHO', 
+                     variables = .x, 
+                     what = 'table', 
+                     pub_styled = TRUE)) %>% 
+    reduce(left_join, by = 'variable') %>% 
+    set_names(c('variable', 'A', 'HM', 'HS'))
+
+# testing, differences between the cohorts or severity strata, symptomatic ----
+  # Mann-Whitney U test or Kruskal-Wallis test
+  
+  insert_msg('Testing for the differences between the cohorts or severity')
+  
+  rec_dist$test_hact <- globals$hact_symptoms %>% 
+    map(~safely(compare_variables)(rec_dist$tbl_hact[c('cohort', .x)] %>% 
+                                     filter(.data[[.x]] != 0), 
+                                   split_factor = 'cohort', 
+                                   variables = .x, 
+                                   what = 'eff_size', 
+                                   types = 'wilcoxon_r', 
+                                   ci = FALSE, 
+                                   pub_styled = TRUE, 
+                                   adj_method = 'none')) %>% 
+    map_dfr(~.x$result) %>% 
+    re_adjust('BH')
+
+  rec_dist$test_covild <- globals$covild_symptoms %>% 
[…]

### 2022-06-02 analysis scripts/symptom_distances.R

@@ -1,0 +1,200 @@
+# Calculates distances (SMC) between particular symptoms for 
+# subsequent time points. Performs MDS to assess the symptom overlap.
+# The relation with the most overlapping symptoms is analyzed by comparing
+# the mean 5-NN SMC distances between the symptoms.
+
+  insert_head()
+  
+# container list -----
+  
+  sympt_dist <- list()
+  
+# globals -----
+  
+  insert_msg('Globals')
+  
+  ## analysis tables
+  
+  sympt_dist$hact_tbl <- mod_tbl[c('north', 'south')] %>% 
+    map(filter, time %in% c(3, 14, 28, 90)) %>%
+    map(dlply, 'time', select, -time) %>% 
+    map(set_names, c('acute', 'sub_acute', 'long', 'pasc')) %>% 
+    unlist(recursive = FALSE) %>% 
+    map(~filter(.x, complete.cases(.x))) %>% 
+    map(column_to_rownames, 'ID') %>% 
+    map(t)
+  
+  sympt_dist$covild_tbl <- mod_tbl$covild %>% 
+    dlply(c('cat_WHO', 'time_numeric'), select, -time_numeric, -cat_WHO) %>% 
+    map(~filter(.x, complete.cases(.x))) %>% 
+    map(~map_dfr(.x, function(x) if(is.factor(x)) as.numeric(x) - 1 else x)) %>% 
+    map(column_to_rownames, 'ID') %>% 
+    map(t)
+  
+  ## n numbers
+  
+  sympt_dist$n_numbers_hact <- sympt_dist$hact_tbl[c(1, 5)] %>% 
+    map_dbl(ncol) %>% 
+    set_names(c('north', 'south'))
+  
+  sympt_dist$n_numbers_covild <- sympt_dist$covild_tbl[c(1, 6, 11)] %>% 
+    map_dbl(ncol) %>% 
+    set_names(c('A', 'HM', 'HS'))
+  
+# Matrices with the SMC distances -----
+  
+  insert_msg('Matrices with the SMC distances')
+  
+  sympt_dist$dist_hact <- sympt_dist$hact_tbl %>% 
+    map(calculate_dist, 'smc')
+  
+  sympt_dist$dist_covild <- sympt_dist$covild_tbl %>% 
+    map(calculate_dist, 'smc')
+  
+# Multi-dimensional scaling -------
+  
+  insert_msg('MDS to visualize the distances')
+  
[…]

### 2022-06-02 analysis.R

@@ -1,0 +1,41 @@
+# Launches particular analysis scripts
+
+  library(plyr)
+  library(tidyverse)
+  library(exda)
+  library(soucer)
+  library(stringi)
+  library(kinet)
+  library(furrr)
+  library(clustTools)
+  library(somKernels)
+  library(ggrepel)
+  library(dbscan)
+  library(arules)
+  library(arulesViz)
+  library(microViz)
+
+  insert_head()
+  
+  source_all('./tools/project_tools.R', 
+             message = TRUE, crash = TRUE)
+  
+# analysis scripts ------
+  
+  insert_msg('Launching the analyses')
+  
+  c('./analysis scripts/recovery_distribution.R', 
+    './analysis scripts/kinetic_modeling.R', 
+    './analysis scripts/symptom_distances.R', 
+    './analysis scripts/apriori.R', 
+    './analysis scripts/pca.R', 
+    './analysis scripts/clust_devel.R', 
+    './analysis scripts/clustering.R', 
+    './analysis scripts/clust_features.R', 
+    './analysis scripts/clust_characteristic.R', 
+    './analysis scripts/hyposmia_rater.R') %>% 
+    source_all(message = TRUE, crash = TRUE)
+  
+# END ----
+  
+  insert_tail()

### 2022-06-02 exec_all.R

@@ -1,58 +1,12 @@
-# This program executes the scripts of the COVID hyposmia project
+# Launches the complete pipeline 
 
-# libraries -----
-
-  require(plyr)
-  require(tidyverse)
-
-# listing figure scripts ----
-
-  executable_scripts <- c('data_import.R', 
-                          'data_analysis.R', 
-                          'render_paper.R')
+  library(soucer)
   
-# error container ----
-  
-  exec_errors <- list()
-  
-# executing the script list ----
-  
-  for(file in executable_scripts) {
-    
-    diagn_output <- try(source(file, encoding = 'UTF-8'), silent = T) ## UTF-8 encoding cause of greek letters 
-                                                                      ##in plots and tables
-    
-    if(class(diagn_output) == 'try-error') {
-      
-      exec_errors <- c(exec_errors, 'execution failed')
-      
-    } else {
-      
-      exec_errors <- c(exec_errors, 'execution successful')
-      
-    }
-
-  }
-  
-  exec_errors <- exec_errors %>% 
-    reduce(c) %>% 
-    tibble(execution_result = .)
-  
-  exec_errors <- exec_errors %>% 
-    mutate(file = executable_scripts) %>% 
-    select(file, execution_result)
-  
-  rm(diagn_output, executable_scripts)
-  
-  message(rep('-', 60))
-  print(exec_errors)
-  message(rep('-', 60))
-  
-# saving execution results on the disc ----
-
-  exec_errors %>% 
-    write_tsv('exec_log.log')
+  print(source_all(c('import.R', 
+                     'exploration.R', 
+                     'analysis.R', 
+                     'paper.R'), 
+                   message = TRUE, 
+                   crash = FALSE))
   
   save.image()
-  
-# END ----

### 2022-06-02 exploration scripts/cohort.R

@@ -1,0 +1,121 @@
+# Characteristic of the study cohorts (HACT and CovILD): baseline, 
+# clinical history, CoV course and recovery
+
+  insert_head()
+  
+# container list ------
+  
+  eda_cohort <- list()
+  
+# globals ------
+  
+  insert_msg('Globals setup')
+  
+  ## analysis table for covild
+  
+  eda_cohort$hact_tbl <- hact[c('north', 'south')] %>% 
+    reduce(rbind)
+  
+  ## test variables for HACT
+  
+  eda_cohort$hact_vars$variable <- globals[c('demo_vars', 
+                                             'clinic_vars', 
+                                             'cov_vars', 
+                                             'psych_var')] %>% 
+    reduce(c)
+  
+  eda_cohort$hact_vars$test_type <- eda_cohort$hact_vars$variable %>% 
+    map_chr(~ifelse(is.numeric(hact$north[[.x]]), 'wilcoxon_r', 'cramer_v'))
+
+  eda_cohort$hact_vars <- as_tibble(eda_cohort$hact_vars)
+  
+  ## test variables for CovILD
+  
+  eda_cohort$covild_vars$variable <- globals$covild_vars
+  
+  eda_cohort$covild_vars$test_type <- eda_cohort$covild_vars$variable %>% 
+    map_chr(~ifelse(is.numeric(covild$data[[.x]]), 'kruskal_test', 'chisq_test'))
+  
+  eda_cohort$covild_vars <- as_tibble(eda_cohort$covild_vars)
+  
+# descriptive stats, HACT -----
+  
+  insert_msg('Descritive stats, HACT cohort')
+  
+  eda_cohort$stats_hact <- explore(eda_cohort$hact_tbl, 
+                                   split_factor = 'cohort', 
+                                   variables = eda_cohort$hact_vars$variable, 
+                                   what = 'table', 
+                                   pub_styled = TRUE) %>% 
+    reduce(left_join, by = 'variable') %>% 
+    set_names(c('variable', 'north', 'south'))
+  
+# testing, HACT ----
+  
+  insert_msg('Testing for the AT/IT differences')
+  
+  eda_cohort$test_hact <- compare_variables(eda_cohort$hact_tbl, 
+                                            split_factor = 'cohort', 
+                                            variables = eda_cohort$hact_vars$variable, 
+                                            what = 'eff_size', 
+                                            types = eda_cohort$hact_vars$test_type, 
+                                            ci = FALSE, 
[…]

### 2022-06-02 exploration scripts/symptoms.R

@@ -1,0 +1,221 @@
+# Frequency of acute, sub-acute and long and PACS symptoms.
+
+  insert_head()
+  
+# container list -----
+  
+  eda_sympt <- list()
+  
+# Globals ------
+  
+  insert_msg('Globals setup')
+  
+  ## analysis tables, HACT
+  
+  eda_sympt$hact_tbl <- mod_tbl[c('north', 'south')] %>% 
+    map2_dfr(., names(.), ~mutate(.x, cohort = .y)) %>% 
+    filter(time %in% c(3, 14, 28, 90)) %>% 
+    dlply('time', select, -ID, -time) %>% 
+    set_names(c('acute', 'sub_acute', 'long', 'pasc')) %>% 
+    map(as_tibble) %>% 
+    map(map_dfr, factor)
+
+  ## analysis tables, covILD
+  
+  eda_sympt$covild_tbl <- mod_tbl$covild %>% 
+    complete_cases %>% 
+    dlply('time_numeric', as_tibble)
+  
+  ## n numbers
+  
+  eda_sympt$hact_n_tags <- paste('\nn =', 
+                                 dlply(eda_sympt$hact_tbl$acute, 
+                                       'cohort', 
+                                       nrow) %>% 
+                                   unlist)
+  
+  eda_sympt$covild_n_tags <- paste('\nn =', 
+                                   dlply(eda_sympt$covild_tbl[[1]], 
+                                         'cat_WHO', 
+                                         nrow) %>% 
+                                     unlist)
+  
+# descriptive stats ------
+  
+  insert_msg('Descriptive stats')
+  
+  eda_sympt$hact_stats <- eda_sympt$hact_tbl %>% 
+    map(explore, 
+        split_factor = 'cohort', 
+        variables = globals$hact_symptoms, 
+        what = 'table', 
+        pub_styled = TRUE) %>% 
+    map(reduce, left_join, by = 'variable') %>% 
+    map(set_names, c('variable', 'AT', 'IT')) %>% 
+    map(map_dfr, stri_replace_all, regex = '0:.*\\n1:\\s{1}', replacement = '') %>% 
+    map(map_dfr, stri_replace_all, fixed = '% (', replacement = '% (n = ') %>% 
+    map(map_dfr, stri_replace, fixed = 'Complete', replacement = 'complete')
+  
+  eda_sympt$covild_stats <- eda_sympt$covild_tbl %>% 
+    map(explore, 
[…]

### 2022-06-02 exploration.R

@@ -1,0 +1,23 @@
+# Exploratory data analysis.
+
+  library(plyr)
+  library(tidyverse)
+  library(exda)
+  library(soucer)
+  library(stringi)
+  
+  insert_head()
+  
+  source_all('./tools/project_tools.R')
+  
+# exploration scripts -----
+  
+  insert_msg('Launching the exploration scripts')
+  
+  c('./exploration scripts/cohort.R', 
+    './exploration scripts/symptoms.R') %>% 
+    source_all(message = TRUE, crash = TRUE)
+  
+# END ------
+  
+  insert_tail()

### 2022-06-02 import.R

@@ -1,0 +1,217 @@
+# imports the pre-cleared Health after COVID (HACT, 
+# observation time at least 28 days) and CovILD project data.
+# Defines common globals, a shared format of the variable dictionaries 
+# and calculates individual times to recovery for specific symptoms.
+# The subset of the HACT study analyzed in the current project has an 
+# observation time of at least 90 days and with acute symptoms.
+
+  library(soucer)
+  library(plyr)
+  library(tidyverse)
+  library(foreign)
+  library(stringi)
+  
+  insert_head()
+  
+  source_all('./tools/project_tools.R', 
+             message = TRUE, crash = TRUE)
+  
+# container lists ------
+  
+  rec_time <- list()
+  mod_tbl <- list()
+  
+# globals ------
+  
+  insert_msg('Globals')
+  
+  source_all('./tools/project_globals.R', 
+             message = TRUE, crash = TRUE)
+  
+# unpacking the data files -----
+  
+  insert_msg('Unpacking')
+  
+  load('./data/covild.RDa')
+  load('./data/hact.RDa')
+  
+  hact[c('north', 'south')] <- hact[c('north', 'south')] %>% 
+    map(filter, 
+        obs_time >= 90, 
+        acute_covid == 'yes') %>% 
+    map(mutate, 
+        comorb_present = ifelse(comorb_absent == 'yes', 'no', 'yes'), 
+        comorb_present = factor(comorb_present, c('no', 'yes')), 
+        incomplete_covelescence = ifelse(complete_covelescence == 'yes', 
+                                         'no', 'yes'), 
+        incomplete_covelescence = factor(incomplete_covelescence, c('no', 'yes')), 
+        bmi_class_before = car::recode(bmi_class_before, 
+                                       "'overweigth' = 'overweight'"), 
+        bmi_class_before = factor(bmi_class_before, 
+                                  c('normal', 'overweight', 'obesity')), 
+        bmi_class_recent = car::recode(bmi_class_recent, 
+                                       "'overweigth' = 'overweight'"), 
+        bmi_class_recent = factor(bmi_class_recent, 
+                                  c('normal', 'overweight', 'obesity')))
+
+# Formatting the dictionaries -----
+  
+  insert_msg('Common format for the variable dictionaries')
+  
+  covild$dict <- covild$dict %>% 
+    mutate(label_short = label)
+  
+  hact$dict  <- hact$dict %>% 
[…]

### 2022-06-02 paper scripts/figures.R

@@ -1,0 +1,252 @@
+# Main text figures 
+
+  insert_head()
+  
+# container list -----
+  
+  figures <- list()
+  
+# Figure 1: CONSORT -------
+  
+  insert_msg('Figure 1: CONSORT')
+  
+  figures$consort <- plot_grid(ggdraw() + 
+    draw_image('./study consort/consort_diagram.png')) %>% 
+    as_figure('figure_1_consort', 
+              w = 180, 
+              h = 110)
+  
+# Figure 2: symptom recovery times -----
+  
+  insert_msg('Figure 2: symptom recovery times')
+  
+  ## emboldening the hyposmia and hypogeusia
+
+  figures$sympt_recovery$upper_panel <- rec_dist$summ_plot_hact %>% 
+    map(~.x + 
+          scale_y_discrete(labels = function(x) embolden_scale(x, c('Hypo/anosmia', 'Hypo/ageusia'))) + 
+          theme(axis.text.y = element_markdown())) %>% 
+    plot_grid(plotlist = ., 
+              ncol = 2, 
+              align = 'hv',
+              axis = 'tblr')
+  
+  figures$sympt_recovery$bottom_panel <- kin_mod$plots_hact %>% 
+    map(~.x[c('anosmia', 'taste_loss')]) %>% 
+    transpose %>% 
+    unlist(recursive = FALSE) %>% 
+    map(~.x + 
+          theme(plot.margin = ggplot2::margin(t = 5, r = 1, 
+                                              b = 0, l = 0, unit = 'mm'))) %>% 
+    plot_grid(plotlist = ., 
+              ncol = 4, 
+              align = 'hv', 
+              axis = 'tblr')
+  
+  figures$sympt_recovery <- plot_grid(figures$sympt_recovery$upper_panel, 
+                                      figures$sympt_recovery$bottom_panel, 
+                                      nrow = 2, 
+                                      rel_heights = c(0.7, 0.3), 
+                                      labels = c('A', 'B'), 
+                                      label_size = 10) %>% 
+    as_figure('figure_2_recovery_times', 
+              w = 180, 
+              h = 220)
+
+# Figure 3: symptom distances -----
+  
+  insert_msg('Figure 3: symptom isolation')
+  
+  figures$sympt_mds <- sympt_dist$mds_plots_hact[c('north.long', 
+                                                   'south.long', 
+                                                   'north.pasc', 
+                                                   'south.pasc')] %>% 
+    map(~.x + 
+          labs(tag = .x$labels$tag %>% 
[…]

### 2022-06-02 paper scripts/manuscript.R

@@ -1,0 +1,87 @@
+# Renders the manuscript figures/tables and the supplementary material
+
+  insert_head()
+  
+# Exporting the figure chunks -----
+  
+  insert_msg('Figure chunks')
+  
+  insert_figure(figures$consort, 
+                figures$sympt_recovery, 
+                figures$sympt_mds, 
+                figures$apriori, 
+                figures$clustering, 
+                figures$sympt_clusters, 
+                figures$base_clusters, 
+                figures$recovery_clusters, 
+                file = './paper/markdown/figure_chunks.Rmd', 
+                ref_names = stri_replace_all(names(figures), 
+                                             fixed = '_', 
+                                             replacement = '-'), 
+                captions = c('Flow diagram of the analysis inclusion process for the CovILD and HACT studies.', 
+                             'Symptom-specific recovery times in ambulatory COVID-19.', 
+                             'Self-reported smell and taste disorders are isolated symptoms of long COVID and PASC.', 
+                             'High extent of co-occurrence of smell and taste disorders in long COVID and PASC.', 
+                             'Clustering of ambulatory COVID-19 individuals by symptom-specific recovery times.', 
+                             'Duration of neurocognitive and respiratory symptoms, fatigue, smell and taste disorders differs between the COVID-19 recovery clusters.', 
+                             'COVID-19 recovery clusters differ in sex distribution, comorbidity and daily medication rates.', 
+                             'Good clinical, mental and psychosocial recovery in individuals with persistent smell and taste disorders'))
+  
+# Exporting the supplement chunks -----
+  
+  insert_msg('Supplement chunks')
+  
+  insert_figure(suppl$hact_freq, 
+                suppl$hact_recovery, 
+                suppl$covild_freq, 
+                suppl$covild_recovery, 
+                suppl$hyposmia_rater100, 
+                suppl$hyposmia_rater360, 
+                suppl$mds_acute, 
+                suppl$clust_devel, 
+                suppl$cov_clusters,
+                suppl$fatigue_clsters, 
+                suppl$neurocog_clusters, 
+                suppl$anosmia_clusters, 
+                suppl$rec_clusters, 
+                suppl$psych_clusters, 
+                file = './paper/markdown/supplement_chunks.Rmd', 
+                ref_names = stri_replace_all(names(suppl), 
+                                             fixed = '_', 
+                                             replacement = '-'), 
[…]

### 2022-06-02 paper scripts/supplements.R

@@ -1,0 +1,336 @@
+# supplementary figures 
+
+  insert_head()
+  
+# container list -----
+  
+  suppl <- list()
+  
+# Figure S1: Frequency of symptoms, HACT -----
+  
+  insert_msg('Figure S1: HACT symptom frequency')
+  
+  suppl$hact_freq <- eda_sympt$hact_bubble %>% 
+    map(~.x + 
+          scale_y_discrete(labels = function(x) embolden_scale(x, 
+                                                               c('anosmia', 'taste_loss'), 
+                                                               translate = TRUE)) + 
+          theme(axis.text.y = element_markdown())) %>% 
+    plot_grid(plotlist = ., 
+              ncol = 2, 
+              align = 'hv', 
+              axis = 'tblr', 
+              labels = LETTERS, 
+              label_size = 10) %>% 
+    as_figure('figure_s1_hact_symp_frequency', 
+              w = 180, 
+              h = 190)
+  
+# Figure S2: HACT recovery symptoms -------
+  
+  insert_msg('Figure S2: fast-recovery symptoms')
+  
+  suppl$hact_recovery <- kin_mod$plots_hact %>% 
+    map(~.x[c('fever', 'dim_appetite', 
+              'joint_pain', 'muscle_pain', 
+              'fatigue', 'breath_short')]) %>% 
+    transpose %>% 
+    unlist(recursive = FALSE) %>% 
+    map(~.x + 
+          scale_y_continuous(limits = c(0, 105), 
+                             breaks = seq(0, 100, by = 25)) + 
+          theme(plot.margin = ggplot2::margin(t = 5, r = 1, 
+                                              b = 0, l = 0, unit = 'mm'))) %>% 
+    plot_grid(plotlist = ., 
+              ncol = 4, 
+              align = 'hv', 
+              axis = 'tblr', 
+              labels = c('A', '', 
+                         'B', '', 
+                         'C', '', 
+                         'D', '', 
+                         'E', '', 
+                         'F', ''), 
+              label_size = 10) %>% 
+    as_figure('figure_s2_hact_kinetic', 
+              w = 180, 
+              h = 180)
+  
+# Figure S3: Frequency of symptoms, CovILD ------
+  
+  insert_msg('Figure S3: frequency of symptoms, CovILD')
+  
+  suppl$covild_freq <- eda_sympt$covild_bubble %>% 
+    map(~.x + 
+          scale_y_discrete(labels = function(x) embolden_scale(x, 
[…]

### 2022-06-02 paper scripts/tables.R

@@ -1,0 +1,184 @@
+# Paper tables
+
+  insert_head()
+  
+# container lists ------
+  
+  paper_tbl <- list()
+  suppl_tbl <- list()
+  
+# globals ------
+  
+  paper_tbl$hact_base_vars <- c('obs_time', 'sex', 'age', 'bmi_class_before', 
+                                'education_class', 'employment_before', 
+                                'comorb_present', 'hypertension', 'heart_circulation', 
+                                'diabetes', 'lung', 'gastrointenstinal', 
+                                'malignancy', 'hay_fever', 'autoimmunity', 
+                                'frequent_flu_like', 'two_plus_infections_antibiotics', 
+                                'depression_burnout', 'insomnia', 'daily_medication')
+  
+# Table 1: baseline characteristic of the HACT study cohorts -----
+  
+  insert_msg('Table 1: baseline, HACT')
+  
+  paper_tbl$hact_baseline <- eda_cohort$cmm_tbl_hact %>% 
+    filter(var_r %in% paper_tbl$hact_base_vars) %>% 
+    mutate(variable = translate_var(var_r, out_value = 'label', dict = hact$dict), 
+           unit = translate_var(var_r, out_value = 'unit', dict = hact$dict), 
+           variable = ifelse(!is.na(unit), 
+                             paste(variable, unit, sep = ', '), 
+                             variable)) %>% 
+    select(variable, north, south, significance, eff_size) %>% 
+    set_names(c('Variable', 'AT', 'IT', 'Significance', 'Effect size'))
+  
+# Table 2: baseline characteristic of the CovILD cohort ------
+  
+  insert_msg('Table 2: baseline, CovILD')
+  
+  paper_tbl$covild_baseline <- eda_cohort$cmm_tbl_covild %>% 
+    filter(var_r != 'no_comorb') %>% 
+    mutate(variable = translate_var(var_r, out_value = 'label', dict = covild$dict), 
+           unit = translate_var(var_r, out_value = 'unit', dict = covild$dict), 
+           variable = ifelse(is.na(unit) | unit == '%', 
+                             variable, 
+                             paste(variable, unit, sep = ', '))) %>% 
+    mutate(variable = car::recode(variable, 
+                                  "'Weight class' = 'BMI at CoV onset'; 
+                                  'CVD' = 'Cardiovascular disease'; 
[…]

### 2022-06-02 paper.R

@@ -1,0 +1,36 @@
+# Paper tables, figures and manuscript text
+
+  library(plyr)
+  library(tidyverse)
+  library(cowplot)
+  library(figur)
+  library(soucer)
+  library(exda)
+  library(writexl)
+  library(rmarkdown)
+  library(knitr)
+  library(bookdown)
+  library(flextable)
+  library(ggtext)
+  library(glue)
+  library(stringi)
+  
+  insert_head()
+  
+  source_all('./tools/project_tools.R', 
+             message = TRUE, crash = TRUE)
+  
+# Paper scripts ----
+  
+  insert_msg('Paper scripts')
+  
+  c('./paper scripts/tables.R', 
+    './paper scripts/figures.R', 
+    './paper scripts/supplements.R', 
+    './paper scripts/manuscript.R') %>% 
+    source_all(message = TRUE, crash = TRUE)
+  
+# END ----
+  
+  insert_tail()
+  

### 2022-06-02 tools/project_globals.R

@@ -1,0 +1,186 @@
+# This script contains project globals
+
+# data container ------
+
+  globals <- list()
+
+# graphics -----
+
+  globals$common_text <- element_text(size = 8,
+                                      face = 'plain',
+                                      color = 'black')
+
+  globals$common_margin <- ggplot2::margin(t = 5,
+                                           l = 4,
+                                           r = 2,
+                                           unit = 'mm')
+
+  globals$common_theme <- theme_classic() + theme(axis.text = globals$common_text,
+                                                  axis.title = globals$common_text,
+                                                  plot.title = element_text(size = 8,
+                                                                            face = 'bold'),
+                                                  plot.subtitle = globals$common_text,
+                                                  plot.tag = element_text(size = 8,
+                                                                          face = 'plain',
+                                                                          color = 'black',
+                                                                          hjust = 0,
+                                                                          vjust = 1),
+                                                  plot.tag.position = 'bottom',
+                                                  legend.text = globals$common_text,
+                                                  legend.title = globals$common_text,
+                                                  strip.text = globals$common_text,
+                                                  strip.background = element_rect(fill = 'gray95',
+                                                                                  color = 'gray80'),
+                                                  plot.margin = globals$common_margin,
+                                                  panel.grid.major = element_line(color = 'gray90'))
+  
+# strata colors ------
+  
+  globals$hact_colors <- c(north = 'steelblue', 
+                           south = 'coral3')
+  
+  globals$hact_labs <- c(north = 'AT', 
+                         south = 'IT')
+  
+  globals$covild_colors <- c(A = 'steelblue', 
+                             HM = 'darkolivegreen3', 
+                             HS = 'coral3', 
+                             cohort = 'plum4')
+  
+  globals$covild_labels <- c(A = 'ambulatory', 
+                             HM = 'moderate', 
+                             HS = 'severe', 
+                             cohort = 'CovILD cohort')
+  
+# symptom variables ------
+  
+  globals$hact_symptoms <- c('fever', 'ague', 'sore_throat', 'running_nose', 
+                             'fatigue', 'dry_cough', 'wet_cough', 'breath_short', 
+                             'dyspnoe', 'chest_pain', 'tachycardia', 'extrasystole', 
+                             'joint_pain', 'bone_pain', 'muscle_pain', 'abdominal_pain', 
[…]

### 2022-06-02 tools/project_tools.R

@@ -1,287 +1,234 @@
 # A medley of functional project tools
 
-  require(plyr)
-  require(tidyverse)
-  require(furrr)
-  require(caret)
-  require(doParallel)
-  require(randomForest)
-  require(grDevices)
-  require(pcaPP)
-  require(rlang)
-  require(cowplot)
-  require(gmodels)
-  require(vcd)
-  
-# data import and transformation -----
-  
-  recode_var <- function(data, old_var, new_var, ID_var = 'ID', time_var = NULL, trans_fun = 'my_identity', args = NULL) {
-    
-    ## renames a variable, if time and id variable names provided, they're included
-    ## in  the output as well
-    ## the transformation function enables
-    ## args specifies the arguments to the transforming function
-    
-    trans_call <- call2(trans_fun, 
-                        data[[old_var]], 
-                        !!!args)
+# kNN plots -----
 
-    new_data <- data %>% 
-      mutate(!!ensym(new_var) := eval(trans_call))
+  draw_knn_course <- function(data, 
+                              time_var = 'time', 
+                              dist_var = 'mean_dist', 
+                              color_var = 'variable', 
+                              label_var = 'variable', 
+                              highlight_sympt = 'anosmia', 
+                              highlight_color = 'steelblue', 
+                              norm_color = 'gray60', 
+                              point_size = 0.5, 
+                              highlight_size = 1, 
+                              line_size = 0.5, 
+                              txt_size = 2.75, 
+                              plot_title = NULL, 
+                              plot_subtitle = NULL, 
+                              plot_tag = NULL, 
+                              x_lab = 'Days post CoV', 
+                              y_lab = 'mean k-NN distance', 
+                              cust_theme = globals$common_theme) {
+    
+    ## plots the course of mean KNN distance in time
+    
+    last_timepoint <- max(data[[time_var]])
+    
+    symptoms <- unique(data[[color_var]])
+    
+    col_scale <- rep(norm_color, length(symptoms)) %>% 
+      set_names(symptoms)
+    
+    col_scale[highlight_sympt] <- highlight_color
+    
+    data <- data %>% 
+      mutate(sympt_lab = ifelse(.data[[time_var]] == last_timepoint, 
+                                .data[[label_var]], 
+                                NA), 
[…]

### 2022-07-13 paper/markdown/figures_and_tables.Rmd
Figure content change
Figure content changed (less supplements...), changes in the legend and supplement text. Update of the figure-making scripts to the newest version of figur package

@@
 subtitle: "Tables and figures"
-author: "HACT and CovILD study teams"
+author: "Distinct smell and taste disorder phenotype of post-acute COVID-19 sequelae"
 date: "`r format(Sys.time(), '%Y-%m-%d')`"
@@
 
-knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, dev = "png", dpi = 600)
+knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, 
+                      dev = "png", dpi = 600, 
+                      tab.cap.style = 'Table Heading')
+
+set_flextable_defaults(font.family = 'Cambria', font.size = 10)
 
@@
 
-```{r tab-hact, tab.cap = 'Baseline characteristic of the HACT study Austria (AT) and Italy (IT) cohorts.'}
+```{r tab-hact, tab.cap = 'Baseline characteristic of the Austria (AT) and Italy (IT) survey study cohorts.'}
 
@@
 P values corrected form multiple testing with Benjamini-Hochberg method.'), ref_symbols = 'a', part = 'header') %>%
-  footnote(3, 1, value = as_paragraph('BMI: body mass index, overweight > 25 kg/m\u00B2, obesity > 30 kg/m\u00B2,'), ref_symbols = 'b', part = 'body') %>% 
+  footnote(3, 1, value = as_paragraph('BMI: body mass index, overweight > 25 kg/m\u00B2, obesity > 30 kg/m\u00B2.'), ref_symbols = 'b', part = 'body') %>% 
   footnote(15, 1, value = as_paragraph('Frequent respiratory infections, > 2 per year.'), ref_symbols = 'c', part = 'body') %>% 
   footnote(16, 1, value = as_paragraph('Frequent bacterial infections with antibiotic therapy, > 2 per year.'), ref_symbols = 'd', part = 'body') %>% 
-  theme_vanilla %>%
-  font(part = 'all', fontname = 'Cambria') %>% 
-  fontsize(size = 10, part = 'all')
+  theme_vanilla
 
@@
   footnote(3, 1, value = as_paragraph('BMI: body mass index, overweight > 25 kg/m\u00B2, obesity > 30 kg/m\u00B2,'), ref_symbols = 'b', part = 'body') %>% 
-  theme_vanilla %>%
-  font(part = 'all', fontname = 'Cambria') %>% 
-  fontsize(size = 10, part = 'all')
+  theme_vanilla
 
@@
 
-```{r fig-consort, fig.width = mm_inch(figures$consort$w), fig.height = mm_inch(figures$consort$h), fig.cap = 'Flow diagram of the analysis inclusion process for the CovILD and HACT studies.'}
+```{r fig-consort, fig.width = mm_inch(figures$consort$w), fig.height = mm_inch(figures$consort$h), fig.cap = 'Flow diagram of the analysis inclusion process for the longitudinal CovILD cohort and the Health after COVID-19 survey study.'}
 
@@
 
-__Figure \@ref(fig:fig-consort). Flow diagram of the analysis inclusion process for the CovILD and HACT studies..__
+__Figure \@ref(fig:fig-consort). Flow diagram of the analysis inclusion process for the longitudinal CovILD cohort and the Health after COVID-19 survey study.__
 
@@
 
-```{r fig-sympt-recovery, fig.width = mm_inch(figures$sympt_recovery$w), fig.height = mm_inch(figures$sympt_recovery$h), fig.cap = 'Symptom-specific recovery times in ambulatory COVID-19.'}
+```{r fig-sympt-recovery, fig.width = mm_inch(figures$sympt_recovery$w), fig.height = mm_inch(figures$sympt_recovery$h), fig.cap = 'Symptom-specific recovery times in ambulatory COVID-19 survey study.'}
 
@@
 
-__Figure \@ref(fig:fig-sympt-recovery). Symptom-specific recovery times in ambulatory COVID-19.__
+__Figure \@ref(fig:fig-sympt-recovery). Symptom-specific recovery times in ambulatory COVID-19 survey study.__
 
-Symptom-specific recovery times were calculated for each participants of the HACT study cohorts (Austria: AT, Italy: IT).
+_Symptom-specific recovery times were calculated for each participants of the survey study cohorts (Austria: AT, Italy: IT)._
 
-__(A)__ Distribution of the recovery times in the individuals with the indicated symptoms present during acute COVID-19.
+_(A) Distribution of the recovery times in the individuals with the indicated symptoms present during acute COVID-19 (first 14 days after clinical onset).
 Diamonds represent median recovery times, tinted ellipses code for interquartile ranges.
-Numbers of complete observations are indicated in the plot captions.
+Numbers of complete observations are indicated in the plot captions._
 
-__(B)__ Percentages of individuals with smell and taste disorders in the AT (Austria) and IT (Italy) HACT study cohorts at particular time points after clinical onset.
-Numbers of complete observations are indicated under the plots.
+_(B) Percentages of individuals with smell and taste disorders in the AT (Austria) and IT (Italy) survey study cohorts at particular time points after clinical onset.
+Numbers of complete observations are indicated under the plots._
 
@@
 
-```{r fig-sympt-mds, fig.width = mm_inch(figures$sympt_mds$w), fig.height = mm_inch(figures$sympt_mds$h), fig.cap = 'Self-reported smell and taste disorders are isolated symptoms of long COVID and PASC.'}
+```{r fig-sympt-mds, fig.width = mm_inch(figures$sympt_mds$w), fig.height = mm_inch(figures$sympt_mds$h), fig.cap = 'Self-reported smell and taste disorders are isolated persistent symptoms of COVID-19.'}
 
@@
 
-__Figure \@ref(fig:fig-sympt-mds). Self-reported smell and taste disorders are isolated symptoms of long COVID and PASC.__
+__Figure \@ref(fig:fig-sympt-mds). Self-reported smell and taste disorders are isolated persistent symptoms of COVID-19.__
 
-Binary symptom occurrence data during for long COVID ($\geq$ 28 days after clinical onset, __A__) and PASC ($\geq$ 90 days, __B__) in the HACT Austria (AT) and Italy (IT) cohorts were subjected to two-dimensional multi-dimensional scaling (MDS) with simple matching distance (SMD) between the symptoms.
+_Symptom data during at $\geq$ 28 days (A) and $\geq$ 3 moths (B) after clinical onset in the Austria (AT) and Italy (IT) survey study cohorts were subjected to two-dimensional multi-dimensional scaling (MDS) with simple matching distance (SMD) between the symptoms.
 MDS coordinates are presented in point plots.
@@
 Percentages of the data set variance associated with the MDS dimensions are indicated in the plot axes.
-Numbers of complete observations are indicated under the plots.
+Numbers of complete observations are indicated under the plots._
 
@@
 
-```{r fig-apriori, fig.width = mm_inch(figures$apriori$w), fig.height = mm_inch(figures$apriori$h), fig.cap = 'Frequent co-occurrence of smell and taste disorders in long COVID and PASC.'}
+```{r fig-apriori, fig.width = mm_inch(figures$apriori$w), fig.height = mm_inch(figures$apriori$h), fig.cap = 'Co-occurrence of smell and taste disorders in post-acute COVID-19 sequelae.'}
 
@@
 
-__Figure \@ref(fig:fig-apriori). Frequent co-occurrence of smell and taste disorders in long COVID and PASC.__
+__Figure \@ref(fig:fig-apriori). Co-occurrence of smell and taste disorders in post-acute COVID-19 sequelae.__
 
-Frequent combinations of symptoms of long COVID (__A__) and PASC (__B__) in the HACT study Austria (AT) and Italy (IT) cohorts were identified with the apriori algorithm.
+_Frequent combinations of symptoms at $\geq$ 28 days (A) and $\geq$ 3 months (B) after clinical onset in the Austria (AT) and Italy (IT) survey study cohorts were identified with the apriori algorithm.
 Symptom combination support and confidence are presented in the plots.
-Point color codes for lift values.
-The hypo/anosmia and hypo/ageusia combinations are labeled.
+Point color codes for lift statistic values.
+The hypo/anosmia and hypo/ageusia combinations are labeled._
 
@@
 
-```{r fig-clustering, fig.width = mm_inch(figures$clustering$w), fig.height = mm_inch(figures$clustering$h), fig.cap = 'Clustering of ambulatory COVID-19 individuals by symptom-specific recovery times.'}
-
-figures$clustering$plot
-
-```
-
-__Figure \@ref(fig:fig-clustering). Clustering of ambulatory COVID-19 individuals by symptom-specific recovery times.__
-
-Individuals of the training Austria (AT) cohort of the HACT study were subjected to clustering in respect to symptom-specific recovery times with the PAM (partitioning around medoids) algorithm and Euclidean distance measure (__Supplementary Figure S8__).
-Cluster assignment in the test Italy (IT) HACT cohort was done with k-NN label propagation algorithm.
-Recovery times for particular COVID-19 symptoms in the COVID-19 recovery clusters are presented as heat maps.
-Numbers of individuals assigned to the recovery clusters are indicated under the plots.
-
-Dim. appetite: diminished appetite, Imp. concentration: impaired concentration, Imp. walk: impaired walk, Imp. FMS: impaired fine motor skills.
-
-\newpage
-
-```{r fig-sympt-clusters, fig.width = mm_inch(figures$sympt_clusters$w), fig.height = mm_inch(figures$sympt_clusters$h), fig.cap = 'Duration of neurocognitive and respiratory symptoms, fatigue, smell and taste disorders differs between the COVID-19 recovery clusters.'}
+```{r fig-sympt-clusters, fig.width = mm_inch(figures$sympt_clusters$w), fig.height = mm_inch(figures$sympt_clusters$h), fig.cap = 'Differing duration of neurocognitive and respiratory symptoms, fatigue, smell and taste disorders defines the COVID-19 recovery clusters.'}
 
@@
 
-__Figure \@ref(fig:fig-sympt-clusters). Duration of neurocognitive and respiratory symptoms, fatigue, smell and taste disorders differs between the COVID-19 recovery clusters.__
+__Figure \@ref(fig:fig-sympt-clusters). Differing duration of neurocognitive and respiratory symptoms, fatigue, smell and taste disorders defines the COVID-19 recovery clusters.__
 
-Semi-supervised clustering was performed as presented in __Figure \@ref(fig:fig-clustering)__ and __Supplementary Figure S8__.
+_The smell and taste disorder recovery (STDR), rapid recovery (RR) and slow recovery (SR) clusters of the survey study participants were identified by semi supervised PAM clustering (partitioning around medoids, Euclidean distance, training cohort: Austria [AT], test cohort: Italy [IT]).
 Mean recovery times in the recovery clusters are presented as lines, 2 $\times$ SEM intervals are displayed as tinted regions.
-Numbers of individuals assigned to the recovery clusters are indicated under the plots.
+Numbers of individuals assigned to the recovery clusters are indicated under the plots._
 
-Dim. appetite: diminished appetite, Imp. concentration: impaired concentration, Imp. walk: impaired walk, Imp. FMS: impaired fine motor skills.
-
-\newpage
-
-```{r fig-base-clusters, fig.width = mm_inch(figures$base_clusters$w), fig.height = mm_inch(figures$base_clusters$h), fig.cap = 'COVID-19 recovery clusters differ in sex distribution, comorbidity and daily medication rates.'}
-
-figures$base_clusters$plot
-
-```
-
-__Figure \@ref(fig:fig-base-clusters). COVID-19 recovery clusters differ in sex distribution, comorbidity and daily medication rates.__
-
-Differences in sex distribution (__A__), frequency of comorbidity (__B__) and daily medication (__C__) between the recovery clusters in the training Austria (AT) and the test Italy (IT) HACT study cohorts (__Figure \@ref(fig:fig-clustering)__, __Supplementary Figure S8__) were assessed by $\chi^2$ test with Cramer V effect size statistic.
-P values were corrected for multiple testing with Benjamini-Hochberg method.
-The frequencies are presented as bar plots.
-Effect size statistics and p values are indicated in the plot caption.
-Numbers of complete observations are displayed under the plots.
+_Dim. appetite: diminished appetite, Imp. concentration: impaired concentration, Imp. walk: impaired walk, Imp. FMS: impaired fine motor skills._
 
@@
 
-Differences in minimum/maximum scaled readouts of clinical and physical recovery, mental health and quality of life scoring at the time of survey completion in the recovery clusters (__Figure \@ref(fig:fig-clustering)__, __Supplementary Figure S8__) are presented in radial plots.
-Yes/no items (incomplete convalescence, weight loss, new medication and need for rehabilitation) were binarized (yes: 1, no: 0) prior to visualization.
+_Differences in minimum/maximum scaled readouts of clinical and physical recovery, mental health and quality of life at the time of survey completion in the recovery clusters (STDR: smell and taste disorder recovery, RR: rapid recovery, SR: slow recovery) in the Austria (AT) and Italy (IT) survey study cohorts.
+Dichotomous items (incomplete convalescence, weight loss, new medication and need for rehabilitation) were binarized (yes: 1, no: 0) prior to visualization.
+Statistical significance for differences between the clusters was assessed by Kruskal-Wallis (numeric readouts) or $\chi^2$ test (dichotomous readouts).
+P values were corrected for multiple testing with Benjamini-Hochberg method.
 Lines represent mean values, 2 $\times$ SEM intervals are displayed as tinted regions.
-Numbers of individuals assigned to the recovery clusters are indicated under the plots.
+Numbers of individuals assigned to the recovery clusters are indicated under the plots._
 
-incomplete recovery: self-reported incomplete recovery, # persist.
-symptoms: number of long COVID symptoms ($\geq$ 28 days after clinical onset), phys.
-performance loss: physical performance loss as compared with the time before COVID-19, QoL score: quality of life impairment score, OMH score: overall mental health impairment score, ANX score: anxiety score, DPR: depression score.
+_incomplete recovery: self-reported incomplete recovery, # persist.
+symptoms: number of symptoms at $\geq$ 28 days after clinical onset, phys.
+performance loss: physical performance loss as compared with the time before COVID-19, QoL score: quality of life impairment score, OMH score: overall mental health impairment score, ANX score: anxiety score, DPR: depression score._

### 2022-07-13 paper/markdown/supplementary_material.Rmd
Figure content change
Figure content changed (less supplements...), changes in the legend and supplement text. Update of the figure-making scripts to the newest version of figur package

@@
 subtitle: "Supplementary Material"
-author: "HACT and CovILD study teams"
+author: "Hralth after COVID-19 in Tyrol and CovILD study teams"
 date: "`r format(Sys.time(), '%Y-%m-%d')`"
@@
 
-The complete list of study variables and stratification scheme is provided in __Supplementary Table S\@ref(tab:tab-hact-vars)__ for the HACT study and in __Supplementary Table S\@ref(tab:tab-covild-vars)__ for the CovILD study.
+The complete list of study variables and stratification scheme is provided in __Supplementary Table S\@ref(tab:tab-hact-vars)__ for the Health after COVID-19 in Tyrol survey study and in __Supplementary Table S\@ref(tab:tab-covild-vars)__ for the CovILD cohort.
 
@@
 
-A total of `r length(globals$hact_symptoms)` was surveyed in the HACT study cohorts (__Supplementary Figure S\@ref(fig:fig-hact-freq)__ and __Supplementary Table S\@ref(tab:tab-hact-vars)__).
-The symptom duration was coded as follows: absent: 0 days, 1 - 3 days: 3 days, up to 1 week: 7 days, up to 2 weeks: 14 days, up to 4 weeks: 28 days, up to 3 months: 90 days, up to 6 months: 90 days, over 6 months: 90 days.
+A total of `r length(globals$hact_symptoms)` was surveyed in the survey study cohorts (__Supplementary Figure S\@ref(fig:fig-hact-freq)__ and __Supplementary Table S\@ref(tab:tab-hact-vars)__).
+The symptom duration was coded as follows: absent: 0 days, 1 - 3 days: 3 days, up to 1 week: 7 days, up to 2 weeks: 14 days, up to 4 weeks: 28 days, up to 3 months: 30 moths, up to 6 months: 3 months, over 6 months: 3 months.
 Acute symptoms were defined as complaints present during the first 14 days after clinical onset of COVID-19.
-Long COVID symptoms were defined as complaints present for $\geq$ 28 days.
-PASC symptoms were defined as manifestations present for $\geq$ 90 days [@Venkatesan2021; @Shah2021].
 
@@
 
-### Rating of physical recovery, mental health and quality of life in the HACT study
+### Rating of physical recovery, mental health and quality of life in the survey study
 
@@
 Likelihood ratio $\lambda$ statistic (full versus null model) was used as an effect size measure.
-Individuals from the HACT or CovILD studies with the complete longitudinal symptom record were included in the kinetic modeling tasks.
+Individuals from the survey study or the CovILD cohort with the complete longitudinal symptom record were included in the kinetic modeling tasks.
 Results of the kinetic modeling were adjusted for multiple comparisons with Benjamini-Hochberg method [@Benjamini1995].
@@
 
-To assess co-occurrence or exclusivity of symptoms, simple matching distances between manifestations of each acute COVID-19, long COVID and PASC in the HACT study cohorts were calculated (package _scrime_ and development package _clustTools_ [https://github.com/PiotrTymoszuk/clustTools]) [@Boriah2008; @Schwender2018].
+To assess co-occurrence or exclusivity of symptoms, simple matching distances between manifestations during the first 14 days, at $\geq$ 28 days and at $\geq$ 3 months after clinical onset in the survey study cohorts were calculated (package _scrime_ and development package _clustTools_ [https://github.com/PiotrTymoszuk/clustTools]) [@Boriah2008; @Schwender2018].
 Subsequently, the distance matrix was subjected to multi-dimensional scaling (MDS, k = 2 dimensions, package _stats_, function _cmdscale()_).
@@
 
-### Apriori analysis of acute COVID-19, long COVID and PASC symptoms in the HACT study
+### Apriori analysis of COVID-19 symptoms in the survey study
 
-Frequent combinations of symptoms of long COVID and PASC in the HACT study cohorts were identified with the apriori algorithm (package _arules_) [@Agrawal1993; @Hahsler2005] with the minimal support cutoff of 0.1, 2 - 10 item transaction length, confidence > 0.8 and lift > 2.
+Frequent combinations of symptoms during the first 14 days, at $\geq$ 28 days and at $\geq$ 3 months after clinical onset in the survey study cohorts were identified with the apriori algorithm (package _arules_) [@Agrawal1993; @Hahsler2005] with the minimal support cutoff of 0.1, 2 - 10 item transaction length, confidence > 0.8 and lift > 2.
 The support statistic were used to estimate the symptom combination frequency.
@@
 The lift statistic was interpreted as a measure of the symptom dependence (lift = 1, symptoms are independent).
-The complete analysis results are presented in __Supplementary Tables S\@ref(tab:tab-apriori-long)__ and __S\@ref(tab:tab-apriori-pasc)__.
 
@@
 
-COVID-19 recovery clusters of the training Autria (AT) HACT cohort participants in respect to symptom-specific recovery times (__Figure 1A__) were defined with the PAM (partitioning around medoids) algorithm and Euclidean distance statistic (packages _cluster_, _philentropy_ and development package _clustTools_ [https://github.com/PiotrTymoszuk/clustTools]) [@Drost2018; @Schubert2019].
+COVID-19 recovery clusters of the training Autria (AT) survey cohort participants in respect to symptom-specific recovery times (__Figure 1A__) were defined with the PAM (partitioning around medoids) algorithm and Euclidean distance statistic (packages _cluster_, _philentropy_ and development package _clustTools_ [https://github.com/PiotrTymoszuk/clustTools]) [@Drost2018; @Schubert2019].
 The set of participants with the complete clustering variable set was included in the analysis.
@@
 Permutation importance of specific clustering variables was investigated by calculating difference in clustering variance (ratio of total between-cluster sum of squares to total sum of squares) between the initial clustering object and the clustering object with the given variable reshuffled at random (package _clustTools_).
-Assignment of the Italy HACT cohort participants to the recovery clusters was accomplished with k-nearest neighbors label propagation algorithm (k = 5) [@Leng2014].
+Assignment of the Italy survey cohort participants to the recovery clusters was accomplished with k-nearest neighbors label propagation algorithm (k = 5) [@Leng2014].
 The clustering efficacy in the training AT cohort and the test IT cohort measured by clustering variance statistic defined above was similar (AT: `r clustTools::var(part_clust$clust_north)$frac_var %>% signif(2)`, IT: `r clustTools::var(part_clust$clust_south)$frac_var %>% signif(2)`).
@@
 
-```{r tab-hact-vars, tab.cap = 'HACT study variables.'}
+```{r tab-hact-vars, tab.cap = 'Survey study variables.'}
 
@@
   width(4, width = 8.7, unit = 'cm') %>% 
-  theme_vanilla %>% 
-  font(part = 'all', fontname = 'Cambria') %>% 
-  fontsize(size = 10, part = 'all')
+  theme_vanilla
 
@@
   width(3, width = 2, unit = 'cm')  %>% 
-  theme_vanilla %>% 
-  font(part = 'all', fontname = 'Cambria') %>% 
-  fontsize(size = 10, part = 'all')
+  theme_vanilla
 
@@
 
-```{r tab-apriori-long, tab.cap = 'Results of apriori analysis for long COVID-19 symptoms in the HACT study .
-Top 10 most frequent symptom combinations (highest support value) in the AT (Austria) and IT (Italy) cohort are presented.
-The full results are available as a supplementary Excel table.'}
-
-suppl_tbl$apriori_long %>% 
-  group_by(Cohort) %>% 
-  top_n(10, Support) %>% 
-  ungroup %>% 
-  arrange(Cohort, Support) %>% 
-  select(-Time) %>% 
-  flextable %>% 
-  width(1, width = 1.5, unit = 'cm') %>% 
-  width(2, width = 9.5, unit = 'cm') %>% 
-  width(3:5, width = 2.2, unit = 'cm')  %>% 
-  theme_vanilla %>% 
-  font(part = 'all', fontname = 'Cambria') %>% 
-  fontsize(size = 10, part = 'all')
-
-```
-
-\newpage
-
-```{r tab-apriori-pasc, tab.cap = 'Results of apriori analysis for PASC symptoms in the HACT study.
-Top 10 most frequent symptom combinations (highest support value) in the AT (Austria) and IT (Italy) cohort are presented.
-The full results are available as a supplementary Excel table.'}
-
-suppl_tbl$apriori_pasc %>% 
-  group_by(Cohort) %>% 
-  top_n(10, Support) %>% 
-  ungroup %>% 
-  arrange(Cohort, Support) %>% 
-  select(-Time) %>% 
-  flextable %>% 
-  width(1, width = 1.5, unit = 'cm') %>% 
-  width(2, width = 9.5, unit = 'cm') %>% 
-  width(3:5, width = 2.2, unit = 'cm')  %>% 
-  theme_vanilla %>% 
-  font(part = 'all', fontname = 'Cambria') %>% 
-  fontsize(size = 10, part = 'all')
-
-```
-
-\newpage
-
-```{r tab-clust-base-at, tab.cap = 'Demographic and baseline clinical characteristic at the COVID-19 onset of the HACT study participants assigned to the recovery clusters, Austria (AT) cohort.'}
+```{r tab-clust-base-at, tab.cap = 'Demographic and baseline clinical characteristic at the COVID-19 onset of the survey study participants assigned to the recovery clusters, Austria (AT) cohort.'}
 
@@
   footnote(16, 1, value = as_paragraph('Frequent bacterial infections with antibiotic therapy, > 2 per year.'), ref_symbols = 'd', part = 'body') %>% 
-  theme_vanilla %>%
-  font(part = 'all', fontname = 'Cambria') %>% 
-  fontsize(size = 10, part = 'all')
+  theme_vanilla
 
@@
 
-```{r tab-clust-base-it, tab.cap = 'Demographic and baseline clinical characteristic at the COVID-19 onset of the HACT study participants assigned to the recovery clusters, Italy (IT) cohort.'}
+```{r tab-clust-base-it, tab.cap = 'Demographic and baseline clinical characteristic at the COVID-19 onset of the survey study participants assigned to the recovery clusters, Italy (IT) cohort.'}
 
@@
 
-```{r tab-clust-reco-at, tab.cap = 'COVID-19 course and recovery in the HACT study participants assigned to the recovery clusters, Austria (AT) cohort.'}
+```{r tab-clust-reco-at, tab.cap = 'COVID-19 course and recovery in the survey study participants assigned to the recovery clusters, Austria (AT) cohort.'}
 
@@
 P values corrected form multiple testing with Benjamini-Hochberg method.'), ref_symbols = 'a', part = 'header') %>%
-  footnote(3, 1, value = as_paragraph('Number of acute COVID-19 symptoms.'), ref_symbols = 'b', part = 'body') %>% 
-  footnote(4, 1, value = as_paragraph('Number of long COVID symptoms.'), ref_symbols = 'c', part = 'body') %>% 
-  footnote(7, 1, value = as_paragraph('Self-reported incomplete recovery.'), ref_symbols = 'd', part = 'body') %>% 
-  footnote(11, 1, value = as_paragraph('ANX: anxiety.'), ref_symbols = 'e', part = 'body') %>% 
-  footnote(12, 1, value = as_paragraph('DPR: depression.'), ref_symbols = 'f', part = 'body') %>% 
-  footnote(14, 1, value = as_paragraph('OMH score: overall mental health impairment score'), ref_symbols = 'g', part = 'body') %>% 
-  footnote(15, 1, value = as_paragraph('QoL score: quality of life impairment score'), ref_symbols = 'h', part = 'body') %>%
-  theme_vanilla %>%
-  font(part = 'all', fontname = 'Cambria') %>% 
-  fontsize(size = 10, part = 'all')
+  footnote(4, 1, value = as_paragraph('Self-reported incomplete recovery.'), ref_symbols = 'd', part = 'body') %>% 
+  footnote(8, 1, value = as_paragraph('ANX: anxiety.'), ref_symbols = 'e', part = 'body') %>% 
+  footnote(9, 1, value = as_paragraph('DPR: depression.'), ref_symbols = 'f', part = 'body') %>% 
+  footnote(11, 1, value = as_paragraph('OMH score: overall mental health impairment score'), ref_symbols = 'g', part = 'body') %>% 
+  footnote(12, 1, value = as_paragraph('QoL score: quality of life impairment score'), ref_symbols = 'h', part = 'body') %>%
+  theme_vanilla
 
@@
 
-```{r tab-clust-reco-it, tab.cap = 'COVID-19 course and recovery in the HACT study participants assigned to the recovery clusters, Italy (AT) cohort.'}
+```{r tab-clust-reco-it, tab.cap = 'COVID-19 course and recovery in the survey study participants assigned to the recovery clusters, Italy (AT) cohort.'}
 
@@
 
-```{r fig-hact-freq, fig.width = mm_inch(suppl$hact_freq$w), fig.height = mm_inch(suppl$hact_freq$h), fig.cap = 'Frequency of COVID-19 symptoms in acute COVID-19, long COVID and PASC in the HACT study.'}
+```{r fig-hact-freq, fig.width = mm_inch(suppl$hact_freq$w), fig.height = mm_inch(suppl$hact_freq$h), fig.cap = 'Frequency of COVID-19 symptoms in the survey study.'}
 
@@
 
-__Supplementary Figure S\@ref(fig:fig-hact-freq). Frequency of COVID-19 symptoms in acute COVID-19, long COVID and PASC in the HACT study.__
+__Supplementary Figure S\@ref(fig:fig-hact-freq). Frequency of COVID-19 symptoms in the survey study.__
 
-Frequency of symptoms of acute COVID-19 (first 14 days post clinical onset), long COVID ($\geq$ 28 days) and PASC (post-acute sequelae of COVID-19, $\geq$ 90 days) in the AT (Austria, __A__) and IT (Italy, __B__) HACT study cohorts expressed as percentages of the cohort.
+_Frequency of symptoms in first 14 days, $\geq$ 28 days and $\geq$ 3 months after clinical onset of COVID-19 in the Austria (AT, A) and Italy (IT, B) survey study cohorts expressed as percentages of the cohort.
 Point size and color represents the percentage.
-Numbers of complete observations are indicated under the plot.
+Numbers of complete observations are indicated under the plot._
 
-Dim. appetite: diminished appetite, Imp. concentration: impaired concentration, Imp. walk: impaired walk, Imp. FMS: impaired fine motor skills.
+_Dim. appetite: diminished appetite, Imp. concentration: impaired concentration, Imp. walk: impaired walk, Imp. FMS: impaired fine motor skills._
 
@@
 
-```{r fig-hact-recovery, fig.width = mm_inch(suppl$hact_recovery$w), fig.height = mm_inch(suppl$hact_recovery$h), fig.cap = 'Kinetic of recovery from leading acute COVID-19 symptoms in the HACT study.'}
+```{r fig-hact-recovery, fig.width = mm_inch(suppl$hact_recovery$w), fig.height = mm_inch(suppl$hact_recovery$h), fig.cap = 'Kinetic of recovery from leading acute COVID-19 symptoms in the survey study.'}
 
@@
 
-__Supplementary Figure S\@ref(fig:fig-hact-recovery). Kinetic of recovery from leading acute COVID-19 symptoms in the HACT study.__
+__Supplementary Figure S\@ref(fig:fig-hact-recovery). Kinetic of recovery from leading acute COVID-19 symptoms in the survey study.__
 
-Percentages of individuals with fever (__A__), diminished appetite (__B__), joint pain (__C__), muscle pain (__D__), fatigue (__E__) and tachypnea (__F__) in the AT (Austria) and IT (Italy) HACT study cohorts at particular time points after clinical onset.
-Numbers of complete observations are indicated under the plots.
+_Percentages of individuals with fever (A), diminished appetite (B), joint pain (C), muscle pain (D), fatigue (E) and tachypnea (F) in the AT (Austria) and IT (Italy) survey study cohorts at particular time points after clinical onset.
+Numbers of complete observations are indicated under the plots._
 
@@
 
-Frequency of symptoms during acute COVID-19 and at the 60-, 100-, 180- and 360-day follow-ups in ambulatory (__A__), moderate (__B__) and severe COVID-19 (__C__) participants of the CovILD study expressed as percentages of individuals with the complete longitudinal data set.
+_Frequency of symptoms during acute COVID-19 and at the 60-, 100-, 180- and 360-day follow-ups in ambulatory (A), moderate (B) and severe COVID-19 (C) participants expressed as percentages of individuals with the complete longitudinal data set.
 Point size and color represents the percentage.
-Numbers of complete observations are indicated under the plots.
+Numbers of complete observations are indicated under the plots._
 
@@
 
-Percentages of individuals with the complete longitudinal data set suffering from smell disorders (__A__), reduced physical performance (__B__) and dyspnea (__C__) in the ambulatory, moderate and severe COVID-19 subsets of the CovILD study during acute COVID-19 and at the 60-, 100-, 180- and 360-day follow-ups.
-Numbers of complete observations are indicated under the plots.
+_Percentages of individuals with the complete longitudinal data set suffering from smell disorders (A), reduced physical performance (B) and dyspnea (C) in the ambulatory, moderate and severe COVID-19 subsets during acute COVID-19 and at the 60-, 100-, 180- and 360-day follow-ups.
+Numbers of complete observations are indicated under the plots._
 
@@
 
-Association of self-reported and sniffing stick test hyposmia rates was investigated with Cohen's $\kappa$ statistic.
+_Association of self-reported and sniffing stick test hyposmia rates was investigated with Cohen's $\kappa$ statistic.
 Statistical significance was assessed with Wald's Z test corrected for multiple testing with Benjamini-Hochberg method.
-Percentages of individuals with self-reported and test hyposmia within the entire cohort (__A__), the ambulatory (__B__), moderate (__C__) and severe (__D__) COVID-19 subsets of the CovILD-19 subsets are presented in bubble plots.
+Percentages of individuals with self-reported and test hyposmia within the entire cohort (A), the ambulatory (B), moderate (C) and severe (D) COVID-19 subsets are presented in bubble plots.
 Point size and color represents the percentage.
-$\kappa$ values with 95$\%$ confidence intervals, p values and numbers of complete observations are indicated in the plot captions.
+$\kappa$ values with 95$\%$ confidence intervals, p values and numbers of complete observations are indicated in the plot captions._
 
@@
 
-```{r fig-mds-acute, fig.width = mm_inch(suppl$mds_acute$w), fig.height = mm_inch(suppl$mds_acute$h), fig.cap = 'Multi-dimensional scaling analysis of acute COVID-19 symptoms in the HACT study.'}
+```{r fig-mds-acute, fig.width = mm_inch(suppl$mds_acute$w), fig.height = mm_inch(suppl$mds_acute$h), fig.cap = 'Multi-dimensional scaling analysis of acute COVID-19 symptoms in the survey study.'}
 
@@
 
-__Supplementary Figure S\@ref(fig:fig-mds-acute). Multi-dimensional scaling analysis of acute COVID-19 symptoms in the HACT study.__
-
-Binary symptom occurrence data for acute COVID-19 (first 14 days after clinical onset) in the HACT study Austria (AT) and Italy (IT) cohorts were subjected to two-dimensional multi-dimensional scaling (MDS) with simple matching distance (SMD) between the symptoms.
+__Supplementary Figure S\@ref(fig:fig-mds-acute). Multi-dimensional scaling analysis of acute COVID-19 symptoms in the survey study.__
+_Symptom data for acute COVID-19 (first 14 days after clinical onset) in the Austria (AT) and Italy (IT) survey study cohorts were subjected to two-dimensional multi-dimensional scaling (MDS) with simple matching distance (SMD) between the symptoms.
 MDS coordinates are presented in point plots.
@@
 
-```{r fig-clust-devel, fig.width = mm_inch(suppl$clust_devel$w), fig.height = mm_inch(suppl$clust_devel$h), fig.cap = 'Definition of the COVID-19 recovery clusters and clustering feature importance in the HACT study.'}
+```{r fig-clust-devel, fig.width = mm_inch(suppl$clust_devel$w), fig.height = mm_inch(suppl$clust_devel$h), fig.cap = 'Definition of the COVID-19 recovery clusters and clustering feature importance in the survey study.'}
 
@@
 
-__Supplementary Figure S\@ref(fig:fig-clust-devel). Definition of the COVID-19 recovery clusters and clustering feature importance in the HACT study.__
+__Supplementary Figure S\@ref(fig:fig-clust-devel). Definition of the COVID-19 recovery clusters and clustering feature importance in the survey study.__
 
-Individuals of the training Austria (AT) cohort of the HACT study were subjected to clustering in respect to symptom-specific recovery times with the PAM (partitioning around medoids) algorithm and Euclidean distance measure.
+_Individuals of the training Austria (AT) study survey cohort were subjected to clustering in respect to symptom-specific recovery times with the PAM (partitioning around medoids) algorithm and Euclidean distance measure._
 
-__(A)__ Comparison of performance of various algorithms (HCl: hierarchical clustering, SOM + HCl: combined self-organizing map and hierarchical clustering, k-means) and distance statistic in clustering of the training data set investigated by clustering variance (ratio of total between-cluster sum of squares to total sum of squares) and cluster assignment accuracy in 10-fold cross-validation (CV).
+_(A) Comparison of performance of various algorithms (HCl: hierarchical clustering, SOM + HCl: combined self-organizing map and hierarchical clustering, k-means) and distance statistic in clustering of the training data set investigated by clustering variance (ratio of total between-cluster sum of squares to total sum of squares) and cluster assignment accuracy in 10-fold cross-validation (CV)._
 
-__(B)__ Determination of the optimal cluster number in the PAM clustering of the training cohort by the bend of the total within-cluster sum of squares curve.
+_(B) Determination of the optimal cluster number in the PAM clustering of the training cohort by the bend of the total within-cluster sum of squares curve._
 
-__(C)__ Permutation importance of the clustering features (symptoms) for clustering of the training cohort expressed as the difference in clustering variance (ratio of total between-cluster sum of squares to total sum of squares) between the initial clustering object and the clustering object with the given variable reshuffled at random.
+_(C) Permutation importance of the clustering features (symptoms) for clustering of the training cohort expressed as the difference in clustering variance (ratio of total between-cluster sum of squares to total sum of squares) between the initial clustering object and the clustering object with the given variable reshuffled at random._
 
@@
 
-```{r fig-cov-clusters, fig.width = mm_inch(suppl$cov_clusters$w), fig.height = mm_inch(suppl$cov_clusters$h), fig.cap = 'Numbers of acute COVID-19 and long COVID symptoms in the HACT study recovery clusters.'}
+```{r fig-clustering, fig.width = mm_inch(suppl$clustering$w), fig.height = mm_inch(suppl$clustering$h), fig.cap = 'Clustering of ambulatory COVID-19 individuals by symptom-specific recovery times.'}
+
+suppl$clustering$plot
+
+```
+
+__Supplementary Figure S\@ref(fig:fig-clustering). Clustering of ambulatory COVID-19 individuals by symptom-specific recovery times.__
+
+_Individuals of the training Austria (AT) survey study cohort were subjected to clustering in respect to symptom-specific recovery times with the PAM (partitioning around medoids) algorithm and Euclidean distance measure (Supplementary Figure S\@ref(fig:fig-clust-devel)).
+Cluster assignment in the test Italy (IT) survey cohort was done with k-NN label propagation algorithm.
+Recovery times for particular COVID-19 symptoms in the COVID-19 recovery clusters are presented as heat maps.
+Numbers of individuals assigned to the recovery clusters are indicated under the plots._
+
+_Dim. appetite: diminished appetite, Imp. concentration: impaired concentration, Imp. walk: impaired walk, Imp. FMS: impaired fine motor skills._
+
+\newpage
+
+```{r fig-cov-clusters, fig.width = mm_inch(suppl$cov_clusters$w), fig.height = mm_inch(suppl$cov_clusters$h), fig.cap = 'Numbers of COVID-19 symptoms in the survey study recovery clusters.'}
 
@@
 
-__Supplementary Figure S\@ref(fig:fig-cov-clusters). Numbers of acute COVID-19 and long COVID symptoms in the HACT study recovery clusters.__
+__Supplementary Figure S\@ref(fig:fig-cov-clusters). Numbers of COVID-19 symptoms in the survey study recovery clusters.__
 
-The recovery clusters were defined in the training Austria cohort (AT, __Supplementary Figure S\@ref(fig:fig-clust-devel)__) and the cluster assignment predicted in the test Italy cohort (IT, __Figure 5__).
-Differences in numbers of acute COVID-19 (first 14 days after clinical onset, __A__) and long COVID symptoms ($\geq$ 28 days, __B__) between the clusters were assessed by Kruskal-Wallis test and $\eta^2$ effect size statistic.
+_The smell and taste disorder recovery (STDR), rapid recovery (RR) and slow recovery (SR) clusters of the survey study participants were identified by semi supervised PAM clustering (partitioning around medoids, Euclidean distance, training cohort: Austria [AT], test cohort: Italy [IT]).
+Differences in numbers of symptoms in the first 14 days (A) and at $\geq$ 28 days (B) after clinical onset between the clusters were assessed by Kruskal-Wallis test and $\eta^2$ effect size statistic.
 P values were corrected for multiple testing with Benjamini-Hochberg method.
@@
 Effect size statistics and p values are indicated in the plot caption.
-Numbers of complete observations are displayed under the plots.
+Numbers of complete observations are displayed under the plots._
 
@@
 
-```{r fig-anosmia-clusters, fig.width = mm_inch(suppl$anosmia_clusters$w), fig.height = mm_inch(suppl$anosmia_clusters$h), fig.cap = 'Recovery from smell and taste disorders in the HACT study recovery clusters'}
+```{r fig-base-clusters, fig.width = mm_inch(suppl$base_clusters$w), fig.height = mm_inch(suppl$base_clusters$h), fig.cap = 'COVID-19 recovery clusters differ in sex distribution, comorbidity and daily medication rates.'}
 
-suppl$anosmia_clusters$plot
+suppl$base_clusters$plot
 
@@
 
-__Supplementary Figure S\@ref(fig:fig-anosmia-clusters). Recovery from smell and taste disorders in the HACT study recovery clusters__
+__Figure \@ref(fig:fig-base-clusters). COVID-19 recovery clusters differ in sex distribution, comorbidity and daily medication rates.__
 
-The recovery clusters were defined in the training Austria cohort (AT, __Supplementary Figure S\@ref(fig:fig-clust-devel)__) and the cluster assignment predicted in the test Italy cohort (IT, __Figure 5__).
-Differences in recovery times for smell (__A__) and taste disorders (__D__) between the clusters were assessed by Kruskal-Wallis test and $\eta^2$ effect size statistic.
+_The smell and taste disorder recovery (STDR), rapid recovery (RR) and slow recovery (SR) clusters of the survey study participants were identified by semi supervised PAM clustering (partitioning around medoids, Euclidean distance, training cohort: Austria [AT], test cohort: Italy [IT]).
+Differences in sex distribution (A), frequency of comorbidity (B) and daily medication (C) between the recovery clusters were assessed by $\chi^2$ test with Cramer V effect size statistic.
 P values were corrected for multiple testing with Benjamini-Hochberg method.
-Symptom counts are presented in violin plots.
-Points represent single observations, orange diamonds with whiskers code for medians and interquartile ranges.
+The frequencies are presented as bar plots.
 Effect size statistics and p values are indicated in the plot caption.
-Numbers of complete observations are displayed under the plots.
-
-\newpage
-
-```{r fig-fatigue-clsters, fig.width = mm_inch(suppl$fatigue_clsters$w), fig.height = mm_inch(suppl$fatigue_clsters$h), fig.cap = 'Recovery from fatigue, tiredness and tachypnea in the HACT study recovery clusters.'}
-
-suppl$fatigue_clsters$plot
-
-```
-
-__Supplementary Figure S\@ref(fig:fig-fatigue-clsters). Recovery from fatigue, tiredness and tachypnea in the HACT study recovery clusters.__
-
-The recovery clusters were defined in the training Austria cohort (AT, __Supplementary Figure S\@ref(fig:fig-clust-devel)__) and the cluster assignment predicted in the test Italy cohort (IT, __Figure 5__).
-Differences in recovery times for fatigue (__A__), day tiredness (__B__) and tachypnea (__C__) between the clusters were assessed by Kruskal-Wallis test and $\eta^2$ effect size statistic.
-P values were corrected for multiple testing with Benjamini-Hochberg method.
-Symptom counts are presented in violin plots.
-Points represent single observations, orange diamonds with whiskers code for medians and interquartile ranges.
-Effect size statistics and p values are indicated in the plot caption.
-Numbers of complete observations are displayed under the plots.
-
-\newpage
-
-```{r fig-neurocog-clusters, fig.width = mm_inch(suppl$neurocog_clusters$w), fig.height = mm_inch(suppl$neurocog_clusters$h), fig.cap = 'Recovery from neurocognitive symptoms of COVID-19 in the HACT study recovery custers.'}
-
-suppl$neurocog_clusters$plot
-
-```
-
-__Supplementary Figure S\@ref(fig:fig-neurocog-clusters). Recovery from neurocognitive symptoms of COVID-19 in the HACT study recovery clusters.__
-
-The recovery clusters were defined in the training Austria cohort (AT, __Supplementary Figure S\@ref(fig:fig-clust-devel)__) and the cluster assignment predicted in the test Italy cohort (IT, __Figure 5__).
-Differences in recovery times for forgetfulness (__A__), concentration impairment (__B__) and confusion (__C__) between the clusters were assessed by Kruskal-Wallis test and $\eta^2$ effect size statistic.
-P values were corrected for multiple testing with Benjamini-Hochberg method.
-Symptom counts are presented in violin plots.
-Points represent single observations, orange diamonds with whiskers code for medians and interquartile ranges.
-Effect size statistics and p values are indicated in the plot caption.
-Numbers of complete observations are displayed under the plots.
-
-\newpage
-
-```{r fig-rec-clusters, fig.width = mm_inch(suppl$rec_clusters$w), fig.height = mm_inch(suppl$rec_clusters$h), fig.cap = 'Physical performance loss, impairment of quality of life and self-perceived complete convalescence in the HACT study recovery clusters.'}
-
-suppl$rec_clusters$plot
-
-```
-
-__Supplementary Figure S\@ref(fig:fig-rec-clusters).
-Physical performance loss, impairment of quality of life and self-perceived complete convalescence at the survey completion in the HACT study recovery clusters.__
-
-The recovery clusters were defined in the training Austria cohort (AT, __Supplementary Figure S\@ref(fig:fig-clust-devel)__) and the cluster assignment predicted in the test Italy cohort (IT, __Figure 5__).
-Differences in self-reported physical performance loss (__A__) and scoring of impairment of quality of life (QoL, __B__) at the survey completion between the clusters were assessed by Kruskal-Wallis test and $\eta^2$ effect size statistic.
-The scoring data were presented in violin plots.
-Points represent single observations, orange diamonds with whiskers code for medians and interquartile ranges.
-Differences in frequency of self-perceived complete convalescence (__C__) between the clusters were investigated by $\chi^2$ test and Cramer V effect size statistic.
-The frequencies are presented as bar plots.
-P values were corrected for multiple testing with Benjamini-Hochberg method.
-Effect size statistics and p values are indicated in the plot caption.
-Numbers of complete observations are displayed under the plots.
-
-\newpage
-
-```{r fig-psych-clusters, fig.width = mm_inch(suppl$psych_clusters$w), fig.height = mm_inch(suppl$psych_clusters$h), fig.cap = 'Rating of anxiety, depression, overall mental health impairment and stress in the HACT study recovery clusters.'}
-
-suppl$psych_clusters$plot
-
-```
-
-__Supplementary Figure S\@ref(fig:fig-psych-clusters). Rating of anxiety, depression, overall mental health impairment and stress at the survey completion in the HACT study recovery clusters.__
-
-The recovery clusters were defined in the training Austria cohort (AT, __Supplementary Figure S\@ref(fig:fig-clust-devel)__) and the cluster assignment predicted in the test Italy cohort (IT, __Figure 5__).
-Differences in rating of anxiety (ANX, __A__), depression (DPR, __B__), stress (__C__) and self-reported overall mental health impairment (OMH, __D__) at the survey completion between the clusters were assessed by Kruskal-Wallis test and $\eta^2$ effect size statistic.
-P values were corrected for multiple testing with Benjamini-Hochberg method.
-The scoring data were presented in violin plots.
-Points represent single observations, orange diamonds with whiskers code for medians and interquartile ranges.
-Effect size statistics and p values are indicated in the plot caption.
-Numbers of complete observations are displayed under the plots.
+Numbers of complete observations are displayed under the plots._
 

### 2022-07-13 analysis scripts/apriori.R

@@ -25,12 +25,14 @@
   ap_sympt$hact_trasactions <- ap_sympt$hact_tbl %>% 
     map(as, 'transactions')
   
-# Calculating the rules -----
-  
-  insert_msg('Rule calculation')
+  ## parallel backend
   
   plan('multisession')
   
+# Calculating the rules -----
+  
+  insert_msg('Rule calculation')
+
   ap_sympt$rules <- ap_sympt$hact_trasactions %>% 
     future_map(apriori, 
                parameter = list(support = 0.1, 
@@ -49,9 +51,7 @@
     future_map(DATAFRAME, 
                .options = furrr_options(seed = TRUE)) %>% 
     future_map(clear_rule_tbl)
-  
-  plan('sequential')
-  
+
 # Plotting the confidence versus support ------
   
   insert_msg('Confidence versus support plots')
@@ -61,12 +61,12 @@
                                          trans_lab = stri_replace(trans_lab, 
                                                                   fixed = ' \u2192 ', 
                                                                   replacement = '\u2192\n')), 
-                                   plot_title = rep(c('Acute CoV', 
-                                                      'Sub-acute CoV', 
-                                                      'Long CoV', 
-                                                      'PASC'), 2), 
-                                   plot_subtitle = c(rep('AT, HACT study', 4), 
-                                                     rep('IT, HACT study', 4))) %>% 
+                                   plot_title = rep(c('0 - 14 days', 
+                                                      '14 days', 
+                                                      '28 days', 
+                                                      '3 months'), 2), 
+                                   plot_subtitle = c(rep('AT, survey study', 4), 
+                                                     rep('IT, survey study', 4))) %>% 
     pmap(draw_conf_supp, 
          by = 'support',
          top_transactions = 7, 
@@ -81,12 +81,12 @@
   
   ap_sympt$top_support_plots <- list(data = ap_sympt$rules_tbl %>% 
                                        map(filter, stri_detect(signature, fixed = 'anosmia')), 
-                                     plot_title = rep(c('Acute CoV: top support', 
-                                                        'Sub-acute CoV: top support', 
-                                                        'Long CoV: top support', 
-                                                        'PASC: top support'), 2), 
-                                     plot_subtitle = c(rep('AT, HACT study', 4), 
-                                                       rep('IT, HACT study', 4)), 
+                                     plot_title = rep(c('0 - 14 days: top support', 
+                                                        '14 days: top support', 
+                                                        '28 days: top support', 
+                                                        '38 days: top support'), 2), 
[…]

### 2022-07-13 analysis scripts/clust_characteristic.R

@@ -50,32 +50,38 @@
                             'phq_depression_score', 
                             'stress_score')
   
+  ## parallel backend
+  
+  plan('multisession')
+  
 # descriptive stats -------
   
   insert_msg('Descriptive stats')
   
   clust_chara$desc_stats <- clust_chara$analysis_tbl %>% 
-    map(~explore(.x, 
-                 split_factor = 'clust_id', 
-                 variables = clust_chara$var_types$variable, 
-                 what = 'table', 
-                 pub_styled = TRUE)) %>% 
+    future_map(~explore(.x, 
+                        split_factor = 'clust_id', 
+                        variables = clust_chara$var_types$variable, 
+                        what = 'table', 
+                        pub_styled = TRUE), 
+               .options = furrr_options(seed = TRUE)) %>% 
     map(reduce, left_join, by = 'variable') %>% 
-    map(set_names, c('variable', 'clust_1', 'clust_2', 'clust_3'))
+    map(set_names, c('variable', names(globals$clust_colors)))
   
 # testing ------
   
   insert_msg('Testing for differences between the clusters')
   
   clust_chara$test_results <- clust_chara$analysis_tbl %>% 
-    map(~compare_variables(.x, 
-                           split_factor = 'clust_id', 
-                           variables = clust_chara$var_types$variable, 
-                           what = 'test', 
-                           types = clust_chara$var_types$type_test, 
-                           ci = FALSE, 
-                           pub_styled = TRUE, 
-                           adj_method = 'BH')) %>% 
+    future_map(~compare_variables(.x, 
+                                  split_factor = 'clust_id', 
+                                  variables = clust_chara$var_types$variable, 
+                                  what = 'test', 
+                                  types = clust_chara$var_types$type_test, 
+                                  ci = FALSE, 
+                                  pub_styled = TRUE, 
+                                  adj_method = 'BH'), 
+               .options = furrr_options(seed = TRUE)) %>% 
     map(mutate, 
         plot_cap = paste(eff_size, significance, sep = ', '))
   
@@ -96,14 +102,15 @@
   insert_msg('Summary plots, significance and effect size')
   
   clust_chara$summ_plots <- clust_chara$analysis_tbl %>% 
-    map(~compare_variables(.x, 
-                           split_factor = 'clust_id', 
-                           variables = clust_chara$var_types$variable, 
-                           what = 'test', 
-                           types = clust_chara$var_types$type_test, 
[…]

### 2022-07-13 analysis scripts/clust_devel.R

@@ -26,51 +26,61 @@
   
   ## hierarchical clustering
   
+  plan('multisession')
+  
   cl_devel$algos[paste0('hcl_', cl_devel$test_dist)] <- 
     cl_devel$test_dist %>% 
-    map(~hcluster(data = cl_devel$analysis_tbl$north, 
+    future_map(~hcluster(data = cl_devel$analysis_tbl$north, 
                   distance_method = .x, 
                   hc_method = 'ward.D2', 
                   k = 3, 
-                  seed = 1234))
+                  seed = 1234), 
+               .options = furrr_options(seed = TRUE))
   
   ## K-Means
   
   cl_devel$algos[paste0('kmeans_', cl_devel$test_dist)] <- 
     cl_devel$test_dist %>% 
-    map(~kcluster(data = cl_devel$analysis_tbl$north, 
-                  distance_method = .x, 
-                  clust_fun = 'kmeans', 
-                  k = 3, 
-                  seed = 1234))
+    future_map(~kcluster(data = cl_devel$analysis_tbl$north, 
+                         distance_method = .x, 
+                         clust_fun = 'kmeans', 
+                         k = 3, 
+                         seed = 1234), 
+               .options = furrr_options(seed = TRUE))
   
   ## PAM
   
   cl_devel$algos[paste0('pam_', cl_devel$test_dist)] <- 
     cl_devel$test_dist %>% 
-    map(~kcluster(data = cl_devel$analysis_tbl$north, 
-                  distance_method = .x, 
-                  clust_fun = 'pam', 
-                  k = 3, 
-                  seed = 1234))
+    future_map(~kcluster(data = cl_devel$analysis_tbl$north, 
+                         distance_method = .x, 
+                         clust_fun = 'pam', 
+                         k = 3, 
+                         seed = 1234),
+               .options = furrr_options(seed = TRUE))
   
   ## combi SOM + HCl clustering
   
   cl_devel$algos[paste0('combi_', cl_devel$test_dist)] <- 
     cl_devel$test_dist %>% 
-    map(~combi_cluster(data = cl_devel$analysis_tbl$north, 
-                       distance_som = .x, 
-                       xdim = cl_devel$som_x, 
-                       ydim = cl_devel$som_x, 
-                       topo = 'hexagonal', 
-                       neighbourhood.fct = 'gaussian', 
-                       toroidal = FALSE, 
-                       rlen = 1500, 
-                       node_clust_fun = hcluster, 
[…]

### 2022-07-13 analysis scripts/clust_features.R

@@ -44,16 +44,19 @@
     map(paste, collapse = ', ') %>% 
     map(~paste0('\n', .x))
   
+  plan('multisession')
+  
 # Descriptive stats ----
   
   insert_msg('Descriptive stats')
 
   clust_ft$clust_ft_stats <- clust_ft$analysis_tbl %>% 
-    map(~explore(.x, 
-                 split_factor = 'clust_id', 
-                 variables = globals$hact_symptoms, 
-                 what = 'table', 
-                 pub_styled = TRUE)) %>% 
+    future_map(~explore(.x, 
+                        split_factor = 'clust_id', 
+                        variables = globals$hact_symptoms, 
+                        what = 'table', 
+                        pub_styled = TRUE), 
+               .options = furrr_options(seed = TRUE)) %>% 
     map(reduce, left_join, by = 'variable') %>% 
     map(set_names, c('variable', 'clust_1', 'clust_2', 'clust_3'))
   
@@ -62,14 +65,15 @@
   insert_msg('Serial testing')
   
   clust_ft$clust_ft_test <- clust_ft$analysis_tbl %>% 
-    map(~compare_variables(.x, 
-                           split_factor = 'clust_id', 
-                           variables = globals$hact_symptoms, 
-                           what = 'test', 
-                           types = 'kruskal_test', 
-                           ci = FALSE, 
-                           pub_styled = TRUE, 
-                           adj_method = 'BH')) %>% 
+    future_map(~compare_variables(.x, 
+                                  split_factor = 'clust_id', 
+                                  variables = globals$hact_symptoms, 
+                                  what = 'test', 
+                                  types = 'kruskal_test', 
+                                  ci = FALSE, 
+                                  pub_styled = TRUE, 
+                                  adj_method = 'BH'), 
+               .options = furrr_options(seed = TRUE)) %>% 
     map(mutate, plot_cap = paste(eff_size, significance, sep = ', '))
   
 # Violin plots for all features -----
@@ -105,14 +109,15 @@
   insert_msg('Plots with summary test results')
   
   clust_ft$clust_ft_summ <- clust_ft$analysis_tbl %>% 
-    map(~compare_variables(.x, 
-                           split_factor = 'clust_id', 
-                           variables = globals$hact_symptoms, 
-                           what = 'test', 
-                           types = 'kruskal_test', 
-                           ci = FALSE, 
-                           pub_styled = FALSE, 
-                           adj_method = 'BH')) %>% 
+    future_map(~compare_variables(.x, 
[…]

### 2022-07-13 analysis scripts/clustering.R

@@ -52,7 +52,7 @@
     map(plot, 
         type = 'heat_map', 
         cust_theme = globals$common_theme) %>% 
-    map2(., c('AT, HACT study', 'IT, HACT study'), 
+    map2(., c('AT, survey study', 'IT, survey study'), 
          ~.x + 
            labs(subtitle = paste(.y, .x$labels$subtitle, sep = ', ')) + 
            theme(axis.text.x = element_blank(), 
@@ -69,7 +69,7 @@
         with = 'data', 
         kdim = 2, 
         cust_theme = globals$common_theme) %>% 
-    map2(., c('AT, HACT study', 'IT, HACT study'), 
+    map2(., c('AT, survey study', 'IT, survey study'), 
          ~.x + 
            scale_fill_manual(values = globals$clust_colors) + 
            labs(subtitle = paste(.y, .x$labels$subtitle, sep = ', ')))
@@ -84,7 +84,7 @@
         with = 'data', 
         kdim = 2, 
         cust_theme = globals$common_theme) %>% 
-    map2(., c('AT, HACT study', 'IT, HACT study'), 
+    map2(., c('AT, survey study', 'IT, survey study'), 
          ~.x + 
            scale_fill_manual(values = globals$clust_colors) + 
            labs(subtitle = paste(.y, .x$labels$subtitle, sep = ', ')))
@@ -99,7 +99,7 @@
         with = 'data', 
         kdim = 2, 
         cust_theme = globals$common_theme) %>% 
-    map2(., c('AT, HACT study', 'IT, HACT study'), 
+    map2(., c('AT, survey study', 'IT, survey study'), 
          ~.x + 
            scale_fill_manual(values = globals$clust_colors) + 
            labs(subtitle = paste(.y, .x$labels$subtitle, sep = ', ')))
@@ -117,7 +117,7 @@
                               fill_color = globals$hact_colors[1], 
                               cust_theme = globals$common_theme, 
                               plot_title = 'Importance of clustering variables', 
-                              plot_subtitle = 'AT, HACT study', 
+                              plot_subtitle = 'AT, survey study', 
                               label = FALSE) + 
     scale_y_discrete(labels = translate_var(globals$hact_symptoms, 
                                             dict = hact$dict))
@@ -128,8 +128,8 @@
   
   part_clust$feature_hm <- list(x_object = part_clust[c('clust_north', 
                                                         'clust_south')], 
-                                plot_subtitle = c('AT, HACT study', 
[…]

### 2022-07-13 analysis scripts/kinetic_modeling.R

@@ -126,7 +126,7 @@
   insert_msg('Recovery kinetic plots, HACT')
 
   kin_mod$plots_hact <- list(mod = kin_mod$models_hact, 
-                             sub = c('AT, HACT study', 'IT, HACT study'), 
+                             sub = c('AT, survey study', 'IT, survey study'), 
                              col = globals$hact_colors) %>% 
     pmap(function(mod, sub, col) list(x = mod, 
                                       plot_title = translate_var(names(mod), 

### 2022-07-13 analysis scripts/pca.R

@@ -18,6 +18,10 @@
     map(~filter(.x, complete.cases(.x))) %>% 
     map(column_to_rownames, 'ID')
   
+  ## parallel backend
+  
+  plan('multisession')
+  
 # PCA -----
   
   insert_msg('PCA and visualizations')
@@ -40,15 +44,15 @@
     pmap(plot,
          type = 'scree', 
          cust_theme = globals$common_theme) %>% 
-    map2(c('AT, HACT study', 
-           'IT, HACT study'), 
+    map2(c('AT, survey study', 
+           'IT, survey study'), 
          ~.x + labs(subtitle = .y))
   
   ## loadings
   
   pca$pca_loadings_plot <- list(x = pca$pca_obj, 
-                                plot_subtitle = c('AT, HACT study', 
-                                                  'IT, HACT study'), 
+                                plot_subtitle = c('AT, survey study', 
+                                                  'IT, survey study'), 
                                 point_color = globals$hact_colors) %>% 
     pmap(plot,
          type = 'loadings', 
@@ -58,8 +62,8 @@
   ## score plots
   
   pca$pca_score_plot <- list(x = pca$pca_obj, 
-                             plot_subtitle = c('AT, HACT study', 
-                                               'IT, HACT study'), 
+                             plot_subtitle = c('AT, survey study', 
+                                               'IT, survey study'), 
                              point_color = globals$hact_colors) %>% 
     pmap(plot,
          type = 'scores', 
@@ -72,8 +76,8 @@
     map(mutate, len = sqrt(comp_1^2 + comp_2^2 + comp_3^2 + comp_4^2 + comp_5^2 + comp_6^2 + comp_7^2 + comp_8^2))
   
   pca$pca_imp_plot <- list(data = pca$pca_imp_plot, 
-                           plot_subtitle = c('AT, HACT study', 
-                                             'IT, HACT study'), 
+                           plot_subtitle = c('AT, survey study', 
+                                             'IT, survey study'), 
                            fill_scale = as.list(globals$hact_colors)) %>% 
     pmap(plot_top, 
          regulation_variable = 'len', 
@@ -99,8 +103,8 @@
   ## score plot
   
   pca$mds_plot <- list(x = pca$mds_obj, 
-                       plot_subtitle = c('AT, HACT study', 
-                                         'IT, HACT study'), 
+                       plot_subtitle = c('AT, survey study', 
+                                         'IT, survey study'), 
[…]

### 2022-07-13 analysis scripts/recovery_distribution.R

@@ -13,17 +13,20 @@
   rec_dist$tbl_hact <- rec_time[c('north', 'south')] %>% 
     map2_dfr(., c('AT', 'IT'), ~mutate(.x, cohort = .y))
   
+  plan('multisession')
+  
 # descriptive stats of the recovery: individuals with the symptom present -----  
   
   insert_msg('Distribution stats for the rec. times, symptomatic  participants')
   
   rec_dist$stats_hact <- globals$hact_symptoms %>% 
-    map_dfr(~explore(rec_dist$tbl_hact[c('cohort', .x)] %>% 
-                       filter(.data[[.x]] != 0), 
-                     split_factor = 'cohort', 
-                     variables = .x, 
-                     what = 'table', 
-                     pub_styled = TRUE)) %>% 
+    future_map_dfr(~explore(rec_dist$tbl_hact[c('cohort', .x)] %>% 
+                              filter(.data[[.x]] != 0), 
+                            split_factor = 'cohort', 
+                            variables = .x, 
+                            what = 'table', 
+                            pub_styled = TRUE), 
+                   .options = furrr_options(seed = TRUE)) %>% 
     reduce(left_join, by = 'variable') %>% 
     set_names(c('variable', 'north', 'south'))
   
@@ -43,15 +46,16 @@
   insert_msg('Testing for the differences between the cohorts or severity')
   
   rec_dist$test_hact <- globals$hact_symptoms %>% 
-    map(~safely(compare_variables)(rec_dist$tbl_hact[c('cohort', .x)] %>% 
-                                     filter(.data[[.x]] != 0), 
-                                   split_factor = 'cohort', 
-                                   variables = .x, 
-                                   what = 'eff_size', 
-                                   types = 'wilcoxon_r', 
-                                   ci = FALSE, 
-                                   pub_styled = TRUE, 
-                                   adj_method = 'none')) %>% 
+    future_map(~safely(compare_variables)(rec_dist$tbl_hact[c('cohort', .x)] %>% 
+                                            filter(.data[[.x]] != 0), 
+                                          split_factor = 'cohort', 
+                                          variables = .x, 
+                                          what = 'eff_size', 
+                                          types = 'wilcoxon_r', 
+                                          ci = FALSE, 
+                                          pub_styled = TRUE, 
+                                          adj_method = 'none'), 
+               .options = furrr_options(seed = TRUE)) %>% 
     map_dfr(~.x$result) %>% 
     re_adjust('BH')
 
[…]

### 2022-07-13 analysis scripts/symptom_distances.R

@@ -79,12 +79,12 @@
          cust_theme = globals$common_theme)
   
   sympt_dist$mds_plots_hact <- list(x = sympt_dist$mds_plots_hact, 
-                                    y = rep(c('Acute CoV', 
-                                              'Sub-acute CoV', 
-                                              'Long CoV', 
-                                              'PASC'), 2), 
-                                    z = c(rep('AT, HACT study', 4), 
-                                          rep('IT, HACT study', 4))) %>% 
+                                    y = rep(c('0 - 14 days', 
+                                              '14 days', 
+                                              '28 days', 
+                                              '3 months'), 2), 
+                                    z = c(rep('AT, survey study', 4), 
+                                          rep('IT, survey study', 4))) %>% 
     pmap(function(x, y, z) x + 
            labs(title = y, 
                 subtitle = z) + 
@@ -172,8 +172,8 @@
   
   sympt_dist$kNN_plots_hact <- list(data = sympt_dist$mean_kNN_hact, 
                                     highlight_color = globals$hact_colors, 
-                                    plot_subtitle = c('AT, HACT study', 
-                                                      'IT, HACT study'), 
+                                    plot_subtitle = c('AT, survey study', 
+                                                      'IT, survey study'), 
                                     plot_tag = paste('\nn =', sympt_dist$n_numbers_hact)) %>% 
     pmap(draw_knn_course, 
          label_var = 'symptom', 

### 2022-07-13 exploration scripts/cohort.R

@@ -6,7 +6,7 @@
 # container list ------
   
   eda_cohort <- list()
-  
+
 # globals ------
   
   insert_msg('Globals setup')
@@ -117,5 +117,5 @@
     format_summ_tbl(dict = covild$dict)
   
 # END -----
-  
+
   insert_tail()

### 2022-07-13 exploration scripts/symptoms.R

@@ -40,17 +40,22 @@
                                          nrow) %>% 
                                      unlist)
   
+  ## parallel backend
+  
+  plan('multisession')
+  
 # descriptive stats ------
   
   insert_msg('Descriptive stats')
   
   eda_sympt$hact_stats <- eda_sympt$hact_tbl %>% 
-    map(explore, 
-        split_factor = 'cohort', 
-        variables = globals$hact_symptoms, 
-        what = 'table', 
-        pub_styled = TRUE) %>% 
-    map(reduce, left_join, by = 'variable') %>% 
+    future_map(explore, 
+               split_factor = 'cohort', 
+               variables = globals$hact_symptoms, 
+               what = 'table', 
+               pub_styled = TRUE, 
+               .options = furrr_options(seed = TRUE)) %>% 
+    future_map(reduce, left_join, by = 'variable') %>% 
     map(set_names, c('variable', 'AT', 'IT')) %>% 
     map(map_dfr, stri_replace_all, regex = '0:.*\\n1:\\s{1}', replacement = '') %>% 
     map(map_dfr, stri_replace_all, fixed = '% (', replacement = '% (n = ') %>% 
@@ -73,14 +78,15 @@
   insert_msg('Testing')
   
   eda_sympt$hact_test <- eda_sympt$hact_tbl %>% 
-    map(compare_variables, 
-        split_factor = 'cohort', 
-        variables = globals$hact_symptoms, 
-        what = 'test',
-        types = 'chisq_test', 
-        ci = FALSE, 
-        adj_method = 'BH', 
-        pub_styled = TRUE)
+    future_map(compare_variables, 
+               split_factor = 'cohort', 
+               variables = globals$hact_symptoms, 
+               what = 'test',
+               types = 'chisq_test', 
+               ci = FALSE, 
+               adj_method = 'BH', 
+               pub_styled = TRUE, 
+               .options = furrr_options(seed = TRUE))
   
   eda_sympt$covild_test <- eda_sympt$covild_tbl %>% 
     map(compare_variables, 
@@ -111,10 +117,11 @@
   insert_msg('Summary of the symptom frequencies')
   
   eda_sympt$hact_sympt_freq <- eda_sympt$hact_tbl %>% 
-    map(explore, 
-        split_factor = 'cohort', 
-        variables = globals$hact_symptoms, 
-        what = 'list') %>% 
+    future_map(explore, 
+               split_factor = 'cohort', 
[…]

### 2022-07-13 exploration.R

@@ -5,6 +5,7 @@
   library(exda)
   library(soucer)
   library(stringi)
+  library(furrr)
   
   insert_head()
   

### 2022-07-13 paper scripts/figures.R

@@ -13,6 +13,8 @@
   figures$consort <- plot_grid(ggdraw() + 
     draw_image('./study consort/consort_diagram.png')) %>% 
     as_figure('figure_1_consort', 
+              ref_name = 'consort', 
+              caption = 'Flow diagram of the analysis inclusion process for the longitudinal CovILD cohort and the Health after COVID-19 survey study.', 
               w = 180, 
               h = 110)
   
@@ -50,6 +52,8 @@
                                       labels = c('A', 'B'), 
                                       label_size = 10) %>% 
     as_figure('figure_2_recovery_times', 
+              ref_name = 'sympt_recovery', 
+              caption = 'Symptom-specific recovery times in ambulatory COVID-19 survey study.', 
               w = 180, 
               h = 220)
 
@@ -72,6 +76,8 @@
               labels = c('A', '', 'B', ''), 
               label_size = 10) %>% 
     as_figure('figure_3_symptom_isolation', 
+              ref_name = 'sympt_mds', 
+              caption = 'Self-reported smell and taste disorders are isolated persistent symptoms of COVID-19.', 
               w = 180, 
               h = 180)
 
@@ -104,36 +110,14 @@
               nrow = 2, 
               rel_heights = c(0.92, 0.08)) %>% 
     as_figure('figure_4_apriori_analysis', 
+              ref_name = 'apriori', 
+              caption = 'Co-occurrence of smell and taste disorders in post-acute COVID-19 sequelae.', 
               w = 180, 
               h = 190)
   
-# Figure 5: clustering of the participants ------
+# Figure 5: symptom duration in the recovery clusters -----
   
-  insert_msg('Figure 5: clustering')
-  
-  figures$clustering <- part_clust$feature_hm %>% 
-    map(~.x + 
-          scale_y_discrete(labels = function(x) embolden_scale(x, 
-                                                               c('anosmia', 'taste_loss'), 
-                                                               translate = TRUE), 
-                           limits = globals$hact_symptom_order) + 
-          theme(axis.text.y = element_markdown(), 
-                legend.position = 'none')) %>% 
-    plot_grid(plotlist = ., 
-              ncol = 2, 
-              align = 'hv', 
-              axis = 'tblr') %>% 
-    plot_grid(get_legend(part_clust$feature_hm[[1]] + 
[…]

### 2022-07-13 paper scripts/manuscript.R

@@ -1,69 +1,7 @@
 # Renders the manuscript figures/tables and the supplementary material
 
   insert_head()
-  
-# Exporting the figure chunks -----
-  
-  insert_msg('Figure chunks')
-  
-  insert_figure(figures$consort, 
-                figures$sympt_recovery, 
-                figures$sympt_mds, 
-                figures$apriori, 
-                figures$clustering, 
-                figures$sympt_clusters, 
-                figures$base_clusters, 
-                figures$recovery_clusters, 
-                file = './paper/markdown/figure_chunks.Rmd', 
-                ref_names = stri_replace_all(names(figures), 
-                                             fixed = '_', 
-                                             replacement = '-'), 
-                captions = c('Flow diagram of the analysis inclusion process for the CovILD and HACT studies.', 
-                             'Symptom-specific recovery times in ambulatory COVID-19.', 
-                             'Self-reported smell and taste disorders are isolated symptoms of long COVID and PASC.', 
-                             'High extent of co-occurrence of smell and taste disorders in long COVID and PASC.', 
-                             'Clustering of ambulatory COVID-19 individuals by symptom-specific recovery times.', 
-                             'Duration of neurocognitive and respiratory symptoms, fatigue, smell and taste disorders differs between the COVID-19 recovery clusters.', 
-                             'COVID-19 recovery clusters differ in sex distribution, comorbidity and daily medication rates.', 
-                             'Good clinical, mental and psychosocial recovery in individuals with persistent smell and taste disorders'))
-  
-# Exporting the supplement chunks -----
-  
-  insert_msg('Supplement chunks')
-  
-  insert_figure(suppl$hact_freq, 
-                suppl$hact_recovery, 
-                suppl$covild_freq, 
-                suppl$covild_recovery, 
-                suppl$hyposmia_rater100, 
-                suppl$hyposmia_rater360, 
-                suppl$mds_acute, 
-                suppl$clust_devel, 
-                suppl$cov_clusters,
-                suppl$fatigue_clsters, 
-                suppl$neurocog_clusters, 
-                suppl$anosmia_clusters, 
-                suppl$rec_clusters, 
-                suppl$psych_clusters, 
-                file = './paper/markdown/supplement_chunks.Rmd', 
-                ref_names = stri_replace_all(names(suppl), 
-                                             fixed = '_', 
-                                             replacement = '-'), 
-                captions = c('Frequency of COVID-19 symptoms in acute COVID-19, long COVID and PASC in the HACT study.', 
[…]

### 2022-07-13 paper scripts/supplements.R

@@ -23,6 +23,8 @@
               labels = LETTERS, 
               label_size = 10) %>% 
     as_figure('figure_s1_hact_symp_frequency', 
+              ref_name = 'hact_freq', 
+              caption = 'Frequency of COVID-19 symptoms in the survey study.', 
               w = 180, 
               h = 190)
   
@@ -53,6 +55,8 @@
                          'F', ''), 
               label_size = 10) %>% 
     as_figure('figure_s2_hact_kinetic', 
+              ref_name = 'hact_recovery', 
+              caption = 'Kinetic of recovery from leading acute COVID-19 symptoms in the survey study.', 
               w = 180, 
               h = 180)
   
@@ -74,6 +78,8 @@
               labels = LETTERS, 
               label_size = 10) %>% 
     as_figure('figure_s3_covild_symp_frequency', 
+              ref_name = 'covild_freq', 
+              caption = 'Symptom frequency in ambulatory, moderate and severe COVID-19 subsets of the CovILD study.', 
               w = 180, 
               h = 220)
   
@@ -96,6 +102,8 @@
                          'C', '', ''), 
               label_size = 10) %>% 
     as_figure('figure_s4_covild_kinetic', 
+              ref_name = 'covild_recovery', 
+              caption = 'Kinetic of recovery from smell disorders, reduced performance and dyspnea in ambulatory, moderate and severe COVID-19 subsets of the CovILD study.', 
               w = 180, 
               h = 180)
   
@@ -118,6 +126,8 @@
               labels = LETTERS, 
               label_size = 10) %>% 
     as_figure('figure_s5_hyposmia_rater100', 
+              re_name = 'hyposmia_rater100', 
+              caption = 'Rates of self-reported hyposmia and hyposmia in sniffing stick test at 3-month post COVID-19 follow-up in the ambulatory, moderate and severe COVID-19 subsets of the CovILD study.', 
               w = 180, 
               h = 160)
   
@@ -140,6 +150,8 @@
               labels = LETTERS, 
               label_size = 10) %>% 
     as_figure('figure_s6_hyposmia_rater360', 
+              ref_name = 'hyposmia_rater360', 
[…]

### 2022-07-13 paper scripts/tables.R

@@ -31,6 +31,12 @@
     select(variable, north, south, significance, eff_size) %>% 
     set_names(c('Variable', 'AT', 'IT', 'Significance', 'Effect size'))
   
+  paper_tbl$hact_baseline <- 
+    as_mdtable(paper_tbl$hact_baseline, 
+                                        label = 'table_1_hact_baseline', 
+                                        ref_name = 'hact', 
+                                        caption = 'Baseline characteristic of the Austria (AT) and Italy (IT) survey study cohorts.')
+  
 # Table 2: baseline characteristic of the CovILD cohort ------
   
   insert_msg('Table 2: baseline, CovILD')
@@ -54,6 +60,11 @@
                 'Severe CoV subset', 
                 'Significance', 
                 'Effect size'))
+  
+  paper_tbl$covild_baseline  <- as_mdtable(paper_tbl$covild_baseline, 
+                                           label = 'table_2_hact_baseline', 
+                                           ref_name = 'covild', 
+                                           caption = 'Baseline characteristic of the CovILD study cohort and the study participants stratified by COVID-19 severity.')
   
 # Supplementary Table S1: HACT study variables ------
   
@@ -82,6 +93,12 @@
                 'Stratification', 
                 'Description'))
   
+  suppl_tbl$hact_vars <- 
+    as_mdtable(suppl_tbl$hact_vars, 
+               label = 'table_s1_hact_vars', 
+               ref_name = 'hact_vars', 
+               caption = 'Survey study variables.')
+  
 # Supplementary Table S2: CovILD study variables -----
   
   insert_msg('Table S2: CovILD study variables')
@@ -97,31 +114,14 @@
     select(variable, label, unit) %>% 
     set_names(c('Variable name', 'Variable label', 'Unit'))
   
-# Supplementary Table S3: Apriori results, long CoV -----
+  suppl_tbl$covild_vars <- as_mdtable(suppl_tbl$covild_vars, 
+                                      label = 'table_s2_covild_vars', 
+                                      ref_name = 'covild_vars', 
+                                      caption = 'CovILD study variables.')
   
-  insert_msg('Table S3: apriori results, long COVID')
+# Supplementary Table S3 - S4: demographic/clinical characteristic, recovery clusters ------
   
-  suppl_tbl$apriori_long <- ap_sympt$rules_tbl[c('north.28', 'south.28')] %>% 
[…]

### 2022-07-13 paper.R

@@ -27,7 +27,8 @@
   c('./paper scripts/tables.R', 
     './paper scripts/figures.R', 
     './paper scripts/supplements.R', 
-    './paper scripts/manuscript.R') %>% 
+    './paper scripts/manuscript.R'
+    ) %>% 
     source_all(message = TRUE, crash = TRUE)
   
 # END ----

### 2023-04-03 analysis scripts/apriori.R
Format of figures and references im Text

@@ -35,10 +35,10 @@
 
   ap_sympt$rules <- ap_sympt$hact_trasactions %>% 
     future_map(apriori, 
-               parameter = list(support = 0.1, 
+               parameter = list(support = 0.15, 
                                 maxlen = 10, 
                                 minlen = 2, 
-                                confidence = 0.8), 
+                                confidence = 0.6), 
                .options = furrr_options(seed = TRUE)) %>% 
     map(~.x[!is.redundant(.x)]) %>% 
     map(~subset(.x, subset = lift > 2))
@@ -70,8 +70,8 @@
     pmap(draw_conf_supp, 
          by = 'support',
          top_transactions = 7, 
-         embolden = c('{Hypo/ageusia}\u2192\n{Hypo/anosmia}', 
-                      '{Hypo/anosmia}\u2192\n{Hypo/ageusia}'), 
+         embolden = c('{Hypogeusia/ageusia}\u2192\n{OD}', 
+                      '{OD}\u2192\n{Hypogeusia/ageusia}'), 
          label_var = 'trans_lab', 
          cust_theme = globals$common_theme)
   
@@ -84,7 +84,7 @@
                                      plot_title = rep(c('0 - 14 days: top support', 
                                                         '14 days: top support', 
                                                         '28 days: top support', 
-                                                        '38 days: top support'), 2), 
+                                                        '90 days: top support'), 2), 
                                      plot_subtitle = c(rep('AT, survey study', 4), 
                                                        rep('IT, survey study', 4)), 
                                      fill_scale = as.list(c(rep(globals$hact_colors[1], 4), 
@@ -103,7 +103,7 @@
                                         plot_title = rep(c('0 - 14 days: top confidence', 
                                                            '14 days: top confidence', 
                                                            '28 days: top confidence', 
-                                                           '3 months: top confidence'), 2), 
+                                                           '90 days: top confidence'), 2), 
                                         plot_subtitle = c(rep('AT, survey study', 4), 
                                                           rep('IT, survey study', 4)), 
                                         fill_scale = as.list(c(rep(globals$hact_colors[1], 4), 
@@ -122,7 +122,7 @@
                                   plot_title = rep(c('0 - 14 days: top lift', 
                                                      '14 days: top lift', 
                                                      '28 days: top lift', 
[…]

### 2023-04-03 analysis scripts/clust_characteristic.R

@@ -164,9 +164,9 @@
                 split_factor = 'clust_id', 
                 scale = 'percent', 
                 cust_theme = globals$common_theme, 
-                x_lab = 'Recovery cluster') %>% 
+                x_lab = 'Cluster') %>% 
            map(~.x + 
-                 scale_x_discrete(limits = c('STDR', 'RR', 'SR')) + 
+                 scale_x_discrete(limits = c('#1', '#2', '#3')) + 
                  labs(tag = .x$labels$tag %>% 
                         stri_replace_all(fixed = '\n', replacement = ', ') %>% 
                         paste0('\n', .))) %>% 
@@ -240,9 +240,9 @@
     map2(., clust_chara$ribbon_labs, 
          ~.x + 
            scale_fill_manual(values = globals$clust_colors, 
-                             name = 'Recovery\ncluster') + 
+                             name = 'Cluster') + 
            scale_color_manual(values = globals$clust_colors, 
-                              name = 'Recovery\ncluster') + 
+                              name = 'Cluster') + 
            scale_y_discrete(labels = .y, 
                             limits = clust_chara$rec_vars) + 
            scale_x_continuous(breaks = seq(0.1, 0.8, by = 0.1)) + 

### 2023-04-03 analysis scripts/clust_features.R

@@ -92,7 +92,7 @@
                 x, 
                 split_factor = 'clust_id', 
                 type = 'violin', 
-                x_lab = 'Recovery cluster', 
+                x_lab = 'Cluster', 
                 y_lab = 'Recovery time, days post CoV', 
                 point_hjitter = 2, 
                 cust_theme = globals$common_theme) %>% 
@@ -166,9 +166,9 @@
     map(~.x + 
           theme(axis.title.y = element_blank()) + 
           scale_fill_manual(values = globals$clust_colors, 
-                            name = 'Recovery\ncluster') + 
+                            name = 'Cluster') + 
           scale_color_manual(values = globals$clust_colors, 
-                             name = 'Recovery\ncluster') + 
+                             name = 'Cluster') + 
           scale_y_discrete(labels = translate_var(globals$hact_symptoms, 
                                                   dict = hact$dict), 
                            limits = globals$hact_symptom_order))

### 2023-04-03 analysis scripts/clustering.R

@@ -19,8 +19,8 @@
   part_clust$clust_north$clust_assignment <- 
     part_clust$clust_north$clust_assignment %>% 
     mutate(clust_id = car::recode(clust_id, 
-                                  "'1' = 'SR'; '2' = 'RR'; '3' = 'STDR'"), 
-           clust_id = factor(clust_id, c('STDR', 'RR', 'SR'))) 
+                                  "'1' = '#3'; '2' = '#2'; '3' = '#1'"), 
+           clust_id = factor(clust_id, c('#1', '#2', '#3'))) 
   
 # cluster projection -----
   
@@ -33,7 +33,7 @@
   
   part_clust$clust_south$clust_assignment <- 
     part_clust$clust_south$clust_assignment %>% 
-    mutate(clust_id = factor(clust_id, c('STDR', 'RR', 'SR'))) 
+    mutate(clust_id = factor(clust_id, c('#1', '#2', '#3'))) 
   
 # Diagnostic of the clustering objects ------
   

### 2023-04-03 analysis scripts/hyposmia_rater.R

@@ -48,14 +48,14 @@
          variables = c('anosmia_sympt', 'sniff_hyposmia'), 
          scale = 'percent', 
          type = 'bar', 
-         x_lab = 'Hypo/anosmia', 
+         x_lab = 'OD', 
          cust_theme = globals$common_theme) %>% 
     map2(.,map(rater$analysis_tbl, nrow), 
          ~.x + 
            labs(subtitle = paste0(.x$labels$subtitle, ', n = ', .y)) + 
            scale_fill_manual(values = c('bisque3', 'steelblue'), 
                              labels = c('self-reported', 'test'), 
-                             name = 'Hyposmia'))
+                             name = 'OD'))
   
 # Plotting the interrater results, bubble -----
   
@@ -71,14 +71,14 @@
          variables = c('anosmia_sympt', 'sniff_hyposmia'), 
          scale = 'percent', 
          type = 'bubble', 
-         x_lab = 'Hypo/anosmia', 
+         x_lab = 'OD', 
          point_alpha = 1, 
          cust_theme = globals$common_theme) %>% 
     map2(.,map(rater$analysis_tbl, nrow), 
          ~.x + 
            labs(subtitle = paste0(.x$labels$subtitle, ', n = ', .y), 
-                x = 'Self-reported hyposmia', 
-                y = 'Test hyposmia') + 
+                x = 'Self-reported OD', 
+                y = 'Test OD') + 
            scale_size(limits = c(0, 71)) + 
            scale_fill_gradient2(low = 'steelblue', 
                                 mid = 'white', 

### 2023-04-03 analysis scripts/kinetic_hyposmia_long.R

@@ -1,0 +1,109 @@
+# Kinetic modeling of the log-term hyposmia in the HACT survey.
+# Participants with an observation time >= 180 days at the 6 month time-point
+# are included
+
+  insert_head()
+  
+# container -------
+  
+  hyp_long <- list()
+  
+# globals -----
+  
+  insert_msg('Globals')
+  
+  hyp_long$symptoms <- c(anosmia = 'anosmia', 
+                         taste_loss = 'taste_loss')
+  
+  hyp_long$analysis_tbl <- hact[c('north', 'south')] %>% 
+    map(filter, obs_time >= 180) %>% 
+    map(select, ID, all_of(hyp_long$symptoms))
+  
+  
+  hyp_long$analysis_tbl <-  hyp_long$analysis_tbl %>% 
+    map(function(cohort) cohort %>% 
+          dlply(.(ID), as_tibble) %>% 
+          map(select, -ID) %>% 
+          map(~map_dfc(.x, function(sympt) switch(as.character(sympt), 
+                                                  absent = c(0, 0, 0, 0, 0, 0), 
+                                                  `1 - 3 days` = c(1, 0, 0, 0, 0, 0), 
+                                                  `up to 1 week` = c(1, 1, 0, 0, 0, 0), 
+                                                  `up to 2 weeks` = c(1, 1, 1, 0, 0, 0), 
+                                                  `up to 4 weeks` = c(1, 1, 1, 1, 0, 0), 
+                                                  `up to 3 months` = c(1, 1, 1, 1, 1, 0), 
+                                                  `up to 6 months` = c(1, 1, 1, 1, 1, 1), 
+                                                  `over 6 months` = c(1, 1, 1, 1, 1, 1)))) %>% 
+          map2_dfr(., names(.), ~mutate(.x, ID = .y, time = c(3, 7, 14, 28, 90, 180)))) %>% 
+    map(as_tibble) %>% 
+    map(complete_cases)
+
+# building the mixed-effect linear models ------
+  
+  insert_msg('Generating the models')
+  
+  hyp_long$models$north <- hyp_long$symptoms %>% 
+    map(~safely(model_kinetic)(data = hyp_long$analysis_tbl$north, 
+                               response = .x, 
+                               time = 'time', 
+                               ID = 'ID', 
+                               family = 'binomial', 
+                               order = 2))
+  
+  hyp_long$models$south <- hyp_long$symptoms %>% 
+    map(~safely(model_kinetic)(data = hyp_long$analysis_tbl$south, 
[…]

### 2023-04-03 analysis.R

@@ -26,6 +26,7 @@
   
   c('./analysis scripts/recovery_distribution.R', 
     './analysis scripts/kinetic_modeling.R', 
+    './analysis scripts/kinetic_hyposmia_long.R', 
     './analysis scripts/symptom_distances.R', 
     './analysis scripts/apriori.R', 
     './analysis scripts/pca.R', 

### 2023-04-03 exec_all.R

@@ -7,6 +7,6 @@
                      'analysis.R', 
                      'paper.R'), 
                    message = TRUE, 
-                   crash = FALSE))
+                   crash = TRUE))
   
   save.image()

### 2023-04-03 exploration scripts/symptoms.R

@@ -170,11 +170,12 @@
           scale_x_discrete(labels = c(acute = '0 - 14 days', 
                                       long = '28 days', 
                                       pasc = '3 months')) + 
-          scale_fill_gradient2(low = 'steelblue', 
-                               mid = 'white', 
-                               high = 'firebrick', 
-                               midpoint = 50, 
-                               limits = c(0, 100)) + 
+          scale_fill_viridis_c(limits = c(0, 100)) + 
+          #scale_fill_gradient2(low = 'steelblue', 
+            #                   mid = 'white', 
+             #                  high = 'firebrick', 
+              #                 midpoint = 50, 
+               #                limits = c(0, 100)) + 
           facet_grid(class ~ ., 
                      scales = 'free', 
                      space = 'free') + 
@@ -215,11 +216,12 @@
                                           'gastro_sympt', 
                                           'fever_sympt'))) + 
           scale_size_area(max_size = 4) + 
-          scale_fill_gradient2(low = 'steelblue', 
-                               mid = 'white', 
-                               high = 'firebrick', 
-                               midpoint = 50, 
-                               limits = c(0, 100)) + 
+          scale_fill_viridis_c(limits = c(0, 100)) + 
+         # scale_fill_gradient2(low = 'steelblue', 
+           #                    mid = 'white', 
+            #                   high = 'firebrick', 
+            #                   midpoint = 50, 
+             #                  limits = c(0, 100)) + 
           guides(fill = FALSE, 
                  size = FALSE))
   

### 2023-04-03 import.R

@@ -212,6 +212,26 @@
               by = 'ID') %>% 
     filter(complete.cases(.))
   
+# manual editing of the variable dictionary -------
+  
+  insert_msg('Manual editing of the variable dictionary')
+  
+  hact$dict <- hact$dict %>% 
+    mutate(label = ifelse(variable == 'anosmia', 
+                          'OD', label), 
+           label_short = ifelse(variable == 'anosmia', 
+                                'OD', label_short), 
+           label = ifelse(variable == 'taste_loss', 
+                          'Hypogeusia/ageusia', label), 
+           label_short = ifelse(variable == 'taste_loss', 
+                                'hypogeusia/ageusia', label_short))
+  
+  covild$dict <- covild$dict %>% 
+    mutate(label = ifelse(variable == 'anosmia_sympt', 
+                          'OD', label), 
+           label_short = ifelse(variable == 'anosmia_sympt', 
+                                'OD', label))
+  
 # END ----
   
   insert_tail()

### 2023-04-03 paper scripts/figures.R

@@ -6,27 +6,15 @@
   
   figures <- list()
   
-# Figure 1: CONSORT -------
+# Figure 1: symptom recovery times -----
   
-  insert_msg('Figure 1: CONSORT')
-  
-  figures$consort <- plot_grid(ggdraw() + 
-    draw_image('./study consort/consort_diagram.png')) %>% 
-    as_figure('figure_1_consort', 
-              ref_name = 'consort', 
-              caption = 'Flow diagram of the analysis inclusion process for the longitudinal CovILD cohort and the Health after COVID-19 survey study.', 
-              w = 180, 
-              h = 110)
-  
-# Figure 2: symptom recovery times -----
-  
-  insert_msg('Figure 2: symptom recovery times')
+  insert_msg('Figure 1: symptom recovery times')
   
   ## emboldening the hyposmia and hypogeusia
 
   figures$sympt_recovery$upper_panel <- rec_dist$summ_plot_hact %>% 
     map(~.x + 
-          scale_y_discrete(labels = function(x) embolden_scale(x, c('Hypo/anosmia', 'Hypo/ageusia'))) + 
+          scale_y_discrete(labels = function(x) embolden_scale(x, c('OD', 'Hypogeusia/ageusia'))) + 
           theme(axis.text.y = element_markdown())) %>% 
     plot_grid(plotlist = ., 
               ncol = 2, 
@@ -51,15 +39,15 @@
                                       rel_heights = c(0.7, 0.3), 
                                       labels = c('A', 'B'), 
                                       label_size = 10) %>% 
-    as_figure('figure_2_recovery_times', 
+    as_figure('figure_1_recovery_times', 
               ref_name = 'sympt_recovery', 
-              caption = 'Symptom-specific recovery times in ambulatory COVID-19 survey study.', 
+              caption = 'Symptom-specific recovery times in the ambulatory COVID-19 survey study.', 
               w = 180, 
               h = 220)
 
-# Figure 3: symptom distances -----
+# Figure 2: symptom distances -----
   
-  insert_msg('Figure 3: symptom isolation')
+  insert_msg('Figure 2: symptom isolation')
   
   figures$sympt_mds <- sympt_dist$mds_plots_hact[c('north.long', 
                                                    'south.long', 
@@ -75,49 +63,41 @@
               axis = 'tblr', 
               labels = c('A', '', 'B', ''), 
               label_size = 10) %>% 
[…]

### 2023-04-03 paper scripts/supplements.R

@@ -6,9 +6,21 @@
   
   suppl <- list()
   
-# Figure S1: Frequency of symptoms, HACT -----
+# Figure S1: CONSORT ------- 
   
-  insert_msg('Figure S1: HACT symptom frequency')
+  insert_msg('Figure S1: CONSORT diagram')
+
+  suppl$consort <- plot_grid(ggdraw() + 
+                                 draw_image('./study consort/consort_diagram.png')) %>% 
+    as_figure('figure_s1_consort', 
+              ref_name = 'consort', 
+              caption = 'Flow diagram of the analysis inclusion process for the longitudinal CovILD cohort and the Health after COVID-19 survey study.', 
+              w = 180, 
+              h = 110)
+    
+# Figure S2: Frequency of symptoms, HACT -----
+  
+  insert_msg('Figure S2: HACT symptom frequency')
   
   suppl$hact_freq <- eda_sympt$hact_bubble %>% 
     map(~.x + 
@@ -22,15 +34,15 @@
               axis = 'tblr', 
               labels = LETTERS, 
               label_size = 10) %>% 
-    as_figure('figure_s1_hact_symp_frequency', 
+    as_figure('figure_s2_hact_symp_frequency', 
               ref_name = 'hact_freq', 
               caption = 'Frequency of COVID-19 symptoms in the survey study.', 
               w = 180, 
               h = 190)
   
-# Figure S2: HACT recovery symptoms -------
+# Figure S3: HACT recovery symptoms -------
   
-  insert_msg('Figure S2: fast-recovery symptoms')
+  insert_msg('Figure S3: fast-recovery symptoms')
   
   suppl$hact_recovery <- kin_mod$plots_hact %>% 
     map(~.x[c('fever', 'dim_appetite', 
@@ -54,15 +66,15 @@
                          'E', '', 
                          'F', ''), 
               label_size = 10) %>% 
-    as_figure('figure_s2_hact_kinetic', 
+    as_figure('figure_s3_hact_kinetic', 
               ref_name = 'hact_recovery', 
               caption = 'Kinetic of recovery from leading acute COVID-19 symptoms in the survey study.', 
               w = 180, 
               h = 180)
   
-# Figure S3: Frequency of symptoms, CovILD ------
+# Figure S4: Frequency of symptoms, CovILD ------
   
-  insert_msg('Figure S3: frequency of symptoms, CovILD')
[…]

### 2023-04-03 paper scripts/tables.R

@@ -138,7 +138,7 @@
         variable, all_of(names(globals$clust_colors)), 
         significance, eff_size) %>% 
     map(set_names, c('Variable', 
-                     names(globals$clust_colors), 
+                     paste('Cluster', names(globals$clust_colors)), 
                      'Significance', 'Effect size'))
   
   suppl_tbl[c('baseline_clusters_north', 
@@ -183,7 +183,7 @@
         variable, all_of(names(globals$clust_colors)), 
         significance, eff_size) %>% 
     map(set_names, c('Variable', 
-                     names(globals$clust_colors), 
+                     paste('Cluster', names(globals$clust_colors)), 
                      'Significance', 'Effect size'))
   
   suppl_tbl[c('cov_clusters_north', 

### 2023-04-03 tools/project_globals.R

@@ -135,9 +135,9 @@
   
 # Clusters -----
   
-  globals$clust_colors <- c('STDR' = 'darkorange3', 
-                            'RR' = 'darkolivegreen3', 
-                            'SR' = 'cadetblue3')
+  globals$clust_colors <- c('#1' = 'darkorange3', 
+                            '#2' = 'darkolivegreen3', 
+                            '#3' = 'cadetblue3')
   
 # demographic and clinical variables,  HACS -----
   

### 2023-04-03 tools/project_tools.R

@@ -132,10 +132,10 @@
     }
     
     data %>% 
-      mutate(LHS_lab = stri_replace(LHS_lab, fixed = ',', replacement = ', '), 
-             RHS_lab = stri_replace(RHS_lab, fixed = ',', replacement = ', '), 
+      mutate(LHS_lab = stri_replace_all(LHS_lab, fixed = ',', replacement = ', '), 
+             RHS_lab = stri_replace_all(RHS_lab, fixed = ',', replacement = ', '), 
              trans_lab = paste(LHS_lab, RHS_lab, sep = ' \u2192 '), 
-             signature = paste(LHS, RHS, sep = ','), 
+             signature = paste(LHS, RHS, sep = ', '), 
              signature = stri_replace_all(signature, regex = '\\{|\\}', replacement = ''), 
              sign_lab = paste(LHS_lab, RHS_lab, sep = ', '), 
              sign_lab = stri_replace_all(sign_lab, regex = '\\{|\\}', replacement = '')) %>% 
@@ -190,7 +190,7 @@
                         point.padding = 0.1, 
                         min.segment.length = 0.2) + 
         scale_alpha_manual(values = c(high = alpha_limits[1], low = alpha_limits[2])) + 
-        guides(alpha = FALSE)
+        guides(alpha = 'none')
       
     } else {
       
@@ -205,7 +205,7 @@
                         point.padding = 0.1, 
                         min.segment.length = 0.2) + 
         scale_alpha_manual(values = c(high = alpha_limits[1], low = alpha_limits[2])) + 
-        guides(alpha = FALSE)
+        guides(alpha = 'none')
       
     }
     

### 2023-04-03 paper/markdown/figures_and_tables.Rmd
Format of figures and references im Text

@@
 bibliography: hyposmia_biblio.bib
-csl: frontiers_medical.csl
+csl: european-archives-of-oto-rhino-laryngology.csl
 header-includes:
@@
 
-```{r fig-consort, fig.width = mm_inch(figures$consort$w), fig.height = mm_inch(figures$consort$h), fig.cap = 'Flow diagram of the analysis inclusion process for the longitudinal CovILD cohort and the Health after COVID-19 survey study.'}
-
-figures$consort$plot
-
-```
-
-__Figure \@ref(fig:fig-consort). Flow diagram of the analysis inclusion process for the longitudinal CovILD cohort and the Health after COVID-19 survey study.__
-
 \newpage
@@
 
-_(B) Percentages of individuals with smell and taste disorders in the AT (Austria) and IT (Italy) survey study cohorts at particular time points after clinical onset.
+_(B) Percentages of individuals with self-reported olfactory dysfunction and taste disorders in the AT (Austria) and IT (Italy) survey study cohorts at particular time points after clinical onset.
 Numbers of complete observations are indicated under the plots._
+
+_OD: self-reported olfactory dysfunction, Imp. concentration: impaired concentration, Dim. appetite: diminished appetite, Imp. walk: impaired walk, Imp. FMS: impaired fine-motor skills._
 
@@
 
-```{r fig-sympt-mds, fig.width = mm_inch(figures$sympt_mds$w), fig.height = mm_inch(figures$sympt_mds$h), fig.cap = 'Self-reported smell and taste disorders are isolated persistent symptoms of COVID-19.'}
+```{r fig-sympt-mds, fig.width = mm_inch(figures$sympt_mds$w), fig.height = mm_inch(figures$sympt_mds$h), fig.cap = 'Self-reported olfactory dysfunction and taste disorders are isolated persistent symptoms of COVID-19.'}
 
@@
 
-__Figure \@ref(fig:fig-sympt-mds). Self-reported smell and taste disorders are isolated persistent symptoms of COVID-19.__
+__Figure \@ref(fig:fig-sympt-mds). Self-reported olfactory dysfunction and taste disorders are isolated persistent symptoms of COVID-19.__
 
@@
 
+_OD: self-reported olfactory dysfunction, Imp. concentration: impaired concentration._
+
 \newpage
 
-```{r fig-apriori, fig.width = mm_inch(figures$apriori$w), fig.height = mm_inch(figures$apriori$h), fig.cap = 'Co-occurrence of smell and taste disorders in post-acute COVID-19 sequelae.'}
+```{r fig-apriori, fig.width = mm_inch(figures$apriori$w), fig.height = mm_inch(figures$apriori$h), fig.cap = 'Co-occurrence of self-reported olfactory dysfunction and other symptoms in post-acute COVID-19 sequelae.'}
 
@@
 
-__Figure \@ref(fig:fig-apriori). Co-occurrence of smell and taste disorders in post-acute COVID-19 sequelae.__
+__Figure \@ref(fig:fig-apriori). Co-occurrence of self-reported olfactory dysfunction and other symptoms in post-acute COVID-19 sequelae.__
 
-_Frequent combinations of symptoms at $\geq$ 28 days (A) and $\geq$ 3 months (B) after clinical onset in the Austria (AT) and Italy (IT) survey study cohorts were identified with the apriori algorithm.
-Symptom combination support and confidence are presented in the plots.
-Point color codes for lift statistic values.
-The hypo/anosmia and hypo/ageusia combinations are labeled._
+_Frequent combinations (>15% of cohort) of self-reported olfactory dysfunction (OD) and other symptoms at $\geq$ 28 days (A) and $\geq$ 3 months (B) after clinical onset in the Austria (AT) and Italy (IT) survey study cohorts were identified with the apriori algorithm.
+Symptom combination frequency (support statistic) and co-occurrence (support statistic) are presented in bubble plots.
+Point size and color corresponds to co-occurrence, points are labeled with percentages of co-occurrence.
+Imp.
+concentration: impaired concentration._
 
@@
 
-```{r fig-sympt-clusters, fig.width = mm_inch(figures$sympt_clusters$w), fig.height = mm_inch(figures$sympt_clusters$h), fig.cap = 'Differing duration of neurocognitive and respiratory symptoms, fatigue, smell and taste disorders defines the COVID-19 recovery clusters.'}
+```{r fig-sympt-clusters, fig.width = mm_inch(figures$sympt_clusters$w), fig.height = mm_inch(figures$sympt_clusters$h), fig.cap = 'Differing duration of neurocognitive and respiratory symptoms, fatigue, olfactory dysfunction and taste disorders defines the COVID-19 recovery clusters.'}
 
@@
 
-__Figure \@ref(fig:fig-sympt-clusters). Differing duration of neurocognitive and respiratory symptoms, fatigue, smell and taste disorders defines the COVID-19 recovery clusters.__
+__Figure \@ref(fig:fig-sympt-clusters). Differing duration of neurocognitive and respiratory symptoms, fatigue, olfactory dysfunction and taste disorders defines the COVID-19 recovery clusters.__
 
-_The smell and taste disorder recovery (STDR), rapid recovery (RR) and slow recovery (SR) clusters of the survey study participants were identified by semi supervised PAM clustering (partitioning around medoids, Euclidean distance, training cohort: Austria [AT], test cohort: Italy [IT]).
+_Clustering of the survey study participants in respect to symptom-specific recovery times was done by semi-supervised PAM algorithm (partitioning around medoids, Euclidean distance, training cohort: Austria [AT], test cohort: Italy [IT]).
 Mean recovery times in the recovery clusters are presented as lines, 2 $\times$ SEM intervals are displayed as tinted regions.
@@
 
-_Dim. appetite: diminished appetite, Imp. concentration: impaired concentration, Imp. walk: impaired walk, Imp. FMS: impaired fine motor skills._
+_OD: self-reported olfactory dysfunction, Dim. appetite: diminished appetite, Imp. concentration: impaired concentration, Imp. walk: impaired walk, Imp. FMS: impaired fine motor skills._
 
@@
 
-_Differences in minimum/maximum scaled readouts of clinical and physical recovery, mental health and quality of life at the time of survey completion in the recovery clusters (STDR: smell and taste disorder recovery, RR: rapid recovery, SR: slow recovery) in the Austria (AT) and Italy (IT) survey study cohorts.
+_Clustering of the survey study participants in respect to symptom-specific recovery times was done by the semi-supervised PAM algorithm (partitioning around medoids, Euclidean distance, training cohort: Austria [AT], test cohort: Italy [IT]).
+Minimum/maximum scaled readouts of clinical and physical recovery, mental health and quality of life at the time of survey completion in the clusters in the Austria (AT) and Italy (IT) survey study cohorts are presented.
 Dichotomous items (incomplete convalescence, weight loss, new medication and need for rehabilitation) were binarized (yes: 1, no: 0) prior to visualization.

### 2023-04-03 paper/markdown/supplementary_material.Rmd
Format of figures and references im Text

@@
 subtitle: "Supplementary Material"
-author: "Hralth after COVID-19 in Tyrol and CovILD study teams"
+author: "Health after COVID-19 in Tyrol and CovILD study teams"
 date: "`r format(Sys.time(), '%Y-%m-%d')`"
@@
 
-The complete list of study variables and stratification scheme is provided in __Supplementary Table S\@ref(tab:tab-hact-vars)__ for the Health after COVID-19 in Tyrol survey study and in __Supplementary Table S\@ref(tab:tab-covild-vars)__ for the CovILD cohort.
+The complete list of study variables and stratification scheme is provided in __Supplementary Table S\@ref(tab:tab-hact-vars)__ for the Health after COVID-19 in Tyrol survey study and in __Supplementary Table S\@ref(tab:tab-covild-vars)__ for the observatory CovILD cohort.
 
@@
 
-A total of `r length(globals$hact_symptoms)` was surveyed in the survey study cohorts (__Supplementary Figure S\@ref(fig:fig-hact-freq)__ and __Supplementary Table S\@ref(tab:tab-hact-vars)__).
-The symptom duration was coded as follows: absent: 0 days, 1 - 3 days: 3 days, up to 1 week: 7 days, up to 2 weeks: 14 days, up to 4 weeks: 28 days, up to 3 months: 30 moths, up to 6 months: 3 months, over 6 months: 3 months.
+A total of self-reported `r length(globals$hact_symptoms)` symptoms were recorded in the survey study cohorts (__Supplementary Figure S\@ref(fig:fig-hact-freq)__ and __Supplementary Table S\@ref(tab:tab-hact-vars)__).
+The symptom duration was coded as follows: absent: 0 days, 1 - 3 days: 3 days, up to 1 week: 7 days, up to 2 weeks: 14 days, up to 4 weeks: 28 days, up to or greater than 3 months: 3 months.
 Acute symptoms were defined as complaints present during the first 14 days after clinical onset of COVID-19.
 
-In the CovILD study, a total of `r length(globals$covild_symptoms)` symptoms (reduced physical performance, hyposmia/anosmia, dyspnea, sleep problems, cough, fever, night sweating, gastrointestinal symptoms) were prospectively recorded with a standardized questionnaire at each of 60-, 100-, 180- and 360-day post COVID-19 follow-up (__Supplementary Figure S\@ref(fig:fig-covild-freq)__ and __Supplementary Table S\@ref(tab:tab-covild-vars)__).
-Acute COVID-19 symptoms were assessed retrospectively [@Sonnweber2020; @Sonnweber2022].
+In the observatory CovILD cohort, a total of self-reported `r length(globals$covild_symptoms)` symptoms (reduced physical performance, olfactory dysfunction, dyspnea, sleep problems, cough, fever, night sweating, gastrointestinal symptoms) were recorded with a standardized questionnaire at each of 60-, 100-, 180- and 360-day post COVID-19 follow-up (__Supplementary Figure S\@ref(fig:fig-covild-freq)__ and __Supplementary Table S\@ref(tab:tab-covild-vars)__).
+Acute COVID-19 symptoms were assessed retrospectively at the 60-day follow-up [@Sonnweber2020; @Sonnweber2022].
 
@@
 
-### Rating of hyposmia with sniffing stick test
+### Rating of olfactory dysfunction with sniffing stick test
 
-Objective hyposmia at the 100-day and 360-day follow-up in the CovILD study participants was investigated with the 16-item sniffing stick test as described [@Rass2021].
-Clinically relevant hyposmia was defined as $\leq$ 12 correct answers [@Rass2021; @Hummel2007].
-In the analysis, participants with the complete answers concerning self-reported hyposmia and complete test results were included.
+Objective olfactory dysfunction at the 100-day and 360-day follow-up in the CovILD study participants was investigated with the 16-item sniffing stick test as described [@Rass2021].
+Clinically relevant olfactory dysfunction was defined as $\leq$ 12 correct answers [@Rass2021; @Hummel2007].
+In the analysis, participants with the complete answers concerning self-reported olfactory dysfunction and complete test results were included.
 
@@
 
-Data transformation and statistical analysis was done with R version 4.0.5 with _tidyverse_ data science environment [@Wickham2019].
+Data transformation and statistical analysis was done with R version 4.2.0 with _tidyverse_ data science environment [@Wickham2019].
 Analysis results were visualized with _ggplot2_ [@Wickham2016], _cowplot_ [@Wilke2019] as well as in-house developed _ExDA_ (https://github.com/PiotrTymoszuk/ExDA) and _figur_ (https://github.com/PiotrTymoszuk/figur) packages.
@@
 
-Since multiple study variables were non-normally distributed as assessed by Shapiro-Wilk test and visual assessment of their distribution (quantile-quantile plots), statistical significance for differences in outcome numeric variables were assessed with Mann-Whitney U test by Wilcoxon r effect size statistic (two groups) or Kruskal-Wallis test with $\eta^2$ effect size statistic.
+Since multiple study variables were non-normally distributed as assessed by Shapiro-Wilk test and visual assessment of their distribution (quantile-quantile plots), statistical significance for differences in outcome numeric variables were assessed with Mann-Whitney U test with r effect size statistic (two groups) or Kruskal-Wallis test with $\eta^2$ effect size statistic.
 Differences in frequency distribution for categorical outcome variables were assessed by $\chi^2$ test with Cramer V effect size statistic.
-Interrater assessment of self-reported and sniffing test hyposmia was accomplished with Cohen's $\kappa$ statistic [@Fleiss1969].
+Inter-rater assessment of self-reported and sniffing test olfactory dysfunction was accomplished with Cohen's $\kappa$ statistic [@Fleiss1969].
 $\kappa$ significance ($\kappa\neq0$) was estimated with Wald Z test.
@@
 
-## Symptom-symptom distances and multi-dimensional scaling
+### Symptom-symptom distances and multi-dimensional scaling
 
@@
 
-Frequent combinations of symptoms during the first 14 days, at $\geq$ 28 days and at $\geq$ 3 months after clinical onset in the survey study cohorts were identified with the apriori algorithm (package _arules_) [@Agrawal1993; @Hahsler2005] with the minimal support cutoff of 0.1, 2 - 10 item transaction length, confidence > 0.8 and lift > 2.
+Frequent combinations of symptoms during the first 14 days, at $\geq$ 28 days and at $\geq$ 3 months after clinical onset in the survey study cohorts were identified with the apriori algorithm (package _arules_) [@Agrawal1993; @Hahsler2005] with the minimal support cutoff of 0.15, 2 - 10 item transaction length, confidence > 0.6 and lift > 2.
 The support statistic were used to estimate the symptom combination frequency.
@@
 
-```{r tab-clust-reco-it, tab.cap = 'COVID-19 course and recovery in the survey study participants assigned to the recovery clusters, Italy (AT) cohort.'}
+```{r tab-clust-reco-it, tab.cap = 'COVID-19 course and recovery in the survey study participants assigned to the recovery clusters, Italy (IT) cohort.'}
 
@@
 
+```{r fig-consort, fig.width = mm_inch(suppl$consort$w), fig.height = mm_inch(suppl$consort$h), fig.cap = 'Flow diagram of the analysis inclusion process for the longitudinal CovILD cohort and the Health after COVID-19 survey study.'}
+
+suppl$consort$plot
+
+```
+
+__Supplementary Figure S\@ref(fig:fig-consort). Flow diagram of the analysis inclusion process for the longitudinal CovILD cohort and the Health after COVID-19 survey study.__
+
+\newpage
+
 ```{r fig-hact-freq, fig.width = mm_inch(suppl$hact_freq$w), fig.height = mm_inch(suppl$hact_freq$h), fig.cap = 'Frequency of COVID-19 symptoms in the survey study.'}
@@
 
-```{r fig-covild-recovery, fig.width = mm_inch(suppl$covild_recovery$w), fig.height = mm_inch(suppl$covild_recovery$h), fig.cap = 'Kinetic of recovery from smell disorders, reduced performance and dyspnea in ambulatory, moderate and severe COVID-19 subsets of the CovILD study.'}
+```{r fig-covild-recovery, fig.width = mm_inch(suppl$covild_recovery$w), fig.height = mm_inch(suppl$covild_recovery$h), fig.cap = 'Kinetic of recovery from olfactory dysfunction, reduced performance and dyspnea in ambulatory, moderate and severe COVID-19 subsets of the CovILD study.'}
 
@@
 __Supplementary Figure S\@ref(fig:fig-covild-recovery).
-Kinetic of recovery from smell disorders, reduced performance and dyspnea in ambulatory, moderate and severe COVID-19 subsets of the CovILD study.__
+Kinetic of recovery from olfactory dysfunction, reduced performance and dyspnea in ambulatory, moderate and severe COVID-19 subsets of the CovILD study.__
 
-_Percentages of individuals with the complete longitudinal data set suffering from smell disorders (A), reduced physical performance (B) and dyspnea (C) in the ambulatory, moderate and severe COVID-19 subsets during acute COVID-19 and at the 60-, 100-, 180- and 360-day follow-ups.
+_Percentages of individuals with the complete longitudinal data set suffering from olfactory dysfunction (OD) (A), reduced physical performance (B) and dyspnea (C) in the ambulatory, moderate and severe COVID-19 subsets during acute COVID-19 and at the 60-, 100-, 180- and 360-day follow-ups.
 Numbers of complete observations are indicated under the plots._
@@
 
-```{r fig-hyposmia-rater100, fig.width = mm_inch(suppl$hyposmia_rater100$w), fig.height = mm_inch(suppl$hyposmia_rater100$h), fig.cap = 'Rates of self-reported hyposmia and hyposmia in sniffing stick test at 3-month post COVID-19 follow-up in the ambulatory, moderate and severe COVID-19 subsets of the CovILD study.'}
+```{r fig-hyposmia-rater100, fig.width = mm_inch(suppl$hyposmia_rater100$w), fig.height = mm_inch(suppl$hyposmia_rater100$h), fig.cap = 'Rates of self-reported olfactory dysfunction and olfactory dysfunction in the sniffing stick test at 3-month post COVID-19 follow-up in the ambulatory, moderate and severe COVID-19 subsets of the CovILD study.'}
 
@@
 __Supplementary Figure S\@ref(fig:fig-hyposmia-rater100).
-Rates of self-reported hyposmia and hyposmia in sniffing stick test at 3-month post COVID-19 follow-up in the ambulatory, moderate and severe COVID-19 subsets of the CovILD study.__
+Rates of self-reported olfactory dysfunction and olfactory dysfunction in the sniffing stick test at 3-month post COVID-19 follow-up in the ambulatory, moderate and severe COVID-19 subsets of the CovILD study.__
 
-_Association of self-reported and sniffing stick test hyposmia rates was investigated with Cohen's $\kappa$ statistic.
+_Association of self-reported and sniffing stick test rates of olfactory dysfunction (OD) was investigated with Cohen's $\kappa$ statistic.
 Statistical significance was assessed with Wald's Z test corrected for multiple testing with Benjamini-Hochberg method.
-Percentages of individuals with self-reported and test hyposmia within the entire cohort (A), the ambulatory (B), moderate (C) and severe (D) COVID-19 subsets are presented in bubble plots.
+Percentages of individuals with self-reported and test OD within the entire cohort (A), the ambulatory (B), moderate (C) and severe (D) COVID-19 subsets are presented in bubble plots.
 Point size and color represents the percentage.
@@
 
-```{r fig-hyposmia-rater360, fig.width = mm_inch(suppl$hyposmia_rater360$w), fig.height = mm_inch(suppl$hyposmia_rater360$h), fig.cap = 'Rates of self-reported hyposmia and hyposmia in sniffing stick test at 1-year post COVID-19 follow-up in the ambulatory, moderate and severe COVID-19 subsets of the CovILD study.'}
+```{r fig-hyposmia-rater360, fig.width = mm_inch(suppl$hyposmia_rater360$w), fig.height = mm_inch(suppl$hyposmia_rater360$h), fig.cap = 'Rates of self-reported olfactory dysfunction and olfactory dysfunction in the sniffing stick test at 1-year post COVID-19 follow-up in the ambulatory, moderate and severe COVID-19 subsets of the CovILD study.'}
 
@@
 __Supplementary Figure S\@ref(fig:fig-hyposmia-rater360).
-Rates of self-reported hyposmia and hyposmia in sniffing stick test at 1-year post COVID-19 follow-up in the ambulatory, moderate and severe COVID-19 subsets of the CovILD study.__
+Rates of self-reported olfactory dysfunction and olfactory dysfunction in the sniffing stick test at 1-year post COVID-19 follow-up in the ambulatory, moderate and severe COVID-19 subsets of the CovILD study.__
 
-_Association of self-reported and sniffing stick test hyposmia rates was investigated with Cohen's $\kappa$ statistic.
+_Association of self-reported and sniffing stick test rates of olfactory dysfunction (OD) was investigated with Cohen's $\kappa$ statistic.
 Statistical significance was assessed with Wald's Z test corrected for multiple testing with Benjamini-Hochberg method.
-Percentages of individuals with self-reported and test hyposmia within the entire cohort (A), the ambulatory (B), moderate (C) and severe (D) COVID-19 subsets are presented in bubble plots.
+Percentages of individuals with self-reported and test OD within the entire cohort (A), the ambulatory (B), moderate (C) and severe (D) COVID-19 subsets are presented in bubble plots.
 Point size and color represents the percentage.
@@
 Percentages of the data set variance associated with the MDS dimensions are indicated in the plot axes.
-Numbers of complete observations are indicated under the plots._
+Numbers of complete observations are indicated under the plots.
+OD: self-reported olfactory dysfunction._
 
@@
 
-_(C) Permutation importance of the clustering features (symptoms) for clustering of the training cohort expressed as the difference in clustering variance (ratio of total between-cluster sum of squares to total sum of squares) between the initial clustering object and the clustering object with the given variable reshuffled at random._
+_(C) Permutation importance of the clustering features (symptoms) for clustering of the training cohort expressed as the difference in clustering variance (ratio of total between-cluster sum of squares to total sum of squares) between the initial clustering object and the clustering object with the given variable reshuffled at random.
+Imp.
+concentration: impaired concentration, OD: self-reported olfactory dysfunction, Dim.
+appetite: diminished appetite, Imp.
+walk: impaired walk, Imp.
+FMS: impaired fine motor skills._
 
@@
 
-```{r fig-cov-clusters, fig.width = mm_inch(suppl$cov_clusters$w), fig.height = mm_inch(suppl$cov_clusters$h), fig.cap = 'Numbers of COVID-19 symptoms in the survey study recovery clusters.'}
+```{r fig-cov-clusters, fig.width = mm_inch(suppl$cov_clusters$w), fig.height = mm_inch(suppl$cov_clusters$h), fig.cap = 'Numbers of COVID-19 symptoms in the survey study clusters.'}
 
@@
 
-__Supplementary Figure S\@ref(fig:fig-cov-clusters). Numbers of COVID-19 symptoms in the survey study recovery clusters.__
+__Supplementary Figure S\@ref(fig:fig-cov-clusters). Numbers of COVID-19 symptoms in the survey study clusters.__
 
-_The smell and taste disorder recovery (STDR), rapid recovery (RR) and slow recovery (SR) clusters of the survey study participants were identified by semi supervised PAM clustering (partitioning around medoids, Euclidean distance, training cohort: Austria [AT], test cohort: Italy [IT]).
+_Clustering of the survey study participants in respect to symptom-specific recovery times was done by the semi-supervised PAM algorithm (partitioning around medoids, Euclidean distance, training cohort: Austria [AT], test cohort: Italy [IT]).
 Differences in numbers of symptoms in the first 14 days (A) and at $\geq$ 28 days (B) after clinical onset between the clusters were assessed by Kruskal-Wallis test and $\eta^2$ effect size statistic.
@@
 
-__Figure \@ref(fig:fig-base-clusters). COVID-19 recovery clusters differ in sex distribution, comorbidity and daily medication rates.__
+__Supplementary Figure S\@ref(fig:fig-base-clusters). COVID-19 recovery clusters differ in sex distribution, comorbidity and daily medication rates.__
 
-_The smell and taste disorder recovery (STDR), rapid recovery (RR) and slow recovery (SR) clusters of the survey study participants were identified by semi supervised PAM clustering (partitioning around medoids, Euclidean distance, training cohort: Austria [AT], test cohort: Italy [IT]).
+_Clustering of the survey study participants in respect to symptom-specific recovery times was done by the semi-supervised PAM algorithm (partitioning around medoids, Euclidean distance, training cohort: Austria [AT], test cohort: Italy [IT]).
 Differences in sex distribution (A), frequency of comorbidity (B) and daily medication (C) between the recovery clusters were assessed by $\chi^2$ test with Cramer V effect size statistic.

### 2023-05-30 analysis scripts/apriori.R
Revision analyses
Analyses requested by the Reviewer. Rationale for the two-cohort analysis approach of the survey study. Clustering tendency. ROC analysis of subjective OD as a readout of objective hyposmia. Inverse distance-weighted kNN as a semi-supervised clustering approach. Multiple runs of permutation importance of the clustering factors. 'Kinetics' of objective OD in the CovILD study

@@ -7,6 +7,12 @@
   
   ap_sympt <- list()
   
+# Parallel backend ------
+  
+  insert_msg('Parallel backend')
+  
+  plan('multisession')
+  
 # globals ------
   
   insert_msg('Globals: analysis tables')
@@ -15,7 +21,8 @@
   
   ap_sympt$hact_tbl <- mod_tbl[c('north', 'south')] %>% 
     map(filter, time %in% c(3, 14, 28, 90)) %>% 
-    map(dlply, 'time', select, -ID, -time) %>% 
+    map(blast, time) %>% 
+    map(map, select, -ID, -time) %>% 
     unlist(recursive = FALSE) %>% 
     map(~filter(.x, complete.cases(.x))) %>% 
     map(as.matrix)
@@ -24,10 +31,6 @@
   
   ap_sympt$hact_trasactions <- ap_sympt$hact_tbl %>% 
     map(as, 'transactions')
-  
-  ## parallel backend
-  
-  plan('multisession')
   
 # Calculating the rules -----
   
@@ -56,17 +59,18 @@
   
   insert_msg('Confidence versus support plots')
   
-  ap_sympt$conf_supp_plots <- list(data = ap_sympt$rules_tbl %>% 
-                                     map(mutate, 
-                                         trans_lab = stri_replace(trans_lab, 
-                                                                  fixed = ' \u2192 ', 
-                                                                  replacement = '\u2192\n')), 
-                                   plot_title = rep(c('0 - 14 days', 
-                                                      '14 days', 
-                                                      '28 days', 
-                                                      '3 months'), 2), 
-                                   plot_subtitle = c(rep('AT, survey study', 4), 
-                                                     rep('IT, survey study', 4))) %>% 
+  ap_sympt$conf_supp_plots <- 
+    list(data = ap_sympt$rules_tbl %>% 
+           map(mutate, 
+               trans_lab = stri_replace(trans_lab, 
+                                        fixed = ' \u2192 ', 
+                                        replacement = '\u2192\n')), 
+         plot_title = rep(c('0 - 14 days', 
+                            '14 days', 
+                            '28 days', 
+                            '3 months'), 2), 
+         plot_subtitle = c(rep('AT, survey study', 4), 
+                           rep('IT, survey study', 4))) %>% 
     pmap(draw_conf_supp, 
          by = 'support',
          top_transactions = 7, 
@@ -79,16 +83,17 @@
   
[…]

### 2023-05-30 analysis scripts/clust_characteristic.R

@@ -6,6 +6,12 @@
   
   clust_chara <- list()
   
+# parallel backend ------
+  
+  insert_msg('Parallel backend')
+  
+  plan('multisession')
+  
 # globals ------
   
   insert_msg('Globals setup')
@@ -23,9 +29,9 @@
                     type_plot = ifelse(is.numeric(hact$north[[.x]]), 
                                        'violin', 'stack'), 
                     y_lab = ifelse(is.numeric(hact$north[[.x]]), 
-                                   translate_var(.x, 
-                                                 out_value = 'axis_lab', 
-                                                 dict = hact$dict), 
+                                   exchange(.x, 
+                                            value = 'axis_lab', 
+                                            dict = hact$dict), 
                                    '% of cluster')))
   
   clust_chara$num_vars <- clust_chara$var_types %>% 
@@ -49,11 +55,7 @@
                             'phq_anxiety_score', 
                             'phq_depression_score', 
                             'stress_score')
-  
-  ## parallel backend
-  
-  plan('multisession')
-  
+
 # descriptive stats -------
   
   insert_msg('Descriptive stats')
@@ -119,23 +121,24 @@
                                    'psych_var')] %>% 
           map(~filter(cohort, variable %in% .x))) %>% 
     unlist(recursive = FALSE) %>% 
-    map(mutate, variable = translate_var(variable, dict = hact$dict))
+    map(mutate, variable = exchange(variable, dict = hact$dict))
   
-  clust_chara$summ_plots <- list(x = clust_chara$summ_plots, 
-                                 plot_title = rep(c('Baseline characteristic', 
-                                                    'Medical history', 
-                                                    'COVID-19 course and recovery', 
-                                                    'Psychosocial rating'), 2), 
-                                 plot_subtitle = c(rep('AT, survey study', 4), 
-                                                   rep('IT, survey study', 4)), 
-                                 point_color = list(c('gray60', globals$hact_colors[1]), 
-                                                    c('gray60', globals$hact_colors[1]), 
-                                                    c('gray60', globals$hact_colors[1]), 
-                                                    c('gray60', globals$hact_colors[1]), 
-                                                    c('gray60', globals$hact_colors[2]), 
[…]

### 2023-05-30 analysis scripts/hyposmia_rater.R

@@ -1,4 +1,4 @@
-# Snbiffing stick versus self-reported hyposmia
+# Sniffing stick versus self-reported hyposmia
 
   insert_head()
   
@@ -10,63 +10,99 @@
   
   insert_msg('Globals setup')
   
+  ## sniffing stick score and subjective hyposmia
+  ## separate datasets for the cohort and severity strata
+  
   rater$analysis_tbl <- list(fup100 = sst$data_100_fup, 
                              fup360 = sst$data_360_fup) %>% 
-    map(~dlply(.x, 'cat_WHO', as_tibble)) %>% 
+    map(blast, cat_WHO) %>% 
     map2(., sst[c('data_100_fup', 'data_360_fup')], 
          ~c(.x, list(cohort = .y))) %>% 
     unlist(recursive = FALSE)
   
+  rater$strata_labs <- 
+    c('Ambulatory, CovILD study', 
+      'Moderate, CovILD study', 
+      'Severe, CovILD study', 
+      'CovILD cohort')
+
 # calculation of the kappas -----
   
   insert_msg('Calculation of Kohen kappas')
   
   rater$test_results <- rater$analysis_tbl %>% 
-    map(~correlate_variables(.x, 
-                             variables = c('anosmia_sympt', 'sniff_hyposmia'), 
-                             what = 'correlation', 
-                             type = 'kappa', 
-                             ci = TRUE, 
-                             pub_styled = TRUE)) %>% 
-    map2_dfr(., names(.), ~mutate(.x, 
-                                  time = stri_split_fixed(.y, pattern = '.', simplify = TRUE)[, 1], 
-                                  subset = stri_split_fixed(.y, pattern = '.', simplify = TRUE)[, 2])) %>% 
+    map(correlate_variables, 
+        variables = c('anosmia_sympt', 'sniff_hyposmia'), 
+        what = 'correlation', 
+        type = 'kappa', 
+        ci = TRUE, 
+        pub_styled = FALSE) %>% 
+    compress(names_to = 'comparison') %>% 
     re_adjust %>% 
-    mutate(plot_cap = paste(eff_size, significance, sep = ', '))
+    mutate(estimate = ifelse(estimate < 0, 0, estimate), 
+           upper_ci = ifelse(upper_ci < 0, 0, upper_ci), 
+           lower_ci = ifelse(lower_ci < 0, 0, lower_ci), 
+           time = stri_split_fixed(comparison, 
+                                   pattern = '.', 
[…]

### 2023-05-30 analysis scripts/kinetic_hyposmia_long.R

@@ -22,7 +22,7 @@
   
   hyp_long$analysis_tbl <-  hyp_long$analysis_tbl %>% 
     map(function(cohort) cohort %>% 
-          dlply(.(ID), as_tibble) %>% 
+          blast(ID) %>% 
           map(select, -ID) %>% 
           map(~map_dfc(.x, function(sympt) switch(as.character(sympt), 
                                                   absent = c(0, 0, 0, 0, 0, 0), 
@@ -33,7 +33,8 @@
                                                   `up to 3 months` = c(1, 1, 1, 1, 1, 0), 
                                                   `up to 6 months` = c(1, 1, 1, 1, 1, 1), 
                                                   `over 6 months` = c(1, 1, 1, 1, 1, 1)))) %>% 
-          map2_dfr(., names(.), ~mutate(.x, ID = .y, time = c(3, 7, 14, 28, 90, 180)))) %>% 
+          map2_dfr(., names(.), 
+                   ~mutate(.x, ID = .y, time = c(3, 7, 14, 28, 90, 180)))) %>% 
     map(as_tibble) %>% 
     map(complete_cases)
 
@@ -88,8 +89,8 @@
                          sub = c('AT, survey study', 'IT, survey study'), 
                          col = globals$hact_colors) %>% 
     pmap(function(mod, sub, col) list(x = mod, 
-                                      plot_title = translate_var(names(mod), 
-                                                                 dict = hact$dict)) %>% 
+                                      plot_title = exchange(names(mod), 
+                                                            dict = hact$dict)) %>% 
            pmap(plot, 
                 type = 'frequency', 
                 plot_subtitle = sub, 
@@ -103,6 +104,11 @@
                  scale_y_continuous(limits = c(0, 100)) + 
                  scale_x_continuous(limits = c(0, 185), 
                                     breaks = c(0, 14, 28, 90, 180))))
+  
+  ## moving n number to the captions
+  
+  hyp_long$plots <- hyp_long$plots %>% 
+    map(~map(.x, tag_to_sub))
   
 # END -----
   

### 2023-05-30 analysis scripts/recovery_distribution.R

@@ -6,14 +6,20 @@
 
   rec_dist <- list()
   
+# Parallel backend -------
+  
+  insert_msg('Parallel backend')
+  
+  plan('multisession')
+  
 # globals -------
   
   insert_msg('Globals setup')
   
   rec_dist$tbl_hact <- rec_time[c('north', 'south')] %>% 
-    map2_dfr(., c('AT', 'IT'), ~mutate(.x, cohort = .y))
-  
-  plan('multisession')
+    compress(names_to = 'cohort') %>% 
+    mutate(cohort = car::recode(cohort, "'north' = 'AT'; 'south' = 'IT'"), 
+           cohort = factor(cohort, c('AT', 'IT')))
   
 # descriptive stats of the recovery: individuals with the symptom present -----  
   
@@ -28,7 +34,7 @@
                             pub_styled = TRUE), 
                    .options = furrr_options(seed = TRUE)) %>% 
     reduce(left_join, by = 'variable') %>% 
-    set_names(c('variable', 'north', 'south'))
+    set_names(c('variable', levels(rec_dist$tbl_hact$cohort)))
   
   rec_dist$stats_covild <- globals$covild_symptoms %>% 
     map_dfr(~explore(rec_time$covild[c('cat_WHO', .x)] %>% 
@@ -38,11 +44,10 @@
                      what = 'table', 
                      pub_styled = TRUE)) %>% 
     reduce(left_join, by = 'variable') %>% 
-    set_names(c('variable', 'A', 'HM', 'HS'))
+    set_names(c('variable', levels(rec_time$covild$cat_WHO)))
 
 # testing, differences between the cohorts or severity strata, symptomatic ----
-  # Mann-Whitney U test or Kruskal-Wallis test
-  
+
   insert_msg('Testing for the differences between the cohorts or severity')
   
   rec_dist$test_hact <- globals$hact_symptoms %>% 
@@ -55,9 +60,7 @@
                                           ci = FALSE, 
                                           pub_styled = TRUE, 
                                           adj_method = 'none'), 
-               .options = furrr_options(seed = TRUE)) %>% 
-    map_dfr(~.x$result) %>% 
-    re_adjust('BH')
+               .options = furrr_options(seed = TRUE))
 
   rec_dist$test_covild <- globals$covild_symptoms %>% 
     map(~safely(compare_variables)(rec_time$covild[c('cat_WHO', .x)] %>% 
@@ -68,14 +71,14 @@
                                    types = 'kruskal_test', 
                                    ci = FALSE, 
[…]

### 2023-05-30 exploration scripts/cohort.R

@@ -14,7 +14,10 @@
   ## analysis table for covild
   
   eda_cohort$hact_tbl <- hact[c('north', 'south')] %>% 
-    reduce(rbind)
+    reduce(rbind) %>% 
+    mutate(cohort = car::recode(cohort, 
+                                "'north' = 'AT'; 'south' = 'IT'"), 
+           cohort = factor(cohort, c('AT', 'IT')))
   
   ## test variables for HACT
   
@@ -24,85 +27,153 @@
                                              'psych_var')] %>% 
     reduce(c)
   
-  eda_cohort$hact_vars$test_type <- eda_cohort$hact_vars$variable %>% 
-    map_chr(~ifelse(is.numeric(hact$north[[.x]]), 'wilcoxon_r', 'cramer_v'))
+  eda_cohort$hact_vars$format <- 
+    hact$north[eda_cohort$hact_vars$variable] %>% 
+    map_lgl(is.numeric) %>% 
+    ifelse('numeric', 'factor')
 
-  eda_cohort$hact_vars <- as_tibble(eda_cohort$hact_vars)
-  
   ## test variables for CovILD
   
   eda_cohort$covild_vars$variable <- globals$covild_vars
   
-  eda_cohort$covild_vars$test_type <- eda_cohort$covild_vars$variable %>% 
-    map_chr(~ifelse(is.numeric(covild$data[[.x]]), 'kruskal_test', 'chisq_test'))
+  eda_cohort$covild_vars$format <- 
+    covild$data[eda_cohort$covild_vars$variable] %>% 
+    map_lgl(is.numeric) %>% 
+    ifelse('numeric', 'factor')
   
-  eda_cohort$covild_vars <- as_tibble(eda_cohort$covild_vars)
+  ## test and plot type
+  
+  eda_cohort[c("hact_vars", "covild_vars")] <- 
+    eda_cohort[c("hact_vars", "covild_vars")] %>% 
+    map(as_tibble) %>% 
+    map2(., list('wilcoxon_r', 'kruskal_etasq'), 
+         ~mutate(.x, 
+                 test_type = ifelse(format == 'numeric', 
+                                    .y, 'cramer_v'), 
+                 plot_type = ifelse(format == 'numeric', 
+                                    'violin', 'stack')))
+  
+  ## lexicons 
+  
+  eda_cohort[c("hact_lexicon", "covild_lexicon")] <- 
+    map2(list(hact$dict, covild$dict), 
[…]

### 2023-05-30 exploration scripts/symptoms.R

@@ -1,4 +1,4 @@
-# Frequency of acute, sub-acute and long and PACS symptoms.
+# Frequency of acute, sub-acute and long and PASC symptoms.
 
   insert_head()
   
@@ -13,33 +13,44 @@
   ## analysis tables, HACT
   
   eda_sympt$hact_tbl <- mod_tbl[c('north', 'south')] %>% 
-    map2_dfr(., names(.), ~mutate(.x, cohort = .y)) %>% 
+    compress(names_to = 'cohort') %>% 
     filter(time %in% c(3, 14, 28, 90)) %>% 
-    dlply('time', select, -ID, -time) %>% 
+    blast(time) %>% 
+    map(select, -ID, -time) %>% 
     set_names(c('acute', 'sub_acute', 'long', 'pasc')) %>% 
     map(as_tibble) %>% 
+    map(map_dfr, car::recode, "0 = 'no'; 1 = 'yes'") %>% 
     map(map_dfr, factor)
 
   ## analysis tables, covILD
   
   eda_sympt$covild_tbl <- mod_tbl$covild %>% 
     complete_cases %>% 
-    dlply('time_numeric', as_tibble)
+    blast(time_numeric)
   
   ## n numbers
   
-  eda_sympt$hact_n_tags <- paste('\nn =', 
-                                 dlply(eda_sympt$hact_tbl$acute, 
-                                       'cohort', 
-                                       nrow) %>% 
-                                   unlist)
+  eda_sympt$hact_n_numbers <- eda_sympt$hact_tbl$acute %>% 
+    blast(cohort) %>% 
+    map_dbl(nrow)
   
-  eda_sympt$covild_n_tags <- paste('\nn =', 
-                                   dlply(eda_sympt$covild_tbl[[1]], 
-                                         'cat_WHO', 
-                                         nrow) %>% 
-                                     unlist)
+  eda_sympt$hact_n_tags <- eda_sympt$hact_n_numbers %>% 
+    paste('n = ', .)
+    
+  eda_sympt$covild_n_numbers <- eda_sympt$covild_tbl[[1]] %>% 
+    blast(cat_WHO) %>% 
+    map_dbl(nrow)
   
+  eda_sympt$covild_n_tags <-  eda_sympt$covild_n_numbers %>% 
+    paste('n = ', .)
+  
+  ## labels for the HACT time points
+  
+  eda_sympt$hact_time_labs <- 
[…]

### 2023-05-30 import.R

@@ -10,6 +10,7 @@
   library(tidyverse)
   library(foreign)
   library(stringi)
+  library(trafo)
   
   insert_head()
   
@@ -98,6 +99,20 @@
                                 fixed = 'Imp. fine motor skills', 
                                 replacement = 'Imp. FMS'))
   
+  ## more informative names for psychometic measures
+  
+  hact$dict[hact$dict$variable == 'life_quality_score', c('label', 'label_short', 'description', 'axis_lab')] <- 
+    tibble(label = 'QoL impairment score', 
+           label_short = 'QoL impairment score', 
+           description = 'Score of impaired quality of life (QoL, 0: excellent QoL; 3: poor QoL)', 
+           axis_lab = 'QoL impairment score')
+  
+  hact$dict[hact$dict$variable == 'mental_health_score', c('label', 'label_short', 'description', 'axis_lab')] <- 
+    tibble(label = 'OMH impairment score', 
+           label_short = 'OMH impairment score', 
+           description = 'Score of impaired self-assessed overall mental health (OMH, 0: excellent QoL; 3: poor QoL)', 
+           axis_lab = 'OMH impairment score')
+  
 # Calculating the times to recovery, HACT -----
   
   insert_msg('Individual times to recovery, HACT')
@@ -124,12 +139,12 @@
   
   rec_time$covild <- covild$data %>% 
     select(ID, time, all_of(globals$covild_symptoms)) %>% 
-    dlply(.(time), as_tibble) %>% 
+    blast(time) %>% 
     map(select, - time) %>% 
     map2_dfr(., c(14, 60, 90, 180, 360), 
              ~map_dfc(.x, 
                       function(x) if(is.factor(x)) ifelse(x == 'yes', .y, 0) else x)) %>% 
-    dlply(.(ID), as_tibble) %>% 
+    blast(ID) %>% 
     map(~map_dfc(.x, 
                  function(x) if(is.numeric(x)) max(x, na.rm = TRUE) else x)) %>% 
     map_dfr(~.x[1, ]) %>% 
@@ -150,7 +165,7 @@
   
   mod_tbl[c('north', 'south')] <- mod_tbl[c('north', 'south')] %>% 
     map(function(cohort) cohort %>% 
-          dlply(.(ID), as_tibble) %>% 
+          blast(ID) %>% 
           map(select, -ID) %>% 
           map(~map_dfc(.x, function(sympt) switch(as.character(sympt), 
                                                   absent = c(0, 0, 0, 0, 0), 

### 2023-05-30 paper scripts/figures.R

@@ -10,16 +10,17 @@
   
   insert_msg('Figure 1: symptom recovery times')
   
+  ## upper panel: median/IQR recovery times
   ## emboldening the hyposmia and hypogeusia
 
   figures$sympt_recovery$upper_panel <- rec_dist$summ_plot_hact %>% 
-    map(~.x + 
-          scale_y_discrete(labels = function(x) embolden_scale(x, c('OD', 'Hypogeusia/ageusia'))) + 
-          theme(axis.text.y = element_markdown())) %>% 
+    map(embolden_elli, c('OD', 'Hypogeusia/ageusia')) %>% 
     plot_grid(plotlist = ., 
               ncol = 2, 
               align = 'hv',
               axis = 'tblr')
+
+  ## bottom panel: recovery curves for hyposmia and hypo/ageusia
   
   figures$sympt_recovery$bottom_panel <- kin_mod$plots_hact %>% 
     map(~.x[c('anosmia', 'taste_loss')]) %>% 
@@ -37,17 +38,44 @@
                                       figures$sympt_recovery$bottom_panel, 
                                       nrow = 2, 
                                       rel_heights = c(0.7, 0.3), 
-                                      labels = c('A', 'B'), 
+                                      labels = c('a', 'b'), 
                                       label_size = 10) %>% 
     as_figure('figure_1_recovery_times', 
               ref_name = 'sympt_recovery', 
-              caption = 'Symptom-specific recovery times in the ambulatory COVID-19 survey study.', 
+              caption = paste('Symptom-specific recovery times in the', 
+                              'ambulatory COVID-19 survey study.'), 
               w = 180, 
               h = 220)
 
-# Figure 2: symptom distances -----
+# Figure 2: subjective and objective hyposmia -------
   
-  insert_msg('Figure 2: symptom isolation')
+  insert_msg('Figure 2: subjective and objective hyposmia')
+  
+  figures$sniff_test <- 
+    plot_grid(rater$hyposmia_rates$plots$`100` + 
+                theme(legend.position = 'bottom'), 
+              rater$kappa_forests$plots$fup100 + 
+                labs(title = ''), 
+              rater$hyposmia_rates$plots$`360` + 
+                theme(legend.position = 'bottom'), 
+              rater$kappa_forests$plots$fup360 + 
+                labs(title = ''), 
+              ncol = 2, 
+              align = 'hv', 
+              axis = 'tblr', 
[…]

### 2023-05-30 paper scripts/manuscript.R

@@ -2,6 +2,12 @@
 
   insert_head()
 
+# Reading the bibliography ------
+  
+  insert_msg('Reading the bibiography')
+  
+  od_bib <- read_bib('./paper/markdown/hyposmia_biblio.bib')
+  
 # Rendering the paper ------
   
   insert_msg('Rendering the paper')
@@ -16,6 +22,15 @@
   insert_msg('Rendering the supplementary material')
   
   render('./paper/markdown/supplementary_material.Rmd', 
+         output_format = word_document2(number_sections = FALSE, 
+                                        reference_docx = 'ms_template.docx'), 
+         output_dir = './paper')
+  
+# Rendering the rebuttal letter -------
+  
+  insert_msg('Rendering the rebuttal letter')
+  
+  render('./paper/markdown/rebuttal_letter.Rmd', 
          output_format = word_document2(number_sections = FALSE, 
                                         reference_docx = 'ms_template.docx'), 
          output_dir = './paper')

### 2023-05-30 paper.R

@@ -27,8 +27,8 @@
   c('./paper scripts/tables.R', 
     './paper scripts/figures.R', 
     './paper scripts/supplements.R', 
-    './paper scripts/manuscript.R'
-    ) %>% 
+    './paper scripts/links.R', 
+    './paper scripts/manuscript.R') %>% 
     source_all(message = TRUE, crash = TRUE)
   
 # END ----

### 2023-05-30 tools/project_tools.R

@@ -1,5 +1,62 @@
 # A medley of functional project tools
 
+# packages ------
+
+  library(plyr)
+  library(tidyverse)
+  library(rlang)
+  library(trafo)
+  library(ggtext)
+  library(clipr)
+
+# Sample size calculation ------
+
+  draw_hopkins <- function(data, 
+                           subset_ids, 
+                           n = floor(length(subset_ids) * 0.5), 
+                           seed = 1234) {
+    
+    ## calculates the hopkins stat for the subset of a dataset
+    #ä# defined by the user-provided rownames
+    
+    hopkins <- get_clust_tendency(data = data[subset_ids, ], 
+                                  n = n, 
+                                  seed = seed)[c('hopkins_stat', 'p_value')]
+    
+    hopkins %>% 
+      as_tibble %>% 
+      mutate(n = nrow(data[subset_ids, ]))
+    
+  }
+  
+# Kinetic plot formatting -----
+  
+  tag_to_sub <- function(plot, replace = FALSE) {
+    
+    if(replace) {
+      
+      plot <- plot + 
+        labs(subtitle = stri_replace(plot$labels$tag, 
+                                     fixed = '\n', 
+                                     replacement = ''))
+      
+    } else {
+      
+      plot <- plot + 
+        labs(subtitle = plot$labels$subtitle %>% 
+               paste(stri_replace(plot$labels$tag, 
+                                  fixed = '\n', 
+                                  replacement = ''), 
+                     sep = ', '))
+      
+    }
+     
+    plot + 
+      theme(plot.tag = element_blank())
+    
+  }
+  
+
 # kNN plots -----
 
   draw_knn_course <- function(data, 
@@ -45,16 +102,18 @@
            aes(x = .data[[time_var]], 
                y = .data[[dist_var]], 
                color = .data[[color_var]])) + 
-      geom_line(aes(size = lsize_var)) + 
+      geom_line(aes(linewidth = lsize_var)) + 
       geom_point(shape = 16, 
                  size = point_size) + 
       geom_text_repel(aes(label = sympt_lab), 
                       size = txt_size) + 
       scale_color_manual(values = col_scale, 
                          name = '') + 
-      scale_size_manual(values = c(highlight_size, line_size)) + 
-      guides(color = FALSE, 
-             size = FALSE) +
+      scale_linewidth_manual(values = c(normal = line_size, 
+                                        high = highlight_size)) + 
[…]

### 2023-05-30 paper/markdown/rebuttal_letter.Rmd
Figures, Tables, Supplementary Material

---
title: "Distinct smell and taste disorder phenotype of post-acute COVID-19 sequelae"
subtitle: "Rebuttal Letter"
author: "Health after COVID-19 in Tyrol and CovILD study teams"
date: "`r format(Sys.time(), '%Y-%m-%d')`"

output: 
  bookdown::word_document2:   
    reference_docx: ms_template.docx

bibliography: hyposmia_biblio.bib
csl: european-archives-of-oto-rhino-laryngology.csl
header-includes:
  \usepackage{longtable}
  \usepackage{tabu}
  \usepackage{caption}
  \usepackage{makecell}
  \usepackage{pdflscape}
  \usepackage{array}
  \usepackage{booktabs}
  \usepackage{threeparttable}
  \usepackage{threeparttablex}
  \usepackage{wrapfig}
  \usepackage{multirow}
  \usepackage[normalem]{ulem}
  \usepackage{colortbl}
  \usepackage{xcolor}
  \usepackage{float} \floatplacement{figure}{H} \floatplacement{table}{H}
  \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}}       \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
---

```{r, setup, include = FALSE}

library(bookdown)

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, 
                      dev = "png", dpi = 600, 
                      tab.cap.style = 'Table Heading')

set_flextable_defaults(font.family = 'Cambria', font.size = 10)

```

\newpage

# Reviewer #1

<span custom-style = 'reviewer'>I had the pleasure of reviewing your manuscript entitled "Distinct smell and taste disorder phenotype of post-acute COVID-19 sequelae". I have several comments in respect to your manuscript:</span>

We thank the Reviewer for the careful lecture of our text and valuable feedback. 
We did our best to address your issues by introducing changes in the text, figures and Supplementary Material. 
The details are listed below (grey: Reviewers issue, black plain: our response, black italic and red: quotation of the altered manuscript text). 
These changes in the main text are marked in red. 
Please mind the new numbering of the main text Figures and Supplementary Figures which were changes to accommodate new analysis results and meet the journal's requirements. 
Additionally, we introduced few changes in the analysis pipeline and text, which, as we believe, make interpretation of our data more straightforward for an interested reader:

* we analyzed kinetic of symptom resolution with Cochran's Q test with Kendall's W effect size statistic instead of second-term modeling and present those statistical testing results in __Supplementary Table S3__ and __Supplementary Table S4__. We use this data to underline the persistent character of self-reported hyposmia in ambulatory COVID-19 and interpret them in __Results/Longitudinal course of COVID-19 symptom resolution__.

* we changed the format of __Tables__ and __Supplementary Tables__ to reduce the number of footnotes.

* to observe the figure number limit of the journal (up to five figures), we moved the figure with apriori analysis results to Supplementary Material

* related to your issue with differences between the AT and IT cohort, we included now additional information on how and where the cohorts were recruited and specified inclusion/exclusion criteria (__Methods/Study design and approval__).

* pertaining to the two-cohort analysis strategy of the survey study, we elaborate semi-supervised clustering procedure, cluster definition in the training AT cohort and quality of cluster assignment in the test IT collective in __Results/Smell and taste disorder phenotype of COVID-19 recovery__.

* we provide a short information on how numeric and categorical variables are presented in the manuscript (__Methods/Statistical analysis__ and __Supplementary Methods__).

* assignment of the test IT cohort participants (survey study) to the recovery clusters defined in the training AT cohort was done with the inverse distance-weighted 7-nearest neighbors classifier. Distance weighting, employed also in our recent papers [@Sonnweber2023; @Pichler2023], slightly improved stability of our clustering structure in cross-validation and semi-supervised clustering (__Supplementary Figure S15__).

## Issue 1

<span custom-style = 'reviewer'>Your abstract is misleading. 
In the introduction it reads as if you compared OD in wild type and alpha-subtype SARS-CoV-2 virus.</span>

Thank you for the comment. 
We accordingly updated the abstract and clarified it for the reader. 
We state now clearly the abstract (__Abstract/Purpose__ and __Methods/Study design and approval__) that the investigated population of COVID-19 patients were infected with the wild-type and alpha-variant of the SARS-CoV-2 pathogen.

## Issue 2

<span custom-style = 'reviewer'>The abbreviations AT/IT are introduced without clarification in the manuscript.</span>

We thank for the thorough review and give the information in the __Methods/Study design and approval__ section and additionally expand the abbreviations in __Results/Study cohorts__.

## Issue 3

<span custom-style = 'reviewer'>The limitations need more thorough discussion. In particular the fact that subjective report of OD is extremely unreliable. In turn, the objective measurements need to be more present throughout the manuscript. I found data on the objective measurements hard to find.</span>

We would like to thank the reviewer for this comment. 
We agree that subjective reports may suffer from some degree of unreliability. 
The discrepancy between subjective and objective OD irrespective of COVID-19 has been shown in other studies [@Oleszkiewicz2019; @Desiato2021], which calls for robust and reliable screening tools to assess for OD. 
To address this issue more prominently, we present rates of objective and subjective OD along with the assessment of inter-rater reliability with Cohen's $\kappa$ in the CovILD study in the new main text __Figure 2__. 
Additionally, we computed sensitivity and specificity of objective DO detection by the self-reported OD and present these data in __Supplementary Figures S9__ and __S10__. 
Finally, we investigated in details individual trajectories of objective OD diagnosed with the Sniffin' Stick test in a subset of the longitudinal CovILD study participants with complete testing results at the three-month and one-year follow-up (n = `r length(sst_kinet$complete_ids)`). 
As presented in __Supplementary Figure S11__, there were no significant differences either in numeric results of the Sniffin' Stick Test or frequency of objective OD in this participant subset. 
Quite surprisingly, a substantial fraction of COVID-19 convalescents without objective OD at the three-month follow-up developed objective OD at the one-year follow-up.

Collectively, in contrary to subjective OD, which displayed a clear resolution trend in the first year after COVID-19 (__Supplementary Figure S8__), we could not detect any significant differences in objective OD rates between the three-month and one-year follow-up (__Figure 2__, __Supplementary Figure S11__). 
Furthermore, self-reported OD was an extremely unreliable (entire CovILD cohort, 
three-month follow-up: $\kappa$ = `r rater$rater_stats %>% filter(comparison == 'fup100.cohort') %>% .$Kappa %>% signif(2)`, 
one-year follow-up: $\kappa$ = `r rater$rater_stats %>% filter(comparison == 'fup360.cohort') %>% .$Kappa %>% signif(2)`) and 
insensitive tool to assess objective OD (entire CovILD cohort, 
three-month follow-up: sensitivity = `r rater$rater_stats %>% filter(comparison == 'fup100.cohort') %>% .$Sens %>% signif(2)`, 
one-year follow-up: sensitivity = `r rater$rater_stats %>% filter(comparison == 'fup360.cohort') %>% .$Sens %>% signif(2)`). 
The discrepancy between the objective and subjective OD in the CovILD cohort was particularly large for the one-year follow-up and severe COVID-19 patients (__Figure 2__, __Supplementary Figures S9__ and __S10__). 
Finally, objective OD may follow a remission/relapse pattern during long term convalescence, in line with literature [@Tipirdamaz2022].
We discuss these findings in more details in __Results/Subjective and objective OD during COVID-19 recovery__ and __Discussion__.

## Issue 4

<span custom-style = 'reviewer'>I would strongly recommend to elaborate on the Sniffin' Sticks, as these are not an internationally used standard. Why not use the UPSIT? And please elaborate what version of the Sniffin' Sticks were used and what the respective cutoff values are</span>

Thank you for your comment. We used the 16-item Sniffin’ Sticks Identification test and defined hyposmia at a cut-off of <13 correctly identified odorants, as per manufacturer criteria. 
Overall, there is wide heterogeneity in the usage of tests elaborating objective hyposmia [@Desiato2021]. 
There are several studies using the Sniffin’ Sticks test including post-viral hyposmia [@Whitcroft2016; @Masala2022; @Damm2014]. We provide now additional details on the method and literature references in __Methods/Procedures and study variables__.

## Issue 5

<span custom-style = 'reviewer'>Why were the results presented separately for Austria and South Tyrol, despite the obvious geographical, social and economical similarities? Please elaborate or present as one.</span>

We are grateful for this interesting and important issue. 
There are multiple study design-inherent, sociodemographic and mathematical arguments for keeping the AT and IT cohorts separate. 
First, the survey was designed as a two cohort study, this implicated differences in recruitment of participants (AT: via media call, IT: via media call and general practitioners), two language versions (German and Italian) and differences in recruitment start. 
The cohorts were recruited independently and the datasets curated by two collaborating teams (AT: Medical University of Innsbruck, IT: Claudiana College of Bozen/Bolzano) [@Sahanic2021]. 
Second, although the survey was conducted in neighboring regions, Austria and Italy pursued different strategies of pandemic management reflected by differing strategies of contact tracing (AT: state authority, IT: medical staff), testing and quarantine duration [@Sahanic2021]. 
Dynamics of case numbers differed substantially between the countries as well and, notably, there was little exchange of Tyrol and South Tyrol inhabitants due to the closed border. 
In our data, there were several differences in baseline characteristics, symptoms and mental health of the AT and IT cohort, which we present in __Supplementary Figures S3__ and __S5__ of the revised manuscript and discuss in more details in __Results/Study cohorts__. 
Third, as a part of exploratory analysis, we checked if the AT and ITR cohort sizes were sufficient for analyses presented in the manuscript. 
As demonstrated in __Supplementary Figure S2__, the minimal sample size required for reliable clustering analysis, which is the most complex analysis task of the  project, was n = 400 as estimated by Hopkins statistic of clustering tendency. 
The sizes of each AT (n = `r nrow(part_clust$analysis_tbl$north)`) and IT cohort (n = `r nrow(part_clust$analysis_tbl$south)`) treated separately were sufficient for reliable clustering, and no differences in clustering results are expected in case of the pooled AT + IT dataset. 
Hence, there is likely none or little qualitative benefit in analysis of the pooled dataset. 
We provide this information in __Methods/Statistical Analysis__ and __Supplementary Methods/Estimation of minimal sample size for clustering analysis__. 
Fourth, multi-cohort setting including a training, in our case AT cohort, and one or more test collectives, in our case the IT cohort, is considered paramount in any statistical analysis, as it allows for bias-free validation of modeling or clustering results. 
This was also suggested by Reviewers of our previous papers with the survey study [@Hufner2022; @Sahanic2021]. 
To accentuate this 'mathematical' aspect of our analysis strategy, we present the readouts of clustering 'goodness' (fraction of clustering variance, distribution of the clusters) in __Supplementary Figure S15AB__ and elaborate on it in __Results/Smell and taste disorder phenotype of COVID-19 recovery__.

## Issue 6

<span custom-style = 'reviewer'>Even though the English is generally acceptable, there are still quite a few typos that should be corrected before the manuscript is suitable for publication.</span>

Thanks for the careful lecture. 
We paid attention for typos and grammar of the revised manuscript.

\newpage

# References

### 2023-05-30 paper/markdown/figures_and_tables.Rmd
Figures, Tables, Supplementary Material

@@
 subtitle: "Tables and figures"
-author: "Distinct smell and taste disorder phenotype of post-acute COVID-19 sequelae"
+author: "Health after COVID-19 in Tyrol and CovILD study team"
 date: "`r format(Sys.time(), '%Y-%m-%d')`"
@@
 
-```{r tab-hact, tab.cap = 'Baseline characteristic of the Austria (AT) and Italy (IT) survey study cohorts.'}
+```{r tab-hact, tab.cap = 'Baseline characteristic of the Austria (AT) and Italy (IT) survey study cohorts.
+Numeric variables are presented as medians with interquartile ranges (IQR) and ranges.
+Categorical variables are presented as percentages and counts within the complete observation set.'}
 
-paper_tbl$hact_baseline %>% 
-  flextable %>% 
+flextable::flextable(paper_tbl$hact_baseline) %>% 
   width(1, width = 3.7, unit = 'cm') %>% 
-  width(2:3, width = 4, unit = 'cm') %>% 
+  width(2:3, width = 4.8, unit = 'cm') %>% 
   width(4, width = 2.6, unit = 'cm') %>% 
   width(5, width = 2.2, unit = 'cm') %>% 
-  footnote(1, 4:5, value = as_paragraph('Categorical variables: \u03C7\u00B2 test with Cramer V effect size statistic.
+  footnote(1, 1, 
+           value = as_paragraph('BMI: body mass index, normal: BMI < 25 kg/m\u00B2, overweight: BMI 25 - 30 kg/m\u00B2, obesity: BMI > 30 kg/m\u00B2; Pre-CoV depression/anxiety: depression or anxiety before COVID-19; Freq.
+resp.
+infections: frequent (> 2 per year) respiratory infections; Freq.
+bact.
+Infections: frequent (> two per year) bacterial infections with antibiotic therapy; Pre-CoV sleep disorders: sleep disorders before COVID-19.'),
+           ref_symbols = 'a', 
+           part = 'header') %>% 
+  footnote(1, 4:5, 
+           value = as_paragraph('Categorical variables: \u03C7\u00B2 test with Cramer V effect size statistic.
 Numeric variables: Mann-Whitney U test with wilcoxon r effect size statistic.
-P values corrected form multiple testing with Benjamini-Hochberg method.'), ref_symbols = 'a', part = 'header') %>%
-  footnote(3, 1, value = as_paragraph('BMI: body mass index, overweight > 25 kg/m\u00B2, obesity > 30 kg/m\u00B2.'), ref_symbols = 'b', part = 'body') %>% 
-  footnote(15, 1, value = as_paragraph('Frequent respiratory infections, > 2 per year.'), ref_symbols = 'c', part = 'body') %>% 
-  footnote(16, 1, value = as_paragraph('Frequent bacterial infections with antibiotic therapy, > 2 per year.'), ref_symbols = 'd', part = 'body') %>% 
+P values corrected form multiple testing with Benjamini-Hochberg method.'),
+           ref_symbols = 'b', 
+           part = 'header') %>% 
   theme_vanilla
@@
 
-```{r tab-covild, tab.cap = 'Baseline characteristic of the CovILD study cohort and the study participants stratified by COVID-19 severity.'}
+```{r tab-covild, tab.cap = 'Baseline characteristic of the CovILD study cohort and the study participants stratified by COVID-19 severity.
+Numeric variables are presented as medians with interquartile ranges (IQR) and ranges.
+Categorical variables are presented as percentages and counts within the complete observation set.'}
 
-paper_tbl$covild_baseline %>% 
-  flextable %>% 
-  width(1, width = 3, unit = 'cm') %>% 
-  width(2:5, width = 2.7, unit = 'cm') %>% 
+flextable::flextable(paper_tbl$covild_baseline) %>% 
+  width(1, width = 2.8, unit = 'cm') %>% 
+  width(2:5, width = 3.3, unit = 'cm') %>% 
   width(6, width = 2.6, unit = 'cm') %>% 
   width(7, width = 2.2, unit = 'cm') %>% 
-  footnote(1, 6:7, value = as_paragraph('Comparison of ambulatory, moderate and severe COVID-19 individuals.
+  footnote(1, 1, 
+           value = as_paragraph('BMI at CoV onset: body mass index at COVID-19 onset, normal: BMI < 25 kg/m\u00B2, overweight: BMI 25 - 30 kg/m\u00B2, obesity: BMI > 30 kg/m\u00B2.'), 
+           part = 'header', 
+           ref_symbols = 'a') %>% 
+  footnote(1, 6:7, 
+           value = as_paragraph('Comparison of ambulatory, moderate and severe COVID-19 individuals.
 Categorical variables: \u03C7\u00B2 test with Cramer V effect size statistic.
 Numeric variables: Kruskal-Wallis test with \u03B7\u00B2 effect size statistic.
-P values corrected form multiple testing with Benjamini-Hochberg method.'), ref_symbols = 'a', part = 'header') %>%
-  footnote(3, 1, value = as_paragraph('BMI: body mass index, overweight > 25 kg/m\u00B2, obesity > 30 kg/m\u00B2,'), ref_symbols = 'b', part = 'body') %>% 
+P values corrected form multiple testing with Benjamini-Hochberg method.'),
+           part = 'header', 
+           ref_symbols = 'b') %>% 
   theme_vanilla
@@
 
-\newpage
-
-```{r fig-sympt-recovery, fig.width = mm_inch(figures$sympt_recovery$w), fig.height = mm_inch(figures$sympt_recovery$h), fig.cap = 'Symptom-specific recovery times in ambulatory COVID-19 survey study.'}
+```{r fig-sympt-recovery, fig.width = figur::convert(figures$sympt_recovery, to = 'in')$w, fig.height = figur::convert(figures$sympt_recovery, to = 'in')$h, fig.cap = 'Symptom-specific recovery times in the ambulatory COVID-19 survey study.'}
 
@@
 
-__Figure \@ref(fig:fig-sympt-recovery). Symptom-specific recovery times in ambulatory COVID-19 survey study.__
+__Figure \@ref(fig:fig-sympt-recovery). Symptom-specific recovery times in the ambulatory COVID-19 survey study.__
 
-_Symptom-specific recovery times were calculated for each participants of the survey study cohorts (Austria: AT, Italy: IT)._
+_Symptom-specific recovery times were calculated for each participant of the survey study cohorts (Austria: AT, Italy: IT)._
 
-_(A) Distribution of the recovery times in the individuals with the indicated symptoms present during acute COVID-19 (first 14 days after clinical onset).
+_(a) Distribution of the recovery times in the individuals with the indicated symptoms present during acute COVID-19 (first 14 days after clinical onset).
 Diamonds represent median recovery times, tinted ellipses code for interquartile ranges.
@@
 
-_(B) Percentages of individuals with self-reported olfactory dysfunction and taste disorders in the AT (Austria) and IT (Italy) survey study cohorts at particular time points after clinical onset.
+_(b) Percentages of individuals with self-reported olfactory dysfunction and taste disorders in the AT and IT survey study cohorts at particular time points after clinical onset.
 Numbers of complete observations are indicated under the plots._
 
-_OD: self-reported olfactory dysfunction, Imp. concentration: impaired concentration, Dim. appetite: diminished appetite, Imp. walk: impaired walk, Imp. FMS: impaired fine-motor skills._
+_OD: self-reported olfactory dysfunction; Imp. concentration: impaired concentration; Dim. appetite: diminished appetite; Imp. walk: impaired walk; Imp. FMS: impaired fine-motor skills._
 
@@
 
-```{r fig-sympt-mds, fig.width = mm_inch(figures$sympt_mds$w), fig.height = mm_inch(figures$sympt_mds$h), fig.cap = 'Self-reported olfactory dysfunction and taste disorders are isolated persistent symptoms of COVID-19.'}
+```{r fig-sniff-test, fig.width = figur::convert(figures$sniff_test, to = 'in')$w, fig.height = figur::convert(figures$sniff_test, to = 'in')$h, fig.cap = 'Rates of subjective and objective hyposmia in the CovILD cohort two months and one year after COVID-19.'}
+
+figures$sniff_test$plot
+
+```
+
+__Figure \@ref(fig:fig-sniff-test). Rates of subjective and objective hyposmia in the CovILD cohort two months and one year after COVID-19.__
+
+_Objective olfactory dysfunction (OD) was diagnosed in CovILD study participants with < 13 correctly identified odorants in the 16-item Sniffin' Sticks Identification Test.
+Frequencies of objective and self-reported olfactory dysfunction were compared at the three-month (a) and one-year follow-up (b) after COVID-19 in the entire cohort and the ambulatory (A), hospitalized moderate COVID-19 (HM) and hospitalized severe COVID-19 (HS) subset of the cohort.
+Concordance between the self-reported and objective olfactory dysfunction was assessed by Cohen's $\kappa$ inter-rater reliability statistic.
+Percentages of individuals with self-reported and objective hyposmia within the cohort or COVID-19 severity strata are presented in bar plots.
+Cohen's $\kappa$ with 95% confidence intervals (CI) are displayed in Forest plots._
+
+\newpage
+
+```{r fig-sympt-mds, fig.width = figur::convert(figures$sympt_mds, to = 'in')$w, fig.height = figur::convert(figures$sympt_mds, to = 'in')$h, fig.cap = 'Self-reported olfactory dysfunction and taste disorders are isolated persistent symptoms of COVID-19.'}
 
@@
 
-_Symptom data during at $\geq$ 28 days (A) and $\geq$ 3 moths (B) after clinical onset in the Austria (AT) and Italy (IT) survey study cohorts were subjected to two-dimensional multi-dimensional scaling (MDS) with simple matching distance (SMD) between the symptoms.
-MDS coordinates are presented in point plots.
+_Symptom data during at 28 days (a) and 3 months (b) after clinical onset in the Austria (AT) and Italy (IT) survey study cohorts were subjected to two-dimensional multi-dimensional scaling (MDS) with simple matching distance (SMD) between the symptoms.
+MDS coordinates are presented in scatter plots.
 Selected data points are labeled with the symptom names.
 Percentages of the data set variance associated with the MDS dimensions are indicated in the plot axes.
-Numbers of complete observations are indicated under the plots._
+Numbers of complete observations are indicated in the plot captions._
 
-_OD: self-reported olfactory dysfunction, Imp. concentration: impaired concentration._
+_OD: self-reported olfactory dysfunction; Imp. concentration: impaired concentration._
 
@@
 
-```{r fig-apriori, fig.width = mm_inch(figures$apriori$w), fig.height = mm_inch(figures$apriori$h), fig.cap = 'Co-occurrence of self-reported olfactory dysfunction and other symptoms in post-acute COVID-19 sequelae.'}
-
-figures$apriori$plot
-
-```
-
-__Figure \@ref(fig:fig-apriori). Co-occurrence of self-reported olfactory dysfunction and other symptoms in post-acute COVID-19 sequelae.__
-
-_Frequent combinations (>15% of cohort) of self-reported olfactory dysfunction (OD) and other symptoms at $\geq$ 28 days (A) and $\geq$ 3 months (B) after clinical onset in the Austria (AT) and Italy (IT) survey study cohorts were identified with the apriori algorithm.
-Symptom combination frequency (support statistic) and co-occurrence (support statistic) are presented in bubble plots.
-Point size and color corresponds to co-occurrence, points are labeled with percentages of co-occurrence.
-Imp.
-concentration: impaired concentration._
-
-\newpage
-
-```{r fig-sympt-clusters, fig.width = mm_inch(figures$sympt_clusters$w), fig.height = mm_inch(figures$sympt_clusters$h), fig.cap = 'Differing duration of neurocognitive and respiratory symptoms, fatigue, olfactory dysfunction and taste disorders defines the COVID-19 recovery clusters.'}
+```{r fig-sympt-clusters, fig.width = figur::convert(figures$sympt_clusters, to = 'in')$w, fig.height = figur::convert(figures$sympt_clusters, to = 'in')$h, fig.cap = 'Differing duration of neurocognitive and respiratory symptoms, fatigue, olfactory dysfunction and taste disorders defines the COVID-19 recovery clusters.'}
 
@@
 Mean recovery times in the recovery clusters are presented as lines, 2 $\times$ SEM intervals are displayed as tinted regions.
-Numbers of individuals assigned to the recovery clusters are indicated under the plots._
+Numbers of individuals assigned to the recovery clusters are indicated in the plot captions._
 
-_OD: self-reported olfactory dysfunction, Dim. appetite: diminished appetite, Imp. concentration: impaired concentration, Imp. walk: impaired walk, Imp. FMS: impaired fine motor skills._
+_OD: self-reported olfactory dysfunction; Dim. appetite: diminished appetite; Imp. concentration: impaired concentration; Imp. walk: impaired walk; Imp. FMS: impaired fine motor skills._
 
@@
 
-```{r fig-recovery-clusters, fig.width = mm_inch(figures$recovery_clusters$w), fig.height = mm_inch(figures$recovery_clusters$h), fig.cap = 'Good clinical and mental health status and quality of life at the survey completion in individuals with persistent smell and taste disorders.'}
+```{r fig-recovery-clusters, fig.width = figur::convert(figures$recovery_clusters, to = 'in')$w, fig.height = figur::convert(figures$recovery_clusters, to = 'in')$h, fig.cap = 'Physical and mental health, and quality of life in the COVID-19 recovery clusters'}
 
@@
 
-__Figure \@ref(fig:fig-recovery-clusters). Good clinical and mental health status and quality of life at the survey completion in individuals with persistent smell and taste disorders.__
+__Figure \@ref(fig:fig-recovery-clusters). Physical and mental health, and quality of life in the COVID-19 recovery clusters__
 
@@
 Dichotomous items (incomplete convalescence, weight loss, new medication and need for rehabilitation) were binarized (yes: 1, no: 0) prior to visualization.
-Statistical significance for differences between the clusters was assessed by Kruskal-Wallis (numeric readouts) or $\chi^2$ test (dichotomous readouts).
+Statistical significance for differences between the clusters was assessed by Kruskal-Wallis with $\eta^2$ effect size statistic (numeric variables) or $\chi^2$ test with Cramer V effect size statistic (categorical variables).
 P values were corrected for multiple testing with Benjamini-Hochberg method.
 Lines represent mean values, 2 $\times$ SEM intervals are displayed as tinted regions.
-Numbers of individuals assigned to the recovery clusters are indicated under the plots._
+Effect sizes and p values are shown in the plots.
+Numbers of individuals assigned to the recovery clusters are indicated in the plot legends._
 
-_incomplete recovery: self-reported incomplete recovery, # persist.
-symptoms: number of symptoms at $\geq$ 28 days after clinical onset, phys.
-performance loss: physical performance loss as compared with the time before COVID-19, QoL score: quality of life impairment score, OMH score: overall mental health impairment score, ANX score: anxiety score, DPR: depression score._
+_incomplete recovery: self-reported incomplete recovery; # persist.
+symptoms: number of symptoms at 28 days after clinical onset; phys.
+performance loss: physical performance loss as compared with the time before COVID-19; QoL impairment score: score of impaired quality of life; OMH impairment score: overall mental health impairment score; ANX score: anxiety score, Patient Health Questionnaire, PHQ-4; DPR: depression score, Patient Health Questionnaire, PHQ-4; stress score: mental stress score, 7 item PHQ stress module._

### 2023-05-30 paper/markdown/supplementary_material.Rmd
Figures, Tables, Supplementary Material

@@
 library(bookdown)
+library(Cairo)
 
 knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, 
-                      dev = "png", dpi = 600, 
+                      dev = "CairoPNG", dpi = 600, 
                       tab.cap.style = 'Table Heading')
@@
 
-A total of self-reported `r length(globals$hact_symptoms)` symptoms were recorded in the survey study cohorts (__Supplementary Figure S\@ref(fig:fig-hact-freq)__ and __Supplementary Table S\@ref(tab:tab-hact-vars)__).
+A total of `r length(globals$hact_symptoms)` self-reported symptoms were recorded in the survey study cohorts 
+(__Supplementary Figure S\@ref(fig:fig-hact-freq)__ and __Supplementary Table S\@ref(tab:tab-hact-vars)__).
 The symptom duration was coded as follows: absent: 0 days, 1 - 3 days: 3 days, up to 1 week: 7 days, up to 2 weeks: 14 days, up to 4 weeks: 28 days, up to or greater than 3 months: 3 months. 
@@
 
-In the observatory CovILD cohort, a total of self-reported `r length(globals$covild_symptoms)` symptoms (reduced physical performance, olfactory dysfunction, dyspnea, sleep problems, cough, fever, night sweating, gastrointestinal symptoms) were recorded with a standardized questionnaire at each of 60-, 100-, 180- and 360-day post COVID-19 follow-up (__Supplementary Figure S\@ref(fig:fig-covild-freq)__ and __Supplementary Table S\@ref(tab:tab-covild-vars)__).
+In the observatory CovILD cohort, a total of `r length(globals$covild_symptoms)` self-reported symptoms 
+(reduced physical performance, olfactory dysfunction, dyspnea, sleep problems, cough, fever, night sweating, gastrointestinal symptoms) were recorded with a standardized questionnaire at each of 60-, 100-, 180- and 360-day post COVID-19 follow-up (__Supplementary Figure S\@ref(fig:fig-covild-freq)__ and __Supplementary Table S\@ref(tab:tab-covild-vars)__).
 Acute COVID-19 symptoms were assessed retrospectively at the 60-day follow-up [@Sonnweber2020; @Sonnweber2022].
@@
 Anxiety (ANX) and depression (DPR) were assessed with the Patient Health Questionnaire (PHQ-4) [@Hufner2022; @Sahanic2021; @Lowe2010]. 
-Psychosocial stress was scored with a modified 7 item PHQ stress module as described before [@Hufner2022; @Sahanic2021; @Grafe2004].
+Mental stress was scored with a modified 7 item PHQ stress module as described before [@Hufner2022; @Sahanic2021; @Grafe2004].
 
-### Rating of olfactory dysfunction with sniffing stick test
+### Rating of olfactory dysfunction with Sniffin' Stick Test
 
-Objective olfactory dysfunction at the 100-day and 360-day follow-up in the CovILD study participants was investigated with the 16-item sniffing stick test as described [@Rass2021].
-Clinically relevant olfactory dysfunction was defined as $\leq$ 12 correct answers [@Rass2021; @Hummel2007].
+Objective olfactory dysfunction at the 100-day and 360-day follow-up in the CovILD study participants was investigated with the 16-item Sniffn' Stick Identification Test as described [@Rass2021; @Whitcroft2016; @Masala2022; @Damm2014].
+In brief, the nasal chemosensory performance was investigated using pen-like odor-dispensing devices for odor identification of 16 common odorants (multiple forced‐choice from four verbal items per test odorant).
+Clinically relevant olfactory dysfunction was defined as < 13 correct answers (points) [@Rass2021; @Hummel2007]. 
 In the analysis, participants with the complete answers concerning self-reported olfactory dysfunction and complete test results were included.
 
-## Statistical analysis
-### Data transformation, descriptive statistic
+## Data analysis
 
-Data transformation and statistical analysis was done with R version 4.2.0 with _tidyverse_ data science environment [@Wickham2019].
-Analysis results were visualized with _ggplot2_ [@Wickham2016], _cowplot_ [@Wilke2019] as well as in-house developed _ExDA_ (https://github.com/PiotrTymoszuk/ExDA) and _figur_ (https://github.com/PiotrTymoszuk/figur) packages.
+### Software
 
-Descriptive statistics including median with interquartile ranges and frequency of complete answers for numeric and categorical variables were calculated with base R functions and _ExDA_ package.
+Analysis was done with R version 4.2.3 (The R Project for Statistical Computing).
 
-### Statistical hypothesis testing
+Import of raw study data was accomplished with the packages _foreign_ [@RCoreTeam] and _readxl_ [@Wickham2022]. 
+Tabular data was handled with the _tidyverse_ package bundle [@Wickham2019], and the _rlang_ [@Henry2022] and [_trafo_](https://github.com/PiotrTymoszuk/trafo) packages. 
+Search and transformation of text data was done with _stringi_ [@Gagolewski2021]. 
+For exploratory data analysis, descriptive statistic and statistical hypothesis testing, packages _rstatix_ [@Kassambara2021], _vcd_ [@Meyer2021], _rcompanion_ [@Mangiafico2022] and [_ExDA_](https://github.com/PiotrTymoszuk/ExDA) were used.
+For modeling of symptom recovery, packages _lme_ [@Bates2015], _lmerTest_ [@Kuznetsova2017] and [_kinet_](https://github.com/PiotrTymoszuk/kinet) were employed. 
+Multi-dimensional scaling was performed with the base R's _stats_ package and [_clustTools_](https://github.com/PiotrTymoszuk/clustTools). 
+Apriori analysis was done with _arules_ [@Hahsler2005]. 
+Clustering analysis was performed with the packages _cluster_ [@Schubert2019], _philentropy_ [@Drost2018], _scrime_ [@Schwender2018], _factoextra_ [@Kassambara2020] and [_clustTools_](https://github.com/PiotrTymoszuk/clustTools).
+Inter-rater analysis and receiver operating characteristic (ROC) were done with the packages _caret_ [@Kuhn2008] and _vcd_ [@Meyer2021]. 
+Parallelization of analysis tasks was accomplished with the package _furrr_ package [@Vaughan2022].
 
-Since multiple study variables were non-normally distributed as assessed by Shapiro-Wilk test and visual assessment of their distribution (quantile-quantile plots), statistical significance for differences in outcome numeric variables were assessed with Mann-Whitney U test with r effect size statistic (two groups) or Kruskal-Wallis test with $\eta^2$ effect size statistic.
+Analysis results were visualized with _ggplot_ [@Wickham2016] (bar, scatter and bubble plots, heat maps of confusion matrices), [_ExDA_](https://github.com/PiotrTymoszuk/ExDA) (ellipse plots, radar, ribbon, stack and violin plots), [_kinet_](https://github.com/PiotrTymoszuk/kinet) (kinetic of symptom resolution),
+_plotROC_ [@Sachs2017] (ROC plots), [_clustTools_](https://github.com/PiotrTymoszuk/clustTools) (multi-dimensional scaling projection/scatter plots, heat maps of clustering features, within-cluster sum of squares).
+HTML elements within the plots were handled with tools provided by the package _ggtext_ [@Wilke2022].
+
+Figures were generated with _cowplot_ [@Wilke2019]. 
+Tables were rendered with the _flextable_ package [@Gohel2022]. 
+Supplementary Material and parts of the manuscript were written in the _rmarkdown_ environment [@Allaire2022] and were rendered as Word documents with the _knitr_ [@Xie2022] and _bookdown_ [@Xie2016] packages.
+Management of figures and tables within the _rmarkdown_ documents was accomplished with the development package [_figur_](https://github.com/PiotrTymoszuk/figur).
+
+### Data import and formatting
+
+Data of the survey study [@Sahanic2021; @Hufner2022] were imported from SPSS files with the _foreign_ package (function `read.spss()`). 
+The CovILD study data set [@Sahanic2023] was imported from an Excel file with the function `read_xlsx()` (package _readxl_). 
+Formatting of the raw data sets was done with in-house developed scripts available from the [project's GitHub repository](https://github.com/PiotrTymoszuk/hyposmia_analysis_pipeline).
+
+### Estimation of minimal sample size for clustering analysis
+
+To assess the minimal sample size of the survey study required for reliable clustering analysis, random subsets of the pooled Austria and Italy cohort COVID-19 symptom data set with differing observation numbers (50, 100, 200, 300, 400, 500, 600, 700, 800, 900 observations, 20 random subsets per sample size) were generated and their clustering tendency was assessed with Hopkins statistic (H).
+Possible H values span from 0 to 1, where H = 0 indicates an ideal uniform distribution and H = 1 suggests a highly clustered data. 
+Beginning from the sample size of n = 400, no improvement of the clustering tendency (plateau) could be observed. 
+This suggest n = 400 as minimal sample size required for reproducible clustering analysis results (__Supplementary Figure S\@ref(fig:fig-sample-size)__). 
+Given the size of the Austria and Italy cohorts 
+(n = `r nrow(part_clust$analysis_tbl$north)` and n = `r nrow(part_clust$analysis_tbl$south)`, respectively) and their H values being very close to the H plateau value in the random subset analysis 
+(Austria: H = `r signif(eda_size$tendency_cohorts$north$hopkins_stat, 2)`, 
+Italy: H = `r signif(eda_size$tendency_cohorts$south$hopkins_stat, 2)`), we inferred that the size of single study cohorts was sufficient for reliable and reproducible clustering analysis and hence abstained from pooling the survey study cohorts.
+
+### Descriptive statistic and effect size
+
+If not indicated otherwise, numeric values are presented in the manuscript text and tables as medians with interquartile ranges (IQR) and ranges. 
+Qualitative variables are presented as percentages and counts of their categories within the complete observation set. 
+Descriptive statistic values were computed with the function `explore()` from the [_ExDA_](https://github.com/PiotrTymoszuk/ExDA) package.
+
+Effects sizes were assessed following the scheme proposed by Cohen [@Cohen2013]. 
+Effect size for comparison of categorical variables was measured with Cramer's V statistic (weak: < 0.3, moderate: 0.3 - 0.5, large effect: $\geq$ 0.5). 
+Effect size for two-group comparisons of numeric variables was assessed with r statistic (weak: < 0.3, moderate: 0.3 - 0.5, large effect: $\geq$ 0.5). 
+Effect size for multi-group comparison of numeric variables was assessed with $\eta^2$ (weak: < 0.13, moderate: 0.13 - 0.26, large effect: $\geq$ 0.26). 
+Effect size for comparison of paired proportions (2 $\times$ 2 contingency table) was assessed with Cohen's g (weak: < 0.15, moderate: 0.15 - 0.25, large effect: $\geq$ 0.25). 
+Effect size for comparison of study participant-matched proportions (multi-dimensional contingency table) was assessed with Kendall's W (weak: < 0.3, moderate: 0.3 - 0.5, large effect: $\geq$ 0.5). 
+Effect size for inter-rate reliability was measures with Cohen's $\kappa$ statistic was interpreted as follows: 0.01 – 0.20 as none to slight, 0.21 – 0.40 as fair, 0.41 – 0.60 as moderate, 0.61 – 0.80 as substantial, and 0.81 – 1.00 as strong [@McHugh2012].
+Effect sizes were computed with the function `compare_variables()` from the development package [_ExDA_](https://github.com/PiotrTymoszuk/ExDA) employing tools introduced by the _rstatix_ and _vcd_ packages (Cramer's V, r, $\eta^2$, Cohen's $\kappa$, Cohen's g) or with the function `cohenG()` from the package _rcompanion_ [@Mangiafico2022].
+
+### Statistical hypothesis testing and inter-rater reliability
+
+Since multiple study variables were non-normally distributed as assessed by Shapiro-Wilk test and visual assessment of their distribution (quantile - quantile plots), statistical significance for differences in outcome numeric variables were assessed with Mann-Whitney U test with r effect size statistic (two groups, independent data), paired Wilcoxon test with r effect size statistic (two groups, paired data) or Kruskal-Wallis test with $\eta^2$ effect size statistic.
 Differences in frequency distribution for categorical outcome variables were assessed by $\chi^2$ test with Cramer V effect size statistic. 
-Inter-rater assessment of self-reported and sniffing test olfactory dysfunction was accomplished with Cohen's $\kappa$ statistic [@Fleiss1969].
-$\kappa$ significance ($\kappa\neq0$) was estimated with Wald Z test.
+Significance of differences of paired proportions (2 $\times$ 2 contingency table) was determined with McNemar test with Cohen's g effect size statistic. 
+Mann-Whitney, Kruskal-Wallis, $\chi^2$ tests were accomplished with the function `compare_variables()` (package [_ExDA_](https://github.com/PiotrTymoszuk/ExDA)) employing algorithms from the package _rstatix_ [@Kassambara2021].
+McNemar tests and Cohen's g were calculated with the functions `mcnemar_test()` (package _rstatix_) [@Kassambara2021] and `cohenG()` (package _rcompanion_) [@Mangiafico2022], respectively. 
+Significance of symptom resolution (longitudinal, participant-matched binary data) was determined by Cochran's Q test with Kendall's W effect size statistic (function `cochran_qtest()`, package _rstatix_) [@Serlin1982; @Kassambara2021].
+
+Inter-rater assessment of self-reported and Sniffin Test olfactory dysfunction was accomplished with Cohen's $\kappa$ statistic [@Fleiss1969; @McHugh2012]. 
+$\kappa$ significance ($\kappa\neq0$) was estimated with Wald Z test (function `correlate_variable()`, package [_ExDA_](https://github.com/PiotrTymoszuk/ExDA) employing the _vcd_ package) [@Meyer2021].
+Additionally, reliability of detection of objective olfactory dysfunction (Sniffin' Stick Test) was additionally assessed by ROC analysis. 
+Overall accuracy, sensitivity and specificity was computed with the functions `defaultSummary()` and `twoClassSummary()` from the package _caret_ [@Kuhn2008].
+
 P values were adjusted for multiple testing with Benjamini-Hochberg method [@Benjamini1995] separately for each analysis task and cohort. 
-Packages _rstatix_ [@Kassambara2021], _vcd_ [@Meyer2021] and _Exda_ (https://github.com/PiotrTymoszuk/ExDA) were used for statistical hypothesis testing.
-
-### Modeling of symptom recovery kinetic
-
-To model recovery kinetics for binary symptom variables (0: absent, 1: present), second-order mixed-effect logistic (categorical features) modeling was applied (packages: _lme4_, _lmerTest_ and development package _kinet_ [https://github.com/PiotrTymoszuk/kinet]) [@Bates2015; @Kuznetsova2017; @Box1978].
-Each model followed the general formula:
-
-$$Response \sim time + time^2 + (1|individual)$$
-
-where $(1|individual)$ indicates the random effect of the individual and $time$ and $time^2$ indicate the first- and second-order time effect terms.
-The first-order term estimate was interpreted as a measure of the recovery speed and the second-order term estimate was used to assess the plateau/rebound effect.
-Significance of the accuracy gain of the full second-order model compared with the nested null model was determined by likelihood ratio test (LRT) versus the nested first-order and null models, respectively.
-Likelihood ratio $\lambda$ statistic (full versus null model) was used as an effect size measure.
-Individuals from the survey study or the CovILD cohort with the complete longitudinal symptom record were included in the kinetic modeling tasks.
-Results of the kinetic modeling were adjusted for multiple comparisons with Benjamini-Hochberg method [@Benjamini1995].
+Effects with p < 0.05 were considered significant.
 
@@
 
-To assess co-occurrence or exclusivity of symptoms, simple matching distances between manifestations during the first 14 days, at $\geq$ 28 days and at $\geq$ 3 months after clinical onset in the survey study cohorts were calculated (package _scrime_ and development package _clustTools_ [https://github.com/PiotrTymoszuk/clustTools]) [@Boriah2008; @Schwender2018].
-Subsequently, the distance matrix was subjected to multi-dimensional scaling (MDS, k = 2 dimensions, package _stats_, function _cmdscale()_).
+To assess co-occurrence or exclusivity of symptoms, simple matching distances between manifestations during the first 14 days, at 28 days and at 3 months after clinical onset in the survey study cohorts were calculated (function `calculate_dist()`, package [_clustTools_](https://github.com/PiotrTymoszuk/clustTools) employing tools from packages _scrime_ and _philentropy_) [@Boriah2008; @Schwender2018; @Drost2018].
+Subsequently, the distance matrix was subjected to multi-dimensional scaling (MDS, k = 2 dimensions, package _stats_, function `cmdscale()`). 
 Association of specific symptoms was assessed by visual analysis of MDS coordinate plots.
@@
 
-Frequent combinations of symptoms during the first 14 days, at $\geq$ 28 days and at $\geq$ 3 months after clinical onset in the survey study cohorts were identified with the apriori algorithm (package _arules_) [@Agrawal1993; @Hahsler2005] with the minimal support cutoff of 0.15, 2 - 10 item transaction length, confidence > 0.6 and lift > 2.
+Frequent combinations of symptoms during the first 14 days, at 28 days and at 3 months after clinical onset in the survey study cohorts were identified with the apriori algorithm (function `apriori()`, package _arules_) [@Agrawal1993; @Hahsler2005] with the minimal support cutoff of 0.15, 2 - 10 item transaction length, confidence > 0.6 and lift > 2.
 The support statistic were used to estimate the symptom combination frequency. 
@@
 The lift statistic was interpreted as a measure of the symptom dependence (lift = 1, symptoms are independent). 
+Frequency of symptom combination in the survey study cohorts and percentage of co-occurrence of the symptoms within the given symptom combination were displayed in bobble plots.
 
@@
 
-COVID-19 recovery clusters of the training Autria (AT) survey cohort participants in respect to symptom-specific recovery times (__Figure 1A__) were defined with the PAM (partitioning around medoids) algorithm and Euclidean distance statistic (packages _cluster_, _philentropy_ and development package _clustTools_ [https://github.com/PiotrTymoszuk/clustTools]) [@Drost2018; @Schubert2019].
-The set of participants with the complete clustering variable set was included in the analysis.
+COVID-19 recovery clusters of the training Austria survey cohort participants in respect to symptom-specific recovery times (__Figure 1A__) were defined with the PAM (partitioning around medoids) algorithm and Euclidean distance statistic [@Drost2018; @Schubert2019].
+The set of participants with the complete clustering variable set (COVID-19 symptom recovery times) was included in the analysis. 
 The symptom recovery times were not subjected to any type of pre-processing. 
-The choice of the clustering procedure was motivated by the analysis of the clustering variance (ratio of the total between-cluster to total sum of squares) and clustering structure stability in 10-fold cross-validation (metric: rate of correct cluster assignment, cluster assignment predicted by k = 5 nearest neighbors label propagation algorithm, package _clustTools_) [@Lange2004; @Leng2014] for several clustering algorithms as presented in __Supplementary Figure S\@ref(fig:fig-clust-devel)A__.
-The optimal number of clusters was determined by the bend of the total within-cluster sum of squares curve (__Supplementary Figure S\@ref(fig:fig-clust-devel)B__, package _factoextra_) [@Kassambara2020].
-Permutation importance of specific clustering variables was investigated by calculating difference in clustering variance (ratio of total between-cluster sum of squares to total sum of squares) between the initial clustering object and the clustering object with the given variable reshuffled at random (package _clustTools_).
-Assignment of the Italy survey cohort participants to the recovery clusters was accomplished with k-nearest neighbors label propagation algorithm (k = 5) [@Leng2014].
-The clustering efficacy in the training AT cohort and the test IT cohort measured by clustering variance statistic defined above was similar (AT: `r clustTools::var(part_clust$clust_north)$frac_var %>% signif(2)`, IT: `r clustTools::var(part_clust$clust_south)$frac_var %>% signif(2)`).
+The clustering objects were generated with the function `kcluster(clust_fun = 'pam')` from the package [_clustTools_](https://github.com/PiotrTymoszuk/clustTools).
+
+The choice of the clustering procedure was motivated by the analysis of the fraction of explained clustering variance (ratio of the total between-cluster to total sum of squares) and clustering structure stability in 10-fold cross-validation (metric: rate of correct cluster assignment, cluster assignment predicted by inverse distance-weighted 7-nearest neighbors label propagation algorithm) [@Lange2004; @Leng2014] for several clustering algorithms as presented in __Supplementary Figure S\@ref(fig:fig-clust-devel)A__.
+The fractions of explained clustering variance and cross-validated cluster assignment accuracy were calculated with the methods `var()` and `cv()` from the package [_clustTools_](https://github.com/PiotrTymoszuk/clustTools).
+
+The optimal number of clusters was determined by the bend of the total within-cluster sum of squares curve 
+(__Supplementary Figure S\@ref(fig:fig-clust-devel)B__, method `plot()`, package [_clustTools_](https://github.com/PiotrTymoszuk/clustTools), employing the genuine _factoextra_ algorithm) [@Kassambara2020].
+Permutation importance of specific clustering variables was investigated by calculating difference in clustering variance (ratio of total between-cluster sum of squares to total sum of squares) between the initial clustering object and the clustering object with the given variable reshuffled at random (function `impact()`, package [_clustTools_](https://github.com/PiotrTymoszuk/clustTools)).
+The permutation importance statistics were computed for 20 random permutations of each clustering variable.
+
+Assignment of the Italy survey cohort participants to the recovery clusters was accomplished with the inverse distance-weighted 7-nearest neighbors label propagation classifier [@Leng2014]. 
+The clustering efficacy in the training Austria cohort and the test Italy cohort measured by clustering variance statistic defined above was similar 
+(Austria: `r clustTools::var(part_clust$clust_north)$frac_var %>% signif(2)`, 
+Italy: `r clustTools::var(part_clust$clust_south)$frac_var %>% signif(2)`), which indicate good reproducibility of the clustering structure developed in the training Austria cohort.
 
@@
 
-The raw data files will be made available upon request. The entire analysis pipeline was published at https://github.com/PiotrTymoszuk/hyposmia_analsis_pipeline.
+The raw data files will be made available upon request. The entire analysis pipeline was published at https://github.com/PiotrTymoszuk/hyposmia_analysis_pipeline.
 
@@
 
-```{r tab-hact-vars, tab.cap = 'Survey study variables.'}
+```{r tab-hact-vars, tab.cap = 'Survey study variables. The table is available as a supplementary Excel sheet.'}
 
-suppl_tbl$hact_vars %>% 
-  select(- Stratification) %>% 
-  flextable %>% 
-  width(1, width = 3.7, unit = 'cm') %>% 
-  width(2, width = 3.9, unit = 'cm') %>% 
-  width(3, width = 1.8, unit = 'cm')  %>% 
-  width(4, width = 8.7, unit = 'cm') %>% 
+suppl_tbl$hact_vars[1, 1] %>% 
+  mutate(`Variable name` = '') %>% 
+  set_names(' ') %>% 
+  flextable
+
+```
+
+\newpage
+
+```{r tab-covild-vars, tab.cap = 'CovILD study variables. The table is available as a supplementary Excel sheet.'}
+
+suppl_tbl$covild_vars[1, 1] %>% 
+  mutate(`Variable name` = '') %>% 
+  set_names(' ') %>% 
+  flextable
+
+```
+
+\newpage
+
+```{r tab-hact-kinetic, tab.cap = 'Results of statistical hypothesis testing for significant recovery of the most frequent COVID-19 symptoms in the Austria (AT) and Italy cohort (IT) of the survey study.'}
+
+flextable::flextable(suppl_tbl$hact_kinetic) %>% 
+  width(1, width = 2.5, unit = 'cm') %>% 
+  width(2, width = 4.5, unit = 'cm') %>% 
+  width(3:4, width = 3, unit = 'cm') %>% 
+  merge_v(1) %>% 
+  footnote(1, 2, 
+           value = as_paragraph('OD: self-reported olfactory dysfunction; Dim. appetite: diminished appetite.'), 
+           part = 'header', 
+           ref_symbols = 'a') %>% 
+  footnote(1, 3:4, 
+           value = as_paragraph("Cochran's Q test with Kendall's W effect size statistic. P values corrected for multiple testing with Benjamini-Hochberg method."), 
+           part = 'header', 
+           ref_symbols = 'b') %>% 
   theme_vanilla
@@
 
-```{r tab-covild-vars, tab.cap = 'CovILD study variables.'}
+```{r tab-covild-kinetic, tab.cap = 'Results of statistical hypothesis testing for significant recovery of COVID-19 symptoms in COVID-19 severity strata of the CovILD study.'}
 
-suppl_tbl$covild_vars %>% 
-  flextable %>% 
-  width(1, width = 4.5, unit = 'cm') %>% 
+flextable::flextable(suppl_tbl$covild_kinetic) %>% 
+  width(1, width = 3, unit = 'cm') %>% 
   width(2, width = 4.5, unit = 'cm') %>% 
-  width(3, width = 2, unit = 'cm')  %>% 
+  width(3:4, width = 3, unit = 'cm') %>% 
+  merge_v(1) %>% 
+  footnote(1, 2, 
+           value = as_paragraph('OD: self-reported olfactory dysfunction.'), 
+           part = 'header', 
+           ref_symbols = 'a') %>% 
+  footnote(1, 3:4, 
+           value = as_paragraph("Cochran's Q test with Kendall's W effect size statistic. P values corrected for multiple testing with Benjamini-Hochberg method."), 
+           part = 'header', 
+           ref_symbols = 'b') %>% 
   theme_vanilla
@@
 
-```{r tab-clust-base-at, tab.cap = 'Demographic and baseline clinical characteristic at the COVID-19 onset of the survey study participants assigned to the recovery clusters, Austria (AT) cohort.'}
+```{r tab-od-kinetic, tab.cap = "Results of the Sniffin' Stick Test in the CovILD study subset with the complete longitudinal follow-up data.
+Numeric variables are presented as medians with interquartile ranges (IQR) and ranges.
+Categorical variables are presented as percentages and counts within the complete observation set."}
 
-suppl_tbl$baseline_clusters_north %>% 
-  flextable %>% 
-  width(1, width = 3.5, unit = 'cm') %>% 
-  width(2:4, width = 3.7, unit = 'cm') %>% 
-  width(5, width = 2.6, unit = 'cm') %>% 
-  width(6, width = 2.2, unit = 'cm') %>% 
-  footnote(1, 5:6, value = as_paragraph('Categorical variables: \u03C7\u00B2 test with Cramer V effect size statistic.
-Numeric variables: Kruskal-Wallis test with \u03B7\u00B2 effect size statistic.
-P values corrected form multiple testing with Benjamini-Hochberg method.'), ref_symbols = 'a', part = 'header') %>%
-  footnote(3, 1, value = as_paragraph('BMI: body mass index, overweight > 25 kg/m\u00B2, obesity > 30 kg/m\u00B2,'), ref_symbols = 'b', part = 'body') %>% 
-  footnote(15, 1, value = as_paragraph('Frequent respiratory infections, > 2 per year.'), ref_symbols = 'c', part = 'body') %>% 
-  footnote(16, 1, value = as_paragraph('Frequent bacterial infections with antibiotic therapy, > 2 per year.'), ref_symbols = 'd', part = 'body') %>% 
+flextable::flextable(suppl_tbl$od_kinetic) %>% 
+  width(1, width = 4, unit = 'cm') %>% 
+  width(2:3, width = 4.5, unit = 'cm') %>% 
+  width(4, width = 2.7, unit = 'cm') %>% 
+  width(5, width = 2.5, unit = 'cm') %>% 
+  footnote(1, 1, 
+           value = as_paragraph('OD: olfactory dysfunction.'), 
+           part = 'header', 
+           ref_symbols = 'a') %>% 
+  footnote(1, 3:4, 
+           value = as_paragraph("Categorical variables: McNemar test with Cohen's g effect size statistic.
+Numeric variables: paired Wilcoxon test with r effect size statistic.
+P values corrected for multiple testing with Benjamini-Hochberg method."),
+           part = 'header', 
+           ref_symbols = 'b') %>% 
   theme_vanilla
@@
 
-```{r tab-clust-base-it, tab.cap = 'Demographic and baseline clinical characteristic at the COVID-19 onset of the survey study participants assigned to the recovery clusters, Italy (IT) cohort.'}
+```{r tab-clust-base-at, tab.cap = 'Demographic and baseline clinical characteristic at the COVID-19 onset of the survey study participants assigned to the recovery clusters, Austria (AT) cohort.
+Numeric variables are presented as medians with interquartile ranges (IQR) and ranges.
+Categorical variables are presented as percentages and counts within the complete observation set.'}
 
-suppl_tbl$baseline_clusters_south %>% 
-  flextable %>% 
+flextable::flextable(suppl_tbl$baseline_clusters_north) %>% 
   width(1, width = 3.5, unit = 'cm') %>% 
@@
   width(6, width = 2.2, unit = 'cm') %>% 
-  footnote(1, 5:6, value = as_paragraph('Categorical variables: \u03C7\u00B2 test with Cramer V effect size statistic.
+  footnote(1, 1, 
+           value = as_paragraph('BMI: body mass index, normal: BMI < 25 kg/m\u00B2, overweight: BMI 25 - 30 kg/m\u00B2, obesity: BMI > 30 kg/m\u00B2; Pre-CoV depression/anxiety: depression or anxiety before COVID-19; Freq.
+resp.
+infections: frequent (> 2 per year) respiratory infections; Freq.
+bact.
+Infections: frequent (> two per year) bacterial infections with antibiotic therapy; Pre-CoV sleep disorders: sleep disorders before COVID-19.'),
+           part = 'header', 
+           ref_symbols = 'a') %>% 
+  footnote(1, 5:6, 
+           value = as_paragraph('Categorical variables: \u03C7\u00B2 test with Cramer V effect size statistic.
 Numeric variables: Kruskal-Wallis test with \u03B7\u00B2 effect size statistic.
-P values corrected form multiple testing with Benjamini-Hochberg method.'), ref_symbols = 'a', part = 'header') %>%
-  footnote(3, 1, value = as_paragraph('BMI: body mass index, overweight > 25 kg/m\u00B2, obesity > 30 kg/m\u00B2,'), ref_symbols = 'b', part = 'body') %>% 
-  footnote(15, 1, value = as_paragraph('Frequent respiratory infections, > 2 per year.'), ref_symbols = 'c', part = 'body') %>% 
-  footnote(16, 1, value = as_paragraph('Frequent bacterial infections with antibiotic therapy, > 2 per year.'), ref_symbols = 'd', part = 'body') %>% 
+P values corrected form multiple testing with Benjamini-Hochberg method.'),
+           part = 'header', 
+           ref_symbols = 'b') %>% 
   theme_vanilla
@@
 
-```{r tab-clust-reco-at, tab.cap = 'COVID-19 course and recovery in the survey study participants assigned to the recovery clusters, Austria (AT) cohort.'}
+```{r tab-clust-base-it, tab.cap = 'Demographic and baseline clinical characteristic at the COVID-19 onset of the survey study participants assigned to the recovery clusters, Italy (IT) cohort.
+Numeric variables are presented as medians with interquartile ranges (IQR) and ranges.
+Categorical variables are presented as percentages and counts within the complete observation set.'}
 
-suppl_tbl$cov_clusters_north %>% 
-  flextable %>% 
+flextable::flextable(suppl_tbl$baseline_clusters_south) %>% 
    width(1, width = 3.5, unit = 'cm') %>% 
@@
   width(6, width = 2.2, unit = 'cm') %>% 
-  footnote(1, 5:6, value = as_paragraph('Categorical variables: \u03C7\u00B2 test with Cramer V effect size statistic.
+  footnote(1, 1, 
+           value = as_paragraph('BMI: body mass index, normal: BMI < 25 kg/m\u00B2, overweight: BMI 25 - 30 kg/m\u00B2, obesity: BMI > 30 kg/m\u00B2; Pre-CoV depression/anxiety: depression or anxiety before COVID-19; Freq.
+resp.
+infections: frequent (> 2 per year) respiratory infections; ;Freq.
+bact.
+Infections: frequent (> two per year) bacterial infections with antibiotic therapy; Pre-CoV sleep disorders: sleep disorders before COVID-19.'),
+           part = 'header', 
+           ref_symbols = 'a') %>% 
+  footnote(1, 5:6, 
+           value = as_paragraph('Categorical variables: \u03C7\u00B2 test with Cramer V effect size statistic.
 Numeric variables: Kruskal-Wallis test with \u03B7\u00B2 effect size statistic.
-P values corrected form multiple testing with Benjamini-Hochberg method.'), ref_symbols = 'a', part = 'header') %>%
-  footnote(4, 1, value = as_paragraph('Self-reported incomplete recovery.'), ref_symbols = 'd', part = 'body') %>% 
-  footnote(8, 1, value = as_paragraph('ANX: anxiety.'), ref_symbols = 'e', part = 'body') %>% 
-  footnote(9, 1, value = as_paragraph('DPR: depression.'), ref_symbols = 'f', part = 'body') %>% 
-  footnote(11, 1, value = as_paragraph('OMH score: overall mental health impairment score'), ref_symbols = 'g', part = 'body') %>% 
-  footnote(12, 1, value = as_paragraph('QoL score: quality of life impairment score'), ref_symbols = 'h', part = 'body') %>%
+P values corrected form multiple testing with Benjamini-Hochberg method.'),
+           part = 'header', 
+           ref_symbols = 'b') %>% 
   theme_vanilla
@@
 
-```{r tab-clust-reco-it, tab.cap = 'COVID-19 course and recovery in the survey study participants assigned to the recovery clusters, Italy (IT) cohort.'}
+```{r tab-clust-reco-at, tab.cap = 'COVID-19 course and recovery in the survey study participants assigned to the recovery clusters, Austria (AT) cohort.
+Numeric variables are presented as medians with interquartile ranges (IQR) and ranges.
+Categorical variables are presented as percentages and counts within the complete observation set.'}
 
-suppl_tbl$cov_clusters_south %>% 
-  flextable %>% 
+flextable::flextable(suppl_tbl$cov_clusters_north) %>% 
   width(1, width = 3.5, unit = 'cm') %>% 
@@
   width(6, width = 2.2, unit = 'cm') %>% 
- footnote(1, 5:6, value = as_paragraph('Categorical variables: \u03C7\u00B2 test with Cramer V effect size statistic.
+  footnote(1, 1, 
+           value = as_paragraph('Incomplete recovery: self-reported incomplete recovery from COVID-19; Physical performance loss: self-rated physical performance loss after COVID-19, before COVID-19: 100%; ANX score: anxiety score, Patient Health Questionnaire, PHQ-4; DPR: depression score, Patient Health Questionnaire, PHQ-4; Stress score: mental stress score; 7 item PHQ stress module; OMH impairment score: score of overall mental health impairment; QoL impairment score: score of impaired quality of life.'),
+           part = 'header', 
+           ref_symbols = 'a') %>% 
+  footnote(1, 5:6, 
+           value = as_paragraph('Categorical variables: \u03C7\u00B2 test with Cramer V effect size statistic.
 Numeric variables: Kruskal-Wallis test with \u03B7\u00B2 effect size statistic.
-P values corrected form multiple testing with Benjamini-Hochberg method.'), ref_symbols = 'a', part = 'header') %>%
-  footnote(4, 1, value = as_paragraph('Self-reported incomplete recovery.'), ref_symbols = 'd', part = 'body') %>% 
-  footnote(8, 1, value = as_paragraph('ANX: anxiety.'), ref_symbols = 'e', part = 'body') %>% 
-  footnote(9, 1, value = as_paragraph('DPR: depression.'), ref_symbols = 'f', part = 'body') %>% 
-  footnote(11, 1, value = as_paragraph('OMH score: overall mental health impairment score'), ref_symbols = 'g', part = 'body') %>% 
-  footnote(12, 1, value = as_paragraph('QoL score: quality of life impairment score'), ref_symbols = 'h', part = 'body') %>%
+P values corrected form multiple testing with Benjamini-Hochberg method.'),
+           part = 'header', 
+           ref_symbols = 'a') %>% 
+  theme_vanilla
+
+```
+
+\newpage
+
+```{r tab-clust-reco-it, tab.cap = 'COVID-19 course and recovery in the survey study participants assigned to the recovery clusters, Italy (IT) cohort.
+Numeric variables are presented as medians with interquartile ranges (IQR) and ranges.
+Categorical variables are presented as percentages and counts within the complete observation set.'}
+
+flextable::flextable(suppl_tbl$cov_clusters_south) %>% 
+   width(1, width = 3.5, unit = 'cm') %>% 
+  width(2:4, width = 3.7, unit = 'cm') %>% 
+  width(5, width = 2.6, unit = 'cm') %>% 
+  width(6, width = 2.2, unit = 'cm') %>% 
+  footnote(1, 1, 
+           value = as_paragraph('Incomplete recovery: self-reported incomplete recovery from COVID-19; Physical performance loss: self-rated physical performance loss after COVID-19, before COVID-19: 100%; ANX score: anxiety score, Patient Health Questionnaire, PHQ-4; DPR: depression score, Patient Health Questionnaire, PHQ-4; Stress score: mental stress score; 7 item PHQ stress module; OMH impairment score: score of overall mental health impairment; QoL impairment score: score of impaired quality of life.'),
+           part = 'header', 
+           ref_symbols = 'a') %>% 
+  footnote(1, 5:6, 
+           value = as_paragraph('Categorical variables: \u03C7\u00B2 test with Cramer V effect size statistic.
+Numeric variables: Kruskal-Wallis test with \u03B7\u00B2 effect size statistic.
+P values corrected form multiple testing with Benjamini-Hochberg method.'),
+           part = 'header', 
+           ref_symbols = 'a') %>% 
   theme_vanilla
@@
 
-```{r fig-consort, fig.width = mm_inch(suppl$consort$w), fig.height = mm_inch(suppl$consort$h), fig.cap = 'Flow diagram of the analysis inclusion process for the longitudinal CovILD cohort and the Health after COVID-19 survey study.'}
+```{r fig-consort, fig.width = figur::convert(suppl$consort, to = 'in')$w, fig.height = figur::convert(suppl$consort, to = 'in')$h, fig.cap = 'Flow diagram of the analysis inclusion process for the observational CovILD cohort and the Health after COVID-19 survey study.'}
 
@@
 
-__Supplementary Figure S\@ref(fig:fig-consort). Flow diagram of the analysis inclusion process for the longitudinal CovILD cohort and the Health after COVID-19 survey study.__
+__Supplementary Figure S\@ref(fig:fig-consort). Flow diagram of the analysis inclusion process for the observational CovILD cohort and the Health after COVID-19 survey study.__
 
@@
 
-```{r fig-hact-freq, fig.width = mm_inch(suppl$hact_freq$w), fig.height = mm_inch(suppl$hact_freq$h), fig.cap = 'Frequency of COVID-19 symptoms in the survey study.'}
+```{r fig-sample-size, fig.width = figur::convert(suppl$sample_size, to = 'in')$w, fig.height = figur::convert(suppl$sample_size, to = 'in')$h, fig.cap = 'Estimation of sample size for clustering analysis with the survey study datasets.'}
+
+suppl$sample_size$plot
+
+```
+
+__Supplementary Figure S\@ref(fig:fig-sample-size). Estimation of sample size for clustering analysis with the survey study datasets.__
+
+_To assess the minimal sample size of the survey study for clustering, random subsets of the pooled Austria (AT) and Italy (IT) COVID-19 symptom data set with differing observation numbers (50, 100, 200, 300, 400, 500, 600, 700, 800, 900 observations, 20 random subsets [draws] per sample size) were generated and their clustering tendency was assessed with Hopkins statistic (H).
+Median H values per sample size with interquartile ranges (IQR) are visualized as boxes.
+Whiskers span over the 150% IQR.
+Single H values are depicted as points.
+Blue line with gray ribbon represents the LOESS trend with 95% confidence interval.
+Dashed lines represent sample sizes of the AT (blue) and IT cohort (blue).
+Sample sizes and H values for the AT and IT collectives are displayed in the plot.
+Note: beginning from the sample size of n = 400, no improvement of the clustering tendency could be observed.
+This suggest n = 400 as minimal sample size required for reproducible clustering analysis results._
+
+\newpage
+
+```{r fig-hact-differences, fig.width = figur::convert(suppl$hact_differences, to = 'in')$w, fig.height = figur::convert(suppl$hact_differences, to = 'in')$h, fig.cap = 'The largest significant differences in demographic, clinical and recovery variables between the Austria and Italy cohorts of the survey study.'}
+
+suppl$hact_differences$plot
+
+```
+
+__Supplementary Figure S\@ref(fig:fig-hact-differences).
+The largest significant differences in demographic, clinical and recovery variables between the Austria and Italy cohorts of the survey study.__
+
+_Differences in numeric variables between the Austria (AT) and Italy (IT) survey study cohorts were assessed by Mann-Whitney test with r effect size statistic.
+Differences in categorical variables were investigated by $\chi^2$ test with Cramer's V effect size statistic.
+P values were corrected for multiple testing with Benjamini-Hochberg method.
+Significant numeric variables are presented in violin plots with medians and interquartile ranges depicted as diamonds and whiskers and single observations visualized as points.
+Percentages of categories of qualitative variables within the AT and IT cohorts are displayed as stack plots.
+Effect sizes and p values are presented in the plot captions.
+Numbers of complete observations are indicated in the plot axes._
+
+_BMI before COVID-19: body mass index before COVID-19, normal: BMI < 25 kg/m\u00B2, overweight: BMI 25 - 30 kg/m\u00B2, obesity: BMI > 30 kg/\u00B2; Freq.
+bact.
+Infections: frequent (> 2 per year) bacterial infection requiring an antibiotic treatment; DPR score: depression score, Patient Health Questionnaire, PHQ-4; ANX score: anxiety score, Patient Health Questionnaire, PHQ-4; QoL impairment score: score of impairment of quality of life._
+
+\newpage
+
+```{r fig-hact-freq, fig.width = figur::convert(suppl$hact_freq, to = 'in')$w, fig.height = figur::convert(suppl$hact_freq, to = 'in')$h, fig.cap = 'Frequency of COVID-19 symptoms in the survey study.'}
 
@@
 
-_Frequency of symptoms in first 14 days, $\geq$ 28 days and $\geq$ 3 months after clinical onset of COVID-19 in the Austria (AT, A) and Italy (IT, B) survey study cohorts expressed as percentages of the cohort.
+_Frequency of symptoms in first 14 days, at 28 days and at three months after clinical onset of COVID-19 in the Austria (AT) and Italy (IT) survey study cohorts expressed as percentages of the cohort.
 Point size and color represents the percentage.
-Numbers of complete observations are indicated under the plot._
+Numbers of complete observations are indicated in the plot captions._
 
-_OD: self-reported olfactory dysfunction, Dim. appetite: diminished appetite, Imp. concentration: impaired concentration, Imp. walk: impaired walk, Imp. FMS: impaired fine motor skills._
+_OD: self-reported olfactory dysfunction; Dim. appetite: diminished appetite; Imp. concentration: impaired concentration; Imp. walk: impaired walk; Imp. FMS: impaired fine motor skills._
 
@@
 
-```{r fig-hact-recovery, fig.width = mm_inch(suppl$hact_recovery$w), fig.height = mm_inch(suppl$hact_recovery$h), fig.cap = 'Kinetic of recovery from leading acute COVID-19 symptoms in the survey study.'}
+```{r fig-hact-signif-sympt, fig.width = figur::convert(suppl$hact_signif_sympt, to = 'in')$w, fig.height = figur::convert(suppl$hact_signif_sympt, to = 'in')$h, fig.cap = 'Significant differences in frequency of COVID-19 symptoms in the survey study.'}
+
+suppl$hact_signif_sympt$plot
+
+```
+
+__Supplementary Figure S\@ref(fig:fig-hact-signif-sympt). Significant differences in frequency of COVID-19 symptoms in the survey study.__
+
+_Differences in frequency of COVID-19-related symptoms between the Austria (AT) and Italy (IT) cohorts of the survey study in the first 14 days, at 28 days and at 3 months after clinical onset were assessed by $\chi^2$ test with Cramer's V effect size statistic.
+P values were adjusted for multiple testing with Benjamini-Hochberg method.
+Significant effects for the 14 and 28 day time points are plotted.
+No significant differences in symptom frequency at the 3-month time point could be observed.
+Symptom percentages within the cohort are displayed in bar plots.
+Effect sizes and p values are displayed in the Y axes.
+Numbers of complete observations are shown in the plot captions._
+
+\newpage
+
+```{r fig-hact-recovery, fig.width = figur::convert(suppl$hact_recovery, to = 'in')$w, fig.height = figur::convert(suppl$hact_recovery, to = 'in')$h, fig.cap = 'Kinetic of recovery from leading acute COVID-19 symptoms in the survey study.'}
 
@@
 
-_Percentages of individuals with fever (A), diminished appetite (B), joint pain (C), muscle pain (D), fatigue (E) and tachypnea (F) in the AT (Austria) and IT (Italy) survey study cohorts at particular time points after clinical onset.
+_Percentages of individuals with fever (a), diminished appetite (b), joint pain (c), muscle pain (d), fatigue (e) and tachypnea (f) in the AT (Austria) and IT (Italy) survey study cohorts at particular time points after clinical onset.
 Numbers of complete observations are indicated under the plots._
@@
 
-```{r fig-covild-freq, fig.width = mm_inch(suppl$covild_freq$w), fig.height = mm_inch(suppl$covild_freq$h), fig.cap = 'Symptom frequency in ambulatory, moderate and severe COVID-19 subsets of the CovILD study.'}
+```{r fig-covild-freq, fig.width = figur::convert(suppl$covild_freq, to = 'in')$w, fig.height = figur::convert(suppl$covild_freq, to = 'in')$h, fig.cap = 'Symptom frequency in ambulatory, moderate and severe COVID-19 subsets of the CovILD study.'}
 
@@
 
-_Frequency of symptoms during acute COVID-19 and at the 60-, 100-, 180- and 360-day follow-ups in ambulatory (A), moderate (B) and severe COVID-19 (C) participants expressed as percentages of individuals with the complete longitudinal data set.
+_Frequency of symptoms during acute COVID-19 and at the 60-, 100-, 180- and 360-day follow-ups in ambulatory, moderate and severe COVID-19 participants expressed as percentages of individuals with the complete longitudinal data set.
 Point size and color represents the percentage.
@@
 
-```{r fig-covild-recovery, fig.width = mm_inch(suppl$covild_recovery$w), fig.height = mm_inch(suppl$covild_recovery$h), fig.cap = 'Kinetic of recovery from olfactory dysfunction, reduced performance and dyspnea in ambulatory, moderate and severe COVID-19 subsets of the CovILD study.'}
+```{r fig-covild-recovery, fig.width = figur::convert(suppl$covild_recovery, to = 'in')$w, fig.height = figur::convert(suppl$covild_recovery, to = 'in')$h, fig.cap = 'Kinetic of recovery from olfactory dysfunction, reduced performance and dyspnea in ambulatory, moderate and severe COVID-19 subsets of the CovILD study.'}
 
@@
 
-_Percentages of individuals with the complete longitudinal data set suffering from olfactory dysfunction (OD) (A), reduced physical performance (B) and dyspnea (C) in the ambulatory, moderate and severe COVID-19 subsets during acute COVID-19 and at the 60-, 100-, 180- and 360-day follow-ups.
+_Percentages of individuals with the complete longitudinal data set suffering from olfactory dysfunction (OD) (a), reduced physical performance (b) and dyspnea (c) in the ambulatory, moderate and severe COVID-19 subsets during acute COVID-19 and at the 60-, 100-, 180- and 360-day follow-ups.
 Numbers of complete observations are indicated under the plots._
@@
 
-```{r fig-hyposmia-rater100, fig.width = mm_inch(suppl$hyposmia_rater100$w), fig.height = mm_inch(suppl$hyposmia_rater100$h), fig.cap = 'Rates of self-reported olfactory dysfunction and olfactory dysfunction in the sniffing stick test at 3-month post COVID-19 follow-up in the ambulatory, moderate and severe COVID-19 subsets of the CovILD study.'}
+```{r fig-hyposmia-rater100, fig.width = figur::convert(suppl$hyposmia_rater100, to = 'in')$w, fig.height = figur::convert(suppl$hyposmia_rater100, to = 'in')$h, fig.cap = 'Rates of self-reported olfactory dysfunction and olfactory dysfunction in the sniffing stick test at 3-month post COVID-19 follow-up in the ambulatory, moderate and severe COVID-19 subsets of the CovILD study.'}
 
@@
 __Supplementary Figure S\@ref(fig:fig-hyposmia-rater100).
-Rates of self-reported olfactory dysfunction and olfactory dysfunction in the sniffing stick test at 3-month post COVID-19 follow-up in the ambulatory, moderate and severe COVID-19 subsets of the CovILD study.__
+Rates of self-reported olfactory dysfunction and olfactory dysfunction in the Sniffin' Stick Test at 3-month post COVID-19 follow-up in the ambulatory, moderate and severe COVID-19 subsets of the CovILD study.__
 
-_Association of self-reported and sniffing stick test rates of olfactory dysfunction (OD) was investigated with Cohen's $\kappa$ statistic.
-Statistical significance was assessed with Wald's Z test corrected for multiple testing with Benjamini-Hochberg method.
-Percentages of individuals with self-reported and test OD within the entire cohort (A), the ambulatory (B), moderate (C) and severe (D) COVID-19 subsets are presented in bubble plots.
-Point size and color represents the percentage.
-$\kappa$ values with 95$\%$ confidence intervals, p values and numbers of complete observations are indicated in the plot captions._
+_Objective olfactory dysfunction (OD) was diagnosed in CovILD study participants with < 13 correctly identified odorants in the 16-item Sniffin' Sticks Identification Test.
+Frequencies of objective and self-reported OD were compared at the two-month follow-up after COVID-19 in the entire cohort and the ambulatory, hospitalized moderate COVID-19 and hospitalized severe COVID-19 patients._
+
+_(a) Rates of objective (test) and self-reported OD presented in heat maps of confusion matrices.
+The overall concordance between the objective and subjective OD was assessed with Cohen's $\kappa$ inter-rater reliability statistic.
+Significance of $\kappa$ was determined by Wald's Z test corrected for multiple testing with Benjamini-Hochberg method.
+$\kappa$ values with 95% confidence intervals, p values and numbers of complete observations are displayed in the plot captions._
+
+_(b) Reliability of detection of objective OD by self-reported OD was assessed by receiver-operating characteristic (ROC).
+Sensitivity and specificity was visualized as ROC curves.
+numbers of complete observations, percentages of objective (test) and self-reported OD are displayed in the plot captions.
+Overall accuracy, sensitivity and specificity were are shown in the plots._
 
@@
 
-```{r fig-hyposmia-rater360, fig.width = mm_inch(suppl$hyposmia_rater360$w), fig.height = mm_inch(suppl$hyposmia_rater360$h), fig.cap = 'Rates of self-reported olfactory dysfunction and olfactory dysfunction in the sniffing stick test at 1-year post COVID-19 follow-up in the ambulatory, moderate and severe COVID-19 subsets of the CovILD study.'}
+```{r fig-hyposmia-rater360, fig.width = figur::convert(suppl$hyposmia_rater360, to = 'in')$w, fig.height = figur::convert(suppl$hyposmia_rater360, to = 'in')$h, fig.cap = "Rates of self-reported olfactory dysfunction and olfactory dysfunction in the Sniffin' Stick Test at 1-year post COVID-19 follow-up in the ambulatory, moderate and severe COVID-19 subsets of the CovILD study."}
 
@@
 __Supplementary Figure S\@ref(fig:fig-hyposmia-rater360).
-Rates of self-reported olfactory dysfunction and olfactory dysfunction in the sniffing stick test at 1-year post COVID-19 follow-up in the ambulatory, moderate and severe COVID-19 subsets of the CovILD study.__
+Rates of self-reported olfactory dysfunction and olfactory dysfunction in the Sniffin' Stick Test at 1-year post COVID-19 follow-up in the ambulatory, moderate and severe COVID-19 subsets of the CovILD study.__
 
-_Association of self-reported and sniffing stick test rates of olfactory dysfunction (OD) was investigated with Cohen's $\kappa$ statistic.
-Statistical significance was assessed with Wald's Z test corrected for multiple testing with Benjamini-Hochberg method.
-Percentages of individuals with self-reported and test OD within the entire cohort (A), the ambulatory (B), moderate (C) and severe (D) COVID-19 subsets are presented in bubble plots.
-Point size and color represents the percentage.
-$\kappa$ values with 95$\%$ confidence intervals, p values and numbers of complete observations are indicated in the plot captions._
+_Objective olfactory dysfunction (OD) was diagnosed in CovILD study participants with < 13 correctly identified odorants in the 16-item Sniffin' Sticks Identification Test.
+Frequencies of objective and self-reported OD were compared at the one-year follow-up after COVID-19 in the entire cohort and the ambulatory, hospitalized moderate COVID-19 and hospitalized severe COVID-19 patients._
+
+_(a) Rates of objective (test) and self-reported OD presented in heat maps of confusion matrices.
+The overall concordance between the objective and subjective OD was assessed with Cohen's $\kappa$ inter-rater reliability statistic.
+Significance of $\kappa$ was determined by Wald's Z test corrected for multiple testing with Benjamini-Hochberg method.
+$\kappa$ values with 95% confidence intervals, p values and numbers of complete observations are displayed in the plot captions._
+
+_(b) Reliability of detection of objective OD by self-reported OD was assessed by receiver-operating characteristic (ROC).
+Sensitivity and specificity was visualized as ROC curves.
+numbers of complete observations, percentages of objective (test) and self-reported OD are displayed in the plot captions.
+Overall accuracy, sensitivity and specificity were are shown in the plots._
 
@@
 
-```{r fig-mds-acute, fig.width = mm_inch(suppl$mds_acute$w), fig.height = mm_inch(suppl$mds_acute$h), fig.cap = 'Multi-dimensional scaling analysis of acute COVID-19 symptoms in the survey study.'}
+```{r fig-od-kinetic, fig.width = figur::convert(suppl$od_kinetic, to = 'in')$w, fig.height = figur::convert(suppl$od_kinetic, to = 'in')$h, fig.cap = 'Individuals trajectories of objective olfactory dysfunction in the CovILD study subset with the complete longitudinal follow-up data.'}
+
+suppl$od_kinetic$plot
+
+```
+
+__Supplementary Figure S\@ref(fig:fig-od-kinetic). Individuals trajectories of objective olfactory dysfunction in the CovILD study subset with the complete longitudinal follow-up data.__
+
+_Objective olfactory dysfunction (OD) assessed in CovILD study participants with the Sniffin' Stick Identification test at the three-month and one-year follow-up after COVID-19.
+Objective olfactory dysfunction (OD) was diagnosed for < 13 correctly identified odorants in the 16-item Sniffin' Sticks Identification Test.
+A subset of the CovILD study participant with the complete Sniffin' Stick Test data for both follow-ups was analyzed.
+_
+
+_(a) Comparison of numeric Sniffin' Stick Test results was done with paired Wilcoxon test with r effect size statistic.
+Results are presented as a before - after plot (left) and box plot (right).
+Medians with interquartile ranges (IQR) are visualized as boxes, whiskers span over 150% IQR.
+Single observations are depicted as points.
+Observations obtained for the same participant are connected with a line.
+Effect sizes, p values are displayed in the plot caption.
+Numbers of complete observations are indicated i the X axis._
+
+_(b) Comparison of frequencies of objective OD expressed as paired proportions was done with McNemar test with Cohen's q effect size statistic.
+Percentages of participants with objective OD at the three-month and one-year follow-up are displayed in an alluvial plot.
+The odds ratio of objective OD (one-year vs three months), effect size, p value and the number of participants are displayed in the plot caption._
+
+\newpage
+
+```{r fig-mds-acute, fig.width = figur::convert(suppl$mds_acute, to = 'in')$w, fig.height = figur::convert(suppl$mds_acute, to = 'in')$h, fig.cap = 'Multi-dimensional scaling analysis of acute COVID-19 symptoms in the survey study.'}
 
@@
 __Supplementary Figure S\@ref(fig:fig-mds-acute). Multi-dimensional scaling analysis of acute COVID-19 symptoms in the survey study.__
+
 _Symptom data for acute COVID-19 (first 14 days after clinical onset) in the Austria (AT) and Italy (IT) survey study cohorts were subjected to two-dimensional multi-dimensional scaling (MDS) with simple matching distance (SMD) between the symptoms.
@@
 Percentages of the data set variance associated with the MDS dimensions are indicated in the plot axes.
-Numbers of complete observations are indicated under the plots.
-OD: self-reported olfactory dysfunction._
+Numbers of complete observations are indicated in the plot captions._
+
+_Dim. appetite: diminished appetite; OD: self-reported olfactory dysfunction._
 
@@
 
-```{r fig-clust-devel, fig.width = mm_inch(suppl$clust_devel$w), fig.height = mm_inch(suppl$clust_devel$h), fig.cap = 'Definition of the COVID-19 recovery clusters and clustering feature importance in the survey study.'}
+```{r fig-apriori, fig.width = figur::convert(suppl$apriori, to = 'in')$w, fig.height = figur::convert(suppl$apriori, to = 'in')$h, fig.cap = 'Co-occurrence of self-reported olfactory dysfunction and other symptoms in post-acute COVID-19 sequelae.'}
+
+suppl$apriori$plot
+
+```
+
+__Supplementary Figure S\@ref(fig:fig-apriori). Co-occurrence of self-reported olfactory dysfunction and other symptoms in post-acute COVID-19 sequelae.__
+
+_Frequent combinations (present in >15% of cohort participants) of self-reported olfactory dysfunction (OD) and other symptoms at 28 days (a) and 3 months (b) after clinical onset in the Austria (AT) and Italy (IT) survey study cohorts were identified with the apriori algorithm.
+Symptom combination frequency and co-occurrence (support statistic) are presented in bubble plots.
+Point size and color corresponds to co-occurrence, points are labeled with percentages of co-occurrence.
+Imp.
+concentration: impaired concentration._
+
+\newpage
+
+```{r fig-clust-devel, fig.width = figur::convert(suppl$clust_devel, to = 'in')$w, fig.height = figur::convert(suppl$clust_devel, to = 'in')$h, fig.cap = 'Definition of the COVID-19 recovery clusters and clustering feature importance in the survey study.'}
 
@@
 
-_Individuals of the training Austria (AT) study survey cohort were subjected to clustering in respect to symptom-specific recovery times with the PAM (partitioning around medoids) algorithm and Euclidean distance measure._
+_Individuals of the training Austria (AT) study survey cohort were clustered in respect to symptom-specific recovery times with the PAM (partitioning around medoids) algorithm and Euclidean distance measure._
 
-_(A) Comparison of performance of various algorithms (HCl: hierarchical clustering, SOM + HCl: combined self-organizing map and hierarchical clustering, k-means) and distance statistic in clustering of the training data set investigated by clustering variance (ratio of total between-cluster sum of squares to total sum of squares) and cluster assignment accuracy in 10-fold cross-validation (CV)._
+_(a) Comparison of performance of various algorithms (HCl: hierarchical clustering, SOM + HCl: combined self-organizing map and hierarchical clustering, k-means) and distance statistic in clustering of the training data set investigated by clustering variance (ratio of total between-cluster sum of squares to total sum of squares) and cluster assignment accuracy in 10-fold cross-validation (CV)._
 
-_(B) Determination of the optimal cluster number in the PAM clustering of the training cohort by the bend of the total within-cluster sum of squares curve._
+_(b) Determination of the optimal cluster number in the PAM clustering of the training cohort by the bend of the total within-cluster sum of squares curve._
 
-_(C) Permutation importance of the clustering features (symptoms) for clustering of the training cohort expressed as the difference in clustering variance (ratio of total between-cluster sum of squares to total sum of squares) between the initial clustering object and the clustering object with the given variable reshuffled at random.
-Imp.
-concentration: impaired concentration, OD: self-reported olfactory dysfunction, Dim.
-appetite: diminished appetite, Imp.
-walk: impaired walk, Imp.
-FMS: impaired fine motor skills._
+_(c) Permutation importance of the clustering features (symptoms) for clustering of the training cohort expressed as the difference in clustering variance (ratio of total between-cluster sum of squares to total sum of squares) between the initial clustering object and the clustering object with the given variable reshuffled at random.
+Importance metrics were computed for 20 random permutations of each clustering factor.
+Median importance metrics with interquartile ranges (IQR) are visualized as boxes, whiskers span over 150% IQR._
+
+_OD: self-reported olfactory dysfunction; Dim. appetite: diminished appetite; Imp. concentration: impaired concentration; Imp. walk: impaired walk; Imp. FMS: impaired fine motor skills._
 
@@
 
-```{r fig-clustering, fig.width = mm_inch(suppl$clustering$w), fig.height = mm_inch(suppl$clustering$h), fig.cap = 'Clustering of ambulatory COVID-19 individuals by symptom-specific recovery times.'}
+```{r fig-clustering, fig.width = figur::convert(suppl$clustering, to = 'in')$w, fig.height = figur::convert(suppl$clustering, to = 'in')$h, fig.cap = 'Clustering of ambulatory COVID-19 individuals in the survey study by symptom-specific recovery times.'}
 
@@
 
-__Supplementary Figure S\@ref(fig:fig-clustering). Clustering of ambulatory COVID-19 individuals by symptom-specific recovery times.__
+__Supplementary Figure S\@ref(fig:fig-clustering). Clustering of ambulatory COVID-19 individuals in the survey study by symptom-specific recovery times.__
 
 _Individuals of the training Austria (AT) survey study cohort were subjected to clustering in respect to symptom-specific recovery times with the PAM (partitioning around medoids) algorithm and Euclidean distance measure (Supplementary Figure S\@ref(fig:fig-clust-devel)).
-Cluster assignment in the test Italy (IT) survey cohort was done with k-NN label propagation algorithm.
-Recovery times for particular COVID-19 symptoms in the COVID-19 recovery clusters are presented as heat maps.
-Numbers of individuals assigned to the recovery clusters are indicated under the plots._
+Cluster assignment in the test Italy (IT) survey cohort was done with an inverse weighted 7-nearest neighbor (7-NN) classification algorithm._
 
-_OD: self-reported olfactory dysfunction, Dim. appetite: diminished appetite, Imp. concentration: impaired concentration, Imp. walk: impaired walk, Imp. FMS: impaired fine motor skills._
+_(a) Fraction of explained clustering variance was computed as a ratio of total between-cluster sum of squares to total sum of squares for the clustering structures in the training AT and test IT cohort.
+Note similar fractions of explained clustering variances in both cohorts, which suggests good reproducibility of the clustering structure._
+
+_(b) Percentages of observations assigned to clusters in the training AT and test IT cohort. Total numbers of complete observations are displayed in the Y axis._
+
+_(c) Recovery times for particular COVID-19 symptoms in the COVID-19 recovery clusters presented as heat maps.
+Numbers of individuals assigned to the recovery clusters are indicated in the plot captions._
+
+_OD: self-reported olfactory dysfunction; Dim. appetite: diminished appetite; Imp. concentration: impaired concentration; Imp. walk: impaired walk; Imp. FMS: impaired fine motor skills._
 
@@
 
-```{r fig-cov-clusters, fig.width = mm_inch(suppl$cov_clusters$w), fig.height = mm_inch(suppl$cov_clusters$h), fig.cap = 'Numbers of COVID-19 symptoms in the survey study clusters.'}
+```{r fig-cov-clusters, fig.width = figur::convert(suppl$cov_clusters, to = 'in')$w, fig.height = figur::convert(suppl$cov_clusters, to = 'in')$h, fig.cap = 'Numbers of COVID-19 symptoms in the survey study recovery clusters.'}
 
@@
 
-__Supplementary Figure S\@ref(fig:fig-cov-clusters). Numbers of COVID-19 symptoms in the survey study clusters.__
+__Supplementary Figure S\@ref(fig:fig-cov-clusters). Numbers of COVID-19 symptoms in the survey study recovery clusters.__
 
 _Clustering of the survey study participants in respect to symptom-specific recovery times was done by the semi-supervised PAM algorithm (partitioning around medoids, Euclidean distance, training cohort: Austria [AT], test cohort: Italy [IT]).
-Differences in numbers of symptoms in the first 14 days (A) and at $\geq$ 28 days (B) after clinical onset between the clusters were assessed by Kruskal-Wallis test and $\eta^2$ effect size statistic.
+Differences in numbers of symptoms in the first 14 days (a) and at 28 days (b) after clinical onset between the clusters were assessed by Kruskal-Wallis test and $\eta^2$ effect size statistic.
 P values were corrected for multiple testing with Benjamini-Hochberg method.
@@
 Points represent single observations, orange diamonds with whiskers code for medians and interquartile ranges.
-Effect size statistics and p values are indicated in the plot caption.
-Numbers of complete observations are displayed under the plots._
+Effect sizes and p values are indicated in the plot caption.
+Numbers of complete observations are displayed in the X axes._
 
@@
 
-```{r fig-base-clusters, fig.width = mm_inch(suppl$base_clusters$w), fig.height = mm_inch(suppl$base_clusters$h), fig.cap = 'COVID-19 recovery clusters differ in sex distribution, comorbidity and daily medication rates.'}
+```{r fig-base-clusters, fig.width = figur::convert(suppl$base_clusters, to = 'in')$w, fig.height = figur::convert(suppl$base_clusters, to = 'in')$h, fig.cap = 'COVID-19 recovery clusters of the survey study differ in age, sex distribution, comorbidity and daily medication rates.'}
 
@@
 
-__Supplementary Figure S\@ref(fig:fig-base-clusters). COVID-19 recovery clusters differ in sex distribution, comorbidity and daily medication rates.__
+__Supplementary Figure S\@ref(fig:fig-base-clusters). COVID-19 recovery clusters of the survey study differ in age, sex distribution, comorbidity and daily medication rates.__
 
 _Clustering of the survey study participants in respect to symptom-specific recovery times was done by the semi-supervised PAM algorithm (partitioning around medoids, Euclidean distance, training cohort: Austria [AT], test cohort: Italy [IT]).
-Differences in sex distribution (A), frequency of comorbidity (B) and daily medication (C) between the recovery clusters were assessed by $\chi^2$ test with Cramer V effect size statistic.
+Differences in age (a), sex distribution (b), frequency of comorbidity (c) and daily medication (d) between the recovery clusters were assessed by Kruskal-Wallis test with $\eta^2$ effect size statistic (age) and $\chi^2$ test with Cramer V effect size statistic (remaining variables).
 P values were corrected for multiple testing with Benjamini-Hochberg method.
 The frequencies are presented as bar plots.
-Effect size statistics and p values are indicated in the plot caption.
-Numbers of complete observations are displayed under the plots._
+Effect sizes and p values are indicated in the plot caption.
+Numbers of complete observations are displayed in the X axes._
 

### 2023-07-19 paper scripts/manuscript.R
Second revision, code

@@ -26,11 +26,13 @@
                                         reference_docx = 'ms_template.docx'), 
          output_dir = './paper')
   
-# Rendering the rebuttal letter -------
+# Rendering the rebuttal letters -------
   
-  insert_msg('Rendering the rebuttal letter')
+  insert_msg('Rendering the rebuttal letters')
   
-  render('./paper/markdown/rebuttal_letter.Rmd', 
+  c('./paper/markdown/rebuttal_letter.Rmd', 
+    './paper/markdown/rebuttal_letter2.Rmd') %>% 
+    walk(render, 
          output_format = word_document2(number_sections = FALSE, 
                                         reference_docx = 'ms_template.docx'), 
          output_dir = './paper')

### 2023-07-19 paper/markdown/rebuttal_letter2.Rmd
Second revision, results

---
title: "Distinct smell and taste disorder phenotype of post-acute COVID-19 sequelae"
subtitle: "Rebuttal Letter, 2^nd^ revision"
author: "Health after COVID-19 in Tyrol and CovILD study teams"
date: "`r format(Sys.time(), '%Y-%m-%d')`"

output: 
  bookdown::word_document2:   
    reference_docx: ms_template.docx

bibliography: hyposmia_biblio.bib
csl: european-archives-of-oto-rhino-laryngology.csl
header-includes:
  \usepackage{longtable}
  \usepackage{tabu}
  \usepackage{caption}
  \usepackage{makecell}
  \usepackage{pdflscape}
  \usepackage{array}
  \usepackage{booktabs}
  \usepackage{threeparttable}
  \usepackage{threeparttablex}
  \usepackage{wrapfig}
  \usepackage{multirow}
  \usepackage[normalem]{ulem}
  \usepackage{colortbl}
  \usepackage{xcolor}
  \usepackage{float} \floatplacement{figure}{H} \floatplacement{table}{H}
  \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}}       \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
---

```{r, setup, include = FALSE}

library(bookdown)

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, 
                      dev = "png", dpi = 600, 
                      tab.cap.style = 'Table Heading')

set_flextable_defaults(font.family = 'Cambria', font.size = 10)

```

\newpage

# Reviewer #1

<span custom-style = 'reviewer'>Thank you for the revised version of the above mentioned manuscript "Distinct smell and taste disorder phenotype of post-acute COVID-19 sequelae". I feel that the revision has helped greatly in improvising the manuscript.</span>

We thank the Reviewer for appreciation of our work and the constructive feedback. 
In the revised text, we addressed the issue concerning the Sniffin' Stick methodology, its strengths, limitation and comparison with other, more popular methods like UPSIT. 
Changes in the main manuscript text introduced during this second round of revision are highlighted in blue.

## Issue 1

<span custom-style = 'reviewer'>There is one point that I feel that still needs to be addressed: The Limitations (and potential strengths) of the Sniffin Sticks. Sniffin Sticks are not an international standard, while the UPSIT has been validated in huge collectives. Nonetheless, the Sniffin Sticks offer dimensions that cannot be addressed with single use test kits like the UPSIT (Discrimination and Thresholds). However, these were not included in the study at hand. This needs clarification and I feel that this has not been addressed appropriately in this revision.</span>

We agree with the Reviewer that among the number of available tools for testing of the olfactory function, the UPSIT (University of Pennsylvania Smell Identification Test) is the one with the best age- and sex-specific normative data [@Doty1984a; @Doty2007] and an excellent test-retest reliability [@Doty1989]. 
It serves also as the 'gold standard' especially in North America and English-speaking countries [@Doty2007]. 
Still, the Sniffin' Stick tool and in particular its 'sub-tool' used here, the Sniffin' Stick Identification test [@Hummel1997], is popular in Europe and especially in German-speaking countries. 
Its features are good test-retest reliability (r = 0.73) and increasing amount of normative data for age groups and genders [@Hummel1997; @Hummel2007; @Oleszkiewicz2019a]. 
Interestingly, 13 out of 16 smells of the Sniffin' Stick Identification Test are common with the 40-item UPSIT. 
This allowed for development of robust conversion tools between these two tools as reported by Lawton and colleagues [@Lawton2016]. 
As noted by Lawton et al., results of the Sniffin' Stick Identification Test and UPSIT were generally highly concordant in a cohort of Parkinson disease patients (r = 0.81) and the olfactory function tended to be rated slightly better with the UPSIT than with the Sniffin' Stick Identification test [@Lawton2016]. 
However, the larger set of test items in the UPSIT (40 smells) as compared with the Sniffin' Stick Identification test (16 smells) may amount to a better sensitivity and test-retest reliability of the UPSIT tool [@Doty2007; @Lawton2016].

To address this methodological issue, we introduced several changes to the revised manuscript:

* We state clearly in __Methods__ and __Supplementary Methods__, that the discrimination and threshold dimensions of the Sniffin' Stick test were not evaluated in the CovILD cohort

* We indicate in __Methods__ and __Supplementary Methods__, that the manufacturer's cutoff for definition of hyposmia applied to the CovILD cohort corresponded approximately to the 25^th^ percentile of correct answers in pooled healthy females and males reported by two large normative studies [@Hummel2007; @Oleszkiewicz2019a]. Accordingly, we state in the __Discussion__ section, that as per definition of hyposmia with the 13 point cutoff, 25% of individuals without any subjective smell impairment would be diagnosed with objective olfactory dysfunction.

* We indicate the significant enrichment of males and elderly participants among the severe COVID-19 subset of the CovILD cohort another factor that may explain the large discrepancy between the subjective and objective olfactory dysfunction in this group [@Doty1984; @Hummel2007; @Oleszkiewicz2019a].

* Furthermore, we state in the main part __Discussion__, that due to the lacking widely accepted definition of objective olfactory dysfunction, its prevalence vary substantially [@Kanavou2021]. In particular, the discrepancy between subjective and objective olfactory dysfunction in the CovILD cohort and, especially, in the severe COVID-19 subset, may be addressed by age- and sex-corrected cutoffs of the Sniffin' Stick Identification Test [@Hummel2007; @Oleszkiewicz2019a].

* We point out in the limitations of out study listed in the final paragraph of __Discussion__, that because of methodological differences and varying definitions, caution is required while comparing the rates of objective OD in the CovILD cohort with other studies.

* In __Discussion__, we compare the Sniffin' Stick Identification Test and UPSIT by pointing out the similarities (forced-choice, shared smells) and inter-convertibility [@Lawton2016] as well as higher sensitivity and reliability of the UPSIT [@Lawton2016; @Doty2007].

\newpage

# References

### 2023-07-19 paper/markdown/supplementary_material.Rmd
Second revision, results

@@
 
-### Rating of olfactory dysfunction with Sniffin' Stick Test
+### Rating of olfactory dysfunction with Sniffin' Stick Identification Test
 
-Objective olfactory dysfunction at the 100-day and 360-day follow-up in the CovILD study participants was investigated with the 16-item Sniffn' Stick Identification Test as described [@Rass2021; @Whitcroft2016; @Masala2022; @Damm2014].
+Objective olfactory dysfunction at the 100-day and 360-day follow-up in the CovILD study participants was investigated with the 16-item Sniffn' Stick Identification Test as described [@Hummel1997; @Rass2021; @Whitcroft2016; @Masala2022; @Damm2014].
+Note that the threshold and discrimination dimensions of the Sniffin' Stick test [@Hummel1997] were not investigated. 
 In brief, the nasal chemosensory performance was investigated using pen-like odor-dispensing devices for odor identification of 16 common odorants (multiple forced‐choice from four verbal items per test odorant).
 Clinically relevant olfactory dysfunction was defined as < 13 correct answers (points) [@Rass2021; @Hummel2007]. 
+The 13 point cutoff of the identification test corresponded approximately to the 25^th^ percentile of correct answers in the identification test in healthy adult subjects of both sexes reported in two large normative studies [@Hummel2007; @Oleszkiewicz2019a].
 In the analysis, participants with the complete answers concerning self-reported olfactory dysfunction and complete test results were included.
@@
 
-```{r fig-hyposmia-rater100, fig.width = figur::convert(suppl$hyposmia_rater100, to = 'in')$w, fig.height = figur::convert(suppl$hyposmia_rater100, to = 'in')$h, fig.cap = 'Rates of self-reported olfactory dysfunction and olfactory dysfunction in the sniffing stick test at 3-month post COVID-19 follow-up in the ambulatory, moderate and severe COVID-19 subsets of the CovILD study.'}
+```{r fig-hyposmia-rater100, fig.width = figur::convert(suppl$hyposmia_rater100, to = 'in')$w, fig.height = figur::convert(suppl$hyposmia_rater100, to = 'in')$h, fig.cap = "Rates of self-reported olfactory dysfunction and olfactory dysfunction in the Sniffin' Stick Identification Test at 3-month post COVID-19 follow-up in the ambulatory, moderate and severe COVID-19 subsets of the CovILD study."}
 
@@
 __Supplementary Figure S\@ref(fig:fig-hyposmia-rater100).
-Rates of self-reported olfactory dysfunction and olfactory dysfunction in the Sniffin' Stick Test at 3-month post COVID-19 follow-up in the ambulatory, moderate and severe COVID-19 subsets of the CovILD study.__
+Rates of self-reported olfactory dysfunction and olfactory dysfunction in the Sniffin' Stick Identification Test at 3-month post COVID-19 follow-up in the ambulatory, moderate and severe COVID-19 subsets of the CovILD study.__
 
